### 2017-08-30 tex/paper.tex
Implement discretization of the log-normal distribution, add a draft for a paper

\documentclass[sigconf]{acmart}

\usepackage[utf8]{inputenc}
\usepackage{booktabs} % For formal tables
\usepackage{algorithmicx}
\usepackage{todonotes}
\usepackage{svg}

\newcommand*{\fullref}[1]{\hyperref[{#1}]{\autoref*{#1} (\nameref*{#1})}} % One single link

% Adds a "given" symbol (vertical bar) that rescales in height (like \left and \right)
\usepackage{mathtools}
\newcommand\givenbase[1][]{\:#1\lvert\:}
\let\given\givenbase
\newcommand\sgiven{\givenbase[\delimsize]}
\DeclarePairedDelimiterX\Basics[1](){\let\given\sgiven #1}
\newcommand\Average{E\Basics}

\newcommand\fnurl[2]{%
\href{#2}{#1}\footnote{\url{#2}}%
}

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{10.475/123_4}

% ISBN
\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El
  Paso, Texas USA} 
\acmYear{2017}
\copyrightyear{2017}

\acmPrice{15.00}

\begin{document}
\title{Let's Play Congestion Control}
\titlenote{Produces the permission block, and
  copyright information}
\subtitle{Extended Abstract}
\subtitlenote{The full version of the author's guide is available as
  \texttt{acmart.pdf} document}

\author{Maximilian Bachl} 
\affiliation{%
 \institution{Institute of Telecommunications}
 \streetaddress{Gußhausstraße 25}
 \postcode{1040}
 \city{Vienna} 
 \country{Austria}}
\email{maximilian.bachl@tuwien.ac.at}

\author{Joachim Fabini} 
\affiliation{%
 \institution{Institute of Telecommunications}
 \streetaddress{Gußhausstraße 25}
 \postcode{1040}
 \city{Vienna} 
 \country{Austria}}
\email{joachim.fabini@tuwien.ac.at}

\author{Tanja Zseby} 
\affiliation{%
 \institution{Institute of Telecommunications}
 \streetaddress{Gußhausstraße 25}
 \postcode{1040}
 \city{Vienna} 
 \country{Austria}}
\email{tanja.zseby@tuwien.ac.at}

% The default list of authors is too long for headers}
\renewcommand{\shortauthors}{Bachl et al.}

\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings.\footnote{This is an abstract footnote}
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

\keywords{Put some comma-separated keywords here}

\maketitle

\section{Introduction}

In the following a time step $t$ corresponds to the reception of an acknowledgement. The beginning of the flow, before any packet is sent, corresponds to time step $0$.

\begin{description}
\item[$\textit{s}_\text{t}$] The state  describes the current ``congestion state''. Various features that indicate congestion are included in it. Each time an acknowledgement is received, the state is updated and the actor network is asked for the next action. 
\item[$\textit{a}_\text{t}$] Based on a given state and the history of previous states, the actor network returns an action $a_t$, which is an integer ($\geq 1$) that stands for the congestion window to be used until the next acknowledgement is received. 
\item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metrics. These are discussed in more detail in \fullref{subsec:value}
\item[$\textit{v}_\text{t}$] The value is a tuple of the rewards estimated by the value network (see \fullref{subsec:value}) for each of the reward metrics. 
\end{description}

\subsection{Value Network}
\label{subsec:value}

The value network estimates the expected reward given a state $s_t$.

Let $r_t$ be the reward that was received at time $t$. $V(s_t; \theta_v)$ designates the expected mean reward in state $s_t$ given the parameters of the value network $theta_v$. Then, with $\gamma$ being the roll-off factor, which designates the influence that future reward has on the moving average, (set to $0.99$), we generically define the expected n-step return as 
\begin{align*}
R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k \frac{V(s_{t+k}; \theta_v)}{1-\gamma}\right)\left(1-\gamma\right),
\end{align*}
where $k$ is upper-bounded by $t_\text{max}$. $R$ is simply a moving average of future rewards. However, usual moving averages take into account values from the past while this one uses values from the future; it runs reversely.

The loss function, which the value network tries to minimize, is the square of the difference of the actual mean reward received and the expected mean reward
\begin{align*}
l_{\text{v}_t} = \left(R_t - V(s_t; \theta_v)\right)^2.
\end{align*}

We actually use four moving averages for four different types of reward: 
\begin{description}
\item[$\textit{R}_\text{packet}$] is the moving average of the packets that the sender sent and that were not lost (so they were acknowledged at some point by the receiver)
\item[$\textit{R}_\text{byte}$] is the moving average of the bytes of the packets that the sender sent and that were not lost
\item[$\textit{R}_\text{delay}$] is the moving average of the round trip time of the packets that the sender sent and that were not lost
\item[$\textit{R}_\text{duration}$] is the moving average of the time between receiving the last packet and receiving this packet (``inter-receive time'') for the packets that the sender sent and that were not lost
\end{description}

The loss function $l_{v_t}$ of the value network is actually the sum of the square of the difference for each of these four moving averages:
\begin{align*}
l_{\text{v}_t} =& \left(R_{\text{packet}_t} - V_\text{packet}(s_t; \theta_v)\right)^2 \\
&+\left(R_{\text{byte}_t} - V_\text{byte}(s_t; \theta_v)\right)^2 \\
&+\left(R_{\text{delay}_t} - V_\text{delay}(s_t; \theta_v)\right)^2 \\
&+\left(R_{\text{duration}_t} - V_\text{duration}(s_t; \theta_v)\right)^2
\end{align*}

\subsection{Actor Network}
\label{subsec:actor}

The actor network outputs two parameters: The mean of a log-normal distribution $\mu_\text{log}$ and its standard deviation $\sigma_\text{log}$.

We calculate the means and standard deviation of the underlying normal distribution as
\begin{align*}
\sigma_\text{normal} &= \sqrt{\log \left( \frac{\sigma_\text{log}}{\mu_\text{log}} +1 \right)} \\
\mu_\text{normal} &= \log\left(\mu_\text{log}\right) - \frac{{\sigma_\text{normal}}^2}{2}.
\end{align*}

Each time an action $a_t$ is requested, a value $X$ is sampled from the current log-normal distribution defined by the parameters $\mu_\text{normal}$ and $\sigma_\text{normal}$: 
\begin{align*}
Z&\sim {\mathcal {N}}(\mu_\text{normal} ,{\sigma_{\text{normal}}}^{2})\\
X&=\exp(Z)
\end{align*}

We can also equally write
\begin{align*}
X \sim \text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right).
\end{align*}

The sampled value $X$ is then discretized by rounding up, which also prevents window sizes smaller than one from occurring:
\begin{align*}
W = \left\lceil X \right\rceil
\end{align*}

$a_t$ is the sampled value $W$ at time $t$ at it is used as the window until the $t+1$, which is the reception of the next acknowledgement.

Let
\begin{align*}
\Sigma R_t = \sum_{i \in \left \{\text{packet}, \text{byte}, \text{delay}, \text{duration}\right \}} R_{i_t}
\end{align*}
and
\begin{align*}
\Sigma v_t = \sum_{i \in \left \{\text{packet}, \text{byte}, \text{delay}, \text{duration}\right \}} v_{i_t}.
\end{align*}

With $H$ being the entropy and $S$ being the skewness, the actor network minimizes the loss
\begin{align*}
l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_a \right)\right)\left(\Sigma R_t - \Sigma v_t\right)\\ 
&+ \beta H\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right)\\
&- \delta S\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right).
\end{align*}

$\beta$ was set to $\frac 1 2 10^{-3}$ and $\delta$ to $10^{-3}$.

Reinforcement learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}). However, in the case of congestion control, it is desirable to perform a new action without having received reward for the previous on. For example, imagine the that you receive two acknowledgements directly after each other. For each of these two acknowledgement an action has to be performed but it does not seem feasible for the second action to wait until the first action has received a reward. The Asynchronous Actor Critic framework as described by \cite{mnih_asynchronous_2016} cannot be applied to congestion control as it assumes that actions and rewards are synchronized.

Thus we propose the Asynchronous Asynchronous Actor Critic framework in which there are two threads within each thread: one thread that generates actions upon receiving a state and one that receives rewards and updates the value and actor networks. \todo[inline]{Insert two algorithm environments that show what happens upon receiving actions and rewards}

\begin{figure}
\includesvg{figures/Reinforcement_learning_diagram}
\caption{The classic reinforcement learning approach.\protect\footnotemark}
\label{fig:reinforcement}
\end{figure}
\footnotetext{\url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}

\bibliographystyle{ACM-Reference-Format}
\bibliography{congestion}

\end{document}

### 2017-08-30 async_deep_reinforce/constants.py

@@ -15,7 +15,7 @@ PRECISION = tf.float32
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
 GAMMA = 0.99 # discount factor for rewards
 # ENTROPY_BETA = 1e-4 # entropy regurarlization constant (continuous actions)
-ENTROPY_BETA = 1e-3
+ENTROPY_BETA = .5e-3
 SKEWNESS_GAMMA = 1e-3
 MAX_TIME_STEP = 10 * 1e7
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
@@ -31,5 +31,4 @@ LAYER_NORMALIZATION = True
 ALPHA = 1.0
 BETA = 1.0
 LOG_LEVEL = logging.INFO
-OFFSET = 0.5
-MINIMUM_STD = 0.5
+MINIMUM_STD = 1e-2

### 2017-08-30 async_deep_reinforce/game_ac_network.py

@@ -10,7 +10,6 @@ from constants import PRECISION
 from constants import LAYER_NORMALIZATION
 from constants import ALPHA
 from constants import BETA
-from constants import OFFSET
 from constants import ENTROPY_BETA
 from constants import SKEWNESS_GAMMA
 from constants import MINIMUM_STD
@@ -70,14 +69,14 @@ class GameACNetwork(object):
 			self.distribution_std = self.pi[1]
 			self.inner_distribution_mean = self.distribution.distribution.mean()
 			self.inner_distribution_std = self.distribution.distribution.stddev()
-			self.chosen_action = tf.round(self.distribution.sample() + OFFSET)
+			self.chosen_action = tf.ceil(self.distribution.sample())
 
 			# policy entropy
 			self.entropy = ENTROPY_BETA * tf.reduce_sum(self.distribution.distribution.mean() + 0.5 * tf.log(2.0*math.pi*math.e*self.distribution.distribution.variance()), axis=1)
 			# self.entropy = ENTROPY_BETA * tf.reduce_sum(0.5 * tf.log(2.0*math.pi*math.e*self.pi[1]*self.pi[1]), axis=1)
 			self.skewness = SKEWNESS_GAMMA * tf.reduce_sum((tf.exp(self.distribution.distribution.variance()) + 2.0)*tf.sqrt(tf.exp(self.distribution.distribution.variance()) - 1.0), axis=1)
-			self.actor_loss = tf.reduce_sum(
-				self.distribution.log_prob(self.a - OFFSET), axis=1) * (self.td_throughput + self.td_delay)
+			# self.actor_loss = tf.reduce_sum(self.distribution.log_prob(self.a - OFFSET), axis=1) * (self.td_throughput + self.td_delay)
+			self.actor_loss = tf.reduce_sum(tf.log(self.distribution.cdf(self.a) - self.distribution.cdf(tf.clip_by_value(self.a - 1.0, tiny, float("inf")))), axis=1) * (self.td_throughput + self.td_delay)
 
 			self.policy_loss = - tf.reduce_sum(self.actor_loss + self.entropy - self.skewness)
 			# self.policy_loss = - tf.reduce_sum(self.actor_loss)

### 2017-08-31 tex/paper.tex
Parameter tuning

@@
 
-Let
-\begin{align*}
-\Sigma R_t = \sum_{i \in \left \{\text{packet}, \text{byte}, \text{delay}, \text{duration}\right \}} R_{i_t}
-\end{align*}
-and
-\begin{align*}
-\Sigma v_t = \sum_{i \in \left \{\text{packet}, \text{byte}, \text{delay}, \text{duration}\right \}} v_{i_t}.
-\end{align*}
+%Let
+%\begin{align*}
+%\Sigma R_t = \sum_{i \in \left \{\text{packet}, \text{byte}, \text{delay}, \text{duration}\right \}} R_{i_t}
+%\end{align*}
+%and
+%\begin{align*}
+%\Sigma v_t = \sum_{i \in \left \{\text{packet}, \text{byte}, \text{delay}, \text{duration}\right \}} v_{i_t}.
+%\end{align*}
 
@@
 \begin{align*}
-l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_a \right)\right)\left(\Sigma R_t - \Sigma v_t\right)\\ 
+l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_a \right)\right)\left(\frac{}{} - \Sigma v_t\right)\\ 
 &+ \beta H\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right)\\
@@
 Reinforcement learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
-However, in the case of congestion control, it is desirable to perform a new action without having received reward for the previous on.
-For example, imagine the that you receive two acknowledgements directly after each other.
-For each of these two acknowledgement an action has to be performed but it does not seem feasible for the second action to wait until the first action has received a reward.
+However, in the case of congestion control, it is desirable to perform a new action without having received a reward for the previous action.
+For example, imagine that you receive two acknowledgements directly after each other.
+For each of these two acknowledgements an action has to be performed but it does not seem feasible for the second action to wait until the first action has received a reward.
 The Asynchronous Actor Critic framework as described by \cite{mnih_asynchronous_2016} cannot be applied to congestion control as it assumes that actions and rewards are synchronized.
-
-Thus we propose the Asynchronous Asynchronous Actor Critic framework in which there are two threads within each thread: one thread that generates actions upon receiving a state and one that receives rewards and updates the value and actor networks.
-\todo[inline]{Insert two algorithm environments that show what happens upon receiving actions and rewards}
 
@@
 \end{figure}
-\footnotetext{\url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}
+\footnotetext{from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}
+
+Thus we propose the Partial Action Actor Critic Learning (PAL) framework (see \autoref{fig:pal}) in which one action generates a number of partial actions ($leq 1$).
+Each partial interaction generates a partial reward upon generating with the environment.
+Upon receiving a partial reward, the agent estimates the current state and triggers a new action.
+When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks.
+
+\begin{figure}
+\includesvg{figures/Reinforcement_learning_diagram_PAL}
+\caption{Partial Action Actor Critic Learning: An action consists of zero or more partial actions which trigger partial rewards upon interacting with the environment.
+Each partial reward updates the state.
+The value and actor networks are updated upon receiving all partial rewards of one action.\protect\footnotemark}
+\label{fig:pal}
+\end{figure}
+\footnotetext{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}
+
+Spam and eggs
 

### 2017-08-31 async_deep_reinforce/a3c.py

@@ -122,28 +122,37 @@ else:
 score_throughput = tf.placeholder(PRECISION)
 score_delay = tf.placeholder(PRECISION)
 entropy = tf.placeholder(PRECISION)
+skewness = tf.placeholder(PRECISION)
 actor_loss = tf.placeholder(PRECISION)
 value_loss = tf.placeholder(PRECISION)
 total_loss = tf.placeholder(PRECISION)
 window = tf.placeholder(PRECISION)
 std = tf.placeholder(PRECISION)
+inner_mean = tf.placeholder(PRECISION)
+inner_std = tf.placeholder(PRECISION)
 tf.summary.scalar("score_throughput", score_throughput)
 tf.summary.scalar("score_delay", score_delay)
 tf.summary.scalar("entropy", entropy)
+tf.summary.scalar("skewness", skewness)
 tf.summary.scalar("actor_loss", actor_loss)
 tf.summary.scalar("value_loss", value_loss)
 tf.summary.scalar("total_loss", total_loss)
 tf.summary.scalar("window", window)
 tf.summary.scalar("std", std)
+tf.summary.scalar("inner_mean", inner_mean)
+tf.summary.scalar("inner_std", inner_std)
 summary_inputs = {
   "score_throughput": score_throughput,
   "score_delay": score_delay,
   "entropy": entropy,
+  "skewness": skewness,
   "actor_loss": actor_loss,
   "value_loss": value_loss,
   "total_loss": total_loss,
   "window": window,
   "std": std,
+  "inner_mean": inner_mean,
+  "inner_std": inner_std
 }
 
 summary_op = tf.summary.merge_all()

### 2017-08-31 async_deep_reinforce/a3c_training_thread.py

@@ -73,13 +73,14 @@ class A3CTrainingThread(object):
       summary_inputs["score_throughput"]: things["score_throughput"],
       summary_inputs["score_delay"]: things["score_delay"],
       summary_inputs["entropy"]: things["entropy"],
+      summary_inputs["skewness"]: things["skewness"],
       summary_inputs["actor_loss"]: things["actor_loss"],
       summary_inputs["value_loss"]: things["value_loss"],
       summary_inputs["total_loss"]: things["total_loss"],
       summary_inputs["window"]: things["window"],
       summary_inputs["std"]: things["std"],
-      # summary_inputs["inner_mean"]: things["inner_mean"],
-      # summary_inputs["inner_std"]: things["inner_std"],
+      summary_inputs["inner_mean"]: things["inner_mean"],
+      summary_inputs["inner_std"]: things["inner_std"],
     })
     summary_writer.add_summary(summary_str, global_t)
     summary_writer.flush()
@@ -279,16 +280,19 @@ class A3CTrainingThread(object):
       }
       feed_dict.update(var_dict)
 
-      entropy, actor_loss, value_loss, total_loss, window, std = self.local_network.run_loss(sess, feed_dict)
+      entropy, skewness, actor_loss, value_loss, total_loss, window, std, inner_mean, inner_std = self.local_network.run_loss(sess, feed_dict)
 
       things = {"score_throughput": normalized_final_score_throughput,
         "score_delay": normalized_final_score_delay, 
         "actor_loss": actor_loss.item(),
         "value_loss": value_loss,
         "entropy": entropy.item(),
+        "skewness": skewness.item(),
         "total_loss": total_loss,
         "window": window.item(),
-        "std": std.item()}
+        "std": std.item(),
+        "inner_mean": inner_mean.item(),
+        "inner_std": inner_std.item()}
       logging.debug(" ".join(map(str,("things", things))))
       self._record_score(sess, summary_writer, summary_op, summary_inputs, things, global_t) # TODO:NOW: is that "not terminal_end" correct?
 

### 2017-08-31 async_deep_reinforce/constants.py

@@ -2,7 +2,7 @@
 import tensorflow as tf
 import logging
 
-LOCAL_T_MAX = 20 # repeat step size
+LOCAL_T_MAX = 40 # repeat step size
 RMSP_ALPHA = 0.99 # decay parameter for RMSProp
 RMSP_EPSILON = 0.1 # epsilon parameter for RMSProp
 CHECKPOINT_DIR = 'checkpoints'
@@ -14,8 +14,8 @@ PRECISION = tf.float32
 
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
 GAMMA = 0.99 # discount factor for rewards
-ENTROPY_BETA = 1e-3
-SKEWNESS_GAMMA = 1e-3
+ENTROPY_BETA = 0.5e-3
+SKEWNESS_GAMMA = 0.5e-3
 MAX_TIME_STEP = 10 * 1e7
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
@@ -31,4 +31,4 @@ BETA = 1.0
 LOG_LEVEL = logging.INFO
 MINIMUM_STD = 1e-2
 # Is this needed?
-ACTOR_FACTOR = 5e-1
+ACTOR_FACTOR = 1e-1

### 2017-08-31 async_deep_reinforce/game_ac_network.py

@@ -13,7 +13,7 @@ from constants import BETA
 from constants import ENTROPY_BETA
 from constants import SKEWNESS_GAMMA
 from constants import MINIMUM_STD
-# from constants import ACTOR_FACTOR
+from constants import ACTOR_FACTOR
 
 # tiny = np.finfo(np.float32).tiny
 tiny = 1e-20
@@ -60,31 +60,27 @@ class GameACNetwork(object):
 			self.td_throughput = tf.placeholder(PRECISION, [None], name="td_throughput")
 			self.td_delay = tf.placeholder(PRECISION, [None], name="td_delay")
 			
-			# normal_variance = tf.log((self.pi[1]*self.pi[1])/(self.pi[0]*self.pi[0])+1.0)
-			# self.distribution = ds.TransformedDistribution(
-			# 		distribution=ds.Normal(loc=tf.log(self.pi[0])-normal_variance/2.0, scale=tf.sqrt(normal_variance), allow_nan_stats=False, validate_args=True),
-			# 		bijector=ds.bijectors.Exp(),
-			# 		name="LogNormalTransformedDistribution")
-
-			self.distribution = ds.Normal(loc=self.pi[0], scale=self.pi[1], allow_nan_stats=False, validate_args=True)
+			normal_variance = tf.log((self.pi[1]*self.pi[1])/(self.pi[0]*self.pi[0])+1.0)
+			self.distribution = ds.TransformedDistribution(
+					distribution=ds.Normal(loc=tf.log(self.pi[0])-normal_variance/2.0, scale=tf.sqrt(normal_variance), allow_nan_stats=False, validate_args=True),
+					bijector=ds.bijectors.Exp(),
+					name="LogNormalTransformedDistribution")
 
 			self.distribution_mean = self.pi[0]
 			self.distribution_std = self.pi[1]
-			# self.inner_distribution_mean = self.distribution.distribution.mean()
-			# self.inner_distribution_std = self.distribution.distribution.stddev()
-			# self.chosen_action = tf.ceil(self.distribution.sample())
-			self.chosen_action = tf.ceil(tf.maximum(self.distribution.sample(), 0.0))
+			self.inner_distribution_mean = self.distribution.distribution.mean()
+			self.inner_distribution_std = self.distribution.distribution.stddev()
+			self.chosen_action = tf.ceil(self.distribution.sample())
 
[…]

### 2017-08-31 async_deep_reinforce/constants.py
Changed parametrization from mean and std to median and std, to fix problem of window growing to infinity

@@ -2,7 +2,7 @@
 import tensorflow as tf
 import logging
 
-LOCAL_T_MAX = 40 # repeat step size
+LOCAL_T_MAX = 20 # repeat step size
 RMSP_ALPHA = 0.99 # decay parameter for RMSProp
 RMSP_EPSILON = 0.1 # epsilon parameter for RMSProp
 CHECKPOINT_DIR = 'checkpoints'
@@ -10,7 +10,7 @@ LOG_FILE = 'tmp/a3c_log'
 INITIAL_ALPHA_LOW = 1e-4    # log_uniform low limit for learning rate
 INITIAL_ALPHA_HIGH = 1e-2   # log_uniform high limit for learning rate
 
-PRECISION = tf.float32
+PRECISION = tf.float64
 
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
 GAMMA = 0.99 # discount factor for rewards
@@ -31,4 +31,4 @@ BETA = 1.0
 LOG_LEVEL = logging.INFO
 MINIMUM_STD = 1e-2
 # Is this needed?
-ACTOR_FACTOR = 1e-1
+ACTOR_FACTOR = 1

### 2017-08-31 async_deep_reinforce/game_ac_network.py

@@ -60,16 +60,22 @@ class GameACNetwork(object):
 			self.td_throughput = tf.placeholder(PRECISION, [None], name="td_throughput")
 			self.td_delay = tf.placeholder(PRECISION, [None], name="td_delay")
 			
-			normal_variance = tf.log((self.pi[1]*self.pi[1])/(self.pi[0]*self.pi[0])+1.0)
+			self.inner_distribution_mean = tf.log(self.pi[0])
+			self.inner_distribution_std = tf.sqrt(tf.log(0.5 * tf.exp(-2. * self.inner_distribution_mean)*(tf.exp(2. * self.inner_distribution_mean)+tf.sqrt(tf.exp(4. * self.inner_distribution_mean) + 4. * tf.exp(2. * self.inner_distribution_mean) * self.pi[1]*self.pi[1]))))
 			self.distribution = ds.TransformedDistribution(
-					distribution=ds.Normal(loc=tf.log(self.pi[0])-normal_variance/2.0, scale=tf.sqrt(normal_variance), allow_nan_stats=False, validate_args=True),
+					distribution=ds.Normal(loc=self.inner_distribution_mean, scale=self.inner_distribution_std, allow_nan_stats=False, validate_args=True),
 					bijector=ds.bijectors.Exp(),
 					name="LogNormalTransformedDistribution")
+			# normal_variance = tf.log((self.pi[1]*self.pi[1])/(self.pi[0]*self.pi[0])+1.0)
+			# self.distribution = ds.TransformedDistribution(
+			# 		distribution=ds.Normal(loc=tf.log(self.pi[0])-normal_variance/2.0, scale=tf.sqrt(normal_variance), allow_nan_stats=False, validate_args=True),
+			# 		bijector=ds.bijectors.Exp(),
+			# 		name="LogNormalTransformedDistribution")
 
-			self.distribution_mean = self.pi[0]
+			self.distribution_median = self.pi[0]
 			self.distribution_std = self.pi[1]
-			self.inner_distribution_mean = self.distribution.distribution.mean()
-			self.inner_distribution_std = self.distribution.distribution.stddev()
+			# self.inner_distribution_mean = self.distribution.distribution.mean()
+			# self.inner_distribution_std = self.distribution.distribution.stddev()
 			self.chosen_action = tf.ceil(self.distribution.sample())
 
[…]

### 2017-08-31 tex/paper.tex
Tuned parameters, fixed formula in TeX

@@
 \begin{align*}
-l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_a \right)\right)\left(\frac{}{} - \Sigma v_t\right)\\ 
+l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_a \right)\right)\\
+&\left( \left(\frac{R_{\text{byte}_t}}{{R_{\text{duration}_t}}} - \frac{v_{\text{byte}_t}}{{v_{\text{duration}_t}}}\right) - \left( \frac{R_{\text{delay}_t}}{{R_{\text{packet}_t}}}- \frac{v_{\text{delay}_t}}{{v_{\text{packet}_t}}} \right)\right)\\
 &+ \beta H\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right)\\

### 2017-08-31 async_deep_reinforce/constants.py

@@ -14,8 +14,8 @@ PRECISION = tf.float64
 
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
 GAMMA = 0.99 # discount factor for rewards
-ENTROPY_BETA = 0.5e-3
-SKEWNESS_GAMMA = 0.5e-3
+ENTROPY_BETA = 1e-4
+SKEWNESS_GAMMA = 1e-4
 MAX_TIME_STEP = 10 * 1e7
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
@@ -31,4 +31,4 @@ BETA = 1.0
 LOG_LEVEL = logging.INFO
 MINIMUM_STD = 1e-2
 # Is this needed?
-ACTOR_FACTOR = 1
+ACTOR_FACTOR = 0.5

### 2017-09-01 async_deep_reinforce/game_ac_network.py
Minor tex changes and sigmoid for std deviation

@@ -281,8 +281,8 @@ class GameACLSTMNetwork(GameACNetwork):
 			# policy (output)
 			self.pi = (
 				tf.nn.softplus(raw_pi_median) + 1.0,
-				tf.nn.softplus(raw_pi_std) + MINIMUM_STD
-				# tf.nn.sigmoid(raw_pi_std)
+				# tf.nn.softplus(raw_pi_std) + MINIMUM_STD
+				tf.nn.sigmoid(raw_pi_std)
 				# tf.constant(0.5, shape=(1,), dtype=PRECISION)
 			)
 

### 2017-09-01 tex/paper.tex
Minor tex changes and sigmoid for std deviation
Realize that too high entropy is the root of all evil...

--- 2017-08-31 tex/paper.tex
@@
 \usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
+
 \usepackage{todonotes}
@@
 
-\newcommand\fnurl[2]{%
-\href{#2}{#1}\footnote{\url{#2}}%
-}
+\makeatletter
+\newcommand{\StateIndent}[1][3]{%
+  \setlength\@tempdima{\algorithmicindent}%
+  \Statex\hskip\dimexpr#1\@tempdima\relax}
+\algdef{S}[WHILE]{WhileNoDo}[1]{\algorithmicwhile\ #1}%
+\makeatother
+
+\newcommand{\algorithmautorefname}{Algorithm}
 
@@
 
-\author{Joachim Fabini} 
-\affiliation{%
- \institution{Institute of Telecommunications}
- \streetaddress{Gußhausstraße 25}
- \postcode{1040}
- \city{Vienna} 
- \country{Austria}}
-\email{joachim.fabini@tuwien.ac.at}
+%\author{Joachim Fabini} 
+%\affiliation{%
+% \institution{Institute of Telecommunications}
+% \streetaddress{Gußhausstraße 25}
+% \postcode{1040}
+% \city{Vienna} 
+% \country{Austria}}
+%\email{joachim.fabini@tuwien.ac.at}
 
@@
 % The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{Bachl et al.}
+\renewcommand{\shortauthors}{Bachl, Zseby}
 
@@
 Let $r_t$ be the reward that was received at time $t$.
-$V(s_t; \theta_v)$ designates the expected mean reward in state $s_t$ given the parameters of the value network $theta_v$.
+$V(s_t; \theta_\text{v})$ designates the expected mean reward in state $s_t$ given the parameters of the value network $theta_v$.
 Then, with $\gamma$ being the roll-off factor, which designates the influence that future reward has on the moving average, (set to $0.99$), we generically define the expected n-step return as
 \begin{align*}
-R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k \frac{V(s_{t+k}; \theta_v)}{1-\gamma}\right)\left(1-\gamma\right),
+R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k \frac{V(s_{t+k}; \theta_\text{v})}{1-\gamma}\right)\left(1-\gamma\right),
 \end{align*}
@@
 \begin{align*}
-l_{\text{v}_t} = \left(R_t - V(s_t; \theta_v)\right)^2.
+l_{\text{v}_t} = \left(R_t - V(s_t; \theta_\text{v})\right)^2.
 \end{align*}
@@
 \begin{align*}
-l_{\text{v}_t} =& \left(R_{\text{packet}_t} - V_\text{packet}(s_t; \theta_v)\right)^2 \\
-&+\left(R_{\text{byte}_t} - V_\text{byte}(s_t; \theta_v)\right)^2 \\
-&+\left(R_{\text{delay}_t} - V_\text{delay}(s_t; \theta_v)\right)^2 \\
-&+\left(R_{\text{duration}_t} - V_\text{duration}(s_t; \theta_v)\right)^2
+l_{\text{v}_t} =& \left(R_{\text{packet}_t} - V_\text{packet}(s_t; \theta_\text{v})\right)^2 \\
+&+\left(R_{\text{byte}_t} - V_\text{byte}(s_t; \theta_\text{v})\right)^2 \\
+&+\left(R_{\text{delay}_t} - V_\text{delay}(s_t; \theta_\text{v})\right)^2 \\
+&+\left(R_{\text{duration}_t} - V_\text{duration}(s_t; \theta_\text{v})\right)^2
 \end{align*}
@@
 
-With $H$ being the entropy and $S$ being the skewness, the actor network minimizes the loss
+With $H$ being the entropy, the actor network minimizes the loss
 \begin{align*}
-l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_a \right)\right)\left(\frac{}{} - \Sigma v_t\right)\\ 
-&+ \beta H\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right)\\
-&- \delta S\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right).
+l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_\text{a} \right)\right)\\
+&\left( \left(\frac{R_{\text{byte}_t}}{{R_{\text{duration}_t}}} - \frac{v_{\text{byte}_t}}{{v_{\text{duration}_t}}}\right) - \left( \frac{R_{\text{delay}_t}}{{R_{\text{packet}_t}}}- \frac{v_{\text{delay}_t}}{{v_{\text{packet}_t}}} \right)\right)\\
+&+ \beta H\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right).
 \end{align*}
@@
 \end{figure}
-\footnotetext{from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}
+\footnotetext{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}
 
@@
 Upon receiving a partial reward, the agent estimates the current state and triggers a new action.
-When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks.
+When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks (see \autoref{alg:pal}).
 
@@
 
-Spam and eggs
+\begin{algorithm}
+\caption{Partial Action Actor Critic Learning -- pseudocode for each learning thread.}
+\label{alg:pal}
+\begin{algorithmic}
+\Loop
+	\State $l_\text{actions} \gets  [] $
+	\State $l_\text{states} \gets  [] $
+	\State $l_\text{values} \gets  [] $
+	\State $l_\text{rewards} \gets  [] $
+	\State $l_\text{estimatedValues} \gets  [] $
+	\State $l_\text{snapshots} \gets  [] $
+	\State $t \gets 0$
+	\State $s_0 \gets \text{initialState}()$
+	\State $\theta_\text{a} \gets \theta_{\text{a}_\text{g}}$
+	\State $\theta_\text{v} \gets \theta_{\text{v}_\text{g}}$
+	\Repeat
+		\State $l_\text{states}.\text{append}(s_t)$
+		\State $l_\text{estimatedValues}.\text{append}(V(s_t; \theta_\text{v}))$
+		\If{$t \bmod t_\text{max} = 0$}
+			\State $\theta_\text{a} \gets \theta_{\text{a}_\text{g}}$
+			\State $\theta_\text{v} \gets \theta_{\text{v}_\text{g}}$
+			\State $l_\text{snapshots}.\text{append}((\theta_\text{a}, \theta_\text{v}))$
+		\EndIf
+		\State Sample $a_t$ from the probability distribution
+		\State $l_\text{actions}.\text{append}(a_t)$
+		\State $l_{\text{partialActions}_t} \gets \text{partialActions}(a_t)$
+		\ForAll{partial actions $a_{\text{p}_{nt}}$ in $l_{\text{partialActions}_t}$}
+			\State Take partial action $a_{\text{p}_{nt}}$
+		\EndFor
+		\State Receive partial reward $r_{\text{p}_{nt'}}$
+		\StateIndent[3] where $t' \leq t$ and $0 \leq n \leq\#(l_{\text{partialActions}_{t'}})$
+		\If{all partial rewards of $a_{t'}$ were received}
+			\State $r_{t'} \gets \text{combination of all partial rewards }r_{\text{p}_{nt'}}$
+			\State $l_\text{rewards}.\text{append}(r_{t'})$
+			\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
+				\State $t_\text{end} = \min\left(t_\text{max}, \#\left( l_\text{rewards} \right)\right)$
+				\State $\theta_\text{backup} \gets \left(\theta_\text{a}, \theta_\text{v}\right)$
+				\State $\theta_\text{a}, \theta_\text{v} \gets l_\text{snapshots}[0]$
+				\State $R_i+1 \gets \frac{l_\text{estimatedValues}[t_\text{end}]}{1-\gamma}]$
+				\For{$i\gets t_\text{end}-1, 0$}
+					\State $R_i \gets r_i + \gamma r_{i+1}$
+					\todo[inline]{Calculate the gradients}
+					\State Calculate the gradients $d\theta_\text{a}$ and $d\theta_\text{v}$
+					\State $\theta_{\text{a}_\text{g}} \gets \theta_{\text{a}_\text{g}} + d\theta_\text{a}$
+					\State $\theta_{\text{v}_\text{g}} \gets \theta_{\text{v}_\text{g}} + d\theta_\text{v}$
+				\EndFor
+				\State $l_\text{actions} \gets l_\text{actions}[t_\text{max}:] $
+				\State $l_\text{states} \gets l_\text{states}[t_\text{max}:] $
+				\State $l_\text{values} \gets l_\text{values}[t_\text{max}:] $
+				\State $l_\text{rewards} \gets l_\text{rewards}[t_\text{max}:] $
+				\State $l_\text{estimatedValues} \gets l_\text{estimatedValues}[t_\text{max}:] $
+				\State $l_\text{snapshots} \gets l_\text{snapshots}[1:] $
+				\State $\theta_\text{a}, \theta_\text{v} = \theta_\text{backup}$
+			\EndIf
+		\EndIf
+		\State Generate $s_{t+1}$ using $r_{\text{p}_{t'}}$
+		\State $t \gets t+1$
+	\Until{reaching the end the episode}
+\EndLoop
+\end{algorithmic}
+\end{algorithm}
 

### 2017-09-01 async_deep_reinforce/constants.py

@@ -10,12 +10,11 @@ LOG_FILE = 'tmp/a3c_log'
 INITIAL_ALPHA_LOW = 1e-4    # log_uniform low limit for learning rate
 INITIAL_ALPHA_HIGH = 1e-2   # log_uniform high limit for learning rate
 
-PRECISION = tf.float64
+PRECISION = tf.float64	
 
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
 GAMMA = 0.99 # discount factor for rewards
-ENTROPY_BETA = 1e-4
-SKEWNESS_GAMMA = 1e-4
+ENTROPY_BETA = 0.
 MAX_TIME_STEP = 10 * 1e7
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
@@ -30,5 +29,3 @@ ALPHA = 1.0
 BETA = 1.0
 LOG_LEVEL = logging.INFO
 MINIMUM_STD = 1e-2
-# Is this needed?
-ACTOR_FACTOR = 0.5

### 2017-09-01 async_deep_reinforce/game_ac_network.py

@@ -11,9 +11,8 @@ from constants import LAYER_NORMALIZATION
 from constants import ALPHA
 from constants import BETA
 from constants import ENTROPY_BETA
-from constants import SKEWNESS_GAMMA
 from constants import MINIMUM_STD
-from constants import ACTOR_FACTOR
+# from constants import OFFSET
 
 # tiny = np.finfo(np.float32).tiny
 tiny = 1e-20
@@ -60,34 +59,29 @@ class GameACNetwork(object):
 			self.td_throughput = tf.placeholder(PRECISION, [None], name="td_throughput")
 			self.td_delay = tf.placeholder(PRECISION, [None], name="td_delay")
 			
-			self.inner_distribution_mean = tf.log(self.pi[0])
-			self.inner_distribution_std = tf.sqrt(tf.log(0.5 * tf.exp(-2. * self.inner_distribution_mean)*(tf.exp(2. * self.inner_distribution_mean)+tf.sqrt(tf.exp(4. * self.inner_distribution_mean) + 4. * tf.exp(2. * self.inner_distribution_mean) * self.pi[1]*self.pi[1]))))
+			normal_variance = tf.log((self.pi[1]*self.pi[1])/(self.pi[0]*self.pi[0])+1.0)
 			self.distribution = ds.TransformedDistribution(
-					distribution=ds.Normal(loc=self.inner_distribution_mean, scale=self.inner_distribution_std, allow_nan_stats=False, validate_args=True),
+					distribution=ds.Normal(loc=tf.log(self.pi[0])-normal_variance/2.0, scale=tf.sqrt(normal_variance), allow_nan_stats=False, validate_args=True),
 					bijector=ds.bijectors.Exp(),
 					name="LogNormalTransformedDistribution")
-			# normal_variance = tf.log((self.pi[1]*self.pi[1])/(self.pi[0]*self.pi[0])+1.0)
-			# self.distribution = ds.TransformedDistribution(
-			# 		distribution=ds.Normal(loc=tf.log(self.pi[0])-normal_variance/2.0, scale=tf.sqrt(normal_variance), allow_nan_stats=False, validate_args=True),
-			# 		bijector=ds.bijectors.Exp(),
-			# 		name="LogNormalTransformedDistribution")
 
-			self.distribution_median = self.pi[0]
+			self.distribution_mean = self.pi[0]
 			self.distribution_std = self.pi[1]
[…]

### 2017-09-04 async_deep_reinforce/constants.py
Lower training rate to make scenario with multiple senders work. It works now

@@ -14,11 +14,11 @@ PRECISION = tf.float64
 
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
 GAMMA = 0.99 # discount factor for rewards
-ENTROPY_BETA = 0.
+ENTROPY_BETA = 1e-4
 MAX_TIME_STEP = 10 * 1e7
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
-N_LSTM_LAYERS = 3
+N_LSTM_LAYERS = 1
 
 STATE_SIZE = 11
 HIDDEN_SIZE = 128
@@ -29,3 +29,4 @@ ALPHA = 1.0
 BETA = 1.0
 LOG_LEVEL = logging.INFO
 MINIMUM_STD = 1e-2
+ACTOR_FACTOR = 0.25

### 2017-09-04 async_deep_reinforce/game_ac_network.py

@@ -12,7 +12,7 @@ from constants import ALPHA
 from constants import BETA
 from constants import ENTROPY_BETA
 from constants import MINIMUM_STD
-# from constants import OFFSET
+from constants import ACTOR_FACTOR
 
 # tiny = np.finfo(np.float32).tiny
 tiny = 1e-20
@@ -73,15 +73,15 @@ class GameACNetwork(object):
 			# self.chosen_action = self.distribution.sample() + OFFSET
 
 			# policy entropy
-			# self.entropy = ENTROPY_BETA * tf.reduce_sum(self.distribution.distribution.mean() + 0.5 * tf.log(2.0*math.pi*math.e*self.distribution.distribution.variance()), axis=1)
-			self.entropy = ENTROPY_BETA * tf.reduce_sum(0.5 * tf.log(2.0*math.pi*math.e*self.pi[1]*self.pi[1]), axis=1)
+			self.entropy = ENTROPY_BETA * tf.reduce_sum(self.distribution.distribution.mean() + 0.5 * tf.log(2.0*math.pi*math.e*self.distribution.distribution.variance()), axis=1)
+			# self.entropy = ENTROPY_BETA * tf.reduce_sum(0.5 * tf.log(2.0*math.pi*math.e*self.pi[1]*self.pi[1]), axis=1)
 			self.skewness = ENTROPY_BETA * tf.reduce_sum((tf.exp(self.distribution.distribution.variance()) + 2.0)*tf.sqrt(tf.exp(self.distribution.distribution.variance()) - 1.0), axis=1)
 			# self.actor_loss = tf.reduce_sum(self.distribution.log_prob(self.a - OFFSET), axis=1) * (self.td_throughput + self.td_delay)
 			self.actor_loss = tf.reduce_sum(tf.log(self.distribution.cdf(self.a) - self.distribution.cdf(tf.clip_by_value(self.a - 1.0, tiny, float("inf")))), axis=1) * (self.td_throughput + self.td_delay)
 
 			# self.policy_loss = - tf.reduce_sum(self.actor_loss + ENTROPY_BETA * self.entropy - SKEWNESS_GAMMA * self.skewness)
-			# self.policy_loss = - tf.reduce_sum(self.actor_loss + self.entropy)
-			self.policy_loss = - tf.reduce_sum(self.actor_loss)
[…]

### 2017-09-04 tex/paper.tex
Lower training rate to make scenario with multiple senders work. It works now
Improve names in LaTeX formulas; finish algorithm typesetting

@@
 \begin{align*}
-l_{\text{v}_t} =& \left(R_{\text{packet}_t} - V_\text{packet}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{byte}_t} - V_\text{byte}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{delay}_t} - V_\text{delay}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{duration}_t} - V_\text{duration}(s_t; \theta_\text{v})\right)^2
+l_{\text{v}_t} =& \left(R_{\text{packet},t} - V_\text{packet}(s_t; \theta_\text{v})\right)^2 \\
+&+\left(R_{\text{byte},t} - V_\text{byte}(s_t; \theta_\text{v})\right)^2 \\
+&+\left(R_{\text{delay},t} - V_\text{delay}(s_t; \theta_\text{v})\right)^2 \\
+&+\left(R_{\text{duration},t} - V_\text{duration}(s_t; \theta_\text{v})\right)^2
 \end{align*}
@@
 l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_\text{a} \right)\right)\\
-&\left( \left(\frac{R_{\text{byte}_t}}{{R_{\text{duration}_t}}} - \frac{v_{\text{byte}_t}}{{v_{\text{duration}_t}}}\right) - \left( \frac{R_{\text{delay}_t}}{{R_{\text{packet}_t}}}- \frac{v_{\text{delay}_t}}{{v_{\text{packet}_t}}} \right)\right)\\
-&+ \beta H\left(\text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right) \right).
+&\left( \left(\frac{R_{\text{byte},t}}{{R_{\text{duration},t}}} - \frac{v_{\text{byte},t}}{{v_{\text{duration},t}}}\right) - \left( \frac{R_{\text{delay},t}}{{R_{\text{packet},t}}}- \frac{v_{\text{delay},t}}{{v_{\text{packet},t}}} \right)\right)\\
+&+ \beta H\left( \pi\left( l_\text{states}[i]; \theta_a \right)\right).
 \end{align*}
@@
 \label{alg:pal}
-\begin{algorithmic}
+\begin{algorithmic}[1]
 \Loop
@@
 	\State $s_0 \gets \text{initialState}()$
-	\State $\theta_\text{a} \gets \theta_{\text{a}_\text{g}}$
-	\State $\theta_\text{v} \gets \theta_{\text{v}_\text{g}}$
+	\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
+	\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
 	\Repeat
@@
 		\If{$t \bmod t_\text{max} = 0$}
-			\State $\theta_\text{a} \gets \theta_{\text{a}_\text{g}}$
-			\State $\theta_\text{v} \gets \theta_{\text{v}_\text{g}}$
+			\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
+			\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
 			\State $l_\text{snapshots}.\text{append}((\theta_\text{a}, \theta_\text{v}))$
@@
 		\State $l_\text{actions}.\text{append}(a_t)$
-		\State $l_{\text{partialActions}_t} \gets \text{partialActions}(a_t)$
-		\ForAll{partial actions $a_{\text{p}_{nt}}$ in $l_{\text{partialActions}_t}$}
-			\State Take partial action $a_{\text{p}_{nt}}$
+		\State $l_{\text{partialActions},t} \gets \text{partialActions}(a_t)$
+		\ForAll{partial actions $a_{{\text{p},n,t}}$ in $l_{\text{partialActions},t}$}
+			\State Take partial action $a_{{\text{p},n,t}}$
 		\EndFor
-		\State Receive partial reward $r_{\text{p}_{nt'}}$
-		\StateIndent[3] where $t' \leq t$ and $0 \leq n \leq\#(l_{\text{partialActions}_{t'}})$
+		\State Receive partial reward $r_{\text{p},n,t'}$
+		\StateIndent[3] where $t' \leq t$ and $0 \leq n \leq\#(l_{\text{partialActions},t'})$
 		\If{all partial rewards of $a_{t'}$ were received}
-			\State $r_{t'} \gets \text{combination of all partial rewards }r_{\text{p}_{nt'}}$
+			\State \Call{computeGradients}{}
+		\EndIf
+		\State Generate $s_{t+1}$ using $r_{\text{p}_{t'}}$
+		\State $t \gets t+1$
+	\Until{reaching the end the episode}
+\EndLoop
+%\algstore{palalg}
+\end{algorithmic}
+\end{algorithm}
+\begin{algorithm}
+\caption{Partial Action Actor Critic Learning -- procedure which computes and applies the gradients.}
+\label{alg:grad}
+\begin{algorithmic}[1]%
+%\algrestore{palalg}
+\Function{computeGradients}{}
+	\State $r_{t'} \gets \text{combination of all partial rewards }r_{\text{p},n,t'}$
 	\State $l_\text{rewards}.\text{append}(r_{t'})$
@@
 		\State $\theta_\text{a}, \theta_\text{v} \gets l_\text{snapshots}[0]$
-				\State $R_i+1 \gets \frac{l_\text{estimatedValues}[t_\text{end}]}{1-\gamma}]$
+		\State $R_{i+1} \gets \frac{l_\text{estimatedValues}[t_\text{end}]}{1-\gamma}$
 		\For{$i\gets t_\text{end}-1, 0$}
 			\State $R_i \gets r_i + \gamma r_{i+1}$
-					\todo[inline]{Calculate the gradients}
-					\State Calculate the gradients $d\theta_\text{a}$ and $d\theta_\text{v}$
-					\State $\theta_{\text{a}_\text{g}} \gets \theta_{\text{a}_\text{g}} + d\theta_\text{a}$
-					\State $\theta_{\text{v}_\text{g}} \gets \theta_{\text{v}_\text{g}} + d\theta_\text{v}$
+			\State $a \gets l_\text{actions}[i]$
+			\State $s \gets l_\text{states}[i]$
+			\State $v \gets l_\text{values}[i]$
+			\State $d\theta_\text{v} \gets \frac{\partial\left(R_i - l_\text{values}[i]\right)}{\partial\theta_v}$
+			\State $d\theta_\text{v} \gets -\frac{\partial\log \left( \pi_W\left( a-1 < W \leq a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
+			\StateIndent[6] $-\frac{\partial\beta H\left( \pi\left( l_\text{states}[i]; \theta_a \right)\right)}{\partial\theta_a}$
+			\State $\theta_{\text{a},\text{g}} \gets \theta_{\text{a},\text{g}} + d\theta_\text{a}$
+			\State $\theta_{\text{v},\text{g}} \gets \theta_{\text{v},\text{g}} + d\theta_\text{v}$
 		\EndFor
@@
 	\EndIf
-		\EndIf
-		\State Generate $s_{t+1}$ using $r_{\text{p}_{t'}}$
-		\State $t \gets t+1$
-	\Until{reaching the end the episode}
-\EndLoop
+\EndFunction
 \end{algorithmic}

### 2017-09-06 async_deep_reinforce/a3c_training_thread.py
Adapt code for use with ns2

@@ -133,6 +133,7 @@ class A3CTrainingThread(object):
     self.estimated_values = self.estimated_values[:-actions_to_remove+1]
 
     # If, for some strange reason, absolutely nothing happened in this episode, don't do anyting...
+    # Or if you're actually in testing mode :)
     if len(self.rewards)>0:
       time_diff = self.process(sess, global_t, summary_writer, summary_op, summary_inputs, time_difference)
     else:

### 2017-09-06 src/unicorn-templates.cc

@@ -6,8 +6,7 @@
 using namespace std;
 
 template <class NextHop>
-void Unicorn::send( const unsigned int id, NextHop & next, const double & tickno,
-		const unsigned int packets_sent_cap )
+void Unicorn::send( const unsigned int id, NextHop & next, const double & tickno, const unsigned int packets_sent_cap )
 {
   // _sent_at_least_once = true;
 
@@ -21,7 +20,8 @@ void Unicorn::send( const unsigned int id, NextHop & next, const double & tickno
     (int(_packets_sent) < _largest_ack + 1 + _the_window ) &&
     (_packets_sent < packets_sent_cap)) {
 
-    Packet p( id, _flow_id, tickno, _packets_sent, _put_actions );
+    Packet p( id, _flow_id, tickno, _packets_sent );
+    _id_to_sent_during_action[id] = _put_actions;
     _outstanding_rewards[_put_actions]["sent"] += 1;
     if (_last_send_time != 0) {
       _outstanding_rewards[_put_actions]["intersend_duration_acc"] += tickno - _last_send_time;

### 2017-09-06 src/unicorn.cc

@@ -22,7 +22,9 @@ Unicorn::Unicorn()
     _put_actions(0),
     _put_rewards(0),
     _outstanding_rewards(),
-    _start_tick(0.0)
+    _start_tick(0.0),
+    _training(true),
+    _id_to_sent_during_action()
 {
   // puts("Creating a Unicorn");
 }
@@ -46,13 +48,15 @@ void Unicorn::packets_received( const vector< Packet > & packets ) {
     const unsigned int lost_since_last_time = (unsigned int) packet.seq_num-_largest_ack-1;
     _memory.lost(lost_since_last_time);
 
-    for (auto it=_outstanding_rewards.begin(); it!=_outstanding_rewards.lower_bound(packet.sent_during_action);) {
+    for (auto it=_outstanding_rewards.begin(); it!=_outstanding_rewards.lower_bound(_id_to_sent_during_action[packet.id]);) {
       const double throughput_final = it->second["received"];
       const double delay_final = it->second["delay_acc"];
       // const double duration = it->second["end_time"] - it->second["start_time"];
       // const double duration = it->second["intersend_duration_acc"];    
       const double duration = it->second["interreceive_duration_acc"];
-      _rainbow.put_reward(_thread_id, throughput_final, delay_final, duration);
+      if (_training) {
+        _rainbow.put_reward(_thread_id, throughput_final, delay_final, duration);
+      }
       _put_rewards += 1;
       it = _outstanding_rewards.erase(it);
     }
@@ -75,6 +79,8 @@ void Unicorn::packets_received( const vector< Packet > & packets ) {
     _largest_ack = packet.seq_num;
 
     get_action(packet.tick_received, packets_sent_in_previous_episode);
+
+    _id_to_sent_during_action.erase(_id_to_sent_during_action]);
   }
 
   // _largest_ack = max( packets.at( packets.size() - 1 ).seq_num, _largest_ack );
@@ -112,6 +118,7 @@ void Unicorn::reset(const double & tickno)
   _put_actions = 0;
   _put_rewards = 0;
   _outstanding_rewards.clear();
+  _id_to_sent_during_action.clear();
   _start_tick = tickno;
 
   assert( _flow_id != 0 );

### 2017-09-06 tex/paper.tex
Change structure of explanation

@@
 
-%\author{Joachim Fabini} 
-%\affiliation{%
-% \institution{Institute of Telecommunications}
-% \streetaddress{Gußhausstraße 25}
-% \postcode{1040}
-% \city{Vienna} 
-% \country{Austria}}
-%\email{joachim.fabini@tuwien.ac.at}
-
 \author{Tanja Zseby} 
@@
 
+\author{Joachim Fabini} 
+\affiliation{%
+ \institution{Institute of Telecommunications}
+ \streetaddress{Gußhausstraße 25}
+ \postcode{1040}
+ \city{Vienna} 
+ \country{Austria}}
+\email{joachim.fabini@tuwien.ac.at}
+
 % The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{Bachl, Zseby}
+\renewcommand{\shortauthors}{Bachl, Zseby, Fabini}
 
@@
 
-\section{Introduction}
+\section{Partial Action Actor Critic Learning}
+\label{sec:pal}
+
+Partial Action Actor Critic Learning (PAL) is a modification of the Asynchronous Advantage Actor Critic algorithm proposed by \citet{mnih_asynchronous_2016}.
+It consists of two Artificial Neural Networks (ANNs), the Actor Network and the Value Network (the Critic part in the abbreviation stands for the Value Network).
+Given a state, the Actor Networks outputs what it deems to be the optimum action in that certain state.
+The Value Network estimates what long-term reward can be expected in this state.
+So an action is considered good if it achieved a long-term reward that is higher than the long-term reward expected by the Value Network and it is considered bad if the reward was lower than expected.
+The long-term reward is implemented as a moving average of future rewards.
+So if a high reward can be achieved right now this is more favorable than if it can be achieved in the future.
+However it can also be beneficial to get a low reward now and instead get a very large one in the future.
+
+\subsection{Algorithm}
+\label{subsec:alg}
+
+In \autoref{alg:pal} we show the code that runs in one of the learning threads.
+As proposed in the literature, we use several concurrent thread to improve training performance.
+A major difference to previous approaches to reinforcement learning is the fact that we allow more actions to be taken before the previous actions' rewards have been received.
+In classical reinforcement learning, an action is always followed by a reward and a reward is always followed by a action.
+In our proposed concept, however, it is possible to take new actions while the previous action hasn't received its reward yet.
+
+\begin{figure}
+\begin{minipage}{\columnwidth}
+\includesvg{figures/Reinforcement_learning_diagram}
+\caption{The classical reinforcement learning approach.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
+\label{fig:reinforcement}
+\end{minipage}
+\end{figure}
+
+Another major difference in PAL is that one action generates a number of partial actions ($\leq 1$) (see \autoref{fig:pal}).
+Each partial interaction generates a partial reward upon generating with the environment.
+Upon receiving a partial reward, the agent estimates the current state and triggers a new action.
+When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks (see \autoref{alg:grad}).
+
+\begin{figure}
+\begin{minipage}{\columnwidth}
+\includesvg{figures/Reinforcement_learning_diagram_PAL}
+\caption{Partial Action Actor Critic Learning: An action consists of zero or more partial actions which trigger partial rewards upon interacting with the environment.
+Each partial reward updates the state.
+The value and actor networks are updated upon receiving all partial rewards of one action.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
+\label{fig:pal}
+\end{minipage}
+\end{figure}
+
+\begin{algorithm}
+\caption{Partial Action Actor Critic Learning -- pseudocode for each learning thread.
+Only $\theta_{\text{a},\text{g}}$ and $\theta_{\text{v},\text{g}}$ are shared between the threads.
+They are initialized randomly in the beginning.}
+\label{alg:pal}
+\begin{algorithmic}[1]
+\Loop
+	\State $l_\text{actions} \gets  [] $
+	\State $l_\text{states} \gets  [] $
+	\State $l_\text{values} \gets  [] $
+	\State $l_\text{rewards} \gets  [] $
+	\State $l_\text{estimatedValues} \gets  [] $
+	\State $l_\text{snapshots} \gets  [] $
+	\State $t \gets 0$
+	\State $s_0 \gets \text{initialState}()$
+	\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
+	\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
+	\Repeat
+		\State $l_\text{states}.\text{append}(s_t)$
+		\State $l_\text{estimatedValues}.\text{append}(V(s_t; \theta_\text{v}))$
+		\If{$t \bmod t_\text{max} = 0$}
+			\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
+			\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
+			\State $l_\text{snapshots}.\text{append}((\theta_\text{a}, \theta_\text{v}))$
+		\EndIf
+		\State Sample $a_t$ from the 
+		\StateIndent[3] Actor Network's probability distribution
+		\State $l_\text{actions}.\text{append}(a_t)$
+		\State $l_{\text{partialActions},t} \gets \text{partialActions}(a_t)$
+		\ForAll{partial actions $a_{{\text{p},n,t}}$ in $l_{\text{partialActions},t}$}
+			\State Take partial action $a_{{\text{p},n,t}}$
+		\EndFor
+		\State Receive partial reward $r_{\text{p},n,t'}$
+		\StateIndent[3] where $t' \leq t$ and $0 \leq n \leq\#(l_{\text{partialActions},t'})$
+		\If{all partial rewards of $a_{t'}$ were received}
+			\State $r_{t'} \gets \text{combination of all partial rewards }r_{\text{p},n,t'}$
+			\State $l_\text{rewards}.\text{append}(r_{t'})$
+			\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
+				\State \Call{computeGradients}{}
+			\EndIf
+		\EndIf
+		\State Generate $s_{t+1}$ using $r_{\text{p}_{t'}}$
+		\State $t \gets t+1$
+	\Until{reaching the end the episode}
+\EndLoop
+%\algstore{palalg}
+\end{algorithmic}
+\end{algorithm}
+\begin{algorithm}
+\caption{Partial Action Actor Critic Learning -- procedure which computes and applies the gradients.}
+\label{alg:grad}
+\begin{algorithmic}[1]%
+%\algrestore{palalg}
+\Function{computeGradients}{}
+	\State $t_\text{end} = \min\left(t_\text{max}, \#\left( l_\text{rewards} \right)\right)$
+	\State $\theta_\text{backup} \gets \left(\theta_\text{a}, \theta_\text{v}\right)$
+	\State $\theta_\text{a}, \theta_\text{v} \gets l_\text{snapshots}[0]$
+	\State $R_{i+1} \gets \frac{l_\text{estimatedValues}[t_\text{end}]}{1-\gamma}$
+	\For{$i\gets t_\text{end}-1, 0$}
+		\State $R_i \gets r_i + \gamma R_{i+1}$
+		\State $a \gets l_\text{actions}[i]$
+		\State $s \gets l_\text{states}[i]$
+		\State $v \gets l_\text{values}[i]$
+		\State $d\theta_\text{v} \gets \frac{\partial\left(R_i - v\right)}{\partial\theta_v}$
+		\State $d\theta_\text{v} \gets -\frac{\partial\log \left( \pi_W\left( a-1 < W \leq a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
+		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta_a}$
+		\State $\theta_{\text{a},\text{g}} \gets \theta_{\text{a},\text{g}} + d\theta_\text{a}$
+		\State $\theta_{\text{v},\text{g}} \gets \theta_{\text{v},\text{g}} + d\theta_\text{v}$
+	\EndFor
+	\State Remove first $t_\text{end}$ elements from
+	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{estimatedValues}$
+	\State Remove first element from $l_\text{snapshots}$
+	\State $\theta_\text{a}, \theta_\text{v} = \theta_\text{backup}$
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Value Network}
+\label{subsec:genericvalue}
+
+The value network estimates the expected long-term reward given a state $s_t$.
+
+Let $r_t$ be the reward that was received at time $t$.
+$V(s_t; \theta_\text{v})$ designates the expected mean reward in state $s_t$ given the parameters of the value network $\theta_v$.
+Then, with $\gamma$ being the roll-off factor, which designates the influence that future reward has on the moving average, (set to $0.99$), we define the average reward as
+\begin{align*}
+R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k \frac{V(s_{t+k}; \theta_\text{v})}{1-\gamma}\right)\left(1-\gamma\right),
+\end{align*}
+where $k$ is upper-bounded by $t_\text{max}$ ($t_\text{max}$ is a fixed hyperparameter that indicates how many steps should be performed before updating the neural network).
+So $R$ is simply a moving average of rewards.
+However, usual moving averages take into account values from the past while this one uses values from the future; it runs reversely.
+
+The loss function, which the value network tries to minimize, is the square of the difference of the actual average reward received and the expected average reward
+\begin{align*}
+l_{\text{v}_t} = \left(R_t - V(s_t; \theta_\text{v})\right)^2.
+\end{align*}
+
+\subsection{Actor Network}
+\label{subsec:genericactor}
+
+Let $a_t$ be the action that was taken at time step $t$, $v_t$ the value that the value network estimated as the future reward given the current state $s_t$ at time step $t$ ($v_t$ is just an abbreviation for $V(s_t; \theta_\text{v})$) and $\theta_a$ the parameters of the actor network.
+Furthermore, let $\beta$ be a factor that specifies the importance of the entropy $H$.
+$\pi$ designates the probability density function, so $\pi\left( a_t \given s_t; \theta_\text{a} \right)$ is the value of the probability density function of taking action $a_t$ in state $s_t$ with the current weights of the actor network $\theta_\text{a}$.
+Then
+\begin{align*}
+l_{\text{a}_t} =& -\log \left( \pi\left( a_t \given s_t; \theta_\text{a} \right)\right)\left( R_t - v_t \right)\\ 
+&- \beta H\left( \pi\left( s_t; \theta_a \right)\right)
+\end{align*}
+is the actor loss that we try to minimize.
+
+\section{Congestion Control specifics}
+
+The classical Reinforcement Learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
+However, in the case of congestion control, it is desirable to perform a new action without having received a reward for the previous action.
+For example, imagine that you receive two acknowledgements directly after each other.
+For each of these two acknowledgements an action has to be performed but it does not seem feasible for the second action to wait until the first action has received a reward.
+The Asynchronous Actor Critic framework as described by \cite{mnih_asynchronous_2016} cannot be applied to congestion control as it assumes that actions and rewards are synchronized and thus we use the proposed Partial Action Actor Critic Learning (see \autoref{sec:pal})
 
@@
 \item[$\textit{a}_\text{t}$] Based on a given state and the history of previous states, the actor network returns an action $a_t$, which is an integer ($\geq 1$) that stands for the congestion window to be used until the next acknowledgement is received.
-\item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metrics. These are discussed in more detail in \fullref{subsec:value}
-\item[$\textit{v}_\text{t}$] The value is a tuple of the rewards estimated by the value network (see \fullref{subsec:value}) for each of the reward metrics. 
+\item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metrics. These are discussed in more detail in \autoref{subsec:value}
+\item[$\textit{v}_\text{t}$] The value is a tuple of the average expected rewards estimated by the value network (see \autoref{subsec:value}) for each of the reward metrics. 
 \end{description}
+
+We actually use four different types of reward: 
+\begin{description}
+\item[$\textit{r}_{\text{packet},t}$] is the number of the packets that the sender sent during time step $t$ and that were not lost (so they were acknowledged at some point by the receiver).
+\item[$\textit{r}_{\text{byte},t}$] is the sum of the bytes of the packets that the sender sent and that were not lost.
+Example: During $t$, three packets were sent with 300, 200, and 1500 bytes each so $\textit{r}_{\text{byte},t} = 2000$
+\item[$\textit{r}_{\text{delay},t}$] is the sum of the round trip times of the packets that the sender sent and that were not lost.
+\item[$\textit{r}_{\text{duration},t}$] is the sum of the time between receiving the last packet and receiving this packet (``inter-receive time'') for the packets that the sender sent and that were not lost.
+\end{description}
+
+%We actually use four moving averages for four different types of reward: 
+%\begin{description}
+%\item[$\textit{R}_\text{packet}$] is the moving average of the packets that the sender sent and that were not lost (so they were acknowledged at some point by the receiver).
+%\item[$\textit{R}_\text{byte}$] is the moving average of the bytes of the packets that the sender sent and that were not lost.
+%\item[$\textit{R}_\text{delay}$] is the moving average of the round trip time of the packets that the sender sent and that were not lost.
+%\item[$\textit{R}_\text{duration}$] is the moving average of the time between receiving the last packet and receiving this packet (``inter-receive time'') for the packets that the sender sent and that were not lost.
+%\end{description}
 
@@
 
-The value network estimates the expected reward given a state $s_t$.
-
-Let $r_t$ be the reward that was received at time $t$.
-$V(s_t; \theta_\text{v})$ designates the expected mean reward in state $s_t$ given the parameters of the value network $theta_v$.
-Then, with $\gamma$ being the roll-off factor, which designates the influence that future reward has on the moving average, (set to $0.99$), we generically define the expected n-step return as
-\begin{align*}
-R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k \frac{V(s_{t+k}; \theta_\text{v})}{1-\gamma}\right)\left(1-\gamma\right),
-\end{align*}
-where $k$ is upper-bounded by $t_\text{max}$.
-$R$ is simply a moving average of future rewards.
-However, usual moving averages take into account values from the past while this one uses values from the future; it runs reversely.
-
-The loss function, which the value network tries to minimize, is the square of the difference of the actual mean reward received and the expected mean reward
-\begin{align*}
-l_{\text{v}_t} = \left(R_t - V(s_t; \theta_\text{v})\right)^2.
-\end{align*}
-
-We actually use four moving averages for four different types of reward: 
-\begin{description}
-\item[$\textit{R}_\text{packet}$] is the moving average of the packets that the sender sent and that were not lost (so they were acknowledged at some point by the receiver)
-\item[$\textit{R}_\text{byte}$] is the moving average of the bytes of the packets that the sender sent and that were not lost
-\item[$\textit{R}_\text{delay}$] is the moving average of the round trip time of the packets that the sender sent and that were not lost
-\item[$\textit{R}_\text{duration}$] is the moving average of the time between receiving the last packet and receiving this packet (``inter-receive time'') for the packets that the sender sent and that were not lost
-\end{description}
-
-The loss function $l_{v_t}$ of the value network is actually the sum of the square of the difference for each of these four moving averages:
+In the case of congestion control, the loss function $l_{v_t}$ of the value network is actually the sum of the square of the difference for each of the four moving averages for each type of reward:
 \begin{align*}
@@
 
-$a_t$ is the sampled value $W$ at time $t$ at it is used as the window until the $t+1$, which is the reception of the next acknowledgement.
+$a_t$ is the sampled value $W$ at time $t$ and it is used as the window until the next time step $t+1$ begins (at the reception of the next acknowledgement).
 
@@
 &\left( \left(\frac{R_{\text{byte},t}}{{R_{\text{duration},t}}} - \frac{v_{\text{byte},t}}{{v_{\text{duration},t}}}\right) - \left( \frac{R_{\text{delay},t}}{{R_{\text{packet},t}}}- \frac{v_{\text{delay},t}}{{v_{\text{packet},t}}} \right)\right)\\
-&+ \beta H\left( \pi\left( l_\text{states}[i]; \theta_a \right)\right).
+&+ \beta H\left( \pi\left( s_t; \theta_a \right)\right).
 \end{align*}
-
-$\beta$ was set to $\frac 1 2 10^{-3}$ and $\delta$ to $10^{-3}$.
-
-Reinforcement learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
-However, in the case of congestion control, it is desirable to perform a new action without having received a reward for the previous action.
-For example, imagine that you receive two acknowledgements directly after each other.
-For each of these two acknowledgements an action has to be performed but it does not seem feasible for the second action to wait until the first action has received a reward.
-The Asynchronous Actor Critic framework as described by \cite{mnih_asynchronous_2016} cannot be applied to congestion control as it assumes that actions and rewards are synchronized.
-
-\begin{figure}
-\includesvg{figures/Reinforcement_learning_diagram}
-\caption{The classic reinforcement learning approach.\protect\footnotemark}
-\label{fig:reinforcement}
-\end{figure}
-\footnotetext{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}
-
-Thus we propose the Partial Action Actor Critic Learning (PAL) framework (see \autoref{fig:pal}) in which one action generates a number of partial actions ($leq 1$).
-Each partial interaction generates a partial reward upon generating with the environment.
-Upon receiving a partial reward, the agent estimates the current state and triggers a new action.
-When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks (see \autoref{alg:pal}).
-
-\begin{figure}
-\includesvg{figures/Reinforcement_learning_diagram_PAL}
-\caption{Partial Action Actor Critic Learning: An action consists of zero or more partial actions which trigger partial rewards upon interacting with the environment.
-Each partial reward updates the state.
-The value and actor networks are updated upon receiving all partial rewards of one action.\protect\footnotemark}
-\label{fig:pal}
-\end{figure}
-\footnotetext{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}
-
-\begin{algorithm}
-\caption{Partial Action Actor Critic Learning -- pseudocode for each learning thread.}
-\label{alg:pal}
-\begin{algorithmic}[1]
-\Loop
-	\State $l_\text{actions} \gets  [] $
-	\State $l_\text{states} \gets  [] $
-	\State $l_\text{values} \gets  [] $
-	\State $l_\text{rewards} \gets  [] $
-	\State $l_\text{estimatedValues} \gets  [] $
-	\State $l_\text{snapshots} \gets  [] $
-	\State $t \gets 0$
-	\State $s_0 \gets \text{initialState}()$
-	\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
-	\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
-	\Repeat
-		\State $l_\text{states}.\text{append}(s_t)$
-		\State $l_\text{estimatedValues}.\text{append}(V(s_t; \theta_\text{v}))$
-		\If{$t \bmod t_\text{max} = 0$}
-			\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
-			\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
-			\State $l_\text{snapshots}.\text{append}((\theta_\text{a}, \theta_\text{v}))$
-		\EndIf
-		\State Sample $a_t$ from the probability distribution
-		\State $l_\text{actions}.\text{append}(a_t)$
-		\State $l_{\text{partialActions},t} \gets \text{partialActions}(a_t)$
-		\ForAll{partial actions $a_{{\text{p},n,t}}$ in $l_{\text{partialActions},t}$}
-			\State Take partial action $a_{{\text{p},n,t}}$
-		\EndFor
-		\State Receive partial reward $r_{\text{p},n,t'}$
-		\StateIndent[3] where $t' \leq t$ and $0 \leq n \leq\#(l_{\text{partialActions},t'})$
-		\If{all partial rewards of $a_{t'}$ were received}
-			\State \Call{computeGradients}{}
-		\EndIf
-		\State Generate $s_{t+1}$ using $r_{\text{p}_{t'}}$
-		\State $t \gets t+1$
-	\Until{reaching the end the episode}
-\EndLoop
-%\algstore{palalg}
-\end{algorithmic}
-\end{algorithm}
-\begin{algorithm}
-\caption{Partial Action Actor Critic Learning -- procedure which computes and applies the gradients.}
-\label{alg:grad}
-\begin{algorithmic}[1]%
-%\algrestore{palalg}
-\Function{computeGradients}{}
-	\State $r_{t'} \gets \text{combination of all partial rewards }r_{\text{p},n,t'}$
-	\State $l_\text{rewards}.\text{append}(r_{t'})$
-	\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
-		\State $t_\text{end} = \min\left(t_\text{max}, \#\left( l_\text{rewards} \right)\right)$
-		\State $\theta_\text{backup} \gets \left(\theta_\text{a}, \theta_\text{v}\right)$
-		\State $\theta_\text{a}, \theta_\text{v} \gets l_\text{snapshots}[0]$
-		\State $R_{i+1} \gets \frac{l_\text{estimatedValues}[t_\text{end}]}{1-\gamma}$
-		\For{$i\gets t_\text{end}-1, 0$}
-			\State $R_i \gets r_i + \gamma r_{i+1}$
-			\State $a \gets l_\text{actions}[i]$
-			\State $s \gets l_\text{states}[i]$
-			\State $v \gets l_\text{values}[i]$
-			\State $d\theta_\text{v} \gets \frac{\partial\left(R_i - l_\text{values}[i]\right)}{\partial\theta_v}$
-			\State $d\theta_\text{v} \gets -\frac{\partial\log \left( \pi_W\left( a-1 < W \leq a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
-			\StateIndent[6] $-\frac{\partial\beta H\left( \pi\left( l_\text{states}[i]; \theta_a \right)\right)}{\partial\theta_a}$
-			\State $\theta_{\text{a},\text{g}} \gets \theta_{\text{a},\text{g}} + d\theta_\text{a}$
-			\State $\theta_{\text{v},\text{g}} \gets \theta_{\text{v},\text{g}} + d\theta_\text{v}$
-		\EndFor
-		\State $l_\text{actions} \gets l_\text{actions}[t_\text{max}:] $
-		\State $l_\text{states} \gets l_\text{states}[t_\text{max}:] $
-		\State $l_\text{values} \gets l_\text{values}[t_\text{max}:] $
-		\State $l_\text{rewards} \gets l_\text{rewards}[t_\text{max}:] $
-		\State $l_\text{estimatedValues} \gets l_\text{estimatedValues}[t_\text{max}:] $
-		\State $l_\text{snapshots} \gets l_\text{snapshots}[1:] $
-		\State $\theta_\text{a}, \theta_\text{v} = \theta_\text{backup}$
-	\EndIf
-\EndFunction
-\end{algorithmic}
-\end{algorithm}
 

### 2017-09-06 tex/paper.tex
Fix spelling mistakes

@@
 It consists of two Artificial Neural Networks (ANNs), the Actor Network and the Value Network (the Critic part in the abbreviation stands for the Value Network).
-Given a state, the Actor Networks outputs what it deems to be the optimum action in that certain state.
+Given a state, the Actor Networks outputs what it deems to be the optimum action to perform in that certain state.
 The Value Network estimates what long-term reward can be expected in this state.
@@
 In \autoref{alg:pal} we show the code that runs in one of the learning threads.
-As proposed in the literature, we use several concurrent thread to improve training performance.
+As proposed in literature, we use several concurrent thread to improve training performance.
 A major difference to previous approaches to reinforcement learning is the fact that we allow more actions to be taken before the previous actions' rewards have been received.
@@
 
-Another major difference in PAL is that one action generates a number of partial actions ($\leq 1$) (see \autoref{fig:pal}).
-Each partial interaction generates a partial reward upon generating with the environment.
-Upon receiving a partial reward, the agent estimates the current state and triggers a new action.
+Another major difference in PAL is that one action generates a number of partial actions ($\geq 1$) (see \autoref{fig:pal}).
+Each partial action generates a partial reward upon interacting with the environment.
+Upon receiving a partial reward, the agent determines the current state and triggers a new action.
 When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks (see \autoref{alg:grad}).
@@
 Let $r_t$ be the reward that was received at time $t$.
-$V(s_t; \theta_\text{v})$ designates the expected mean reward in state $s_t$ given the parameters of the value network $\theta_v$.
+$V(s_t; \theta_\text{v})$ designates the expected mean reward in state $s_t$ given the parameters (neural network weights) of the value network $\theta_v$.
 Then, with $\gamma$ being the roll-off factor, which designates the influence that future reward has on the moving average, (set to $0.99$), we define the average reward as
@@
 \begin{align*}
-l_{\text{v}_t} = \left(R_t - V(s_t; \theta_\text{v})\right)^2.
+l_{\text{v},t} = \left(R_t - V(s_t; \theta_\text{v})\right)^2.
 \end{align*}
@@
 
-Let $a_t$ be the action that was taken at time step $t$, $v_t$ the value that the value network estimated as the future reward given the current state $s_t$ at time step $t$ ($v_t$ is just an abbreviation for $V(s_t; \theta_\text{v})$) and $\theta_a$ the parameters of the actor network.
+Let $a_t$ be the action that was taken at time step $t$, $v_t$ the value that the value network estimated as the future reward given the current state $s_t$ at time step $t$ ($v_t$ is just an abbreviation for $V(s_t; \theta_\text{v})$) and $\theta_a$ the parameters (neural network weights) of the actor network.
 Furthermore, let $\beta$ be a factor that specifies the importance of the entropy $H$.
-$\pi$ designates the probability density function, so $\pi\left( a_t \given s_t; \theta_\text{a} \right)$ is the value of the probability density function of taking action $a_t$ in state $s_t$ with the current weights of the actor network $\theta_\text{a}$.
+$\pi$ designates the probability density function, which means that $\pi\left( a_t \given s_t; \theta_\text{a} \right)$ is the value of the probability density function of taking action $a_t$ in state $s_t$ with the current weights of the actor network $\theta_\text{a}$.
 Then
 \begin{align*}
-l_{\text{a}_t} =& -\log \left( \pi\left( a_t \given s_t; \theta_\text{a} \right)\right)\left( R_t - v_t \right)\\ 
+l_{\text{a},t} =& -\log \left( \pi\left( a_t \given s_t; \theta_\text{a} \right)\right)\left( R_t - v_t \right)\\ 
 &- \beta H\left( \pi\left( s_t; \theta_a \right)\right)
@@
 \item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metrics. These are discussed in more detail in \autoref{subsec:value}
-\item[$\textit{v}_\text{t}$] The value is a tuple of the average expected rewards estimated by the value network (see \autoref{subsec:value}) for each of the reward metrics. 
+\item[$\textit{v}_\text{t}$] The value is a tuple of the expected average reward estimated by the value network (see \autoref{subsec:value}) for each of the reward metrics. 
 \end{description}
@@
 
-In the case of congestion control, the loss function $l_{v_t}$ of the value network is actually the sum of the square of the difference for each of the four moving averages for each type of reward:
+In the case of congestion control, the loss function $l_{\text{v},t}$ of the value network is actually the sum of the square of the difference for each of the four moving averages for each type of reward:
 \begin{align*}
-l_{\text{v}_t} =& \left(R_{\text{packet},t} - V_\text{packet}(s_t; \theta_\text{v})\right)^2 \\
+l_{\text{v},t} =& \left(R_{\text{packet},t} - V_\text{packet}(s_t; \theta_\text{v})\right)^2 \\
 &+\left(R_{\text{byte},t} - V_\text{byte}(s_t; \theta_\text{v})\right)^2 \\
@@
 \begin{align*}
-l_{\text{a}_t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_\text{a} \right)\right)\\
+l_{\text{a},t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_\text{a} \right)\right)\\
 &\left( \left(\frac{R_{\text{byte},t}}{{R_{\text{duration},t}}} - \frac{v_{\text{byte},t}}{{v_{\text{duration},t}}}\right) - \left( \frac{R_{\text{delay},t}}{{R_{\text{packet},t}}}- \frac{v_{\text{delay},t}}{{v_{\text{packet},t}}} \right)\right)\\

### 2017-09-28 async_deep_reinforce/a3c.py
Completed transition to performing window updates, not learning the window itself

@@ -226,7 +226,7 @@ def call_process_action(thread_id, state):
   logging.debug(" ".join(map(str,("call_process_action", thread_id, state))))
   global sess, global_t, summary_writer, summary_op, summary_inputs
   chosen_action = training_threads[thread_id].action_step(sess, state)
-  return int(chosen_action)
+  return chosen_action
 
 def call_process_reward(thread_id, reward_throughput, reward_delay, duration):
   logging.debug(" ".join(map(str,("call_process_reward", thread_id, reward_throughput, reward_delay, duration))))

### 2017-09-28 async_deep_reinforce/a3c_training_thread.py

@@ -20,7 +20,7 @@ from constants import DELAY_MULTIPLIER
 import logging
 logging.basicConfig(level=LOG_LEVEL)
 
-LOG_INTERVAL = 10
+LOG_INTERVAL = 1
 
 class A3CTrainingThread(object):
   def __init__(self,
@@ -72,6 +72,7 @@ class A3CTrainingThread(object):
     return learning_rate
 
   def _record_score(self, sess, summary_writer, summary_op, summary_inputs, things, global_t):
+    # print("window in _record_score", self.windows, self.time_differences)
     summary_str = sess.run(summary_op, feed_dict={
       summary_inputs["score_throughput"]: things["score_throughput"],
       summary_inputs["score_delay"]: things["score_delay"],
@@ -229,8 +230,7 @@ class A3CTrainingThread(object):
 
     batch_si = []
     batch_ai = []
-    batch_td_throughput = []
-    batch_td_delay = []
+    batch_td = []
     batch_R_duration = []
     batch_R_packets = []
     batch_R_accumulated_delay = []
@@ -242,18 +242,17 @@ class A3CTrainingThread(object):
 
       R_packets = (ri[0] + GAMMA*R_packets)
       # TODO: In theory it should be np.log(R_bytes/R_duration) but remy doesn't have bytes
-      td_throughput = (np.log(R_packets/R_duration) - np.log(Vi[0]/Vi[2]))
+      td = (np.log(R_packets/R_duration) - np.log(Vi[0]/Vi[2]))
 
       R_accumulated_delay = (ri[1] + GAMMA*R_accumulated_delay)
       # R_delay = R_accumulated_delay/R_packets
       # td_delay = -(np.log(R_accumulated_delay/R_packets/DELAY_MULTIPLIER) - np.log(Vi[1]/Vi[0]/DELAY_MULTIPLIER))
-      td_delay = -DELAY_MULTIPLIER*(R_accumulated_delay/R_packets - Vi[1]/Vi[0])
+      td -= DELAY_MULTIPLIER*(R_accumulated_delay/R_packets - Vi[1]/Vi[0])
       # td_delay = -(R_accumulated_delay/R_packets - Vi[1]/Vi[0])
 
       batch_si.append(si)
       batch_ai.append(ai)
-      batch_td_throughput.append(td_throughput)
-      batch_td_delay.append(td_delay)
+      batch_td.append(td)
       batch_R_duration.append(R_duration*(1-GAMMA))
[…]

### 2017-09-28 async_deep_reinforce/game_ac_network.py

@@ -54,8 +54,7 @@ class GameACNetwork(object):
 			self.a = tf.placeholder(PRECISION, [None, 1], name="a")
 
 			# temporary difference (R-V) (input for policy)
-			self.td_throughput = tf.placeholder(PRECISION, [None], name="td_throughput")
-			self.td_delay = tf.placeholder(PRECISION, [None], name="td_delay")
+			self.td = tf.placeholder(PRECISION, [None], name="td")
 
 			self.distribution = ds.Normal(loc=self.pi[0], scale=self.pi[1], allow_nan_stats=False, validate_args=True)
 
@@ -79,7 +78,7 @@ class GameACNetwork(object):
 			self.entropy = ENTROPY_BETA * tf.reduce_sum(0.5 * tf.log(2.0*math.pi*math.e*self.pi[1]*self.pi[1]), axis=1)
 			# self.skewness = ENTROPY_BETA * tf.reduce_sum((tf.exp(self.distribution.distribution.variance()) + 2.0)*tf.sqrt(tf.exp(self.distribution.distribution.variance()) - 1.0), axis=1)
 			self.skewness = tf.constant(0.0, dtype=PRECISION)
-			self.actor_loss = tf.reduce_sum(self.distribution.log_prob(self.a), axis=1) * (self.td_throughput + self.td_delay)
+			self.actor_loss = tf.reduce_sum(self.distribution.log_prob(self.a), axis=1) * (self.td)
 			# self.actor_loss = tf.reduce_sum(tf.log(self.distribution.cdf(self.a) - self.distribution.cdf(tf.clip_by_value(self.a - 1.0, tiny, float("inf")))), axis=1) * (self.td_throughput + self.td_delay)
 
 			# self.policy_loss = - tf.reduce_sum(self.actor_loss + ENTROPY_BETA * self.entropy - SKEWNESS_GAMMA * self.skewness)
@@ -171,14 +170,14 @@ class GameACNetwork(object):
 
 	# weight initialization based on muupan's code
 	# https://github.com/muupan/async-rl/blob/master/a3c_ale.py
-	def _fc_variable(self, weight_shape, factor=1.0):
+	def _fc_variable(self, weight_shape, factor=1.0, bias_offset=0):
 		input_channels  = weight_shape[0]
 		output_channels = weight_shape[1]
 		d = 1.0 / np.sqrt(input_channels)
[…]

### 2017-09-28 src/rainbow.cc

@@ -96,7 +96,8 @@ double Rainbow::get_action(const long unsigned int thread_id, const vector<doubl
 		PyErr_Print();
 	}
 
-	int action = (int) PyLong_AsLong(pActionReturnValue);
+	double action = PyFloat_AsDouble(pActionReturnValue);
+	// printf("action: %f\n", action);
 	Py_DECREF(pActionReturnValue);
 	Py_DECREF(pArgs);
 	Py_DECREF(pState);

### 2017-09-28 src/unicorn.cc

@@ -187,6 +187,7 @@ void Unicorn::get_action(const double& tickno, const int& packets_sent_in_this_e
   // printf("%lu: action is: %f\n", _thread_id, action);
   _put_actions += 1;
 
+  // printf("%lu: window=%f, action=%f\n", _thread_id, _the_window, action);
   // _the_window = window(_the_window, action.window_increment, action.window_multiple);
   _the_window = std::min(std::max(_the_window + action, MIN_WINDOW), MAX_WINDOW);
   // if (!_training) printf("%lu: window: %d\n", _thread_id, _the_window);
@@ -210,7 +211,7 @@ void Unicorn::finishFlow() {
   // const bool at_least_one_packet_sent = true;
   // printf("%lu: finish, _packets_sent: %u\n", _thread_id, _packets_sent);
   // _rainbow.finish(_thread_id, {_memory.field(0), _memory.field(1), _memory.field(2), _memory.field(3), _memory.field(6), (double)_the_window/WINDOW_NORMALIZER}, at_least_one_packet_sent);
-  printf("%lu: window=%f\n", _thread_id, _the_window);
+  printf("%lu: Finishing, window=%f\n", _thread_id, _the_window);
   _rainbow.finish(_thread_id, _outstanding_rewards.size(), _memory._last_tick_received-_start_tick, _the_window);
   // printf("%lu: actions: %lu, rewards: %lu, outstanding_rewards: %lu\n", _thread_id, _put_actions, _put_rewards, _outstanding_rewards.size());
   // _put_actions -= _outstanding_rewards.size();

### 2017-09-28 async_deep_reinforce/a3c.py
Fixed logging, remove old junk

@@ -122,42 +122,33 @@ with tf.device(device):
   score_throughput = tf.placeholder(PRECISION)
   score_delay = tf.placeholder(PRECISION)
   entropy = tf.placeholder(PRECISION)
-  skewness = tf.placeholder(PRECISION)
   actor_loss = tf.placeholder(PRECISION)
   value_loss = tf.placeholder(PRECISION)
   total_loss = tf.placeholder(PRECISION)
   window = tf.placeholder(PRECISION)
   window_increase = tf.placeholder(PRECISION)
   std = tf.placeholder(PRECISION)
-  inner_mean = tf.placeholder(PRECISION)
-  inner_std = tf.placeholder(PRECISION)
   speed = tf.placeholder(PRECISION)
   tf.summary.scalar("score_throughput", score_throughput)
   tf.summary.scalar("score_delay", score_delay)
   tf.summary.scalar("entropy", entropy)
-  tf.summary.scalar("skewness", skewness)
   tf.summary.scalar("actor_loss", actor_loss)
   tf.summary.scalar("value_loss", value_loss)
   tf.summary.scalar("total_loss", total_loss)
   tf.summary.scalar("window", window)
   tf.summary.scalar("window_increase", window_increase)
   tf.summary.scalar("std", std)
-  tf.summary.scalar("inner_mean", inner_mean)
-  tf.summary.scalar("inner_std", inner_std)
   tf.summary.scalar("speed", speed)
   summary_inputs = {
     "score_throughput": score_throughput,
     "score_delay": score_delay,
     "entropy": entropy,
-    "skewness": skewness,
     "actor_loss": actor_loss,
     "value_loss": value_loss,
     "total_loss": total_loss,
     "window": window,
     "window_increase": window_increase,
     "std": std,
-    "inner_mean": inner_mean,
-    "inner_std": inner_std,
     "speed": speed
   }
 

### 2017-09-28 async_deep_reinforce/a3c_training_thread.py

@@ -77,15 +77,12 @@ class A3CTrainingThread(object):
       summary_inputs["score_throughput"]: things["score_throughput"],
       summary_inputs["score_delay"]: things["score_delay"],
       summary_inputs["entropy"]: things["entropy"],
-      summary_inputs["skewness"]: things["skewness"],
       summary_inputs["actor_loss"]: things["actor_loss"],
       summary_inputs["value_loss"]: things["value_loss"],
       summary_inputs["total_loss"]: things["total_loss"],
       summary_inputs["window"]: things["window"],
       summary_inputs["window_increase"]: things["window_increase"],
       summary_inputs["std"]: things["std"],
-      summary_inputs["inner_mean"]: things["inner_mean"],
-      summary_inputs["inner_std"]: things["inner_std"],
       summary_inputs["speed"]: things["speed"]
     })
     summary_writer.add_summary(summary_str, global_t)
@@ -325,20 +322,17 @@ class A3CTrainingThread(object):
         }
         feed_dict.update(var_dict)
 
-        entropy, skewness, actor_loss, value_loss, total_loss, window_increase, std, inner_mean, inner_std = self.local_network.run_loss(sess, feed_dict)
+        entropy, actor_loss, value_loss, total_loss, window_increase, std = self.local_network.run_loss(sess, feed_dict)
 
         things = {"score_throughput": normalized_final_score_throughput,
           "score_delay": normalized_final_score_delay,
           "actor_loss": actor_loss.item(),
           "value_loss": value_loss,
           "entropy": entropy.item(),
-          "skewness": skewness.item(),
           "total_loss": total_loss,
           "window_increase": window_increase.item(),
           "window": self.windows[0],
           "std": std.item(),
-          "inner_mean": inner_mean.item(),
-          "inner_std": inner_std.item(),
           "speed": steps_per_sec}
         # logging.debug(" ".join(map(str,("things", things))))
         self._record_score(sess, summary_writer, summary_op, summary_inputs, things, global_t)

### 2017-09-28 async_deep_reinforce/constants.py

@@ -20,7 +20,7 @@ PRECISION = tf.float64
 
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
 GAMMA = 0.99 # discount factor for rewards
-ENTROPY_BETA = 1e-4
+ENTROPY_BETA = 1e-5
 MAX_TIME_STEP = 10 * 1e7
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True

### 2017-09-28 async_deep_reinforce/game_ac_network.py

@@ -66,22 +66,15 @@ class GameACNetwork(object):
 
 			self.distribution_mean = self.pi[0]
 			self.distribution_std = self.pi[1]
-			# self.inner_distribution_mean = self.distribution.distribution.mean()
-			# self.inner_distribution_std = self.distribution.distribution.stddev()
-			self.inner_distribution_mean = self.pi[0]
-			self.inner_distribution_std = self.pi[1]
 			# self.chosen_action = tf.ceil(self.distribution.sample())
 			self.chosen_action = self.distribution.sample()
 
 			# policy entropy
 			# self.entropy = ENTROPY_BETA * tf.reduce_sum(self.distribution.distribution.mean() + 0.5 * tf.log(2.0*math.pi*math.e*self.distribution.distribution.variance()), axis=1)
 			self.entropy = ENTROPY_BETA * tf.reduce_sum(0.5 * tf.log(2.0*math.pi*math.e*self.pi[1]*self.pi[1]), axis=1)
-			# self.skewness = ENTROPY_BETA * tf.reduce_sum((tf.exp(self.distribution.distribution.variance()) + 2.0)*tf.sqrt(tf.exp(self.distribution.distribution.variance()) - 1.0), axis=1)
-			self.skewness = tf.constant(0.0, dtype=PRECISION)
 			self.actor_loss = tf.reduce_sum(self.distribution.log_prob(self.a), axis=1) * (self.td)
 			# self.actor_loss = tf.reduce_sum(tf.log(self.distribution.cdf(self.a) - self.distribution.cdf(tf.clip_by_value(self.a - 1.0, tiny, float("inf")))), axis=1) * (self.td_throughput + self.td_delay)
 
-			# self.policy_loss = - tf.reduce_sum(self.actor_loss + ENTROPY_BETA * self.entropy - SKEWNESS_GAMMA * self.skewness)
 			self.policy_loss = - ACTOR_FACTOR * tf.reduce_sum(self.actor_loss + self.entropy)
 			# self.policy_loss = - ACTOR_FACTOR * tf.reduce_sum(self.actor_loss)
 
@@ -342,7 +335,7 @@ class GameACLSTMNetwork(GameACNetwork):
 
 	def run_loss(self, sess, feed_dict):
 		# We don't have to roll back the LSTM state here as it is restored in the "process" function of a3c_training_thread.py anyway and because run_loss is only called there.
[…]

### 2017-09-28 src/unicorn.cc

@@ -211,7 +211,7 @@ void Unicorn::finishFlow() {
   // const bool at_least_one_packet_sent = true;
   // printf("%lu: finish, _packets_sent: %u\n", _thread_id, _packets_sent);
   // _rainbow.finish(_thread_id, {_memory.field(0), _memory.field(1), _memory.field(2), _memory.field(3), _memory.field(6), (double)_the_window/WINDOW_NORMALIZER}, at_least_one_packet_sent);
-  printf("%lu: Finishing, window=%f\n", _thread_id, _the_window);
+  // printf("%lu: Finishing, window=%f\n", _thread_id, _the_window);
   _rainbow.finish(_thread_id, _outstanding_rewards.size(), _memory._last_tick_received-_start_tick, _the_window);
   // printf("%lu: actions: %lu, rewards: %lu, outstanding_rewards: %lu\n", _thread_id, _put_actions, _put_rewards, _outstanding_rewards.size());
   // _put_actions -= _outstanding_rewards.size();

### 2017-10-02 async_deep_reinforce/a3c.py
Bug fixes and decrease precision to float instead of double

@@ -177,7 +177,7 @@ def create_training_thread(training):
     created_thread = A3CTrainingThread(global_thread_index, global_network, initial_learning_rate,
                                         learning_rate_input,
                                         grad_applier, MAX_TIME_STEP,
-                                        device, training)
+                                        device, training, cooperative)
     training_threads[global_thread_index] = created_thread
     return_index = global_thread_index
     global_thread_index += 1
@@ -206,7 +206,8 @@ def create_training_thread(training):
   created_thread.episode_reward_throughput = 0
   created_thread.episode_reward_delay = 0
   created_thread.local_network.reset_state()
-  sess.run( created_thread.sync )
+  if cooperative:
+    sess.run( created_thread.sync )
 
   return return_index
 

### 2017-10-02 async_deep_reinforce/a3c_training_thread.py

@@ -18,6 +18,8 @@ from constants import LOG_LEVEL
 from constants import DELAY_MULTIPLIER
 
 import logging
+
+logging.info(" ".join(map(str,("DELAY_MULTIPLIER", DELAY_MULTIPLIER))))
 logging.basicConfig(level=LOG_LEVEL)
 
 LOG_INTERVAL = 1
@@ -31,9 +33,11 @@ class A3CTrainingThread(object):
                grad_applier,
                max_global_time_step,
                device,
-               training):
+               training,
+               cooperative):
 
     self.training = training
+    self.cooperative = cooperative
     self.thread_index = thread_index
     self.learning_rate_input = learning_rate_input
     self.max_global_time_step = max_global_time_step
@@ -50,9 +54,14 @@ class A3CTrainingThread(object):
         aggregation_method=None,
         colocate_gradients_with_ops=False)
 
-    self.apply_gradients = grad_applier.apply_gradients(
-      zip(self.gradients, global_network.get_vars())
-    )
+      if self.cooperative:
+        self.apply_gradients = grad_applier.apply_gradients(
+          zip(self.gradients, global_network.get_vars())
+        )
+      else:
+        self.apply_gradients = grad_applier.apply_gradients(
+          zip(self.gradients, local_network.get_vars())
+        )
 
     self.sync = self.local_network.sync_from(global_network)
     self.episode_count = 0
@@ -101,13 +110,14 @@ class A3CTrainingThread(object):
         self.time_differences.append(None)
         self.windows.append(None)
         # Sync for the next iteration
-        sess.run( self.sync )
+        if self.cooperative:
+          sess.run( self.sync )
         self.start_lstm_states.append(self.local_network.lstm_state_out)
         self.variable_snapshots.append(sess.run(self.local_network.get_vars()))
 
     pi_, action, value_ = self.local_network.run_policy_action_and_value(sess, state)
 
-    logging.debug(" ".join(map(str,(self.thread_index,"pi_values:",pi_))))
+    # logging.debug(" ".join(map(str,(self.thread_index,"pi_values:",pi_))))
 
     if self.training:
       self.states.append(state)
@@ -184,9 +194,6 @@ class A3CTrainingThread(object):
     # self.start_lstm_states = []
     # self.variable_snapshots = []
     self.local_network.reset_state()
-
-    if not self.training:
-      sess.run(self.sync)
 
   def process(self, sess, global_t, summary_writer, summary_op, summary_inputs, time_difference=None):
     if self.local_t <= 0:

### 2017-10-02 async_deep_reinforce/constants.py

@@ -17,7 +17,7 @@ LEARNING_RATE_MULTIPLIER = 0.5e-1
 INITIAL_ALPHA_LOW = 1e-4*LEARNING_RATE_MULTIPLIER    # log_uniform low limit for learning rate
 INITIAL_ALPHA_HIGH = 1e-2*LEARNING_RATE_MULTIPLIER   # log_uniform high limit for learning rate
 
-PRECISION = tf.float64
+PRECISION = tf.float32
 
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
 GAMMA = 0.99 # discount factor for rewards
@@ -26,7 +26,7 @@ STD_BIAS_OFFSET = -3
 MAX_TIME_STEP = 10**7
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
-N_LSTM_LAYERS = 4
+N_LSTM_LAYERS = 3
 
 STATE_SIZE = 11
 HIDDEN_SIZE = 256

### 2017-10-02 tex/paper.tex
Simplified formula
Update tex and figure

@@
 Another major difference in PAL is that one action generates a number of partial actions ($\geq 1$) (see \autoref{fig:pal}).
-Each partial action generates a partial reward upon interacting with the environment.
-Upon receiving a partial reward, the agent determines the current state and triggers a new action.
+Each partial action generates feedback upon interacting with the environment.
+Upon receiving feedback for a partial action, the agent determines the current state and triggers a new action.
 When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks (see \autoref{alg:grad}).
@@
 		\EndFor
-		\State Receive partial reward $r_{\text{p},n,t'}$
+		\State Receive feedback $r_{\text{p},n,t'}$
 		\StateIndent[3] where $t' \leq t$ and $0 \leq n \leq\#(l_{\text{partialActions},t'})$
-		\If{all partial rewards of $a_{t'}$ were received}
-			\State $r_{t'} \gets \text{combination of all partial rewards }r_{\text{p},n,t'}$
+		\If{all feedback of $a_{t'}$ was received}
+			\State $r_{t'} \gets \text{reward w.r.t all feedback }r_{\text{p},n,t'}$
 			\State $l_\text{rewards}.\text{append}(r_{t'})$
@@
 		\State $d\theta_\text{v} \gets \frac{\partial\left(R_i - v\right)}{\partial\theta_v}$
-		\State $d\theta_\text{v} \gets -\frac{\partial\log \left( \pi_W\left( a-1 < W \leq a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
+		\State $d\theta_\text{v} \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
 		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta_a}$
@@
 \begin{align*}
-R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k \frac{V(s_{t+k}; \theta_\text{v})}{1-\gamma}\right)\left(1-\gamma\right),
+R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta_\text{v})\right),
 \end{align*}
@@
 \item[$\textit{a}_\text{t}$] Based on a given state and the history of previous states, the actor network returns an action $a_t$, which is an integer ($\geq 1$) that stands for the congestion window to be used until the next acknowledgement is received.
-\item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metrics. These are discussed in more detail in \autoref{subsec:value}
+\item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metric. These are discussed in more detail in \autoref{subsec:value}
 \item[$\textit{v}_\text{t}$] The value is a tuple of the expected average reward estimated by the value network (see \autoref{subsec:value}) for each of the reward metrics. 
@@
 
-The actor network outputs two parameters: The mean of a log-normal distribution $\mu_\text{log}$ and its standard deviation $\sigma_\text{log}$.
+The actor network outputs two parameters: The mean of a normal distribution $\mu$ and its standard deviation $\sigma$.
 
-We calculate the means and standard deviation of the underlying normal distribution as
+Each time an action $a_t$ is requested, a value $X$ is sampled from the current normal distribution defined by the parameters $\mu$ and $\sigma$: 
 \begin{align*}
-\sigma_\text{normal} &= \sqrt{\log \left( \frac{\sigma_\text{log}}{\mu_\text{log}} +1 \right)} \\
-\mu_\text{normal} &= \log\left(\mu_\text{log}\right) - \frac{{\sigma_\text{normal}}^2}{2}.
+X&\sim {\mathcal {N}}(\mu ,\sigma^{2})\\
 \end{align*}
 
-Each time an action $a_t$ is requested, a value $X$ is sampled from the current log-normal distribution defined by the parameters $\mu_\text{normal}$ and $\sigma_\text{normal}$: 
-\begin{align*}
-Z&\sim {\mathcal {N}}(\mu_\text{normal} ,{\sigma_{\text{normal}}}^{2})\\
-X&=\exp(Z)
-\end{align*}
-
-We can also equally write
-\begin{align*}
-X \sim \text{Lognormal}\left( \mu_\text{normal}, {\sigma_{\text{normal}}}^{2}\right).
-\end{align*}
-
-The sampled value $X$ is then discretized by rounding up, which also prevents window sizes smaller than one from occurring:
-\begin{align*}
-W = \left\lceil X \right\rceil
-\end{align*}
-
-$a_t$ is the sampled value $W$ at time $t$ and it is used as the window until the next time step $t+1$ begins (at the reception of the next acknowledgement).
-
-%Let
-%\begin{align*}
-%\Sigma R_t = \sum_{i \in \left \{\text{packet}, \text{byte}, \text{delay}, \text{duration}\right \}} R_{i_t}
-%\end{align*}
-%and
-%\begin{align*}
-%\Sigma v_t = \sum_{i \in \left \{\text{packet}, \text{byte}, \text{delay}, \text{duration}\right \}} v_{i_t}.
-%\end{align*}
+$a_t$ is the sampled value $X$ at time $t$ and the window is incremented by $a_t$; however, the window can never be smaller than 1. At the beginning of a flow the window is 1 as well.
 
@@
 \begin{align*}
-l_{\text{a},t} =& -\log \left( \pi_W\left( a_t-1 < W \leq a_t \given s_t; \theta_\text{a} \right)\right)\\
+l_{\text{a},t} =& -\log \left( \pi \left( a_t \given s_t ; \theta_\text{a} \right) \right)\\
 &\left( \left(\frac{R_{\text{byte},t}}{{R_{\text{duration},t}}} - \frac{v_{\text{byte},t}}{{v_{\text{duration},t}}}\right) - \left( \frac{R_{\text{delay},t}}{{R_{\text{packet},t}}}- \frac{v_{\text{delay},t}}{{v_{\text{packet},t}}} \right)\right)\\

### 2017-10-03 tex/paper.tex
Added petri net to paper

@@
 \usepackage{svg}
+
+\def\xcolorversion{2.00}
+\def\xkeyvalversion{1.8}
+
+\usepackage{pgf}
+\usepackage{tikz}
+\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}
 
@@
 
+\begin{figure}
+
+\begin{tikzpicture}[node distance=2cm,>=stealth',bend angle=45,auto]
+
+  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
+  \tikzstyle{red place}=[place,draw=red!75,fill=red!20]
+  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
+  			  fill=black!20,minimum size=4mm]
+
+%  \tikzstyle{every label}=[blue!75]
+
+  \begin{scope}
+    \node [place,tokens=1] (s1) at (2,4) [label=Received feedback] {};
+    \node [place,tokens=1] (s2) at (2,0) [label={[align=center]below:Open\\congestion\\window}] {};
+
+    \node [transition] (e2) at (0,2) [label=left:Take action] {}
+      edge [post,bend right] node {$a$}(s2)
+      edge [pre,bend left] node {1} (s1);
+
+    \node [transition] (e1) at (4,2) [label={[align=left]right:Send packet \& \\ receive ACK}] {}
+      edge [pre,bend left] node {1} (s2)
+      edge [post,bend right] node {1} (s2)
+      edge [post, bend right] node {1} (s1);
+
+  \end{scope}
+
+\end{tikzpicture}
+
+\caption{A petri net describing the congestion control mechanism.
+We start with one token in the state \textit{open congestion window}, which means that initially we can send one packet.
+We start with another token in the state \textit{Received feedback}, which means that we can take an action right in the beginning.
+The concept is the following: If there is at least one token in the \textit{open congestion window}, send a packet.
+The acknowledgement (or timeout if the packet gets lost) for the packet is received by the \textit{open congestion window}, meaning that when we get an ACK we can send another packet (The \textit{open congestion window} signifies the congestion window minus the packets that are currently unacknowledged.).
+Furthermore the state \textit{Received feedback} receives each ACK/timeout and takes an action.
+\textit{Take action} adds \textit{a} tokens to the \textit{open congestion window}, where $a$ is a real number (possibly also negative); thus in each state there can also be a real number of tokens (e.g.~ 2.34 tokens are possible in the \textit{open congestion window}).
+However, we define that the congestion window can never be smaller than 1.}
+\label{fig:petri}
+\end{figure}
+
 The classical Reinforcement Learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
@@
 For each of these two acknowledgements an action has to be performed but it does not seem feasible for the second action to wait until the first action has received a reward.
-The Asynchronous Actor Critic framework as described by \cite{mnih_asynchronous_2016} cannot be applied to congestion control as it assumes that actions and rewards are synchronized and thus we use the proposed Partial Action Actor Critic Learning (see \autoref{sec:pal})
+The Asynchronous Actor Critic framework as described by \cite{mnih_asynchronous_2016} cannot be applied to congestion control as it assumes that actions and rewards are synchronized and thus we use the proposed Partial Action Actor Critic Learning (see \autoref{sec:pal}).
+We describe the concept of using reinforcement learning for congestion control with a petri net (see \autoref{fig:petri}).
 

### 2017-10-04 tex/paper.tex
Add new config files

@@
   \begin{scope}
-    \node [place,tokens=1] (s1) at (2,4) [label=Received feedback] {};
-    \node [place,tokens=1] (s2) at (2,0) [label={[align=center]below:Open\\congestion\\window}] {};
+    \node [place,tokens=1] (s1) [label=Received feedback] {};
+    \node [place] (s2) [below of=s1, label=Network] {};
+    \node [place,tokens=1] (s3) [below of=s2, label={[align=center]below:Open\\congestion\\window}] {};
 
-    \node [transition] (e2) at (0,2) [label=left:Take action] {}
-      edge [post,bend right] node {$a$}(s2)
+    \node [transition] (e2) [left of=s2, label=left:Take action] {}
+      edge [post,bend right] node {$a$}(s3)
       edge [pre,bend left] node {1} (s1);

### 2017-10-04 async_deep_reinforce/a3c.py

@@ -247,7 +247,7 @@ def call_process_reward(thread_id, reward_throughput, reward_delay, duration):
   diff_global_t = training_threads[thread_id].reward_step(sess, global_t, summary_writer, summary_op, summary_inputs, reward_throughput, reward_delay, duration)
   global_t += diff_global_t
   if global_t >= MAX_TIME_STEP:
-    print("Reached maximum time step, saving and terminating...")
+    logging.info("Reached maximum time step, saving and terminating...")
     save_session()
     sys.exit()
 
@@ -257,7 +257,7 @@ def call_process_finished(thread_id, actions_to_remove, time_difference, window)
   training_threads[thread_id].final_step(sess, global_t, summary_writer, summary_op, summary_inputs, actions_to_remove, time_difference, window)
 
 def save_session():
-  logging.debug("save_session")
+  # logging.debug("save_session")
   global global_t, sess, CHECKPOINT_DIR
   if not os.path.exists(CHECKPOINT_DIR):
     os.mkdir(CHECKPOINT_DIR)

### 2017-10-04 src/aimd.cc

@@ -25,7 +25,7 @@ void Aimd::packets_received( const vector< Packet > & packets ) {
     loss_detected = ( not loss_detected ) ?
                     ( packet.seq_num > _largest_ack + 1 ): /* At least lost one packet */
                     ( true ); /* Bypass seq_num check if you saw a loss already */
-    _largest_ack = max( _largest_ack, packet.seq_num );
+    _largest_ack = max( _largest_ack, (int) packet.seq_num );
 
     _packets_received++;
     if ( packet.flow_id != _flow_id ) {

### 2017-10-04 src/fish.cc

@@ -26,7 +26,7 @@ Fish::Fish( const FinTree & fins, const unsigned int s_prng_seed, const bool s_t
 void Fish::packets_received( const vector< Packet > & packets ) {
     _packets_received += packets.size();
     _memory.packets_received( packets, _flow_id, _largest_ack );
-    _largest_ack = max( packets.at( packets.size() - 1 ).seq_num, _largest_ack );
+    _largest_ack = max( (int) packets.at( packets.size() - 1 ).seq_num, _largest_ack );
 
     const Fin & current_fin( _fins.use_fin( _memory, _track ) );
     _update_lambda( current_fin.lambda() );

### 2017-10-04 src/rat.cc

@@ -23,7 +23,7 @@ void Rat::packets_received( const vector< Packet > & packets ) {
   _packets_received += packets.size();
   /* Assumption: There is no reordering */
   _memory.packets_received( packets, _flow_id, _largest_ack );
-  _largest_ack = max( packets.at( packets.size() - 1 ).seq_num, _largest_ack );
+  _largest_ack = max( (int) packets.at( packets.size() - 1 ).seq_num, _largest_ack );
 
   const Whisker & current_whisker( _whiskers.use_whisker( _memory, _track ) );
 

### 2017-10-04 src/unicorn.cc

@@ -201,7 +201,7 @@ void Unicorn::get_action(const double& tickno, const int& packets_sent_in_this_e
 
   // printf("%lu: window=%f, action=%f\n", _thread_id, _the_window, action);
   // _the_window = window(_the_window, action.window_increment, action.window_multiple);
-  _the_window = std::min(std::max(_the_window + action, MIN_WINDOW_UNICORN), MAX_WINDOW_UNICORN);
+  _the_window = std::max(_the_window + action, MIN_WINDOW_UNICORN);
   // if (!_training) printf("%lu: window: %d\n", _thread_id, _the_window);
   // printf("%lu: after update: window=%f, action=%f\n", _thread_id, _the_window, action);
   for (auto it=_outstanding_rewards.begin(); it!=_outstanding_rewards.end(); it++) {

### 2017-10-08 async_deep_reinforce/a3c_training_thread.py
Update config files

@@ -71,11 +71,11 @@ class A3CTrainingThread(object):
     return self.local_network.get_vars()
 
   def _anneal_learning_rate(self, global_time_step):
-    # learning_rate = self.initial_learning_rate * (self.max_global_time_step - global_time_step) / self.max_global_time_step
-    # if learning_rate < 0.0:
-    #   learning_rate = 0.0
-    # return learning_rate
-    return self.initial_learning_rate
+    learning_rate = self.initial_learning_rate * (self.max_global_time_step - global_time_step) / self.max_global_time_step
+    if learning_rate < 0.0:
+      learning_rate = 0.0
+    return learning_rate
+    # return self.initial_learning_rate
 
   def _record_score(self, sess, summary_writer, summary_op, summary_inputs, things, global_t):
     # print("window in _record_score", self.windows, self.time_differences)

### 2017-10-08 async_deep_reinforce/constants.py

@@ -30,7 +30,7 @@ INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate
 GAMMA = 0.99 # discount factor for rewards
 ENTROPY_BETA = 1e-4
 STD_BIAS_OFFSET = -1
-MAX_TIME_STEP = 1e9
+MAX_TIME_STEP = 1e7
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
 N_LSTM_LAYERS = 3

### 2017-10-08 async_deep_reinforce/game_ac_network.py

@@ -231,9 +231,9 @@ class GameACLSTMNetwork(GameACNetwork):
 			self.W_hidden_to_action_std_fc, self.b_hidden_to_action_std_fc = self._fc_variable([HIDDEN_SIZE, 1], factor=1.0, bias_offset=STD_BIAS_OFFSET)
 
 			# weight for value output layer
-			self.W_hidden_to_value_packets_fc, self.b_hidden_to_value_packets_fc = self._fc_variable([HIDDEN_SIZE, 1], bias_range=(0, 0), bias_offset=PACKETS_BIAS_OFFSET)
-			self.W_hidden_to_value_delay_fc, self.b_hidden_to_value_delay_fc = self._fc_variable([HIDDEN_SIZE, 1], bias_range=(0, 0), bias_offset=DELAY_BIAS_OFFSET)
-			self.W_hidden_to_value_duration_fc, self.b_hidden_to_value_duration_fc = self._fc_variable([HIDDEN_SIZE, 1], bias_range=(0, 0), bias_offset=INTER_PACKET_ARRIVAL_TIME_OFFSET)
+			self.W_hidden_to_value_packets_fc, self.b_hidden_to_value_packets_fc = self._fc_variable([HIDDEN_SIZE, 1], bias_offset=PACKETS_BIAS_OFFSET)
+			self.W_hidden_to_value_delay_fc, self.b_hidden_to_value_delay_fc = self._fc_variable([HIDDEN_SIZE, 1], bias_offset=DELAY_BIAS_OFFSET)
+			self.W_hidden_to_value_duration_fc, self.b_hidden_to_value_duration_fc = self._fc_variable([HIDDEN_SIZE, 1], bias_offset=INTER_PACKET_ARRIVAL_TIME_OFFSET)
 
 			# state (input)
 			self.s = tf.placeholder(PRECISION, [None, STATE_SIZE])

### 2017-10-08 tex/paper.tex
Update config files
Add a petri net and an example to the tex file.

--- 2017-10-02 tex/paper.tex
@@
 
-\usepackage{todonotes}
+%\usepackage{todonotes}
 \usepackage{svg}
+
+\def\xcolorversion{2.00}
+\def\xkeyvalversion{1.8}
+
+\usepackage{pgf}
+\usepackage{tikz}
+\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}
+
+\newcommand\note[2]{{\color{#1}#2}}
+\newcommand\todo[1]{{\note{red}{TODO: #1}}}
 
@@
 Only $\theta_{\text{a},\text{g}}$ and $\theta_{\text{v},\text{g}}$ are shared between the threads.
-They are initialized randomly in the beginning.}
+They are initialized randomly in the beginning.
+\todo{Update algorithm}}
 \label{alg:pal}
@@
 
+\begin{figure}
+
+\newcommand{\nodedistance}{2}
+\newcommand{\nd}[1]{\numexpr #1 * \nodedistance \relax}
+
+\begin{tikzpicture}[node distance=\nodedistance,>=stealth',bend angle=45,auto]
+
+  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
+  \tikzstyle{red place}=[place,draw=red!75,fill=red!20]
+  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
+  			  fill=black!20,minimum size=4mm]
+
+%  \tikzstyle{every label}=[blue!75]
+
+  \begin{scope}
+    \node [place,tokens=1] (s1) at (\nd{1}, \nd{3}) [label=Received feedback] {};
+    \node [place] (s2) at (\nd{1}, \nd{2}) [label={[align=center]left:Received\\ACKs}] {};
+    \node [place] (s3) at (\nd{3}, \nd{2}) [label={[align=center]above:Packets\\in the\\network}] {};
+    \node [place,tokens=1] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open\\congestion\\window\\(cwnd)}] {};
+
+    \node [transition] (e1) at (\nd{0}, \nd{1}) [label=left:Take action] {}
+      edge [post,bend right] node {$a$}(s4)
+      edge [pre,bend left] node {1} (s1);
+
+    \node [transition] (e2) at (\nd{2}, \nd{1}) [label=Send packet] {}
+      edge [pre,bend left] node {1} (s4)
+      edge [post,bend right] node {1} (s3);
+
+    \node [transition] (e3) at (\nd{2}, \nd{2}) [label=below:Receive ACK] {}
+      edge [pre] node {1} (s3)
+      edge [post] node {1} (s2);
+
+    \node [transition] (e4) at (\nd{1}, \nd{1}) [label={[align=center]left:Update\\open\\cwnd}] {}
+      edge [pre] node {1} (s2)
+      edge [post] node {1} (s4);
+
+    \node [transition] (e5) at (\nd{2}, \nd{3}) [label={[align=center]right:Receive feedback}] {}
+      edge [pre, bend left] node {1} (s2)
+      edge [post] node {1} (s1);
+
+  \end{scope}
+
+\end{tikzpicture}
+
+\caption{A petri net describing the congestion control mechanism.
+We start with one token in the state \textit{Open congestion window}, which means that initially we can send one packet.
+We start with another token in the state \textit{Received feedback}, which means that we can take an action right in the beginning.
+The concept is the following: If there is at least one token in the \textit{Open congestion window}, send a packet.
+The acknowledgement (or timeout if the packet gets lost) for the packet is received by the \textit{Open congestion window}, meaning that when we get an ACK we can send another packet (The \textit{Open congestion window} signifies the congestion window minus the packets that are currently unacknowledged.).
+Furthermore the state \textit{Received feedback} receives each ACK/timeout and takes an action.
+\textit{Take action} adds \textit{a} tokens to the \textit{Open congestion window}, where $a$ is a real number (possibly also negative); thus in each state there can also be a real number of tokens (e.g.~ 2.34 tokens are possible in the \textit{Open congestion window}).
+However, we define that the congestion window can never be smaller than 1.
+\todo{Draw as 3 times 3 square}}
+\label{fig:petri}
+\end{figure}
+
 The classical Reinforcement Learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
@@
 l_{\text{a},t} =& -\log \left( \pi \left( a_t \given s_t ; \theta_\text{a} \right) \right)\\
-&\left( \left(\frac{R_{\text{byte},t}}{{R_{\text{duration},t}}} - \frac{v_{\text{byte},t}}{{v_{\text{duration},t}}}\right) - \left( \frac{R_{\text{delay},t}}{{R_{\text{packet},t}}}- \frac{v_{\text{delay},t}}{{v_{\text{packet},t}}} \right)\right)\\
+&\left( \left(\log\left(\frac{R_{\text{byte},t}}{{R_{\text{duration},t}}}\right) - \log\left(\frac{v_{\text{byte},t}}{{v_{\text{duration},t}}}\right)\right) - \left( \frac{R_{\text{delay},t}}{{R_{\text{packet},t}}}- \frac{v_{\text{delay},t}}{{v_{\text{packet},t}}} \right)\right)\\
 &+ \beta H\left( \pi\left( s_t; \theta_a \right)\right).
 \end{align*}
+
+\subsection{Example}
+
+\begin{enumerate}
+\item Receive an ACK
+\item Update the internal state to reflect the information that this ACK provides (for example, the state consists of the round trip time experienced by this packet, the time since we receive the last ACK etc.)
+\item Sample an increase for the congestion window from the Actor Network (e.g. $0.24$) and add it to the current congestion window.
+\item 
+\begin{enumerate}
+\item If this was the last ACK, which completes a previous action (e.g.
+a previous action resulted in $3$ packets being sent; so if this is the ACK for the third (and last) packet of this previous action we got all the ACKs to compute the reward.
+)
+\item If we now have $t_{max}$ rewards (set to 20) then we also update the neural network. 
+\end{enumerate}
+
+\end{enumerate}
+
+\subsection{Convergence}
+
+\todo{Add formulas which ensures convergence.}
 

### 2017-10-08 tex/paper.tex
Spelling mistakes in TeX

@@
 In \autoref{alg:pal} we show the code that runs in one of the learning threads.
-As proposed in literature, we use several concurrent thread to improve training performance.
+As proposed in literature, we use several concurrent threads to improve training performance.
 A major difference to previous approaches to reinforcement learning is the fact that we allow more actions to be taken before the previous actions' rewards have been received.
@@
 
-Another major difference in PAL is that one action generates a number of partial actions ($\geq 1$) (see \autoref{fig:pal}).
+Another major difference in PAL is that one action generates a number of partial actions ($\geq 0$) (see \autoref{fig:pal}).
 Each partial action generates feedback upon interacting with the environment.
@@
 Each time an acknowledgement is received, the state is updated and the actor network is asked for the next action.
-\item[$\textit{a}_\text{t}$] Based on a given state and the history of previous states, the actor network returns an action $a_t$, which is an integer ($\geq 1$) that stands for the congestion window to be used until the next acknowledgement is received.
+\item[$\textit{a}_\text{t}$] Based on a given state and the history of previous states, the actor network returns an action $a_t$, which is a real number ($\geq 1$) that stands for the congestion window to be used until the next acknowledgement is received.
 \item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metric. These are discussed in more detail in \autoref{subsec:value}
@@
 \item If this was the last ACK, which completes a previous action (e.g.
-a previous action resulted in $3$ packets being sent; so if this is the ACK for the third (and last) packet of this previous action we got all the ACKs to compute the reward.
-)
+a previous action resulted in $3$ packets being sent; so if this is the ACK for the third (and last) packet of this previous action we got all the ACKs to compute the reward.)
 \item If we now have $t_{max}$ rewards (set to 20) then we also update the neural network. 

### 2017-10-09 tex/paper.tex
Started adding constraints to paper

@@
 
-\todo{Add formulas which ensures convergence.}
+We use
+
+\begin{align*}
+\text{Reward} = \log(\text{throughput}) - \delta\text{delay}
+\end{align*}
+
+as our metric of reward, where $\delta$ is a parameter that allows us to change the importance of the throughput vs.~the delay.
+We take the logarithm of the throughput to ensure that flows which have small throughput have a higher incentive to increase it as flows which already have a large one (because the derivative of the logarithm decreases with increasing value).
+We take minus the delay to punish inducing delay.
+We do not take the logarithm because we think that it is equally bad if a sender behind a satellite connection with a minimum RTT of $500\,$ms adds $1\,$ms of delay as if a sender connected using Ethernet with a minimum RTT of $10\,$ms adds $1\,$ms of delay.
+
+There are two conditions that we have to take into account to ensure convergence:
+\begin{enumerate}
+\item \begin{align*}
+\left(\frac{\partial}{\partial w}\text{throughput}\right)\left(1\right) > \left(\frac{\partial}{\partial w}\delta\text{delay}\right)\left(1\right)
+\end{align*}
+\item \begin{align*}
+\frac{\partial}{\partial w}\left(\delta\right)
+\end{align*}
+\end{enumerate}
+
+\todo{Finish second constraint for convergence.}
+
+\todo{Add figure to visually explain that}
+
+\section{Results}
+
+We train on the following scenario:
+
+\begin{table}[h]
+\centering
+\begin{tabular}{rll}
+\toprule
+Parameter & Value & Distribution \\
+\midrule
+Two-way propagation delay & $100\,$ms & constant \\
+Bottleneck bandwidth & $3\,$Mbit/s & constant \\
+Number of senders & $2$ & constant \\
+Flow length & $\infty$ & constant \\
+Simulation duration & $50\,$s--$150\,$s & uniform \\
+Buffer size & $\infty$ & constant \\
+Stochastic loss prob. & $0\%$ & constant \\
+\bottomrule
+\end{tabular}
+\caption{The parameters of our very first simple experiment}
+\label{tab:smallexperiment}
+\end{table}
+
+\todo{Insert results when comparing to regular TCP}
 

### 2017-10-09 async_deep_reinforce/constants.py

@@ -4,6 +4,9 @@ import logging
 import os
 import math
 
+def inverse_softplus(x):
+	return math.log(math.exp(x) - 1)
+
 LOCAL_T_MAX = 20 # repeat step size
 RMSP_ALPHA = 0.99 # decay parameter for RMSProp
 RMSP_EPSILON = 0.1 # epsilon parameter for RMSProp
@@ -29,22 +32,23 @@ PRECISION = tf.float32
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
 GAMMA = 0.99 # discount factor for rewards
 ENTROPY_BETA = 1e-4
-STD_BIAS_OFFSET = -1
-MAX_TIME_STEP = 1e7
+STD_BIAS_OFFSET = inverse_softplus(0.3)
+MAX_TIME_STEP = 1e6
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
 N_LSTM_LAYERS = 3
 
-def inverse_softplus(x):
-	return math.log(math.exp(x) - 1)
-
 SECONDS_NORMALIZER = 1e-2
 
-DELAY = 150*SECONDS_NORMALIZER
+DELAY = 100*SECONDS_NORMALIZER
 BIAS_OFFSET = 1
 PACKETS_BIAS_OFFSET = inverse_softplus(BIAS_OFFSET)
 DELAY_BIAS_OFFSET = inverse_softplus(DELAY)
 INTER_PACKET_ARRIVAL_TIME_OFFSET = inverse_softplus(1.0/DELAY)
+
+PACKETS_BIAS_OFFSET = 0
+DELAY_BIAS_OFFSET = 0
+INTER_PACKET_ARRIVAL_TIME_OFFSET = 0
 
 STATE_SIZE = 11
 HIDDEN_SIZE = 256

### 2017-10-09 async_deep_reinforce/game_ac_network.py

@@ -227,7 +227,7 @@ class GameACLSTMNetwork(GameACNetwork):
 			self.lstm = tf.contrib.rnn.MultiRNNCell([GameACLSTMNetwork.create_cell(HIDDEN_SIZE, LAYER_NORMALIZATION) for i in range(N_LSTM_LAYERS)], state_is_tuple=True)
 
 			# weight for policy output layer
-			self.W_hidden_to_action_mean_fc, self.b_hidden_to_action_mean_fc = self._fc_variable([HIDDEN_SIZE, 1], factor=1e-4, bias_range=(0, 0), bias_offset=1e-2)
+			self.W_hidden_to_action_mean_fc, self.b_hidden_to_action_mean_fc = self._fc_variable([HIDDEN_SIZE, 1], factor=1e-4, bias_offset=1e-4)
 			self.W_hidden_to_action_std_fc, self.b_hidden_to_action_std_fc = self._fc_variable([HIDDEN_SIZE, 1], factor=1.0, bias_offset=STD_BIAS_OFFSET)
 
 			# weight for value output layer

### 2017-10-10 async_deep_reinforce/constants.py
Improved simulation training

@@ -50,6 +50,9 @@ PACKETS_BIAS_OFFSET = 0
 DELAY_BIAS_OFFSET = 0
 INTER_PACKET_ARRIVAL_TIME_OFFSET = 0
 
+INITIAL_WINDOW_INCREASE_BIAS_OFFSET = 1e-3
+INITIAL_WINDOW_INCREASE_WEIGHT_FACTOR = 1e-3
+
 STATE_SIZE = 11
 HIDDEN_SIZE = 256
 # ACTION_SIZE = 1 # action size

### 2017-10-10 async_deep_reinforce/game_ac_network.py

@@ -15,6 +15,8 @@ from constants import VALUE_FACTOR
 from constants import PACKETS_BIAS_OFFSET
 from constants import DELAY_BIAS_OFFSET
 from constants import INTER_PACKET_ARRIVAL_TIME_OFFSET
+from constants import INITIAL_WINDOW_INCREASE_BIAS_OFFSET
+from constants import INITIAL_WINDOW_INCREASE_WEIGHT_FACTOR
 
 # tiny = np.finfo(np.float32).tiny
 tiny = 1e-20
@@ -227,7 +229,7 @@ class GameACLSTMNetwork(GameACNetwork):
 			self.lstm = tf.contrib.rnn.MultiRNNCell([GameACLSTMNetwork.create_cell(HIDDEN_SIZE, LAYER_NORMALIZATION) for i in range(N_LSTM_LAYERS)], state_is_tuple=True)
 
 			# weight for policy output layer
-			self.W_hidden_to_action_mean_fc, self.b_hidden_to_action_mean_fc = self._fc_variable([HIDDEN_SIZE, 1], factor=1e-4, bias_offset=1e-4)
+			self.W_hidden_to_action_mean_fc, self.b_hidden_to_action_mean_fc = self._fc_variable([HIDDEN_SIZE, 1], factor=INITIAL_WINDOW_INCREASE_WEIGHT_FACTOR, bias_offset=INITIAL_WINDOW_INCREASE_BIAS_OFFSET)
 			self.W_hidden_to_action_std_fc, self.b_hidden_to_action_std_fc = self._fc_variable([HIDDEN_SIZE, 1], factor=1.0, bias_offset=STD_BIAS_OFFSET)
 
 			# weight for value output layer

### 2017-10-10 tex/al_dente.tex
Add Al-Dente stuff

\documentclass[11pt]{beamer}
\usetheme{Pittsburgh}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%\author{}
%\title{}
%\setbeamercovered{transparent} 
%\setbeamertemplate{navigation symbols}{} 
%\logo{} 
%\institute{} 
%\date{} 
%\subject{} 
\begin{document}

%\begin{frame}
%\titlepage
%\end{frame}

%\begin{frame}
%\tableofcontents
%\end{frame}

\begin{frame}{•}

\end{frame}

\end{document}

### 2017-10-10 tex/paper.tex
Tuned constants

@@
   \begin{scope}
-    \node [place,tokens=1] (s1) at (\nd{1}, \nd{3}) [label=Received feedback] {};
+    \node [place,tokens=1] (s1) at (\nd{0}, \nd{1}) [label=Received feedback] {};
     \node [place] (s2) at (\nd{1}, \nd{2}) [label={[align=center]left:Received\\ACKs}] {};
-    \node [place] (s3) at (\nd{3}, \nd{2}) [label={[align=center]above:Packets\\in the\\network}] {};
+    \node [place] (s3) at (\nd{2}, \nd{1}) [label={[align=center]above:Packets\\in the\\network}] {};
     \node [place,tokens=1] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open\\congestion\\window\\(cwnd)}] {};

### 2017-10-10 async_deep_reinforce/a3c.py

@@ -115,11 +115,11 @@ checkpoint = tf.train.get_checkpoint_state(CHECKPOINT_DIR)
 saver = tf.train.Saver()
 if checkpoint and checkpoint.model_checkpoint_path:
   saver.restore(sess, checkpoint.model_checkpoint_path)
-  print("checkpoint loaded:", checkpoint.model_checkpoint_path)
+  logging.info("checkpoint loaded:", checkpoint.model_checkpoint_path)
   tokens = checkpoint.model_checkpoint_path.split("-")
   # set global step
   global_t = int(tokens[-1])
-  print(">>> global step set:", global_t)
+  logging.info(">>> global step set:", global_t)
   # set wall time
   wall_t_fname = CHECKPOINT_DIR + '/' + 'wall_t.' + str(global_t)
   with open(wall_t_fname, 'r') as f:
@@ -128,7 +128,7 @@ if checkpoint and checkpoint.model_checkpoint_path:
   with open(current_datetime_fname, 'r') as f:
     current_datetime = f.read()
 else:
-  print("Could not find old checkpoint")
+  logging.info("Could not find old checkpoint")
   # set wall time
   wall_t = 0.0
   current_datetime = datetime.datetime.now().isoformat()[:-7]

### 2017-10-10 async_deep_reinforce/constants.py

@@ -40,7 +40,7 @@ N_LSTM_LAYERS = 3
 
 SECONDS_NORMALIZER = 1e-2
 
-DELAY = 100*SECONDS_NORMALIZER
+DELAY = 150*SECONDS_NORMALIZER
 BIAS_OFFSET = 1
 PACKETS_BIAS_OFFSET = inverse_softplus(BIAS_OFFSET)
 DELAY_BIAS_OFFSET = inverse_softplus(DELAY)
@@ -51,7 +51,7 @@ DELAY_BIAS_OFFSET = 0
 INTER_PACKET_ARRIVAL_TIME_OFFSET = 0
 
 INITIAL_WINDOW_INCREASE_BIAS_OFFSET = 1e-3
-INITIAL_WINDOW_INCREASE_WEIGHT_FACTOR = 1e-3
+INITIAL_WINDOW_INCREASE_WEIGHT_FACTOR = 1e-4
 
 STATE_SIZE = 11
 HIDDEN_SIZE = 256

### 2017-10-10 async_deep_reinforce/constants.py
Add new config file

@@ -21,7 +21,7 @@ if environ.get('checkpoints') is not None:
 else:
 	CHECKPOINT_DIR = ABSOLUTE_PATH+'checkpoints'
 LOG_FILE = ABSOLUTE_PATH+'tmp/a3c_log'
-ACTOR_FACTOR = 1e-1
+ACTOR_FACTOR = 0.5e-1
 VALUE_FACTOR = 1e0
 GENERAL_FACTOR = 1e-3
 INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate

### 2017-10-10 tex/al_dente.tex
Parameter tuning

@@
 \documentclass[11pt]{beamer}
-\usetheme{Pittsburgh}
+\usetheme{Boadilla}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{amssymb}
-%\author{}
-%\title{}
+\usepackage{booktabs}
+\usepackage{svg}
+\def\xcolorversion{2.00}
+\def\xkeyvalversion{1.8}
+\usepackage{pgf}
+\usepackage{tikz}
+\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}
+
+\usepackage{mathtools}
+\newcommand\givenbase[1][]{\:#1\lvert\:}
+\let\given\givenbase
+\newcommand\sgiven{\givenbase[\delimsize]}
+\DeclarePairedDelimiterX\Basics[1](){\let\given\sgiven #1}
+\newcommand\Average{E\Basics}
+
+%\author{Maximilian Bachl}
+\title{Congestion Control}
 %\setbeamercovered{transparent} 
@@
 
-%\begin{frame}
-%\titlepage
-%\end{frame}
+\begin{frame}
+\titlepage
+\end{frame}
 
@@
 
-\begin{frame}{•}
+\begin{frame}{Congestion Control}
+\begin{itemize}
+\item Why Congestion Control? Before congestion control was invented: Everyone sent as much as they pleased $\rightarrow$ \textit{Congestion Collapse}.
+\item Goal: Estimate available bandwidth. Don't send too much, don't send too little.
+\item Method: Keep a \textit{Congestion Window}
+\item E.g. Congestion Window of 5 means that we can have up to 5 packets somewhere in the network.
+\end{itemize}
+\end{frame}
 
+\begin{frame}{How does Congestion Control work nowadays?}
+Simplified:
+\begin{itemize}
+\item Start with Congestion Window being 1 in the beginning of a flow.
+\item Upon receiving an acknowledgement for a previously sent packet, increase the Congestion Window (cwnd): $$\text{cwnd} = \frac{1}{\text{cwnd}}$$
+\item Recently also a bit more sophisticated $\rightarrow$ \textit{CUBIC} etc. 
+\item When there is packet loss then we sent too much (buffer of a router on the way overflew) $\rightarrow$ congestion $\rightarrow$ decrease Congestion Window
+\item Currently the most common thing to do is $$\text{cwnd} = \frac{\text{cwnd}}{2}$$
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Problems}
+\begin{itemize}
+\item Only decrease window on loss $\rightarrow$ then it's already too late! We should decrease before when buffers of routers fill up and latency increases!
+\item On wireless connections stochastic packet loss is common $\rightarrow$ TCP thinks it's congestion.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Potential Solution}
+\begin{itemize}
+\item Let's build some machine learning thing!
+\item Solutions already exist $\rightarrow$ \textit{TCP ex Machina} by Winstein and Balakrishnan (2013).
+\item They simulate networks and learn an optimum congestion control more or less by using a brute force algorithm.
+\item Example: Use networks with 1 to 5 senders, RTT from 10 to 100\,ms.
+Use one set of congestion control rules for 1000 simulations.
+Then change some parameters and check if it improved (actually they do it in a smarter way)
+\item Problem: Training has to be done offline. But it would be nice to have a Congestion Control that learns in real time, online!
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Reinforcement Learning}
+\centering
+\includesvg{figures/Reinforcement_learning_diagram}
+\end{frame}
+
+\begin{frame}{Problems with RL}
+\begin{itemize}
+\item Let's say an action is increasing the Congestion Window
+\item We can calculate the reward when we get back all the ACKs of packets that were sent during an action
+\item This means we can't do anything until receiving each action's reward. But that takes at least one Round Trip Time! Doesn't work.
+\item Example: We increase the cwnd by 2 packets. So we can send two packets. To evaluate if this action was good we have to get the ACKs of these two packets, which happens after one RTT!
+In the meantime another ACK could have arrived. What do we do? Not defined with RL...
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Partial Actions}
+\centering
+\includesvg{figures/Reinforcement_learning_diagram_PAL}
+\end{frame}
+
+\begin{frame}{Partial Actions: Example}
+\begin{itemize}
+\item We increase the window by $0.3$ (action). This allows us to send two packets (partial actions).  
+\item We receive the ACK for the first packet (feedback). We update the state and perform a new action.
+\item We receive the second ACK. Again, we update the state and perform an action. However, because we got all partial rewards of the previous action, we can calculate the reward and update our agent.
+\end{itemize}
+\textbf{Key point}: We can update the state without receiving the full reward yet.
+\end{frame}
+
+\begin{frame}{Asynchronous Actor Critic}
+\begin{itemize}
+\item A Deep Learning framework for reinforcement learning. Used for learning how to play video games.
+\item Maximize long term reward: \begin{align*}
+R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta_\text{v})\right),
+\end{align*}
+\item The \textbf{Critic} tries to estimate how much (long-term) reward he can expect considering the current state.
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Asynchronous Actor Critic -- Actor}
+\begin{itemize}
+\item The \textbf{Actor} tries to perform an action that is better than what the Critic would expect.
+\item It outputs two things:
+\begin{itemize}
+\item What it thinks is the best action (e.g. increase the window by $0.3$)
+\item A standard deviation to experiment a little bit (e.g. $0.45$)
+\end{itemize}
+\item So the actor outputs a distribution from which actions are sampled. Thanks to the standard deviation we experiment and don't get stuck with suboptimal actions. 
+\end{itemize}
+\end{frame}
+
+\begin{frame}{Preliminary results -- Experiment characteristics}
+\centering
+\begin{tabular}{rll}
+\toprule
+Parameter & Value & Distribution \\
+\midrule
+Two-way propagation delay & $150\,$ms & constant \\
+Bottleneck bandwidth & $15\,$Mbit/s & constant \\
+Number of senders & $8$ & constant \\
+Flow length & $100\,$kB & exponential \\
+Time between flows & $0.5\,$s & exponential \\
+Simulation duration & $100\,$s & constant \\
+Buffer size & $1000$ packets & constant \\
+Stochastic loss prob. & $0\%$ & constant \\
+\bottomrule
+\end{tabular}
+\end{frame}
+
+\begin{frame}{Preliminary results -- Comparison}
+\centering
+\begin{tabular}{rll}
+\toprule
+Parameter & Value & Distribution \\
+\midrule
+Two-way propagation delay & $150\,$ms & constant \\
+Bottleneck bandwidth & $15\,$Mbit/s & constant \\
+Number of senders & $8$ & constant \\
+Flow length & $100\,$kB & exponential \\
+Time between flows & $0.5\,$s & exponential \\
+Simulation duration & $100\,$s & constant \\
+Buffer size & $1000$ packets & constant \\
+Stochastic loss prob. & $0\%$ & constant \\
+\bottomrule
+\end{tabular}
+\end{frame}
+
+\newcommand{\nodedistance}{2}
+\newcommand{\nd}[1]{\numexpr #1 * \nodedistance \relax}
+\begin{frame}{Petri net}
+
+\centering
+\begin{tikzpicture}[node distance=\nodedistance,>=stealth',bend angle=45,auto]
+
+  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
+  \tikzstyle{red place}=[place,draw=red!75,fill=red!20]
+  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
+  			  fill=black!20,minimum size=4mm]
+
+%  \tikzstyle{every label}=[blue!75]
+
+  \begin{scope}
+    \node [place,tokens=1] (s1) at (\nd{0}, \nd{1}) [label={[align=center]left:Receive\\ feedback}] {};
+    \node [place] (s2) at (\nd{1}, \nd{2}) [label={[align=center]above:Received\\ ACKs}] {};
+    \node [place] (s3) at (\nd{2}, \nd{1}) [label={[align=center]right:Packets\\ in the\\ network}] {};
+    \node [place,tokens=1] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};
+
+    \node [transition] (e1) at (\nd{0}, \nd{0}) [label=left:Take action] {}
+      edge [post] node {$a$}(s4)
+      edge [pre] node {1} (s1);
+
+    \node [transition] (e2) at (\nd{2}, \nd{0}) [label={[align=center]right:Send\\ packet}] {}
+      edge [pre] node {1} (s4)
+      edge [post] node {1} (s3);
+
+    \node [transition] (e3) at (\nd{2}, \nd{2}) [label={[align=center]right:Receive\\ ACK}] {}
+      edge [pre] node {1} (s3)
+      edge [post] node {1} (s2);
+
+    \node [transition] (e4) at (\nd{1}, \nd{1}) [label={[align=center]left:Update\\open\\cwnd}] {}
+      edge [pre] node {1} (s2)
+      edge [post] node {1} (s4);
+
+    \node [transition] (e5) at (\nd{0}, \nd{2}) [label={[align=center]left:Received\\ feedback}] {}
+      edge [pre] node {1} (s2)
+      edge [post] node {1} (s1);
+
+  \end{scope}
+
+\end{tikzpicture}
 \end{frame}

### 2017-10-10 tex/paper.tex
Parameter tuning

@@
   \begin{scope}
-    \node [place,tokens=1] (s1) at (\nd{0}, \nd{1}) [label=Received feedback] {};
-    \node [place] (s2) at (\nd{1}, \nd{2}) [label={[align=center]left:Received\\ACKs}] {};
-    \node [place] (s3) at (\nd{2}, \nd{1}) [label={[align=center]above:Packets\\in the\\network}] {};
-    \node [place,tokens=1] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open\\congestion\\window\\(cwnd)}] {};
+    \node [place,tokens=1] (s1) at (\nd{0}, \nd{1}) [label={[align=center]left:Receive\\ feedback}] {};
+    \node [place] (s2) at (\nd{1}, \nd{2}) [label={[align=center]above:Received\\ ACKs}] {};
+    \node [place] (s3) at (\nd{2}, \nd{1}) [label={[align=center]right:Packets\\ in the\\ network}] {};
+    \node [place,tokens=1] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};
 
-    \node [transition] (e1) at (\nd{0}, \nd{1}) [label=left:Take action] {}
-      edge [post,bend right] node {$a$}(s4)
-      edge [pre,bend left] node {1} (s1);
+    \node [transition] (e1) at (\nd{0}, \nd{0}) [label=left:Take action] {}
+      edge [post] node {$a$}(s4)
+      edge [pre] node {1} (s1);
 
-    \node [transition] (e2) at (\nd{2}, \nd{1}) [label=Send packet] {}
-      edge [pre,bend left] node {1} (s4)
-      edge [post,bend right] node {1} (s3);
+    \node [transition] (e2) at (\nd{2}, \nd{0}) [label={[align=center]right:Send\\ packet}] {}
+      edge [pre] node {1} (s4)
+      edge [post] node {1} (s3);
 
-    \node [transition] (e3) at (\nd{2}, \nd{2}) [label=below:Receive ACK] {}
+    \node [transition] (e3) at (\nd{2}, \nd{2}) [label={[align=center]right:Receive\\ ACK}] {}
       edge [pre] node {1} (s3)
@@
 
-    \node [transition] (e5) at (\nd{2}, \nd{3}) [label={[align=center]right:Receive feedback}] {}
-      edge [pre, bend left] node {1} (s2)
+    \node [transition] (e5) at (\nd{0}, \nd{2}) [label={[align=center]left:Received\\ feedback}] {}
+      edge [pre] node {1} (s2)
       edge [post] node {1} (s1);

### 2017-10-10 async_deep_reinforce/constants.py

@@ -33,7 +33,7 @@ INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate
 GAMMA = 0.99 # discount factor for rewards
 ENTROPY_BETA = 1e-4
 STD_BIAS_OFFSET = inverse_softplus(0.3)
-MAX_TIME_STEP = 1e6
+MAX_TIME_STEP = 2e6
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
 N_LSTM_LAYERS = 3
@@ -50,7 +50,7 @@ PACKETS_BIAS_OFFSET = 0
 DELAY_BIAS_OFFSET = 0
 INTER_PACKET_ARRIVAL_TIME_OFFSET = 0
 
-INITIAL_WINDOW_INCREASE_BIAS_OFFSET = 1e-3
+INITIAL_WINDOW_INCREASE_BIAS_OFFSET = 1e-2
 INITIAL_WINDOW_INCREASE_WEIGHT_FACTOR = 1e-4
 
 STATE_SIZE = 11

### 2017-10-11 async_deep_reinforce/a3c.py
Add new config files that are less conservative and allow for more delay

@@ -115,11 +115,11 @@ checkpoint = tf.train.get_checkpoint_state(CHECKPOINT_DIR)
 saver = tf.train.Saver()
 if checkpoint and checkpoint.model_checkpoint_path:
   saver.restore(sess, checkpoint.model_checkpoint_path)
-  logging.info("checkpoint loaded:", checkpoint.model_checkpoint_path)
+  logging.info(" ".join(map(str,("checkpoint loaded:", checkpoint.model_checkpoint_path))))
   tokens = checkpoint.model_checkpoint_path.split("-")
   # set global step
   global_t = int(tokens[-1])
-  logging.info(">>> global step set:", global_t)
+  logging.info(" ".join(map(str,(">>> global step set:", global_t))))
   # set wall time
   wall_t_fname = CHECKPOINT_DIR + '/' + 'wall_t.' + str(global_t)
   with open(wall_t_fname, 'r') as f:
@@ -239,6 +239,7 @@ def delete_training_thread(thread_id):
   # del training_threads[thread_id]
 
 def call_process_action(thread_id, state, tickno, window):
+  print("state", state)
   # logging.debug(" ".join(map(str,("call_process_action", thread_id, state, tickno))))
   global sess, global_t, summary_writer, summary_op, summary_inputs
   chosen_action = training_threads[thread_id].action_step(sess, state, tickno, window)

### 2017-10-11 async_deep_reinforce/a3c_training_thread.py

@@ -109,7 +109,10 @@ class A3CTrainingThread(object):
         self.start_lstm_states.append(self.local_network.lstm_state_out)
         self.variable_snapshots.append(sess.run(self.local_network.get_vars()))
 
-    pi_, action, value_ = self.local_network.run_policy_action_and_value(sess, state)
+    if self.training:
+      action, value_ = self.local_network.run_action_and_value(sess, state)
+    else:
+      action = self.local_network.run_action(sess, state)
 
     # logging.debug(" ".join(map(str,(self.thread_index,"pi_values:",pi_))))
 
@@ -331,7 +334,8 @@ class A3CTrainingThread(object):
 
     # if len(ticknos) == 0:
     #   print(self.thread_index, "actions", self.actions, "rewards", self.rewards, "values", self.values, "estimated_values", self.estimated_values, "ticknos", self.ticknos)
-    if final or self.local_t % LOG_INTERVAL == 0:
+    # if final or self.local_t % LOG_INTERVAL == 0:
+    if final:
       if ticknos[-1]-ticknos[0] > 0 and self.episode_reward_throughput > 0:
         # print(ticknos)
         # print(self.episode_reward_throughput, ticknos[0], ticknos[-1])

### 2017-10-11 async_deep_reinforce/constants.py

@@ -21,9 +21,9 @@ if environ.get('checkpoints') is not None:
 else:
 	CHECKPOINT_DIR = ABSOLUTE_PATH+'checkpoints'
 LOG_FILE = ABSOLUTE_PATH+'tmp/a3c_log'
-ACTOR_FACTOR = 0.5e-1
+ACTOR_FACTOR = 1e-3
 VALUE_FACTOR = 1e0
-GENERAL_FACTOR = 1e-3
+GENERAL_FACTOR = 1e-2
 INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate
 INITIAL_ALPHA_HIGH = 1e0*GENERAL_FACTOR   # log_uniform high limit for learning rate
 

### 2017-10-11 async_deep_reinforce/game_ac_network.py

@@ -309,17 +309,30 @@ class GameACLSTMNetwork(GameACNetwork):
 	# 	# pi_out: (1,3), v_out: (1)
 	# 	return ((pi_out[0][0], pi_out[1][0]), (v_packets_out[0], v_accumulated_delay_out[0], v_duration_out[0]))
 
-	def run_policy_action_and_value(self, sess, s_t):
+	def run_action_and_value(self, sess, s_t):
 		# This run_policy_and_value() is used when forward propagating.
 		# so the step size is 1.
-		pi_out, action_out, v_packets_out, v_accumulated_delay_out, v_duration_out, self.lstm_state_out = sess.run(
-			[self.pi, self.chosen_action, self.v_packets, self.v_accumulated_delay, self.v_duration, self.lstm_state],
+		action_out, v_packets_out, v_accumulated_delay_out, v_duration_out, self.lstm_state_out = sess.run(
+			[self.chosen_action, self.v_packets, self.v_accumulated_delay, self.v_duration, self.lstm_state],
 			feed_dict = {self.s : [s_t],
 			self.initial_lstm_state : self.lstm_state_out,
 			self.step_size : [1]}
 		)
 		# pi_out: (1,3), v_out: (1)
-		return ((pi_out[0][0], pi_out[1][0]), action_out[0],(v_packets_out[0], v_accumulated_delay_out[0], v_duration_out[0]))
+		return (action_out[0],(v_packets_out[0], v_accumulated_delay_out[0], v_duration_out[0]))
+
+	# Misleading name: Actually returns the mean of the distribution returned by the actor.
+	def run_action(self, sess, s_t):
+		# This run_policy_and_value() is used when forward propagating.
+		# so the step size is 1.
+		pi_out, self.lstm_state_out = sess.run(
+			[self.pi, self.lstm_state],
+			feed_dict = {self.s : [s_t],
+			self.initial_lstm_state : self.lstm_state_out,
+			self.step_size : [1]}
+		)
+		# pi_out: (1,3), v_out: (1)
+		return pi_out[0][0]
 
 	def run_value(self, sess, s_t):
 		# This run_value() is used for calculating V for bootstrapping at the

### 2017-10-11 src/unicorn.cc

@@ -110,7 +110,7 @@ void Unicorn::packets_received( const vector< remy::Packet > & packets ) {
 
 void Unicorn::reset(const double & tickno)
 {
-  // printf("%lu: Fucking resetting\n", _thread_id);
+  printf("%lu: Fucking resetting\n", _thread_id);
   _rainbow._training = _training;
   // assert(false);
   // printf("%lu: Resetting\n", _thread_id);

### 2017-10-11 async_deep_reinforce/constants.py
Add config files that don't really care about the delay whatsoeverand (hopefully) find the right parameters finally

@@ -21,7 +21,7 @@ if environ.get('checkpoints') is not None:
 else:
 	CHECKPOINT_DIR = ABSOLUTE_PATH+'checkpoints'
 LOG_FILE = ABSOLUTE_PATH+'tmp/a3c_log'
-ACTOR_FACTOR = 1e-3
+ACTOR_FACTOR = 1e-2
 VALUE_FACTOR = 1e0
 GENERAL_FACTOR = 1e-2
 INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate

### 2017-10-11 async_deep_reinforce/constants.py
Make actor learning rate a bit slower

@@ -21,7 +21,7 @@ if environ.get('checkpoints') is not None:
 else:
 	CHECKPOINT_DIR = ABSOLUTE_PATH+'checkpoints'
 LOG_FILE = ABSOLUTE_PATH+'tmp/a3c_log'
-ACTOR_FACTOR = 1e-2
+ACTOR_FACTOR = 0.5e-2
 VALUE_FACTOR = 1e0
 GENERAL_FACTOR = 1e-2
 INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate

### 2017-10-11 async_deep_reinforce/a3c.py
Remove debug messages

@@ -239,7 +239,7 @@ def delete_training_thread(thread_id):
   # del training_threads[thread_id]
 
 def call_process_action(thread_id, state, tickno, window):
-  print("state", state)
+  # print("state", state)
   # logging.debug(" ".join(map(str,("call_process_action", thread_id, state, tickno))))
   global sess, global_t, summary_writer, summary_op, summary_inputs
   chosen_action = training_threads[thread_id].action_step(sess, state, tickno, window)

### 2017-10-11 src/unicorn.cc
Remove debug messages

@@ -110,7 +110,7 @@ void Unicorn::packets_received( const vector< remy::Packet > & packets ) {
 
 void Unicorn::reset(const double & tickno)
 {
-  printf("%lu: Fucking resetting\n", _thread_id);
+  // printf("%lu: Fucking resetting\n", _thread_id);
   _rainbow._training = _training;
   // assert(false);
   // printf("%lu: Resetting\n", _thread_id);

### 2017-11-09 async_deep_reinforce/constants.py
Good config

@@ -42,7 +42,7 @@ STD_BIAS_OFFSET = 0
 MAX_TIME_STEP = 5e6
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
-N_LSTM_LAYERS = 3
+N_LSTM_LAYERS = 1
 
 # Don't know if this normalization stuff makes sense. Better disable it when possible...
 SECONDS_NORMALIZER = 1e-2

### 2017-11-09 tex/paper.tex
Updated tex file

@@
 \usepackage{algorithm}
+\usepackage[inline]{enumitem}
 
@@
 \textit{Take action} adds \textit{a} tokens to the \textit{Open congestion window}, where $a$ is a real number (possibly also negative); thus in each state there can also be a real number of tokens (e.g.~ 2.34 tokens are possible in the \textit{Open congestion window}).
-However, we define that the congestion window can never be smaller than 1.
-\todo{Draw as 3 times 3 square}}
+However, we define that the congestion window can never be smaller than 1.}
 \label{fig:petri}
@@
 
-\subsection{Convergence}
-
-We use
-
-\begin{align*}
-\text{Reward} = \log(\text{throughput}) - \delta\text{delay}
-\end{align*}
-
-as our metric of reward, where $\delta$ is a parameter that allows us to change the importance of the throughput vs.~the delay.
+%\subsection{Convergence}
+%
+%We use 
+%
+%\begin{align*}
+%\text{Reward} = \log(\text{throughput}) - \delta\text{delay}
+%\end{align*}
+%
+%as our metric of reward, where $\delta$ is a parameter that allows us to change the importance of the throughput vs.~the delay.
 We take the logarithm of the throughput to ensure that flows which have small throughput have a higher incentive to increase it as flows which already have a large one (because the derivative of the logarithm decreases with increasing value).
@@
 We do not take the logarithm because we think that it is equally bad if a sender behind a satellite connection with a minimum RTT of $500\,$ms adds $1\,$ms of delay as if a sender connected using Ethernet with a minimum RTT of $10\,$ms adds $1\,$ms of delay.
+%
+%There are two conditions that we have to take into account to ensure convergence:
+%\begin{enumerate}
+%\item \begin{align*}
+%\left(\frac{\partial}{\partial w}\text{throughput}\right)\left(1\right) > \left(\frac{\partial}{\partial w}\delta\text{delay}\right)\left(1\right)
+%\end{align*}
+%\item \begin{align*}
+%\frac{\partial}{\partial w}\left(\delta\right)
+%\end{align*}
+%\end{enumerate}
+%
+%\todo{Finish second constraint for convergence.}
+%
+%\todo{Add figure to visually explain that}
 
-There are two conditions that we have to take into account to ensure convergence:
-\begin{enumerate}
-\item \begin{align*}
-\left(\frac{\partial}{\partial w}\text{throughput}\right)\left(1\right) > \left(\frac{\partial}{\partial w}\delta\text{delay}\right)\left(1\right)
+\subsection{Convergence}
+
+Assume we have $N$ senders, each with its specific baseline delay (this encompasses all delays excluding only the queueing delay) $d_i$.
+Each sender keeps a congestion window $b_i$.
+The link speed at the bottleneck is designated as $c$.
+
+%We define the excessive window as 
+%\begin{align*}
+%b_\text{excess} = \max\left( \left( \sum_{j=1}^N \frac{b_j}{d_j} \right) - c, 0 \right)
+%\end{align*}
+%and the queueing delay as 
+%\begin{align*}
+%d_q = \frac{b_\text{excess}}{c}.
+%\end{align*}
+
+With $b = (b_i)$ we observe the following relation:
+
+\begin{align*}
+c = \sum_{j=1}^N \frac{b_j}{d_j + d_q(b)}
 \end{align*}
-\item \begin{align*}
-\frac{\partial}{\partial w}\left(\delta\right)
+
+Then we define the utility function for each sender $i$ as
+\begin{align*}
+V_i(b) \coloneqq \begin{cases}
+\log\left(\frac{b_i}{d_i}\right) - \delta_i d_i & \text{if } \sum_{j=1}^N \frac{b_i}{d_i} \leq c\\
+\log\left(\frac{b_i}{d_i + d_q(b)}\right) - \delta_i (d_i + d_q(b)) & \text{otherwise}
+\end{cases}
 \end{align*}
-\end{enumerate}
 
-\todo{Finish second constraint for convergence.}
+\begin{align*}
+&\log\left(\frac{b_i}{d_i + d_q}\right) - \delta_i (d_i + d_q) \\
+=& \log\left(b_i\right) - \log\left(d_i + d_q\right) - \delta_i (d_i + d_q)
+\end{align*}
 
-\todo{Add figure to visually explain that}
+\begin{align*}
+\frac{\partial V_i}{\partial b_i} = \begin{cases}
+\frac{1}{b_i} & \text{if } \sum_{j=1}^N \frac{b_i}{d_i} \leq c\\
+\frac{1}{b_i} - \frac{\partial d_q}{\partial b_i}\left(b_i\right)\frac{1}{d_i+d_q\left(b_i\right)} - \delta\frac{\partial d_q}{b_i}\left(b_i\right) & \text{otherwise}
+\end{cases}
+\end{align*}
+
+%\begin{align*}
+%G(b) = \begin{pmatrix}
+%-\frac{1}{b_1^2} & & \\
+%& \ddots & \\
+%& & -\frac{1}{b_N^2}
+%\end{pmatrix} \text{if } \text{if } \sum_{j=1}^N \frac{b_i}{d_i} \leq c
+%\end{align*}
+
+To show diagonal strict concavity we have to show that d.s.c.~is fulfilled \begin{enumerate*} \item for two distinct points lying in $\sum_{j=1}^N \frac{b_i}{d_i} \leq c$ \item for two distinct points lying in $\sum_{j=1}^N \frac{b_i}{d_i} > c$ and \item when one point is in one part of the function and the other in the other one\end{enumerate*}.
+
+\todo{Insert what d.s.c~actually is and quote Altman's paper}
+
+%\begin{enumerate}
+%\item (b^1 - 
+%\item
+%\item
+%\end{enumerate}
 

### 2017-11-09 async_deep_reinforce/constants.py
Fix bug in configure file and now allowing layers to be set over the command line

@@ -42,7 +42,7 @@ STD_BIAS_OFFSET = 0
 MAX_TIME_STEP = 5e6
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
-N_LSTM_LAYERS = 1
+N_LSTM_LAYERS = int(environ.get('layers')) if environ.get('layers') is not None else 1
 
 # Don't know if this normalization stuff makes sense. Better disable it when possible...
 SECONDS_NORMALIZER = 1e-2

### 2017-11-21 tex/al_dente.tex
Add new config files that are less conservative and allow for more delay
Updated AL-DENTE-presentation

@@
 \begin{itemize}
-\item Start with Congestion Window being 1 in the beginning of a flow.
+\item Congestion Window 1 in the beginning of a flow.
 \item Upon receiving an acknowledgement for a previously sent packet, increase the Congestion Window (cwnd): $$\text{cwnd} = \frac{1}{\text{cwnd}}$$
@@
 \item When there is packet loss then we sent too much (buffer of a router on the way overflew) $\rightarrow$ congestion $\rightarrow$ decrease Congestion Window
-\item Currently the most common thing to do is $$\text{cwnd} = \frac{\text{cwnd}}{2}$$
+\item The most common thing to do is $$\text{cwnd} = \frac{\text{cwnd}}{2}$$
 \end{itemize}
@@
 \begin{itemize}
-\item Only decrease window on loss $\rightarrow$ then it's already too late! We should decrease before when buffers of routers fill up and latency increases!
+\item Only decrease window on loss $\rightarrow$ That's too late! Decrease before when buffers of routers fill up and latency increases!
 \item On wireless connections stochastic packet loss is common $\rightarrow$ TCP thinks it's congestion.
@@
 \begin{itemize}
-\item We increase the window by $0.3$ (action). This allows us to send two packets (partial actions).  
+\item Let's say the window is 1.9. We increase the window by $0.3$ (action). This allows us to send two packets (partial actions).  
 \item We receive the ACK for the first packet (feedback). We update the state and perform a new action.
@@
 \end{align*}
-\item The \textbf{Critic} tries to estimate how much (long-term) reward he can expect considering the current state.
+\item The \textbf{Critic} tries to estimate how much (long-term) reward one can expect considering the current state.
 \end{itemize}
@@
 \end{itemize}
-\item So the actor outputs a distribution from which actions are sampled. Thanks to the standard deviation we experiment and don't get stuck with suboptimal actions. 
+\item So the actor outputs a normal distribution from which actions are sampled. Thanks to the standard deviation we experiment and don't get stuck with suboptimal actions. 
 \end{itemize}
@@
 \midrule
-Two-way propagation delay & $150\,$ms & constant \\
+Two-way RTT when queue is 0 & $150\,$ms & constant \\
 Bottleneck bandwidth & $15\,$Mbit/s & constant \\
@@
 Time between flows & $0.5\,$s & exponential \\
-Simulation duration & $100\,$s & constant \\
-Buffer size & $1000$ packets & constant \\
+Simulation duration & $500\,$s & constant \\
+Buffer size & $100$ packets & constant \\
 Stochastic loss prob. & $0\%$ & constant \\
@@
 \begin{frame}{Preliminary results -- Comparison}
-\centering
-\begin{tabular}{rll}
-\toprule
-Parameter & Value & Distribution \\
-\midrule
-Two-way propagation delay & $150\,$ms & constant \\
-Bottleneck bandwidth & $15\,$Mbit/s & constant \\
-Number of senders & $8$ & constant \\
-Flow length & $100\,$kB & exponential \\
-Time between flows & $0.5\,$s & exponential \\
-Simulation duration & $100\,$s & constant \\
-Buffer size & $1000$ packets & constant \\
-Stochastic loss prob. & $0\%$ & constant \\
-\bottomrule
-\end{tabular}
+Stuff goes here!
 \end{frame}
@@
     \node [place] (s3) at (\nd{2}, \nd{1}) [label={[align=center]right:Packets\\ in the\\ network}] {};
-    \node [place,tokens=1] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};
+    \node [place] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};
 
@@
 
+\newcommand{\nnfontsize}{\tiny}
+
+\begin{frame}{Neural Network}
+
+\centering
+\begin{tikzpicture}[node distance=1.2cm, >=stealth',bend angle=45,auto]
+
+  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
+  \tikzstyle{input place}=[circle,thick,draw=olive!75,fill=olive!20,minimum size=6mm]
+  \tikzstyle{output place}=[circle,thick,draw=magenta!75,fill=magenta!20,minimum size=6mm]
+  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
+  			  fill=black!20,minimum size=4mm]
+
+%  \tikzstyle{every label}=[blue!75]
+
+  \begin{scope}
+    \node [input place] (s1) [label={[align=center]left:\nnfontsize State (11)}] {};
+    \node [place] (s2) [above right of=s1, label={[align=center]above:\nnfontsize Lin.~(11,32)}] {};
+
+  \path[->] (s1) edge node {} (s2);
+
+    \node [place] (s3) [right of=s2, label={[align=center]below:\nnfontsize LSTM (32)}] {};
+
+  \path[->] (s2) edge node {} (s3); 
+  \path[->] (s3) edge [loop above] node {} (s3);
+
+    \node [place] (s4) [right of=s3, label={[align=center]below:\nnfontsize LSTM (32)}] {};
+
+  \path[->] (s3) edge node {} (s4);
+  \path[->] (s4) edge [loop above] node {} (s4);
+
+    \node [place] (s7) [right of=s4, label={[align=center]below:\nnfontsize Lin.~(32,1)}] {};
+
+  \path[->] (s4) edge node {} (s7);
+
+    \node [output place] (s8) [right of=s7, label={[align=center]below:\nnfontsize $\sigma$~(1)}] {};
+
+  \path[->] (s7) edge node {} (s8);
+
+    \node [place] (s5) [above of=s7, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+
+  \path[->] (s4) edge node {} (s5);
+
+    \node [output place] (s6) [right of=s5, label={[align=center]above:\nnfontsize $\mu$ (1)}] {};
+
+  \path[->] (s5) edge node {} (s6);
+
+    \node [place] (s10) [below right of=s1, label={[align=center]below:\nnfontsize Lin.~(11,32)}] {};
+
+  \path[->] (s1) edge node {} (s10);
+
+    \node [place] (s11) [right of=s10, label={[align=center]above:\nnfontsize LSTM (32)}] {};
+
+  \path[->] (s10) edge node {} (s11);
+  \path[->] (s11) edge [loop below] node {} (s11);
+
+    \node [place] (s12) [right of=s11, label={[align=center]above:\nnfontsize LSTM (32)}] {};
+
+  \path[->] (s11) edge node {} (s12);
+  \path[->] (s12) edge [loop below] node {} (s12);
+
+    \node [place] (s13) [right of=s12, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+
+  \path[->] (s12) edge node {} (s13);
+
+    \node [output place] (s17) [right of=s13, label={[align=center]above:\nnfontsize $R_\text{packets}$ (1)}] {};
+
+  \path[->] (s13) edge node {} (s17);
+
+    \node [place] (s15) [below of=s13, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+
+  \path[->] (s12) edge node {} (s15);
+
+    \node [output place] (s19) [right of=s15, label={[align=center]above:\nnfontsize $R_\text{delay}$ (1)}] {};
+
+  \path[->] (s15) edge node {} (s19);
+
+    \node [place] (s16) [below of=s15, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+
+  \path[->] (s12) edge node {} (s16);
+
+    \node [output place] (s20) [right of=s16, label={[align=center]above:\nnfontsize $R_\text{lost}$ (1)}] {};
+
+  \path[->] (s16) edge node {} (s20);
+
+    \node [place] (s14) [below of=s16, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+
+  \path[->] (s12) edge node {} (s14);
+
+    \node [output place] (s21) [right of=s14, label={[align=center]above:\nnfontsize $R_\text{duration}$ (1)}] {};
+
+  \path[->] (s14) edge node {} (s21);
+
+  \end{scope}
+
+\end{tikzpicture}
+\end{frame}
+
 \end{document}

### 2017-11-22 async_deep_reinforce/a3c_training_thread.py
Change some conig things and add comments

@@ -17,6 +17,7 @@ from constants import LOCAL_T_MAX
 from constants import LOG_LEVEL
 from constants import SECONDS_NORMALIZER
 from constants import STATE_SIZE
+from constants import SIGMOID_ALPHA
 
 import logging
 
@@ -24,8 +25,8 @@ logging.basicConfig(level=LOG_LEVEL)
 
 LOG_INTERVAL = 200
 
-def inverse_sigmoid(alpha, x):
-  return 1/(1+math.exp(alpha*x))
+def inverse_sigmoid(x):
+  return 1/(1+math.exp(x))
 
 class A3CTrainingThread(object):
   def __init__(self,
@@ -177,7 +178,7 @@ class A3CTrainingThread(object):
         self.windows += nones_to_add
         self.ticknos += nones_to_add
       # TODO: Is this useful? I guess only the `local_t' is actually needed...
-      elif len(self.actions) <= 0:
+      else:
         self.local_t = 0
         self.episode_count += 1
         self.episode_reward_throughput = 0
@@ -288,7 +289,12 @@ class A3CTrainingThread(object):
       # td = R_packets/R_duration - Vi[0]/Vi[2] - self.delay_delta*(R_accumulated_delay/R_packets - Vi[1]/Vi[0])
       # td = R_packets/R_duration - Vi[0]/(Vi[2]) - self.delay_delta*(R_accumulated_delay/R_packets - Vi[1]/Vi[0]) - (R_lost/R_duration - Vi[3]/(Vi[2]))
       # td = inverse_sigmoid(self.delay_delta, R_lost/(R_packets+R_lost))*R_packets/R_duration - inverse_sigmoid(self.delay_delta, Vi[3]/(Vi[0]+Vi[3]))*Vi[0]/Vi[2] - (R_lost/R_duration - Vi[3]/(Vi[2]))
-      td = inverse_sigmoid(self.delay_delta, R_lost/(R_packets+R_lost) - 0.05)*R_packets/R_duration - inverse_sigmoid(self.delay_delta, Vi[3]/(Vi[0]+Vi[3]) - 0.05)*Vi[0]/Vi[2] - (R_lost/R_duration - Vi[3]/(Vi[2])) - (R_accumulated_delay/R_packets - Vi[1]/Vi[0])
+
[…]

### 2017-11-22 async_deep_reinforce/constants.py

@@ -27,9 +27,9 @@ else:
 logging.info(" ".join(map(str,("CHECKPOINT_DIR:",CHECKPOINT_DIR))))
 
 LOG_FILE = ABSOLUTE_PATH+'tmp/a3c_log'
-ACTOR_FACTOR = 0.5e0
+ACTOR_FACTOR = 1e0
 VALUE_FACTOR = 1e0
-GENERAL_FACTOR = 1e-4
+GENERAL_FACTOR = 1e-2
 # INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate
 # INITIAL_ALPHA_HIGH = 1e0*GENERAL_FACTOR   # log_uniform high limit for learning rate
 INITIAL_RATE = GENERAL_FACTOR
@@ -41,7 +41,7 @@ GAMMA = 0.99 # discount factor for rewards
 ENTROPY_BETA = 1e-4
 # STD_BIAS_OFFSET = inverse_softplus(0.3)
 STD_BIAS_OFFSET = 0
-MAX_TIME_STEP = 5e6
+MAX_TIME_STEP = 1e6
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
 N_LSTM_LAYERS = int(environ.get('layers')) if environ.get('layers') is not None else 2
@@ -69,7 +69,9 @@ LOST_OFFSET = inverse_softplus(1e-3)
 INITIAL_WINDOW_INCREASE_BIAS_OFFSET = 0
 INITIAL_WINDOW_INCREASE_WEIGHT_FACTOR = 1
 
-STATE_SIZE = int(environ.get('state_size')) if environ.get('state_size') is not None else 11
+STATE_SIZE = int(environ.get('state_size')) if environ.get('state_size') is not None else 12
 HIDDEN_SIZE = int(environ.get('hidden_size')) if environ.get('hidden_size') is not None else 32
 # ACTION_SIZE = 1 # action size
 LAYER_NORMALIZATION = False
+
+SIGMOID_ALPHA = 100.0

### 2017-11-22 src/rainbow.cc

@@ -9,6 +9,7 @@
 
 using namespace std;
 
+// TODO: I guess one needs another global lock so that it cannot happen that two threads concurrently call this method?
 Rainbow& Rainbow::getInstance() {
 	// FIXME: I guess one should check for a NULL pointer
 	const char* cooperative_string = getenv("cooperative");

### 2017-11-22 src/unicorn-templates.cc

@@ -29,6 +29,7 @@ void Unicorn::send( const unsigned int id, NextHop & next, const double & tickno
     if (_last_send_time == 0) {
       _memory._last_tick_sent = tickno;
       _memory._last_tick_received = tickno;
+      _memory._lost_since_last_time = 0;
       _flow_to_last_received[_flow_id] = tickno;
     }
 

### 2017-11-22 src/unicorn.cc

@@ -156,6 +156,12 @@ void Unicorn::reset(const double & tickno)
     // _put_rewards += 1;
     finishFlow();
   }
+
+  // FIXME: Then why exactly do we also do this in unicorn-template.cc?
+  if (_active_flows.find(_flow_id) != _active_flows.end()) {
+    _active_flows.erase(_flow_id);
+  }
+
   // if (_put_actions != _put_rewards) {
   //   printf("%lu: _put_actions: %lu, _put_rewards: %lu\n", _thread_id, _put_actions, _put_rewards);
   // }
@@ -229,6 +235,7 @@ void Unicorn::get_action(const double& tickno, const int& packets_sent_in_this_e
       (double) packets_sent_in_this_episode,
       _memory._send*NORMALIZER,
       _memory._rec*NORMALIZER,
+      (double) _memory._lost_since_last_time,
       // _memory.field(2),
       // _memory.field(4),
       // (double) _the_window,

### 2017-11-22 tex/al_dente.tex

@@
 Stuff goes here!
+
 \end{frame}
@@
   \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
-  \tikzstyle{red place}=[place,draw=red!75,fill=red!20]
   \tikzstyle{transition}=[rectangle,thick,draw=black!75,
@@
   \tikzstyle{output place}=[circle,thick,draw=magenta!75,fill=magenta!20,minimum size=6mm]
+   \tikzstyle{lstm place}=[circle,thick,draw=red!75,fill=red!20,minimum size=6mm]
+   \tikzstyle{softplus place}=[circle,thick,draw=green!75,fill=green!20,minimum size=6mm]
+
   \tikzstyle{transition}=[rectangle,thick,draw=black!75,
@@
   \begin{scope}
-    \node [input place] (s1) [label={[align=center]left:\nnfontsize State (11)}] {};
-    \node [place] (s2) [above right of=s1, label={[align=center]above:\nnfontsize Lin.~(11,32)}] {};
+    \node [input place] (s1) [label={[align=center]left:\nnfontsize State (12)}] {};
+    \node [place] (s2) [above right of=s1, label={[align=center]above:\nnfontsize Linear}] {};
 
@@
 
-    \node [place] (s3) [right of=s2, label={[align=center]below:\nnfontsize LSTM (32)}] {};
+    \node [lstm place] (s3) [right of=s2, label={[align=center]below:\nnfontsize LSTM (32)}] {};
 
@@
 
-    \node [place] (s4) [right of=s3, label={[align=center]below:\nnfontsize LSTM (32)}] {};
+    \node [lstm place] (s4) [right of=s3, label={[align=center]below:\nnfontsize LSTM (32)}] {};
 
@@
 
-    \node [place] (s7) [right of=s4, label={[align=center]below:\nnfontsize Lin.~(32,1)}] {};
+    \node [softplus place] (s7) [right of=s4, label={[align=center]below:{\nnfontsize Softplus}}] {};
 
@@
 
-    \node [place] (s5) [above of=s7, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+    \node [place] (s5) [above of=s7, label={[align=center]above:\nnfontsize Linear}] {};
 
@@
 
-    \node [place] (s10) [below right of=s1, label={[align=center]below:\nnfontsize Lin.~(11,32)}] {};
+    \node [place] (s10) [below right of=s1, label={[align=center]below:\nnfontsize Linear}] {};
 
@@
 
-    \node [place] (s11) [right of=s10, label={[align=center]above:\nnfontsize LSTM (32)}] {};
+    \node [lstm place] (s11) [right of=s10, label={[align=center]above:\nnfontsize LSTM (32)}] {};
 
@@
 
-    \node [place] (s12) [right of=s11, label={[align=center]above:\nnfontsize LSTM (32)}] {};
+    \node [lstm place] (s12) [right of=s11, label={[align=center]above:\nnfontsize LSTM (32)}] {};
 
@@
 
-    \node [place] (s13) [right of=s12, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+    \node [softplus place] (s13) [right of=s12, label={[align=center]above:\nnfontsize Softplus}] {};
 
@@
 
-    \node [output place] (s17) [right of=s13, label={[align=center]above:\nnfontsize $R_\text{packets}$ (1)}] {};
+    \node [output place] (s17) [right of=s13, label={[align=center]above:\nnfontsize $V_\text{packets}$ (1)}] {};
 
@@
 
-    \node [place] (s15) [below of=s13, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+    \node [softplus place] (s15) [below of=s13, label={[align=center]above:\nnfontsize Softplus}] {};
 
@@
 
-    \node [output place] (s19) [right of=s15, label={[align=center]above:\nnfontsize $R_\text{delay}$ (1)}] {};
+    \node [output place] (s19) [right of=s15, label={[align=center]above:\nnfontsize $V_\text{delay}$ (1)}] {};
 
@@
 
-    \node [place] (s16) [below of=s15, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+    \node [softplus place] (s16) [below of=s15, label={[align=center]above:\nnfontsize Softplus}] {};
 
@@
 
-    \node [output place] (s20) [right of=s16, label={[align=center]above:\nnfontsize $R_\text{lost}$ (1)}] {};
+    \node [output place] (s20) [right of=s16, label={[align=center]above:\nnfontsize $V_\text{lost}$ (1)}] {};
 
@@
 
-    \node [place] (s14) [below of=s16, label={[align=center]above:\nnfontsize Lin.~(32,1)}] {};
+    \node [softplus place] (s14) [below of=s16, label={[align=center]above:\nnfontsize Softplus}] {};
 
@@
 
-    \node [output place] (s21) [right of=s14, label={[align=center]above:\nnfontsize $R_\text{duration}$ (1)}] {};
+    \node [output place] (s21) [right of=s14, label={[align=center]above:\nnfontsize $V_\text{duration}$ (1)}] {};
 
@@
 
+\begin{frame}{Reward 1}
+\begin{align*}
+\frac{R_\text{packets}}{R_\text{duration}}\text{Sigmoid}_\alpha\left(\frac{R_\text{lost}}{R_\text{lost} + R_\text{packets}} - 0.05\right) - \frac{R_\text{lost}}{R_\text{duration}} - \frac{R_\text{delay}}{R_\text{packets}}
+\end{align*}
+\end{frame}
+
+\begin{frame}{Reward 2}
+\begin{align*}
+R_\text{packets}\text{Sigmoid}_\alpha\left(\frac{R_\text{lost}}{R_\text{lost} + R_\text{packets}} - 0.05\right) - R_\text{lost} - \frac{R_\text{delay}}{R_\text{packets}}
+\end{align*}
+\end{frame}
+
 \end{document}

### 2017-12-02 tex/al_dente.tex
Update paper

--- 2017-10-10 tex/al_dente.tex
@@
 \begin{frame}{Preliminary results -- Comparison}
-\centering
-\begin{tabular}{rll}
-\toprule
-Parameter & Value & Distribution \\
-\midrule
-Two-way propagation delay & $150\,$ms & constant \\
-Bottleneck bandwidth & $15\,$Mbit/s & constant \\
-Number of senders & $8$ & constant \\
-Flow length & $100\,$kB & exponential \\
-Time between flows & $0.5\,$s & exponential \\
-Simulation duration & $100\,$s & constant \\
-Buffer size & $1000$ packets & constant \\
-Stochastic loss prob. & $0\%$ & constant \\
-\bottomrule
-\end{tabular}
+Stuff goes here!
+
 \end{frame}
 
-\newcommand{\nodedistance}{2}
-\newcommand{\nd}[1]{\numexpr #1 * \nodedistance \relax}
 \begin{frame}{Petri net}
 
-\centering
-\begin{tikzpicture}[node distance=\nodedistance,>=stealth',bend angle=45,auto]
+\input{petri_net.tex}
 
-  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
-  \tikzstyle{red place}=[place,draw=red!75,fill=red!20]
-  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
-  			  fill=black!20,minimum size=4mm]
+\end{frame}
 
-%  \tikzstyle{every label}=[blue!75]
+\begin{frame}{Neural Network}
 
-  \begin{scope}
-    \node [place,tokens=1] (s1) at (\nd{0}, \nd{1}) [label={[align=center]left:Receive\\ feedback}] {};
-    \node [place] (s2) at (\nd{1}, \nd{2}) [label={[align=center]above:Received\\ ACKs}] {};
-    \node [place] (s3) at (\nd{2}, \nd{1}) [label={[align=center]right:Packets\\ in the\\ network}] {};
-    \node [place,tokens=1] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};
+\input{neural_network.tex}
 
-    \node [transition] (e1) at (\nd{0}, \nd{0}) [label=left:Take action] {}
-      edge [post] node {$a$}(s4)
-      edge [pre] node {1} (s1);
+\end{frame}
 
-    \node [transition] (e2) at (\nd{2}, \nd{0}) [label={[align=center]right:Send\\ packet}] {}
-      edge [pre] node {1} (s4)
-      edge [post] node {1} (s3);
+\begin{frame}{Reward 1}
+\begin{align*}
+\frac{R_\text{packets}}{R_\text{duration}}\text{Sigmoid}_\alpha\left(\frac{R_\text{lost}}{R_\text{lost} + R_\text{packets}} - 0.05\right) - \frac{R_\text{lost}}{R_\text{duration}} - \frac{R_\text{delay}}{R_\text{packets}}
+\end{align*}
+\end{frame}
 
-    \node [transition] (e3) at (\nd{2}, \nd{2}) [label={[align=center]right:Receive\\ ACK}] {}
-      edge [pre] node {1} (s3)
-      edge [post] node {1} (s2);
-
-    \node [transition] (e4) at (\nd{1}, \nd{1}) [label={[align=center]left:Update\\open\\cwnd}] {}
-      edge [pre] node {1} (s2)
-      edge [post] node {1} (s4);
-
-    \node [transition] (e5) at (\nd{0}, \nd{2}) [label={[align=center]left:Received\\ feedback}] {}
-      edge [pre] node {1} (s2)
-      edge [post] node {1} (s1);
-
-  \end{scope}
-
-\end{tikzpicture}
+\begin{frame}{Reward 2}
+\begin{align*}
+R_\text{packets}\text{Sigmoid}_\alpha\left(\frac{R_\text{lost}}{R_\text{lost} + R_\text{packets}} - 0.05\right) - R_\text{lost} - \frac{R_\text{delay}}{R_\text{packets}}
+\end{align*}
 \end{frame}

### 2017-12-02 tex/neural_network.tex
Update paper

\newcommand{\nnfontsize}{\tiny}

\centering
\begin{tikzpicture}[node distance=1.2cm, >=stealth',bend angle=45,auto]

  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
  \tikzstyle{input place}=[circle,thick,draw=olive!75,fill=olive!20,minimum size=6mm]
  \tikzstyle{output place}=[circle,thick,draw=magenta!75,fill=magenta!20,minimum size=6mm]
   \tikzstyle{lstm place}=[circle,thick,draw=red!75,fill=red!20,minimum size=6mm]
   \tikzstyle{softplus place}=[circle,thick,draw=green!75,fill=green!20,minimum size=6mm]

  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
  			  fill=black!20,minimum size=4mm]

%  \tikzstyle{every label}=[blue!75]

  \begin{scope}
    \node [input place] (s1) [label={[align=center]left:\nnfontsize State (12)}] {};
    \node [place] (s2) [above right of=s1, label={[align=center]above:\nnfontsize Linear}] {};

  \path[->] (s1) edge node {} (s2);

    \node [lstm place] (s3) [right of=s2, label={[align=center]below:\nnfontsize LSTM (32)}] {};

  \path[->] (s2) edge node {} (s3); 
  \path[->] (s3) edge [loop above] node {} (s3);

    \node [lstm place] (s4) [right of=s3, label={[align=center]below:\nnfontsize LSTM (32)}] {};

  \path[->] (s3) edge node {} (s4);
  \path[->] (s4) edge [loop above] node {} (s4);

    \node [softplus place] (s7) [right of=s4, label={[align=center]below:{\nnfontsize Softplus}}] {};

  \path[->] (s4) edge node {} (s7);

    \node [output place] (s8) [right of=s7, label={[align=center]below:\nnfontsize $\sigma$~(1)}] {};

  \path[->] (s7) edge node {} (s8);

    \node [place] (s5) [above of=s7, label={[align=center]above:\nnfontsize Linear}] {};

  \path[->] (s4) edge node {} (s5);

    \node [output place] (s6) [right of=s5, label={[align=center]above:\nnfontsize $\mu$ (1)}] {};

  \path[->] (s5) edge node {} (s6);

    \node [place] (s10) [below right of=s1, label={[align=center]below:\nnfontsize Linear}] {};

  \path[->] (s1) edge node {} (s10);

    \node [lstm place] (s11) [right of=s10, label={[align=center]above:\nnfontsize LSTM (32)}] {};

  \path[->] (s10) edge node {} (s11);
  \path[->] (s11) edge [loop below] node {} (s11);

    \node [lstm place] (s12) [right of=s11, label={[align=center]above:\nnfontsize LSTM (32)}] {};

  \path[->] (s11) edge node {} (s12);
  \path[->] (s12) edge [loop below] node {} (s12);

    \node [softplus place] (s13) [right of=s12, label={[align=center]above:\nnfontsize Softplus}] {};

  \path[->] (s12) edge node {} (s13);

    \node [output place] (s17) [right of=s13, label={[align=center]above:\nnfontsize $V_\text{packets}$ (1)}] {};

  \path[->] (s13) edge node {} (s17);

    \node [softplus place] (s15) [below of=s13, label={[align=center]above:\nnfontsize Softplus}] {};

  \path[->] (s12) edge node {} (s15);

    \node [output place] (s19) [right of=s15, label={[align=center]above:\nnfontsize $V_\text{delay}$ (1)}] {};

  \path[->] (s15) edge node {} (s19);

    \node [softplus place] (s16) [below of=s15, label={[align=center]above:\nnfontsize Softplus}] {};

  \path[->] (s12) edge node {} (s16);

    \node [output place] (s20) [right of=s16, label={[align=center]above:\nnfontsize $V_\text{lost}$ (1)}] {};

  \path[->] (s16) edge node {} (s20);

    \node [softplus place] (s14) [below of=s16, label={[align=center]above:\nnfontsize Softplus}] {};

  \path[->] (s12) edge node {} (s14);

    \node [output place] (s21) [right of=s14, label={[align=center]above:\nnfontsize $V_\text{duration}$ (1)}] {};

  \path[->] (s14) edge node {} (s21);

%    \node [softplus place] (s17) [below of=s14, label={[align=center]above:\nnfontsize Softplus}] {};
%    
%  \path[->] (s12) edge node {} (s17);
%    
%    \node [output place] (s22) [right of=s17, label={[align=center]above:\nnfontsize $V_\text{byte}$ (1)}] {};   
%    
%  \path[->] (s17) edge node {} (s22);

  \end{scope}

\end{tikzpicture}

### 2017-12-02 tex/paper.tex
Update paper

@@
 \begin{document}
-\title{Let's Play Congestion Control}
+\title{Some title}
 \titlenote{Produces the permission block, and
@@
 
-\maketitle
+\settopmatter{printfolios=true}
 
-\section{Partial Action Actor Critic Learning}
-\label{sec:pal}
+\maketitle%
 
-Partial Action Actor Critic Learning (PAL) is a modification of the Asynchronous Advantage Actor Critic algorithm proposed by \citet{mnih_asynchronous_2016}.
-It consists of two Artificial Neural Networks (ANNs), the Actor Network and the Value Network (the Critic part in the abbreviation stands for the Value Network).
-Given a state, the Actor Networks outputs what it deems to be the optimum action to perform in that certain state.
+\section{Introduction}
+
+In recent years, there have been several approaches to make congestion control learnable, most notably \textit{Remy} \citep{winstein_tcp_2013}, which learns an optimum congestion control w.r.t.~a specific utility function for specified network scenarios in an off-line fashion.
+Furthermore there is \textit{Performance-oriented Congestion Control} (PCC) \citep{dong_pcc:_2015}, which repeatedly tries different rates during periods of approximately one round-trip time (RTT) and chooses the better rate w.r.t.~a specific utility function.
+
+We make use of a new Reinforcement Learning (RL) based approach that can learn a congestion control algorithm on-line and is able to perform actions on a per acknowledgement basis.
+
+\subsection{Actor Critic Learning}
+\label{subsec:ac}
+
+To this end we propose \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) which is a modification of the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline in this section.
+
+There are two Artificial Neural Networks (ANNs), the Actor Network and the Value Network (the Critic part in the abbreviation stands for the Value Network).
+Given a state, the Actor Network outputs what it deems to be the optimum action to perform in that certain state.
 The Value Network estimates what long-term reward can be expected in this state.
@@
 
-\subsection{Algorithm}
-\label{subsec:alg}
+\subsubsection{Value Network}
+\label{subsubsec:genericvalue}
 
-In \autoref{alg:pal} we show the code that runs in one of the learning threads.
-As proposed in literature, we use several concurrent threads to improve training performance.
-A major difference to previous approaches to reinforcement learning is the fact that we allow more actions to be taken before the previous actions' rewards have been received.
-In classical reinforcement learning, an action is always followed by a reward and a reward is always followed by a action.
-In our proposed concept, however, it is possible to take new actions while the previous action hasn't received its reward yet.
+The Value Network outputs the expected long-term reward $V(s_{t}; \theta_\text{v})$ given a state $s_t$ at time step $t$ and the parameters (neural network weights) of the value network $\theta_v$.
+
+With $r_t$ being the reward that was received at time $t$ and $\gamma$ being the roll-off factor, which stands for the influence that future reward has on the moving average (commonly set to $0.99$), we define the expected long-term reward as
+\begin{align*}
+R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta_\text{v})\right)\left( 1-\gamma \right),
+\end{align*}
+where $k$ is upper-bounded by $t_\text{max}$ ($t_\text{max}$ is a fixed hyperparameter that indicates how many rewards should be received before updating the neural network).
+So $R_t$ is simply a moving average of rewards at time step $t$.
+However, usual moving averages take into account values from the past while this one uses values from the future.
+In practice this means that we collect $t_\text{max}$ rewards, compute the expected long-term reward starting at each time step, update the neural network and start the same procedure again.
+
+%\todo{Explanation not so clear?}
+
+The loss function, which the value network tries to minimize at each time step, is the square of the difference of the actual long-term reward received and the expected long-term reward
+\begin{align*}
+l_{\text{v},t} = \left(R_t - V(s_t; \theta_\text{v})\right)^2.
+\end{align*}
+
+\subsubsection{Actor Network}
+\label{subsubsec:genericactor}
+
+The Actor Network outputs a probability distribution from which the action $a_t$ at time step $t$ is randomly sampled.
+We use the mean $\mu$ and the standard deviation $\sigma$, to parametrize a normal distribution.
+The main idea is that the network learns to output the right mean at the right time step to maximize the future reward and that it uses the standard deviation to try out new actions, which could yield a better than expected reward.
+
+With $v_t$ being the value that the value network estimated as the future reward given the current state $s_t$ at time step $t$ ($v_t$ is just an abbreviation for $V(s_t; \theta_\text{v})$) and $\theta_a$ the parameters of the actor network and with $\beta$ being a factor that specifies the importance of the entropy $H$, $\pi$ designating the probability density function, meaning that $\pi\left( a_t \given s_t; \theta_\text{a} \right)$ is the value of the probability density function of taking action $a_t$ in state $s_t$ with the current weights of the actor network $\theta_\text{a}$, we define the loss that the actor network aims to minimize as follows:
+\begin{align*}
+l_{\text{a},t} =& -\log \left( \pi\left( a_t \given s_t; \theta_\text{a} \right)\right)\left( R_t - v_t \right)\\ 
+&- \beta H\left( \pi\left( s_t; \theta_a \right)\right).
+\end{align*}
+
+\section{Method}
+\subsection{Partial Action Learning}
+\label{subsec:pal}
+
+The key difference between PAL and previous approaches to Reinforcement Learning is that in classical Reinforcement Learning, an action is always followed by a reward and a reward is always followed by a action.
+In our proposed concept, however, it is possible to take new actions while previous actions haven't received their rewards yet.
+
+Another major difference in PAL is that one action generates a number of partial actions ($\geq 0$) (see \autoref{fig:pal}).
+Each partial action generates feedback upon interacting with the environment.
+Upon receiving feedback for a partial action, the agent determines the current state and triggers a new action.
+When all feedbacks of one action were received, the agent combines them to form the reward and updates the value and actor networks.
+
+In \autoref{alg:pal} we show the code that runs in each of the agents (in the congestion control scenario, one agent corresponds to a sender).
+It is possible to have several agents which share a set of weights but one can also use separate weights for each agent, which is more realistic in case of congestion control in the Internet, as different senders cannot easily share a set of neural network weights over the Internet.
 
@@
 \includesvg{figures/Reinforcement_learning_diagram}
-\caption{The classical reinforcement learning approach.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
+\caption{The classical Reinforcement Learning approach.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
 \label{fig:reinforcement}
@@
 
-Another major difference in PAL is that one action generates a number of partial actions ($\geq 0$) (see \autoref{fig:pal}).
-Each partial action generates feedback upon interacting with the environment.
-Upon receiving feedback for a partial action, the agent determines the current state and triggers a new action.
-When all partial rewards of one action were received, the agent combines them to form the reward and updates the value and actor networks (see \autoref{alg:grad}).
-
 \begin{figure}
@@
 \includesvg{figures/Reinforcement_learning_diagram_PAL}
-\caption{Partial Action Actor Critic Learning: An action consists of zero or more partial actions which trigger partial rewards upon interacting with the environment.
-Each partial reward updates the state.
-The value and actor networks are updated upon receiving all partial rewards of one action.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
+\caption{Partial Action Learning: An action consists of zero or more partial actions which trigger feedback upon interacting with the environment.
+Each feedback updates the state.
+The value and actor networks are updated upon receiving all feedback of one action.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
 \label{fig:pal}
@@
 \begin{algorithm}
-\caption{Partial Action Actor Critic Learning -- pseudocode for each learning thread.
-Only $\theta_{\text{a},\text{g}}$ and $\theta_{\text{v},\text{g}}$ are shared between the threads.
-They are initialized randomly in the beginning.
-\todo{Update algorithm}}
+\caption{Partial Action Learning -- pseudocode for each agent.
+It is possible that the agents share the global weights $\theta_{\text{a},\text{g}}$ and $\theta_{\text{v},\text{g}}$, which can be reasonable when learning off-line.
+Otherwise an individual copy of them is kept by each agent.
+All weights are initialized randomly in the beginning.
+LSTM cells are used in both the actor and the value network.}
 \label{alg:pal}
@@
 	\State $l_\text{snapshots} \gets  [] $
+	\State $l_\text{hiddenStates} \gets  [] $
 	\State $t \gets 0$
 	\State $s_0 \gets \text{initialState}()$
+	\State $h_\text{a} \gets \text{initialHiddenState}()$
+	\State $h_\text{v} \gets \text{initialHiddenState}()$
 	\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
@@
 	\Repeat
-		\State $l_\text{states}.\text{append}(s_t)$
 		\State $l_\text{estimatedValues}.\text{append}(V(s_t; \theta_\text{v}))$
@@
 			\State $l_\text{snapshots}.\text{append}((\theta_\text{a}, \theta_\text{v}))$
+			\State $l_\text{hiddenStates}.\text{append}((h_\text{a}, h_\text{v}))$
 		\EndIf
+		\State $l_\text{states}.\text{append}(s_t)$
 		\State Sample $a_t$ from the 
@@
 		\State $l_\text{actions}.\text{append}(a_t)$
+		\State $l_\text{values}.\text{append}(V(s_t; \theta_\text{v}))$
 		\State $l_{\text{partialActions},t} \gets \text{partialActions}(a_t)$
-		\ForAll{partial actions $a_{{\text{p},n,t}}$ in $l_{\text{partialActions},t}$}
-			\State Take partial action $a_{{\text{p},n,t}}$
+		\ForAll{partial actions $a_{{\text{p},i,t}}$ in $l_{\text{partialActions},t}$}
+			\State Take partial action $a_{{\text{p},i,t}}$
 		\EndFor
-		\State Receive feedback $r_{\text{p},n,t'}$
-		\StateIndent[3] where $t' \leq t$ and $0 \leq n \leq\#(l_{\text{partialActions},t'})$
+		\State Wait for the next feedback $r_{\text{p},j,t'}$
+		\StateIndent[3] where $t' \leq t$ and $0 \leq j \leq\#(l_{\text{partialActions},t'})$
 		\If{all feedback of $a_{t'}$ was received}
-			\State $r_{t'} \gets \text{reward w.r.t all feedback }r_{\text{p},n,t'}$
+			\State $r_{t'} \gets \text{reward w.r.t all feedback }r_{\text{p},k,t'}$
+					\StateIndent[4] where $0 \leq k \leq\#(l_{\text{partialActions},t'})$
 			\State $l_\text{rewards}.\text{append}(r_{t'})$
@@
 \begin{algorithm}
-\caption{Partial Action Actor Critic Learning -- procedure which computes and applies the gradients.}
+\caption{Partial Action Learning -- procedure which computes and applies the gradients.}
 \label{alg:grad}
@@
 	\State $\theta_\text{backup} \gets \left(\theta_\text{a}, \theta_\text{v}\right)$
+	\State $h_\text{backup} \gets \left(h_\text{a}, h_\text{v}\right)$
 	\State $\theta_\text{a}, \theta_\text{v} \gets l_\text{snapshots}[0]$
-	\State $R_{i+1} \gets \frac{l_\text{estimatedValues}[t_\text{end}]}{1-\gamma}$
+	\State $h_\text{a}, h_\text{v} \gets l_\text{hiddenStates}[0]$
+	\State $R_{i+1} \gets$ last element of $l_\text{estimatedValues}$%
 	\For{$i\gets t_\text{end}-1, 0$}
-		\State $R_i \gets r_i + \gamma R_{i+1}$
+		\State $R_i \gets \left(r_i + \gamma R_{i+1}\right)\left(1-\gamma\right)$
 		\State $a \gets l_\text{actions}[i]$
@@
 		\State $v \gets l_\text{values}[i]$
+		\State $d\theta_\text{a} \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
+		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta_a}$
 		\State $d\theta_\text{v} \gets \frac{\partial\left(R_i - v\right)}{\partial\theta_v}$
-		\State $d\theta_\text{v} \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
-		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta_a}$
 		\State $\theta_{\text{a},\text{g}} \gets \theta_{\text{a},\text{g}} + d\theta_\text{a}$
@@
 	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{estimatedValues}$
-	\State Remove first element from $l_\text{snapshots}$
+	\State Remove the first element from $l_\text{snapshots}$
+	\State Remove the first element from $l_\text{hiddenStates}$
 	\State $\theta_\text{a}, \theta_\text{v} = \theta_\text{backup}$
+	\State $h_\text{a}, h_\text{v} = h_\text{backup}$
 \EndFunction
@@
 
-\subsection{Value Network}
-\label{subsec:genericvalue}
-
-The value network estimates the expected long-term reward given a state $s_t$.
-
-Let $r_t$ be the reward that was received at time $t$.
-$V(s_t; \theta_\text{v})$ designates the expected mean reward in state $s_t$ given the parameters (neural network weights) of the value network $\theta_v$.
-Then, with $\gamma$ being the roll-off factor, which designates the influence that future reward has on the moving average, (set to $0.99$), we define the average reward as
-\begin{align*}
-R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta_\text{v})\right),
-\end{align*}
-where $k$ is upper-bounded by $t_\text{max}$ ($t_\text{max}$ is a fixed hyperparameter that indicates how many steps should be performed before updating the neural network).
-So $R$ is simply a moving average of rewards.
-However, usual moving averages take into account values from the past while this one uses values from the future; it runs reversely.
-
-The loss function, which the value network tries to minimize, is the square of the difference of the actual average reward received and the expected average reward
-\begin{align*}
-l_{\text{v},t} = \left(R_t - V(s_t; \theta_\text{v})\right)^2.
-\end{align*}
-
-\subsection{Actor Network}
-\label{subsec:genericactor}
-
-Let $a_t$ be the action that was taken at time step $t$, $v_t$ the value that the value network estimated as the future reward given the current state $s_t$ at time step $t$ ($v_t$ is just an abbreviation for $V(s_t; \theta_\text{v})$) and $\theta_a$ the parameters (neural network weights) of the actor network.
-Furthermore, let $\beta$ be a factor that specifies the importance of the entropy $H$.
-$\pi$ designates the probability density function, which means that $\pi\left( a_t \given s_t; \theta_\text{a} \right)$ is the value of the probability density function of taking action $a_t$ in state $s_t$ with the current weights of the actor network $\theta_\text{a}$.
-Then
-\begin{align*}
-l_{\text{a},t} =& -\log \left( \pi\left( a_t \given s_t; \theta_\text{a} \right)\right)\left( R_t - v_t \right)\\ 
-&- \beta H\left( \pi\left( s_t; \theta_a \right)\right)
-\end{align*}
-is the actor loss that we try to minimize.
-
-\section{Congestion Control specifics}
+\subsection{Congestion Control specifics}
 
@@
 
-\newcommand{\nodedistance}{2}
-\newcommand{\nd}[1]{\numexpr #1 * \nodedistance \relax}
-
-\begin{tikzpicture}[node distance=\nodedistance,>=stealth',bend angle=45,auto]
-
-  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
-  \tikzstyle{red place}=[place,draw=red!75,fill=red!20]
-  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
-  			  fill=black!20,minimum size=4mm]
-
-%  \tikzstyle{every label}=[blue!75]
-
-  \begin{scope}
-    \node [place,tokens=1] (s1) at (\nd{0}, \nd{1}) [label={[align=center]left:Receive\\ feedback}] {};
-    \node [place] (s2) at (\nd{1}, \nd{2}) [label={[align=center]above:Received\\ ACKs}] {};
-    \node [place] (s3) at (\nd{2}, \nd{1}) [label={[align=center]right:Packets\\ in the\\ network}] {};
-    \node [place,tokens=1] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};
-
-    \node [transition] (e1) at (\nd{0}, \nd{0}) [label=left:Take action] {}
-      edge [post] node {$a$}(s4)
-      edge [pre] node {1} (s1);
-
-    \node [transition] (e2) at (\nd{2}, \nd{0}) [label={[align=center]right:Send\\ packet}] {}
-      edge [pre] node {1} (s4)
-      edge [post] node {1} (s3);
-
-    \node [transition] (e3) at (\nd{2}, \nd{2}) [label={[align=center]right:Receive\\ ACK}] {}
-      edge [pre] node {1} (s3)
-      edge [post] node {1} (s2);
-
-    \node [transition] (e4) at (\nd{1}, \nd{1}) [label={[align=center]left:Update\\open\\cwnd}] {}
-      edge [pre] node {1} (s2)
-      edge [post] node {1} (s4);
-
-    \node [transition] (e5) at (\nd{0}, \nd{2}) [label={[align=center]left:Received\\ feedback}] {}
-      edge [pre] node {1} (s2)
-      edge [post] node {1} (s1);
-
-  \end{scope}
-
-\end{tikzpicture}
+\input{petri_net.tex}
 
 \caption{A petri net describing the congestion control mechanism.
-We start with one token in the state \textit{Open congestion window}, which means that initially we can send one packet.
-We start with another token in the state \textit{Received feedback}, which means that we can take an action right in the beginning.
+We start with one token in the state \textit{Received feedback}, which means that we can take an action right in the beginning.
 The concept is the following: If there is at least one token in the \textit{Open congestion window}, send a packet.
-The acknowledgement (or timeout if the packet gets lost) for the packet is received by the \textit{Open congestion window}, meaning that when we get an ACK we can send another packet (The \textit{Open congestion window} signifies the congestion window minus the packets that are currently unacknowledged.).
-Furthermore the state \textit{Received feedback} receives each ACK/timeout and takes an action.
+The token goes to the network and after some time the acknowledgement (or timeout if the packet gets lost) for the packet is received in the \textit{Received ACK} state.
+From here one token goes to the \textit{Open congestion window} meaning that when we get an ACK (or timeout) we can send another packet (The \textit{Open congestion window} signifies the congestion window minus the packets that are currently unacknowledged.).
+Furthermore the state \textit{Received feedback} receives a token upon receiving an ACK/timeout.
 \textit{Take action} adds \textit{a} tokens to the \textit{Open congestion window}, where $a$ is a real number (possibly also negative); thus in each state there can also be a real number of tokens (e.g.~ 2.34 tokens are possible in the \textit{Open congestion window}).
-However, we define that the congestion window can never be smaller than 1.}
+However, we define that the congestion window can never be smaller than 1 (This means that the sum of all tokens in the \textit{Open congestion window.}, \textit{Received ACK} and \textit{Packets in the network} states can never be smaller than 1.}
 \label{fig:petri}
@@
 
-The classical Reinforcement Learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
+The motivation for PAL is that classical Reinforcement Learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
 However, in the case of congestion control, it is desirable to perform a new action without having received a reward for the previous action.
-For example, imagine that you receive two acknowledgements directly after each other.
-For each of these two acknowledgements an action has to be performed but it does not seem feasible for the second action to wait until the first action has received a reward.
-The Asynchronous Actor Critic framework as described by \cite{mnih_asynchronous_2016} cannot be applied to congestion control as it assumes that actions and rewards are synchronized and thus we use the proposed Partial Action Actor Critic Learning (see \autoref{sec:pal}).
-We describe the concept of using reinforcement learning for congestion control with a petri net (see \autoref{fig:petri}).
+For example, imagine that we receive two acknowledgements directly after one other.
+For each of these two acknowledgements an action has to be performed but it does not seem feasible for the second action to wait until the first action has received a reward as it takes one RTT until an acknowledgement (a reward) is received.
+Thus the Asynchronous Actor Critic framework as described by \cite{mnih_asynchronous_2016} cannot be applied to congestion control as it assumes that actions and rewards are synchronized and so we have to use Partial Action Learning (see \autoref{subsec:pal}).
+We describe the overall workings of PAL for congestion control using a petri net (see \autoref{fig:petri}).
 
-In the following a time step $t$ corresponds to the reception of an acknowledgement. The beginning of the flow, before any packet is sent, corresponds to time step $0$.
+To use PAL for congestion control we first have to define the correct semantics for this specific use case and we have to explicitly state how the state, reward etc.~are defined.
+Furthermore, we have to define how the actor and value network explicitely work in case of congestion control.
+In the following a time step $t$ corresponds to the reception of an acknowledgement.
+The beginning of the flow, before any packet is sent, corresponds to time step $0$.
 
 \begin{description}
-\item[$\textit{s}_\text{t}$] The state  describes the current ``congestion state''.
-Various features that indicate congestion are included in it.
+\item[$\textit{s}_\text{t}$] The state  describes the current ``congestion state''. The following features are included in it:
+\begin{itemize}
+\item the round-trip time of the last packet
+\item the current congestion window
+\item the time between the last two packets that were sent
+\item the time between the last two packets that were received
+\item the number of packets that were lost since the last acknowledgement was received
+\end{itemize}
 Each time an acknowledgement is received, the state is updated and the actor network is asked for the next action.
-\item[$\textit{a}_\text{t}$] Based on a given state and the history of previous states, the actor network returns an action $a_t$, which is a real number ($\geq 1$) that stands for the congestion window to be used until the next acknowledgement is received.
-\item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metric. These are discussed in more detail in \autoref{subsec:value}
-\item[$\textit{v}_\text{t}$] The value is a tuple of the expected average reward estimated by the value network (see \autoref{subsec:value}) for each of the reward metrics. 
+\item[$\textit{a}_\text{t}$] Based on a given state and the history of previous states (because we use LSTM cells and thus can also consider previous states), the actor network returns an action $a_t$, which is a real number that stands for the change of the congestion window.
+\item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metric. 
+%These are discussed in more detail in \autoref{subsubsec:value}. 
+For each reward metric there is also an output of the value network that predicts the expected long-term average of this reward metric given the current state. 
+%\item[$\textit{v}_\text{t}$] The value is a tuple of the expected average reward estimated by the value network (see \autoref{subsubsec:value}) for each of the reward metrics. 
 \end{description}
 
-We actually use four different types of reward: 
+We actually use the following types of reward: 
 \begin{description}
 \item[$\textit{r}_{\text{packet},t}$] is the number of the packets that the sender sent during time step $t$ and that were not lost (so they were acknowledged at some point by the receiver).
-\item[$\textit{r}_{\text{byte},t}$] is the sum of the bytes of the packets that the sender sent and that were not lost.
-Example: During $t$, three packets were sent with 300, 200, and 1500 bytes each so $\textit{r}_{\text{byte},t} = 2000$
+%\item[$\textit{r}_{\text{byte},t}$] is the sum of the bytes of the packets that the sender sent and that were not lost.
+Example: At time $t'$, three packets were sent with 300, 200, and 1500 bytes each so $\textit{r}_{\text{byte},t'} = 2000$
 \item[$\textit{r}_{\text{delay},t}$] is the sum of the round trip times of the packets that the sender sent and that were not lost.
 \item[$\textit{r}_{\text{duration},t}$] is the sum of the time between receiving the last packet and receiving this packet (``inter-receive time'') for the packets that the sender sent and that were not lost.
+\item[$\textit{r}_{\text{lost},t}$] is the sum of packets that were lost between receiving the last packet and receiving this packet.
 \end{description}
@@
 
-\subsection{Value Network}
-\label{subsec:value}
+The overall structure of the neural network for both the value and the actor network is depicted in \autoref{fig:neural_network}.
 
-In the case of congestion control, the loss function $l_{\text{v},t}$ of the value network is actually the sum of the square of the difference for each of the four moving averages for each type of reward:
+\begin{figure}
+
+\input{neural_network.tex}
+
+\caption{An overview of the complete neural network being used.
+Ocher stands for the input, blue for linear layers, red for LSTM layers, green for linear layers with a softplus activation function and purple for the outputs of the neural network.
+The numbers in parentheses next to each label stand  for the number of neurons being used in that layer.}
+\label{fig:neural_network}
+\end{figure}
+
+\subsubsection{Value Network}
+\label{subsubsec:value}
+
+In the case of congestion control, the loss function $l_{\text{v},t}$ of the value network is actually the sum of the squares of the difference for each of the expected long-term averages and the empirically found averages for each reward metric:
 \begin{align*}
 l_{\text{v},t} =& \left(R_{\text{packet},t} - V_\text{packet}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{byte},t} - V_\text{byte}(s_t; \theta_\text{v})\right)^2 \\
+%&+\left(R_{\text{byte},t} - V_\text{byte}(s_t; \theta_\text{v})\right)^2 \\
 &+\left(R_{\text{delay},t} - V_\text{delay}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{duration},t} - V_\text{duration}(s_t; \theta_\text{v})\right)^2
+&+\left(R_{\text{duration},t} - V_\text{duration}(s_t; \theta_\text{v})\right)^2 \\
+&+\left(R_{\text{lost},t} - V_\text{lost}(s_t; \theta_\text{v})\right)^2
 \end{align*}
 
-\subsection{Actor Network}
-\label{subsec:actor}
+\subsubsection{Actor Network}
+\label{subsubsec:actor}
 
@@
 
-Each time an action $a_t$ is requested, a value $X$ is sampled from the current normal distribution defined by the parameters $\mu$ and $\sigma$: 
-\begin{align*}
-X&\sim {\mathcal {N}}(\mu ,\sigma^{2})\\
-\end{align*}
+Each time an action $a_t$ is requested, it is sampled from the current normal distribution defined by the parameters $\mu$ and $\sigma$.
 
-$a_t$ is the sampled value $X$ at time $t$ and the window is incremented by $a_t$; however, the window can never be smaller than 1. At the beginning of a flow the window is 1 as well.
+Then the window is incremented by $a_t$; however, the window can never be smaller than 1. At the beginning of a flow the window starts with 1 as well.
 
 With $H$ being the entropy, the actor network minimizes the loss
+%\begin{align*}
+%l_{\text{a},t} =& -\log \left( \pi \left( a_t \given s_t ; \theta_\text{a} \right) \right)\\
+%&\left( \left(\log\left(\frac{R_{\text{byte},t}}{{R_{\text{duration},t}}}\right) - \log\left(\frac{v_{\text{byte},t}}{{v_{\text{duration},t}}}\right)\right) - \left( \frac{R_{\text{delay},t}}{{R_{\text{packet},t}}}- \frac{v_{\text{delay},t}}{{v_{\text{packet},t}}} \right)\right)\\
+%&+ \beta H\left( \pi\left( s_t; \theta_a \right)\right).
+
 \begin{align*}
 l_{\text{a},t} =& -\log \left( \pi \left( a_t \given s_t ; \theta_\text{a} \right) \right)\\
-&\left( \left(\log\left(\frac{R_{\text{byte},t}}{{R_{\text{duration},t}}}\right) - \log\left(\frac{v_{\text{byte},t}}{{v_{\text{duration},t}}}\right)\right) - \left( \frac{R_{\text{delay},t}}{{R_{\text{packet},t}}}- \frac{v_{\text{delay},t}}{{v_{\text{packet},t}}} \right)\right)\\
-&+ \beta H\left( \pi\left( s_t; \theta_a \right)\right).
+& \left( U_{\text{measured},t} - U_{\text{expected},t} \right)\\ 
+&- \beta H\left( \pi\left( s_t; \theta_a \right)\right)
 \end{align*}
+where $U$ can be any utility function defined based on some of the reward metrics.
+In the above formulation, one considers how much better the actual experienced Utility was compared to the expected one.
+The Utility is formed from the previously defined expected long-term values.
+
+\subsection{Utility function}
+
+As the utility function one can choose an arbitrary function based on the reward metrics previously defined.
+A safe choice is to use PCC's \citep{dong_pcc:_2015} reward function as its convergence has been previously proven.
+It is defined as follows:
+
+\begin{align*}
+U_t = r_{\text{received},t}\text{Sigmoid}_\alpha\left(\frac{r_{\text{loss},t}}{r_{\text{sent},t}} - 0.05\right) + r_{\text{loss},t}
+\end{align*}
+where 
+\begin{align*}
+\text{Sigmoid}_\alpha(x) = \frac{1}{1+e^{\alpha x}},
+\end{align*}
+$r_\text{received}$ is the actual throughput (in packets per second), $r_\text{loss}$ is the loss rate (in packets per second) and $r_\text{sent}$ is the sending rate (in packets per second).
+Thus $\frac{r_\text{loss}}{r_\text{sent}}$ is the ratio of data getting lost compared to those being sent, which is a number between 0 and 1.
+
+Intuitively it means that one takes the actual throughput (packets that get received by the receiver and do not get lost) times the sigmoid function of the loss ratio minus 0.05 plus the loss rate.
+The sigmoid function essentially acts as a cutoff threshold.
+As soon as the loss rate rises above 5\%, the Sigmoid function decreases very quickly and thus diminishes the Utility.
+
+It is also possible to set the cutoff threshold to 0, which means that losing packets is heavily punished; or one can increase the threshold to allow for more loss.
+
+$r_{\text{received},t}$ can be calculated as $\frac{R_\text{packet},t}{R_\text{duration},t}$, $r_{\text{loss},t}$ can be calculated as $\frac{R_\text{lost},t}{R_\text{duration},t}$ and $r_{\text{sent},t}$ can be calculated as $\frac{{R_\text{packet},t} + {R_\text{lost},t}}{R_\text{duration},t}$.
+
+Actually, we assume equally sized packets for all of the above methodology, however, it would also easily be possible to consider the number of bytes per packet as well: One would have to add an additional reward metric of the average number of bytes per packet and as well add one more output to the value network.
 
 \subsection{Example}
+
+In this section, an example of the overall procedure is outlined.
 
@@
 \item If this was the last ACK, which completes a previous action (e.g.
-a previous action resulted in $3$ packets being sent; so if this is the ACK for the third (and last) packet of this previous action we got all the ACKs to compute the reward.)
-\item If we now have $t_{max}$ rewards (set to 20) then we also update the neural network. 
+a previous action resulted in $3$ packets being sent; so if this is the ACK for the third (and last) packet of this previous action, we got all the ACKs to compute the reward), combine all partial rewards to form the actual reward for that action.
+\item If we now have $t_\text{max}$ rewards (set to 20) then we also update the neural network. 
 \end{enumerate}
@@
 
-\subsection{Convergence}
-
-Assume we have $N$ senders, each with its specific baseline delay (this encompasses all delays excluding only the queueing delay) $d_i$.
+%\subsection{Convergence}
+%
+%Assume we have $N$ senders, each with its specific baseline delay (this encompasses all delays excluding only the queueing delay) $d_i$.
 Each sender keeps a congestion window $b_i$.
 The link speed at the bottleneck is designated as $c$.
-
-%We define the excessive window as 
+%
+%%We define the excessive window as 
+%%\begin{align*}
+%%b_\text{excess} = \max\left( \left( \sum_{j=1}^N \frac{b_j}{d_j} \right) - c, 0 \right)
+%%\end{align*}
+%%and the queueing delay as 
+%%\begin{align*}
+%%d_q = \frac{b_\text{excess}}{c}.
+%%\end{align*}
+%
+%With $b = (b_i)$ we observe the following relation:
+%
 %\begin{align*}
-%b_\text{excess} = \max\left( \left( \sum_{j=1}^N \frac{b_j}{d_j} \right) - c, 0 \right)
+%c = \sum_{j=1}^N \frac{b_j}{d_j + d_q(b)}
 %\end{align*}
-%and the queueing delay as 
+%
+%Then we define the utility function for each sender $i$ as
 %\begin{align*}
-%d_q = \frac{b_\text{excess}}{c}.
+%V_i(b) \coloneqq \begin{cases}
+%\log\left(\frac{b_i}{d_i}\right) - \delta_i d_i & \text{if } \sum_{j=1}^N \frac{b_i}{d_i} \leq c\\
+%\log\left(\frac{b_i}{d_i + d_q(b)}\right) - \delta_i (d_i + d_q(b)) & \text{otherwise}
+%\end{cases}
 %\end{align*}
-
-With $b = (b_i)$ we observe the following relation:
-
-\begin{align*}
-c = \sum_{j=1}^N \frac{b_j}{d_j + d_q(b)}
-\end{align*}
-
-Then we define the utility function for each sender $i$ as
-\begin{align*}
-V_i(b) \coloneqq \begin{cases}
-\log\left(\frac{b_i}{d_i}\right) - \delta_i d_i & \text{if } \sum_{j=1}^N \frac{b_i}{d_i} \leq c\\
-\log\left(\frac{b_i}{d_i + d_q(b)}\right) - \delta_i (d_i + d_q(b)) & \text{otherwise}
-\end{cases}
-\end{align*}
-
-\begin{align*}
-&\log\left(\frac{b_i}{d_i + d_q}\right) - \delta_i (d_i + d_q) \\
-=& \log\left(b_i\right) - \log\left(d_i + d_q\right) - \delta_i (d_i + d_q)
-\end{align*}
-
-\begin{align*}
-\frac{\partial V_i}{\partial b_i} = \begin{cases}
-\frac{1}{b_i} & \text{if } \sum_{j=1}^N \frac{b_i}{d_i} \leq c\\
-\frac{1}{b_i} - \frac{\partial d_q}{\partial b_i}\left(b_i\right)\frac{1}{d_i+d_q\left(b_i\right)} - \delta\frac{\partial d_q}{b_i}\left(b_i\right) & \text{otherwise}
-\end{cases}
-\end{align*}
+%
+%\begin{align*}
+%&\log\left(\frac{b_i}{d_i + d_q}\right) - \delta_i (d_i + d_q) \\
+%=& \log\left(b_i\right) - \log\left(d_i + d_q\right) - \delta_i (d_i + d_q)
+%\end{align*}
+%
+%\begin{align*}
+%\frac{\partial V_i}{\partial b_i} = \begin{cases}
+%\frac{1}{b_i} & \text{if } \sum_{j=1}^N \frac{b_i}{d_i} \leq c\\
+%\frac{1}{b_i} - \frac{\partial d_q}{\partial b_i}\left(b_i\right)\frac{1}{d_i+d_q\left(b_i\right)} - \delta\frac{\partial d_q}{b_i}\left(b_i\right) & \text{otherwise}
+%\end{cases}
+%\end{align*}
 
@@
 
-To show diagonal strict concavity we have to show that d.s.c.~is fulfilled \begin{enumerate*} \item for two distinct points lying in $\sum_{j=1}^N \frac{b_i}{d_i} \leq c$ \item for two distinct points lying in $\sum_{j=1}^N \frac{b_i}{d_i} > c$ and \item when one point is in one part of the function and the other in the other one\end{enumerate*}.
-
-\todo{Insert what d.s.c~actually is and quote Altman's paper}
+%To show diagonal strict concavity we have to show that d.s.c.~is fulfilled \begin{enumerate*} \item for two distinct points lying in $\sum_{j=1}^N \frac{b_i}{d_i} \leq c$ \item for two distinct points lying in $\sum_{j=1}^N \frac{b_i}{d_i} > c$ and \item when one point is in one part of the function and the other in the other one\end{enumerate*}.
+%
+%\todo{Insert what d.s.c~actually is and quote Altman's paper}
 
@@
 
-\section{Results}
-
-We train on the following scenario:
-
-\begin{table}[h]
-\centering
-\begin{tabular}{rll}
-\toprule
-Parameter & Value & Distribution \\
-\midrule
-Two-way propagation delay & $100\,$ms & constant \\
-Bottleneck bandwidth & $3\,$Mbit/s & constant \\
-Number of senders & $2$ & constant \\
-Flow length & $\infty$ & constant \\
-Simulation duration & $50\,$s--$150\,$s & uniform \\
-Buffer size & $\infty$ & constant \\
-Stochastic loss prob. & $0\%$ & constant \\
-\bottomrule
-\end{tabular}
-\caption{The parameters of our very first simple experiment}
-\label{tab:smallexperiment}
-\end{table}
-
-\todo{Insert results when comparing to regular TCP}
+%\section{Results}
+%
+%We train on the following scenario:
+% 
+%\begin{table}[h]
+%\centering
+%\begin{tabular}{rll}
+%\toprule
+%Parameter & Value & Distribution \\
+%\midrule
+%Two-way propagation delay & $100\,$ms & constant \\
+%Bottleneck bandwidth & $3\,$Mbit/s & constant \\
+%Number of senders & $2$ & constant \\
+%Flow length & $\infty$ & constant \\
+%Simulation duration & $50\,$s--$150\,$s & uniform \\
+%Buffer size & $\infty$ & constant \\
+%Stochastic loss prob. & $0\%$ & constant \\
+%\bottomrule
+%\end{tabular}
+%\caption{The parameters of our very first simple experiment}
+%\label{tab:smallexperiment}
+%\end{table}
+%
+%\todo{Insert results when comparing to regular TCP}
 

### 2017-12-02 tex/petri_net.tex
Update paper

\newcommand{\nodedistance}{2}
\newcommand{\nd}[1]{\numexpr #1 * \nodedistance \relax}

\centering
\begin{tikzpicture}[node distance=\nodedistance,>=stealth',bend angle=45,auto]

  \tikzstyle{place}=[circle,thick,draw=blue!75,fill=blue!20,minimum size=6mm]
  \tikzstyle{transition}=[rectangle,thick,draw=black!75,
  			  fill=black!20,minimum size=4mm]

%  \tikzstyle{every label}=[blue!75]

  \begin{scope}
    \node [place,tokens=1] (s1) at (\nd{0}, \nd{1}) [label={[align=center]left:Receive\\ feedback}] {};
    \node [place] (s2) at (\nd{1}, \nd{2}) [label={[align=center]above:Received\\ ACK}] {};
    \node [place] (s3) at (\nd{2}, \nd{1}) [label={[align=center]right:Packets\\ in the\\ network}] {};
    \node [place] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};

    \node [transition] (e1) at (\nd{0}, \nd{0}) [label=left:Take action] {}
      edge [post] node {$a$}(s4)
      edge [pre] node {1} (s1);

    \node [transition] (e2) at (\nd{2}, \nd{0}) [label={[align=center]right:Send\\ packet}] {}
      edge [pre] node {1} (s4)
      edge [post] node {1} (s3);

    \node [transition] (e3) at (\nd{2}, \nd{2}) [label={[align=center]right:Receive\\ ACK}] {}
      edge [pre] node {1} (s3)
      edge [post] node {1} (s2);

    \node [transition] (e4) at (\nd{1}, \nd{1}) [label={[align=center]left:Update\\open\\cwnd}] {}
      edge [pre] node {1} (s2)
      edge [post] node {1} (s4);

    \node [transition] (e5) at (\nd{0}, \nd{2}) [label={[align=center]left:Received\\ feedback}] {}
      edge [pre] node {1} (s2)
      edge [post] node {1} (s1);

  \end{scope}

\end{tikzpicture}

### 2017-12-04 async_deep_reinforce/a3c_training_thread.py
Hopefully commit all changes

@@ -246,7 +246,7 @@ class A3CTrainingThread(object):
 
     R_packets_initial, R_accumulated_delay_initial, R_duration_initial, R_lost_initial = R_packets, R_accumulated_delay, R_duration, R_lost
 
-    R_packets, R_accumulated_delay, R_duration, R_lost = (R_packets)/(1-GAMMA), (R_accumulated_delay)/(1-GAMMA), (R_duration)/(1-GAMMA), (R_lost)/(1-GAMMA)
+    # R_packets, R_accumulated_delay, R_duration, R_lost = (R_packets)/(1-GAMMA), (R_accumulated_delay)/(1-GAMMA), (R_duration)/(1-GAMMA), (R_lost)/(1-GAMMA)
     # logging.debug(" ".join(map(str,("exp(R_packets)", R_packets, "exp(R_accumulated_delay)", R_accumulated_delay, "exp(R_duration)", R_duration))))
     assert(np.isfinite(R_duration))
     assert(np.isfinite(R_packets))
@@ -268,15 +268,17 @@ class A3CTrainingThread(object):
 
     # compute and accmulate gradients
     for(ai, ri, si, Vi) in zip(actions, rewards, states, values):
+      # FIXME: Make sure that it actually works with how the roll-off factor gets normalized.
+      assert(false)
 
-      R_duration = (ri[2]*SECONDS_NORMALIZER + GAMMA*R_duration)
+      R_duration = (ri[2]*SECONDS_NORMALIZER + GAMMA*R_duration)*(1-GAMMA)
 
-      R_packets = (ri[0] + GAMMA*R_packets)
-      R_lost = (ri[3] + GAMMA*R_lost)
+      R_packets = (ri[0] + GAMMA*R_packets)*(1-GAMMA)
+      R_lost = (ri[3] + GAMMA*R_lost)*(1-GAMMA)
       # TODO: In theory it should be np.log(R_bytes/R_duration) but remy doesn't have bytes
       # td = (np.log(R_packets/R_duration) - np.log(Vi[0]/Vi[2]))
 
-      R_accumulated_delay = (ri[1]*SECONDS_NORMALIZER + GAMMA*R_accumulated_delay)
+      R_accumulated_delay = (ri[1]*SECONDS_NORMALIZER + GAMMA*R_accumulated_delay)*(1-GAMMA)
[…]

### 2017-12-04 async_deep_reinforce/constants.py

@@ -29,7 +29,7 @@ logging.info(" ".join(map(str,("CHECKPOINT_DIR:",CHECKPOINT_DIR))))
 LOG_FILE = ABSOLUTE_PATH+'tmp/a3c_log'
 ACTOR_FACTOR = 1e0
 VALUE_FACTOR = 1e0
-GENERAL_FACTOR = 1e-3
+GENERAL_FACTOR = 1e-4
 # INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate
 # INITIAL_ALPHA_HIGH = 1e0*GENERAL_FACTOR   # log_uniform high limit for learning rate
 INITIAL_RATE = GENERAL_FACTOR

### 2017-12-04 src/input-configrange-unicorn.cc

@@ -118,38 +118,38 @@ bool update_config_with_uint32(RemyBuffers::ConfigRangeUnicorn & range,
 // If the input arugments are invalid, or if mandatory is true and no input arguments
 // are found, calls exit(1).
 // If all input arguments were found, updates the range and returns true.
-bool update_config_with_double(RemyBuffers::ConfigRangeUnicorn & range,
-  void (RemyBuffers::ConfigRangeUnicorn::*set_fn)(unsigned int), int argc, char *argv[],
-  string arg_name, bool mandatory) {
+// bool update_config_with_double(RemyBuffers::ConfigRangeUnicorn & range,
+//   void (RemyBuffers::ConfigRangeUnicorn::*set_fn)(unsigned int), int argc, char *argv[],
+//   string arg_name, bool mandatory) {
 
-bool found = false;
-double value = NAN;
-int arg_name_size = arg_name.size();
-for ( int i = 1; i < argc; i++ ) {
-  string arg( argv[i] );
-  if ( arg.substr( 0, arg_name_size+1 ) == arg_name + "=" ) {
-    found = true;
-    try {
-      value = stod( arg.substr( arg_name_size+1 ) );
-    } catch ( invalid_argument ) {
-      fprintf( stderr, "Could not parse %s argument: %s", arg_name.c_str(), arg.c_str() );
-      exit(1);
-    }
-    break;
-  }
-}
-if ( !found && mandatory ) {
-  fprintf( stderr, "Please provide an argument for %s\n", arg_name.c_str() );
-  exit(1);
-}
-if ( found ) {
-  fprintf( stderr, "Setting %s to %f\n", arg_name.c_str(), value );
-  (range.*set_fn)(value);
-  return true;
-}
-fprintf( stderr, "No argument found for %s\n", arg_name.c_str() );
-return false;
-}
+// bool found = false;
+// double value = NAN;
+// int arg_name_size = arg_name.size();
+// for ( int i = 1; i < argc; i++ ) {
+//   string arg( argv[i] );
+//   if ( arg.substr( 0, arg_name_size+1 ) == arg_name + "=" ) {
+//     found = true;
+//     try {
+//       value = stod( arg.substr( arg_name_size+1 ) );
+//     } catch ( invalid_argument ) {
+//       fprintf( stderr, "Could not parse %s argument: %s", arg_name.c_str(), arg.c_str() );
+//       exit(1);
+//     }
+//     break;
[…]

### 2017-12-07 tex/neural_network.tex
Small changes in tex, not visible

@@
-\newcommand{\nnfontsize}{\tiny}
-
 \centering
-\begin{tikzpicture}[node distance=1.2cm, >=stealth',bend angle=45,auto]
+\begin{tikzpicture}[every label/.append style={font=\tiny}, node distance=1.2cm, >=stealth',bend angle=45,auto]
 
@@
   \begin{scope}
-    \node [input place] (s1) [label={[align=center]left:\nnfontsize State (12)}] {};
-    \node [place] (s2) [above right of=s1, label={[align=center]above:\nnfontsize Linear}] {};
+    \node [input place] (s1) [label={[align=center]left: State (12)}] {};
+    \node [place] (s2) [above right of=s1, label={[align=center]above: Linear}] {};
 
@@
 
-    \node [lstm place] (s3) [right of=s2, label={[align=center]below:\nnfontsize LSTM (32)}] {};
+    \node [lstm place] (s3) [right of=s2, label={[align=center]below: LSTM (32)}] {};
 
@@
 
-    \node [lstm place] (s4) [right of=s3, label={[align=center]below:\nnfontsize LSTM (32)}] {};
+    \node [lstm place] (s4) [right of=s3, label={[align=center]below: LSTM (32)}] {};
 
@@
 
-    \node [softplus place] (s7) [right of=s4, label={[align=center]below:{\nnfontsize Softplus}}] {};
+    \node [softplus place] (s7) [right of=s4, label={[align=center]below:{ Softplus}}] {};
 
@@
 
-    \node [output place] (s8) [right of=s7, label={[align=center]below:\nnfontsize $\sigma$~(1)}] {};
+    \node [output place] (s8) [right of=s7, label={[align=center]below: $\sigma$~(1)}] {};
 
@@
 
-    \node [place] (s5) [above of=s7, label={[align=center]above:\nnfontsize Linear}] {};
+    \node [place] (s5) [above of=s7, label={[align=center]above: Linear}] {};
 
@@
 
-    \node [output place] (s6) [right of=s5, label={[align=center]above:\nnfontsize $\mu$ (1)}] {};
+    \node [output place] (s6) [right of=s5, label={[align=center]above: $\mu$ (1)}] {};
 
@@
 
-    \node [place] (s10) [below right of=s1, label={[align=center]below:\nnfontsize Linear}] {};
+    \node [place] (s10) [below right of=s1, label={[align=center]below: Linear}] {};
 
@@
 
-    \node [lstm place] (s11) [right of=s10, label={[align=center]above:\nnfontsize LSTM (32)}] {};
+    \node [lstm place] (s11) [right of=s10, label={[align=center]above: LSTM (32)}] {};
 
@@
 
-    \node [lstm place] (s12) [right of=s11, label={[align=center]above:\nnfontsize LSTM (32)}] {};
+    \node [lstm place] (s12) [right of=s11, label={[align=center]above: LSTM (32)}] {};
 
@@
 
-    \node [softplus place] (s13) [right of=s12, label={[align=center]above:\nnfontsize Softplus}] {};
+    \node [softplus place] (s13) [right of=s12, label={[align=center]above: Softplus}] {};
 
@@
 
-    \node [output place] (s17) [right of=s13, label={[align=center]above:\nnfontsize $V_\text{packets}$ (1)}] {};
+    \node [output place] (s17) [right of=s13, label={[align=center]above: $V_\text{packets}$ (1)}] {};
 
@@
 
-    \node [softplus place] (s15) [below of=s13, label={[align=center]above:\nnfontsize Softplus}] {};
+    \node [softplus place] (s15) [below of=s13, label={[align=center]above: Softplus}] {};
 
@@
 
-    \node [output place] (s19) [right of=s15, label={[align=center]above:\nnfontsize $V_\text{delay}$ (1)}] {};
+    \node [output place] (s19) [right of=s15, label={[align=center]above: $V_\text{delay}$ (1)}] {};
 
@@
 
-    \node [softplus place] (s16) [below of=s15, label={[align=center]above:\nnfontsize Softplus}] {};
+    \node [softplus place] (s16) [below of=s15, label={[align=center]above: Softplus}] {};
 
@@
 
-    \node [output place] (s20) [right of=s16, label={[align=center]above:\nnfontsize $V_\text{lost}$ (1)}] {};
+    \node [output place] (s20) [right of=s16, label={[align=center]above: $V_\text{lost}$ (1)}] {};
 
@@
 
-    \node [softplus place] (s14) [below of=s16, label={[align=center]above:\nnfontsize Softplus}] {};
+    \node [softplus place] (s14) [below of=s16, label={[align=center]above: Softplus}] {};
 
@@
 
-    \node [output place] (s21) [right of=s14, label={[align=center]above:\nnfontsize $V_\text{duration}$ (1)}] {};
+    \node [output place] (s21) [right of=s14, label={[align=center]above: $V_\text{duration}$ (1)}] {};
 
@@
 
-%    \node [softplus place] (s17) [below of=s14, label={[align=center]above:\nnfontsize Softplus}] {};
+%    \node [softplus place] (s17) [below of=s14, label={[align=center]above: Softplus}] {};
 %    
@@
 %    
-%    \node [output place] (s22) [right of=s17, label={[align=center]above:\nnfontsize $V_\text{byte}$ (1)}] {};   
+%    \node [output place] (s22) [right of=s17, label={[align=center]above: $V_\text{byte}$ (1)}] {};   
 %    

### 2017-12-07 tex/petri_net.tex
Small changes in tex, not visible

@@
-\newcommand{\nodedistance}{2}
-\newcommand{\nd}[1]{\numexpr #1 * \nodedistance \relax}
-
 \centering
-\begin{tikzpicture}[node distance=\nodedistance,>=stealth',bend angle=45,auto]
+\begin{tikzpicture}[x=2cm,y=2cm,>=stealth',bend angle=45,auto]
 
@@
   \begin{scope}
-    \node [place,tokens=1] (s1) at (\nd{0}, \nd{1}) [label={[align=center]left:Receive\\ feedback}] {};
-    \node [place] (s2) at (\nd{1}, \nd{2}) [label={[align=center]above:Received\\ ACK}] {};
-    \node [place] (s3) at (\nd{2}, \nd{1}) [label={[align=center]right:Packets\\ in the\\ network}] {};
-    \node [place] (s4) at (\nd{1}, \nd{0}) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};
+    \node [place,tokens=1] (s1) at (0, 1) [label={[align=center]left:Receive\\ feedback}] {};
+    \node [place] (s2) at (1, 2) [label={[align=center]above:Received\\ ACK}] {};
+    \node [place] (s3) at (2, 1) [label={[align=center]right:Packets\\ in the\\ network}] {};
+    \node [place] (s4) at (1, 0) [label={[align=center]below:Open congestion\\ window (cwnd)}] {};
 
-    \node [transition] (e1) at (\nd{0}, \nd{0}) [label=left:Take action] {}
+    \node [transition] (e1) at (0, 0) [label=left:Take action] {}
       edge [post] node {$a$}(s4)
@@
 
-    \node [transition] (e2) at (\nd{2}, \nd{0}) [label={[align=center]right:Send\\ packet}] {}
+    \node [transition] (e2) at (2, 0) [label={[align=center]right:Send\\ packet}] {}
       edge [pre] node {1} (s4)
@@
 
-    \node [transition] (e3) at (\nd{2}, \nd{2}) [label={[align=center]right:Receive\\ ACK}] {}
+    \node [transition] (e3) at (2, 2) [label={[align=center]right:Receive\\ ACK}] {}
       edge [pre] node {1} (s3)
@@
 
-    \node [transition] (e4) at (\nd{1}, \nd{1}) [label={[align=center]left:Update\\open\\cwnd}] {}
+    \node [transition] (e4) at (1, 1) [label={[align=center]left:Update\\open\\cwnd}] {}
       edge [pre] node {1} (s2)
@@
 
-    \node [transition] (e5) at (\nd{0}, \nd{2}) [label={[align=center]left:Received\\ feedback}] {}
+    \node [transition] (e5) at (0, 2) [label={[align=center]left:Received\\ feedback}] {}
       edge [pre] node {1} (s2)

### 2017-12-12 tex/al_dente.tex
Revise slides

@@
 \item Congestion Window 1 in the beginning of a flow.
-\item Upon receiving an acknowledgement for a previously sent packet, increase the Congestion Window (cwnd): $$\text{cwnd} = \frac{1}{\text{cwnd}}$$
+\item Upon receiving an acknowledgement (ACK) for a previously sent packet, increase the Congestion Window (cwnd): $$\text{cwnd} = \frac{1}{\text{cwnd}}$$
 \item Recently also a bit more sophisticated $\rightarrow$ \textit{CUBIC} etc. 
@@
 
+\begin{frame}{Other Potential Solution}
+\begin{itemize}
+\item \textit{PCC: Re-architecting Congestion Control for Consistent High Performance (PCC)} by Dong et al.~(2013).
+\item Sender uses rate $r_1$ for some time and then $b_2$ for some time. 
+\item Which rate was better? 
+\item Use better rate and start experimenting again 
+\item Problem: cannot react quickly as it has to wait for a whole measurement period until it changes its action. 
+\end{itemize}
+\end{frame}
+
 \begin{frame}{Reinforcement Learning}
@@
 \begin{itemize}
-\item Let's say an action is increasing the Congestion Window
-\item We can calculate the reward when we get back all the ACKs of packets that were sent during an action
-\item This means we can't do anything until receiving each action's reward. But that takes at least one Round Trip Time! Doesn't work.
+\item An action is increasing the Congestion Window
+\item We can calculate the reward after receiving all ACKs
+\item That takes at least one Round Trip Time
+\item In the mean time other packets could have arrived
 \item Example: We increase the cwnd by 2 packets. So we can send two packets. To evaluate if this action was good we have to get the ACKs of these two packets, which happens after one RTT!
@@
 \begin{itemize}
-\item Let's say the window is 1.9. We increase the window by $0.3$ (action). This allows us to send two packets (partial actions).  
-\item We receive the ACK for the first packet (feedback). We update the state and perform a new action.
-\item We receive the second ACK. Again, we update the state and perform an action. However, because we got all partial rewards of the previous action, we can calculate the reward and update our agent.
+\item Example: the window is 1.9
+\item Increase the window by $0.3$ (action)
+\item Send two packets (partial actions).  
+\item Receive the ACK for the first packet (feedback). Update the state and perform a new action.
+\item Receive the second ACK. Again, we update the state and perform an action. However, because we got all partial rewards of the previous action, we can calculate the reward and update our agent.
 \end{itemize}
@@
 
-\begin{frame}{Reward 2}
-\begin{align*}
-R_\text{packets}\text{Sigmoid}_\alpha\left(\frac{R_\text{lost}}{R_\text{lost} + R_\text{packets}} - 0.05\right) - R_\text{lost} - \frac{R_\text{delay}}{R_\text{packets}}
-\end{align*}
-\end{frame}
-
 \end{document}

### 2017-12-13 async_deep_reinforce/a3c_training_thread.py
New config files

@@ -356,7 +356,7 @@ class A3CTrainingThread(object):
       # td = inverse_sigmoid(self.delay_delta, R_sent/(R_packets+R_sent) - 0.05)*R_packets/R_duration - inverse_sigmoid(self.delay_delta, Vi[3]/(Vi[0]+Vi[3]) - 0.05)*Vi[0]/Vi[2] - (R_sent/R_duration - Vi[3]/(Vi[2])) - (R_accumulated_delay/R_packets - Vi[1]/Vi[0])
 
       # PCC
-      td = inverse_sigmoid(((R_sent - R_packets)/R_sent) - self.delay_delta)*R_packets/R_duration - inverse_sigmoid(((Vi[3]-Vi[0])/Vi[3]) - self.delay_delta)*Vi[0]/Vi[2] - ((R_sent - R_packets)/R_duration - (Vi[3] - Vi[0])/Vi[2]) #- (R_accumulated_delay/R_packets - Vi[1]/Vi[0])
+      # td = inverse_sigmoid(((R_sent - R_packets)/R_sent) - self.delay_delta)*R_packets/R_duration - inverse_sigmoid(((Vi[3]-Vi[0])/Vi[3]) - self.delay_delta)*Vi[0]/Vi[2] - ((R_sent - R_packets)/R_duration - (Vi[3] - Vi[0])/Vi[2]) #- (R_accumulated_delay/R_packets - Vi[1]/Vi[0])
 
       # td = R_packets*(1-GAMMA)*inverse_sigmoid(SIGMOID_ALPHA * R_sent/(R_packets+R_sent) - self.delay_delta) - Vi[0]*inverse_sigmoid(SIGMOID_ALPHA * Vi[3]/(Vi[0]+Vi[3]) - self.delay_delta) - (R_sent*(1-GAMMA) - Vi[3]) - (R_accumulated_delay/R_packets - Vi[1]/Vi[0])
 
@@ -365,7 +365,9 @@ class A3CTrainingThread(object):
 
       # td = R_packets/R_duration/(R_accumulated_delay/R_packets) - Vi[0]/Vi[2]/(Vi[1]/Vi[0]) - (R_accumulated_delay/R_packets - Vi[1]/Vi[0])
 
-      # td = (np.log(R_packets/R_duration) - np.log(Vi[0]/Vi[2])) - self.delay_delta/SECONDS_NORMALIZER*(R_accumulated_delay/R_packets - Vi[1]/Vi[0])
+
[…]

### 2017-12-13 async_deep_reinforce/constants.py

@@ -27,7 +27,7 @@ else:
 logging.info(" ".join(map(str,("CHECKPOINT_DIR:",CHECKPOINT_DIR))))
 
 LOG_FILE = ABSOLUTE_PATH+'tmp/a3c_log'
-ACTOR_FACTOR = 1e-1
+ACTOR_FACTOR = 1e0
 VALUE_FACTOR = 1e0
 GENERAL_FACTOR = 1e-4
 # INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate

### 2017-12-13 tex/al_dente.tex

@@
 
+%\begin{frame}{Asynchronous Actor Critic}
+%\begin{itemize}
+%\item A Deep Learning framework for reinforcement learning. Used for learning how to play video games.
+%\item Maximize long term reward: \begin{align*}
+%R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta_\text{v})\right),
+%\end{align*}
+%\item The \textbf{Critic} tries to estimate how much (long-term) reward one can expect considering the current state.
+%\end{itemize}
+%\end{frame}
+
 \begin{frame}{Asynchronous Actor Critic}
@@
 \item A Deep Learning framework for reinforcement learning. Used for learning how to play video games.
-\item Maximize long term reward: \begin{align*}
-R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta_\text{v})\right),
+\item Maximize expected long term reward: \begin{align*}
+R_t = \gamma r_t + (1-\gamma) R_{t+1}
 \end{align*}
@@
 
+\makeatletter
+\begin{frame}{Exponential decay with $\gamma = 0.01$}
+    \global\beamer@shrinktrue
+    \gdef\beamer@shrinkframebox{
+        \setbox\beamer@framebox=\vbox to\beamer@frametextheight{
+            \centering
+            \includegraphics[height=\beamer@frametextheight]{figures/exponential}
+        }
+    }
+\end{frame}
+\makeatother
+
+\begin{frame}{Asynchronous Actor Critic -- How to choose $\gamma$}
+\begin{itemize}
+\item In literature people just set $\gamma = 0.01$. 
+\item Problem: The larger the congestion window the more often we'll receive new packets!
+\item $\frac{2}{k+1}$ is equivalent to a moving average considering the next $k$ elements.
+\item Choose $\frac{2}{w_t+2}$ where $w_t$ is the congestion window at time step $t$.
+\end{itemize}
+\end{frame}
+
 \begin{frame}{Asynchronous Actor Critic -- Actor}
 \begin{itemize}
-\item The \textbf{Actor} tries to perform an action that is better than what the Critic would expect.
+\item Wants to performs an action that are better than what the Critic would expect.
 \item It outputs two things:
@@
 \end{itemize}
-\item So the actor outputs a normal distribution from which actions are sampled. Thanks to the standard deviation we experiment and don't get stuck with suboptimal actions. 
+\item So we get a normal distribution from which we sample.
+\item Thanks to the standard deviation we experiment and don't get stuck with suboptimal actions. 
 \end{itemize}
-\end{frame}
-
-\begin{frame}{Preliminary results -- Experiment characteristics}
-\centering
-\begin{tabular}{rll}
-\toprule
-Parameter & Value & Distribution \\
-\midrule
-Two-way RTT when queue is 0 & $150\,$ms & constant \\
-Bottleneck bandwidth & $15\,$Mbit/s & constant \\
-Number of senders & $8$ & constant \\
-Flow length & $100\,$kB & exponential \\
-Time between flows & $0.5\,$s & exponential \\
-Simulation duration & $500\,$s & constant \\
-Buffer size & $100$ packets & constant \\
-Stochastic loss prob. & $0\%$ & constant \\
-\bottomrule
-\end{tabular}
-\end{frame}
-
-\begin{frame}{Preliminary results -- Comparison}
-Stuff goes here!
-
-\end{frame}
-
-\begin{frame}{Petri net}
-
-\input{petri_net.tex}
-
 \end{frame}
@@
 
-\begin{frame}{Reward 1}
+\makeatletter
+\begin{frame}{Softplus function}
+    \global\beamer@shrinktrue
+    \gdef\beamer@shrinkframebox{
+        \setbox\beamer@framebox=\vbox to\beamer@frametextheight{
+            \centering
+            \includegraphics[height=\beamer@frametextheight]{figures/softplus}
+        }
+    }
+\end{frame}
+\makeatother
+
+\begin{frame}{Utility function}
 \begin{align*}
-\frac{R_\text{packets}}{R_\text{duration}}\text{Sigmoid}_\alpha\left(\frac{R_\text{lost}}{R_\text{lost} + R_\text{packets}} - 0.05\right) - \frac{R_\text{lost}}{R_\text{duration}} - \frac{R_\text{delay}}{R_\text{packets}}
+U_t = \text{throughput} \cdot \text{Sigmoid}_\alpha\left(\text{loss rate} - 0.05 \right) - \text{sending rate}\cdot\text{loss rate}
 \end{align*}
@@
 
+\begin{frame}{Utility function}
+\begin{align*}
+U_t = \frac{r_{\text{received},t}}{r_{\text{duration},t}}\text{Sigmoid}_\alpha\left(\frac{r_{\text{sent},t} - r_{\text{received},t}}{r_{\text{sent},t}} - 0.05\right) - \frac{r_{\text{sent},t} - r_{\text{received},t}}{r_{\text{duration},t}}
+\end{align*}
+\end{frame}
+
+\begin{frame}{Sigmoid function}
+\begin{align*}
+\text{Sigmoid}_\alpha(x) = \frac{1}{1+e^{\alpha (x-0.05)}}
+\end{align*}
+\end{frame}
+
+\makeatletter
+\begin{frame}{Sigmoid function for $\alpha = 100$ with $0.05$ cutoff}
+    \global\beamer@shrinktrue
+    \gdef\beamer@shrinkframebox{
+        \setbox\beamer@framebox=\vbox to\beamer@frametextheight{
+            \centering
+            \includegraphics[height=\beamer@frametextheight]{figures/sigmoid}
+        }
+    }
+\end{frame}
+\makeatother
+
+\begin{frame}{Petri net}
+
+\input{petri_net.tex}
+
+\end{frame}
+
+%\begin{frame}{Preliminary results -- Experiment characteristics}
+%\centering
+%\begin{tabular}{rll}
+%\toprule
+%Parameter & Value & Distribution \\
+%\midrule
+%Two-way RTT when queue is 0 & $150\,$ms & constant \\
+%Bottleneck bandwidth & $15\,$Mbit/s & constant \\
+%Number of senders & $8$ & constant \\
+%Flow length & $100\,$kB & exponential \\
+%Time between flows & $0.5\,$s & exponential \\
+%Simulation duration & $500\,$s & constant \\
+%Buffer size & $100$ packets & constant \\
+%Stochastic loss prob. & $0\%$ & constant \\
+%\bottomrule
+%\end{tabular}
+%\end{frame}
+
+\begin{frame}{Some data}
+\includegraphics[width=\textwidth]{figures/window}
+\end{frame}
+
+\begin{frame}{Some data}
+\includegraphics[width=\textwidth]{figures/score_lost}
+\end{frame}
+
+\begin{frame}{Some data}
+\includegraphics[width=\textwidth]{figures/window_increase}
+\end{frame}
+
 \end{document}

### 2018-01-12 src/unicorn.cc
Changes

@@ -248,11 +248,11 @@ void Unicorn::get_action(const double& tickno) {
   const double action = _rainbow.get_action(
     _thread_id,
     {
-      // // // _memory._send,
+      _memory._send,
       _memory._rec_send_ewma,
       _memory._slow_rec_send_ewma,
 
-      // // // _memory._rec,
+      _memory._rec,
       _memory._rec_rec_ewma,
       _memory._slow_rec_rec_ewma,
 
@@ -260,15 +260,15 @@ void Unicorn::get_action(const double& tickno) {
       // _memory._rec_rec_ewma > 0 ? 1.0/_memory._rec_rec_ewma : _memory._rec_rec_ewma,
       // _memory._slow_rec_rec_ewma > 0 ? 1.0/_memory._slow_rec_rec_ewma : _memory._slow_rec_rec_ewma,
 
-      // _memory._rtt,
+      _memory._rtt,
       _memory._rtt_ewma,
       _memory._slow_rtt_ewma,
 
-      // _memory._loss,
+      _memory._loss,
       _memory._loss_ewma,
       _memory._slow_loss_ewma,
 
-      // _memory._window,
+      _memory._window,
       _memory._window_ewma,
       _memory._slow_window_ewma,
     },

### 2018-01-12 async_deep_reinforce/a3c_training_thread.py
Changes

@@ -13,8 +13,6 @@ from functools import reduce
 
 from game_ac_network import GameACLSTMNetwork
 
-from constants import GAMMA
-from constants import GAMMA_FACTOR
 # gamma_current, gamma_future = 1./(1+GAMMA), GAMMA/(1.+GAMMA)
 # from constants import LOCAL_T_MAX
 from constants import LOG_LEVEL
@@ -337,7 +335,7 @@ class A3CTrainingThread(object):
       # assert(False)
       # The GAMMA_FACTOR increases the influence that following observations have on this one.
 
-      GAMMA = (1 - 2/(GAMMA_FACTOR*A3CTrainingThread.get_actual_window(wi+ai) + 1))
+      GAMMA = (1 - 2/(A3CTrainingThread.get_actual_window(wi+ai) + 1))
 
       # R_duration = ((1-GAMMA)*ri[2] + GAMMA*R_duration)
       # R_packets = ((1-GAMMA)*ri[0] + GAMMA*R_packets)

### 2018-01-12 async_deep_reinforce/constants.py

@@ -20,9 +20,8 @@ RMSP_EPSILON = 0.1 # epsilon parameter for RMSProp
 # FIXME: Super ugly to hardcode the path!!!
 # ABSOLUTE_PATH = os.path.join(os.path.expanduser('~'),"repos/remy/")
 ABSOLUTE_PATH = os.path.join(os.getcwd())
-from os import environ
-if environ.get('checkpoints') is not None:
-	CHECKPOINT_DIR = os.path.join(ABSOLUTE_PATH,environ.get('checkpoints'))
+if os.environ.get('checkpoints') is not None:
+	CHECKPOINT_DIR = os.path.join(ABSOLUTE_PATH,os.environ.get('checkpoints'))
 else:
 	CHECKPOINT_DIR = os.path.join(ABSOLUTE_PATH,'checkpoints')
 if not os.path.exists(CHECKPOINT_DIR):
@@ -30,7 +29,7 @@ if not os.path.exists(CHECKPOINT_DIR):
 
 logging.info(" ".join(map(str,("CHECKPOINT_DIR:",CHECKPOINT_DIR))))
 
-LOG_FILE = os.path.join(ABSOLUTE_PATH,'tmp/a3c_log')
+LOG_FILE = os.path.join(ABSOLUTE_PATH,'a3c_log')
 logging.info(" ".join(map(str,("LOG_DIR:",LOG_FILE))))
 ACTOR_FACTOR = 1e0
 VALUE_FACTOR = 1e0
@@ -42,8 +41,6 @@ INITIAL_RATE = GENERAL_FACTOR
 PRECISION = tf.float32
 
 INITIAL_ALPHA_LOG_RATE = 0.4226 # log_uniform interpolate rate for learning rate (around 7 * 10^-4)
-GAMMA = 0.99 # discount factor for rewards
-GAMMA_FACTOR = 1
 ENTROPY_BETA = 1e-4
 # STD_BIAS_OFFSET = 0.3
 STD_BIAS_OFFSET = inverse_softplus(0.5)
@@ -53,10 +50,10 @@ STD_BIAS_MINIMUM = 0.0
 MAX_TIME_STEP = 1e8
 # GRAD_NORM_CLIP = 40.0 # gradient norm clipping
 USE_GPU = False # To use GPU, set True
-N_LSTM_LAYERS = int(environ.get('layers')) if environ.get('layers') is not None else 2
[…]

### 2018-01-12 tex/paper.tex
Changes to the paper
Tex

@@
 \begin{abstract}
-This paper provides a sample of a \LaTeX\ document which conforms,
-somewhat loosely, to the formatting guidelines for
-ACM SIG Proceedings.\footnote{This is an abstract footnote}
+This paper proposes a new method of congestion control in packet-switched networks that uses on-line reinforcement learning to determine the optimum congestion window size at a given point in time with respect to a chosen utility function.
+
+To this end we use an Artifical Neural Network based approach that can be initialized with a previously trained neural network or with random weights (without any preknowledge).
+When being used without any preknowledge the approach converges to a stable solution within the order of minutes and then outperforms existing congestion control algorithms with respect to throughput.
+
+To our knowledge (\todo{Make sure this is really true!})
+
 \end{abstract}
@@
 Ocher stands for the input, blue for linear layers, red for LSTM layers, green for linear layers with a softplus activation function and purple for the outputs of the neural network.
-The numbers in parentheses next to each label stand  for the number of neurons being used in that layer.}
+The numbers in parentheses next to each label stand for the number of width of that layer (i.e.~the number of neurons except for the inputs and outputs).}
 \label{fig:neural_network}

### 2018-01-13 tex/paper.tex
Paper

@@
 %\usepackage{todonotes}
+\usepackage{graphicx}
+\graphicspath{figures}
 \usepackage{svg}
@@
 
-To this end we use an Artifical Neural Network based approach that can be initialized with a previously trained neural network or with random weights (without any preknowledge).
+To this end we use an Artificial Neural Network based approach that can be initialized with a previously trained neural network or with random weights (without any preknowledge).
 When being used without any preknowledge the approach converges to a stable solution within the order of minutes and then outperforms existing congestion control algorithms with respect to throughput.
+
+As in case of congestion control feedback used for reinforcement learning is naturally always delayed (due to the round trip time), we propose a novel approach to reinforcement learning that can handle delayed rewards as well as rewards that are returned in parts one after another.
 

### 2018-01-14 tex/paper.tex
Added related work

--- 2017-12-02 tex/paper.tex
@@
 \begin{abstract}
-This paper provides a sample of a \LaTeX\ document which conforms,
-somewhat loosely, to the formatting guidelines for
-ACM SIG Proceedings.\footnote{This is an abstract footnote}
+This paper proposes a novel method of Congestion Control that uses on-line Reinforcement Learning (RL) to maintain the optimum congestion window with respect to a given utility function.
+
+To this end we use an Artificial Neural Network (ANN) based approach that can be initialized with a previously trained neural network or with random weights (without any prior knowledge).
+As in congestion control the rewards used in the learning process naturally come delayed and bit by bit, we propose a novel formulation of reinforcement learning that allows us to deal with delayed and partial rewards.
+
+We show that our method converges to a stable, close-to-optimum solution within the order of minutes and then commonly outperforms existing congestion control algorithms in typical networks.
+Thus, for the first time, we demonstrate that ANN based Reinforcement Learning without any preknowledge can feasibly be done on-line and can compete with hand-crafted solutions given long enough episodes.
+
 \end{abstract}
@@
 <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
+<concept_id>10003033.10003039.10003048</concept_id>
+<concept_desc>Networks~Transport protocols</concept_desc>
 <concept_significance>500</concept_significance>
@@
 <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
+<concept_id>10003752.10010070.10010071.10010261.10010275</concept_id>
+<concept_desc>Theory of computation~Multi-agent reinforcement learning</concept_desc>
+<concept_significance>500</concept_significance>
 </concept>
 <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
+<concept_id>10010520.10010521.10010542.10010294</concept_id>
+<concept_desc>Computer systems organization~Neural networks</concept_desc>
+<concept_significance>300</concept_significance>
 </concept>
@@
 
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
+\ccsdesc[500]{Networks~Transport protocols}
+\ccsdesc[500]{Theory of computation~Multi-agent reinforcement learning}
+\ccsdesc[300]{Computer systems organization~Neural networks}
 
-\keywords{Put some comma-separated keywords here}
+\keywords{congestion control, machine learning, reinforcement learning, artificial neural networks}
 
@@
 
-In recent years, there have been several approaches to make congestion control learnable, most notably \textit{Remy} \citep{winstein_tcp_2013}, which learns an optimum congestion control w.r.t.~a specific utility function for specified network scenarios in an off-line fashion.
-Furthermore there is \textit{Performance-oriented Congestion Control} (PCC) \citep{dong_pcc:_2015}, which repeatedly tries different rates during periods of approximately one round-trip time (RTT) and chooses the better rate w.r.t.~a specific utility function.
+Recent advances in Artificial Neural Network (ANN) based Reinforcement Learning (RL) for the purpose of playing video games raise the question if it is also possible to formulate sending Internet data in a fashion similar to a video game.
+The rules of the ``Game of Congestion Control'' can be summarized approximately as follows: Each player wants to send \begin{enumerate*}
+\item as many packets as possible
+\item with as little delay as possible
+\item while losing as few packets as possible
+\end{enumerate*}.
+As, for instance, for some applications overall throughput is more important than little packet loss, each application can define its preference using a custom utility function.
+Clearly, it is also desirable from an overall point of view that all participating players play the game in a way that does not give them an unfair advantage over others.
+When several players (from now on called senders) share an Internet link under the same conditions, they should receive an equal share of the reward.
 
-We make use of a new Reinforcement Learning (RL) based approach that can learn a congestion control algorithm on-line and is able to perform actions on a per acknowledgement basis.
+When performing Congestion Control (CG) \todo{What should be capitalized and what shouldn't?
+Capitalize RL, CG, ANN?}, each sender uses a set of observed environment conditions to determine what action he should take to maximize his reward in the future.
+The environment conditions can be any metrics that the sender can obtain from the network, for example the mean round-trip time of the last received packets, the packet loss rate etc.
+An action is a change to the sending rate: For example, if a sender perceives an increase in the loss rate it might be advisable to lower the sending rate.
+
+From this formulation of Congestion Control, one can observe a couple of unique features of the problem that seem to prohibit the use of RL: \begin{itemize}
+\item When playing a video game, if one pushes a button, one can see the consequences of an action immediately.
+However, in Congestion Control, the rewards are always \textbf{delayed} by one round-trip time.
+\item Due to the delay, by the time an action receives its reward, probably other actions have already been carried out in the mean time.
+Thus, contrary to classical RL, \textbf{action and rewards are out of sync}.
+\item \textbf{Actions and rewards are not atomic}: If an action causes three packets to be sent, we consider these three packets being sent three \textit{partial actions}.
+In case these packets are transmitted correctly, the receiver will send three acknowledgements.
+Each of these acknowledgements is a \textit{partial reward} that already unveils new information on the current state of the network and enables the sender to take an action based on this new information.
+However, only if all partial rewards have been received, the sender can assemble the overall reward of the corresponding action and can perform an update of the underlying RL logic.
+\end{itemize}
+
+To address these problems, we develop a new formulation of RL called \textit{Partial Action Learning} (PAL).
+PAL is a superset of Reinforcement Learning: If one uses PAL for a learning problem without delay, asynchronicity and partial actions/partial rewards, one gets classical RL.
+
+While using PAL to train an optimum congestion control for a specific range of network scenarios in an off-line fashion is possible, it is something that has already been done previously in an approach called \textit{Remy} \citep{winstein_tcp_2013}.
+The only seizable advantage that PAL could provide here is increased training speed.
+Thus we want to show that it is  not only possible to learn congestion control by using machine learning but that it is even possible to do so on-line and without any preknowledge about the network environment.
+
+Using a utility function that encourages high throughput and discourages packet loss (adapted from \citep{dong_pcc:_2015}) we see that our machine learning based approach can learn congestion control that maximizes this objective given a couple of minutes of time (without any pre-training) \ref{fig:demo}, which -- to our knowledge -- is the first demonstration that neural network based on-line reinforcement learning is possible.
+
+\begin{figure}
+\begin{minipage}{\columnwidth}
+\includegraphics[width=\columnwidth]{figures/1_4}
+\includegraphics[width=\columnwidth]{figures/2_2}
+\includegraphics[width=\columnwidth]{figures/4_1}
+\caption{The throughput (solid line) and the loss rate (dashed line) when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1, 2 and 4 senders (from top to bottom).
+One can see that the time to convergence increases as the number of senders increases, which is caused by the fact that the more senders use the link, the fewer packets each sender receives per unit of time for training.
+Also, the complexity of the learning problem increases with more senders as the environment becomes more dynamic and unpredictable.}
+\label{fig:demo}
+\end{minipage}
+\end{figure}
+
+After having learned how to perform congestion control in a certain network scenario, our approach can compete with established algorithms and achieve superior throughput with packet loss at an average of less than 5\%, while still gradually improving and adapting its way of performing congestion control.
+
+\section{Related Work}
+
+\subsection{Congestion Control}
+
+Congestion control has been implemented in TCP since the 1980s as a series of congestion collapses dramatically decreased applications' throughput in the Internet, which lead to the development of the Tahoe and Reno algorithms \cite{jacobson_congestion_1988}.
+These approaches as well as most others maintain a congestion window which stands for the maximum amount of data that are allowed to be traveling in the network without having been acknowledged by the receiver.
+In general, successful transmission and acknowledgement of data leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995} \todo{How much detail is necessary in the related work section?}) lead to the congestion window being lowered according to a set of fixed rules.
+Algorithms such as Cubic and Compound \cite{ha_cubic:_2008, tan_compound_2006} improve congestion control specifically with respect to so-called long fat networks with a high round-trip time and bandwidth.
+Proportional rate reduction \cite{dukkipati_proportional_2011} aims to make the reduction of the rate that occurs in case of congestion more smooth and steady in time to avoid the bursty behavior that occurred upon loss in previous TCP congestion control variants.
+On the other hand, BBR \cite{cardwell_bbr:_2016} does not use a set of fixed rules such as previous congestion control algorithms but instead estimates a model of the network path by using measurements of RTT and throughput and then adjusts its sending rate so that it uses the maximum bandwidth according to its network model while trying not to fill up queues.
+A similar approach is PCC \cite{dong_pcc:_2015}, which also uses measurements to find an optimum sending rate with respect to a defined utility function (e.g.
+maximizing throughput while minimizing packet loss).
+
+Besides the aforementioned human-designed algorithms there have been attempts to use machine learning to improve congestion control.
+\citet{geurts_machine_2004} train a classifier off-line to determine whether packet loss is caused by congestion or by the link layer.
+If it is caused by the link layer, TCP does not alter the window.
+Otherwise -- if a packet was lost due to congestion according to the classifier -- it uses a traditional TCP congestion control algorithm.
+\citet{winstein_tcp_2013} train a machine learning solution called \textit{Remy} that finds an optimum congestion control for a given range of a range of network parameters.
+For instance, one could find an optimum congestion control for networks with a RTT of 50-100\,ms and link speeds of 100-500\,Mbit/s.
+After a lengthy training procedure Remy finds an optimum congestion control algorithm for the specified networks on a per acknowledgement basis.
+
+Our goal is to find an optimum congestion control algorithm on a per acknowledgement basis similar to Remy that, however, can be trained on-line.
+Such a solution could be used in a purely on-line fashion, in a off-line fashion like Remy or a combination of both: One pre-trains a generic congestion control algorithm that works reasonably well for every network that then gets refined during on-line training according to the current network circumstances.
+
+To this end we use Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn a wide range of video games and commonly outperform human players.
+In particular it is a good choice for congestion control as it can be easily adapted to a wide range of problems and has been proven to deliver good performance.
+Furthermore, as we will show, it is conceptually possible to use it for on-line training although -- to our knowledge -- this has not been done until now.
 
@@
 
-To this end we propose \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) which is a modification of the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline in this section.
+The \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) framework is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline in this section.
 
@@
 So an action is considered good if it achieved a long-term reward that is higher than the long-term reward expected by the Value Network and it is considered bad if the reward was lower than expected.
-The long-term reward is implemented as a moving average of future rewards.
+The long-term reward is implemented as an exponentially weighted moving average of future rewards.
 So if a high reward can be achieved right now this is more favorable than if it can be achieved in the future.
@@
 \begin{minipage}{\columnwidth}
-\includesvg{figures/Reinforcement_learning_diagram}
+\includesvg{Reinforcement_learning_diagram}
 \caption{The classical Reinforcement Learning approach.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
@@
 \begin{minipage}{\columnwidth}
-\includesvg{figures/Reinforcement_learning_diagram_PAL}
+\includesvg{Reinforcement_learning_diagram_PAL}
 \caption{Partial Action Learning: An action consists of zero or more partial actions which trigger feedback upon interacting with the environment.

### 2018-01-14 tex/neural_network.tex
Al Dente finished

@@
 
-    \node [lstm place] (s3) [right of=s2, label={[align=center]below: LSTM (32)}] {};
+    \node [lstm place] (s3) [right of=s2, label={[align=center]below: GRU (32)}] {};
 
@@
 
-    \node [lstm place] (s4) [right of=s3, label={[align=center]below: LSTM (32)}] {};
+    \node [lstm place] (s4) [right of=s3, label={[align=center]below: GRU (32)}] {};
 
@@
 
-    \node [softplus place] (s7) [right of=s4, label={[align=center]below:{ Softplus}}] {};
+    \node [place] (s7) [right of=s4, label={[align=center]below:{ Linear}}] {};
 
@@
 
-    \node [lstm place] (s11) [right of=s10, label={[align=center]above: LSTM (32)}] {};
+    \node [lstm place] (s11) [right of=s10, label={[align=center]above: GRU (32)}] {};
 
@@
 
-    \node [lstm place] (s12) [right of=s11, label={[align=center]above: LSTM (32)}] {};
+    \node [lstm place] (s12) [right of=s11, label={[align=center]above: GRU (32)}] {};
 
@@
 
-    \node [softplus place] (s13) [right of=s12, label={[align=center]above: Softplus}] {};
+    \node [place] (s13) [right of=s12, label={[align=center]above: Linear}] {};
 
@@
 
-    \node [softplus place] (s15) [below of=s13, label={[align=center]above: Softplus}] {};
+    \node [place] (s15) [below of=s13, label={[align=center]above: Linear}] {};
 
@@
 
-    \node [softplus place] (s16) [below of=s15, label={[align=center]above: Softplus}] {};
+    \node [place] (s16) [below of=s15, label={[align=center]above: Linear}] {};
 
@@
 
-    \node [output place] (s20) [right of=s16, label={[align=center]above: $V_\text{lost}$ (1)}] {};
+    \node [output place] (s20) [right of=s16, label={[align=center]above: $V_\text{sent}$ (1)}] {};
 
@@
 
-    \node [softplus place] (s14) [below of=s16, label={[align=center]above: Softplus}] {};
+    \node [place] (s14) [below of=s16, label={[align=center]above: Linear}] {};
 
@@
 
-%    \node [softplus place] (s17) [below of=s14, label={[align=center]above: Softplus}] {};
+%    \node [place] (s17) [below of=s14, label={[align=center]above: Linear}] {};
 %    

### 2018-01-14 tex/paper.tex

--- 2018-01-12 tex/paper.tex
@@
 \begin{abstract}
-This paper proposes a new method of congestion control in packet-switched networks that uses on-line reinforcement learning to determine the optimum congestion window size at a given point in time with respect to a chosen utility function.
+This paper proposes a novel method of Congestion Control that uses on-line Reinforcement Learning (RL) to maintain the optimum congestion window with respect to a given utility function.
 
-To this end we use an Artifical Neural Network based approach that can be initialized with a previously trained neural network or with random weights (without any preknowledge).
-When being used without any preknowledge the approach converges to a stable solution within the order of minutes and then outperforms existing congestion control algorithms with respect to throughput.
+To this end we use an Artificial Neural Network (ANN) based approach that can be initialized with a previously trained neural network or with random weights (without any prior knowledge).
+As in congestion control the rewards used in the learning process naturally come delayed and bit by bit, we propose a novel formulation of reinforcement learning that allows us to deal with delayed and partial rewards.
 
-To our knowledge (\todo{Make sure this is really true!})
+We show that our method converges to a stable, close-to-optimum solution within the order of minutes and then commonly outperforms existing congestion control algorithms in typical networks.
+Thus, for the first time, we demonstrate that ANN based Reinforcement Learning without any preknowledge can feasibly be done on-line and can compete with hand-crafted solutions given long enough episodes.
 
@@
 
-In recent years, there have been several approaches to make congestion control learnable, most notably \textit{Remy} \citep{winstein_tcp_2013}, which learns an optimum congestion control w.r.t.~a specific utility function for specified network scenarios in an off-line fashion.
-Furthermore there is \textit{Performance-oriented Congestion Control} (PCC) \citep{dong_pcc:_2015}, which repeatedly tries different rates during periods of approximately one round-trip time (RTT) and chooses the better rate w.r.t.~a specific utility function.
+Recent advances in Artificial Neural Network (ANN) based Reinforcement Learning for the purpose of playing video games raise the question if it is also possible to formulate sending Internet data in a fashion similar to a video game.
+The rules of the ``Game of Congestion Control'' can be summarized approximately as follows: Each player wants to send \begin{enumerate*}
+\item as many packets as possible
+\item with as little delay as possible
+\item while losing as few packets as possible
+\end{enumerate*}.
+As, for instance, for some applications overall throughput is more important than little packet loss, each application can define preferences using a custom utility function.
+Clearly, it is also desirable from an overall point of view that all participating players play the game in a way that does not give them an unfair advantage over others.
+When several players (from now on called senders) share an Internet link under the same conditions, they should receive an equal share of the reward.
 
-We make use of a new Reinforcement Learning (RL) based approach that can learn a congestion control algorithm on-line and is able to perform actions on a per acknowledgement basis.
+When performing Congestion Control (CG), each sender uses a set of observed environment conditions to determine what action he should take to maximize his reward in the future.
+The environment conditions can be any metrics that the sender can obtain from the network, for example the mean round-trip time of the last received packets, the packet loss rate etc.
+An action is a change to the sending rate: For example, if a sender perceives an increase in the loss rate it might be advisable to lower the sending rate.
+
+From this formulation of Congestion Control, one can observe a couple of unique features of the problem that seem to prohibit the use of Reinforcement Learning: \begin{itemize}
+\item When playing a video game, if one pushes a button, one can see the consequences of an action immediately.
+However, in Congestion Control, the rewards are always \textbf{delayed} by one round-trip time.
+\item Due to the delay, by the time an action receives its reward, probably other actions have already been carried out in the mean time.
+Thus, contrary to classical RL, \textbf{action and rewards are out of sync}.
+\item \textbf{Actions and rewards are not atomic}: If an action causes three packets to be sent, we consider these three packets being sent three \textit{partial actions}.
+In case these packets are transmitted correctly, the receiver will send three acknowledgements.
+Each of these acknowledgements is a \textit{partial reward}, that already unveils new information on the current state of the network and enables the sender to take an action based on this new information.
+However, only if all partial rewards have been received, the sender can assemble the overall reward of the corresponding action and can perform an update of the underlying Reinforcement Learning logic.
+\end{itemize}
+
+To address these problems, we develop a new formulation of RL called \textit{Partial Action Learning} (PAL).
+PAL is a superset of Reinforcement Learning: If one uses PAL for a learning problem without delay, asynchronicity and partial actions/partial rewards, one gets classical RL.
+
+While using PAL to train an optimum congestion control for a specific range of network scenarios in an off-line fashion is possible, it is something that has already been done previously in an approach called \textit{Remy} \citep{winstein_tcp_2013}.
+The only seizable advantage that PAL could provide here is increased speed.
+Thus we want to show that it is  not only possible to learn congestion control by using machine learning but that it is even possible on-line and without any preknowledge about the network environment.
+
+\todo{Figures here!}
+
+Using a utility function that encourages high throughput and discourages packet loss (adapted from \citep{dong_pcc:_2015}) we see that our machine learning based approach can learn congestion control that maximizes this objective given a couple of minutes of time (without any pre-training).
+
+After this initial exploration, our approach can compete with established algorithms and achieve superior throughput while keeping packet loss at an average of less than 1\%.
 

### 2018-01-15 src/unicorn-templates.cc
Fixed some timeout problems

@@ -18,7 +18,7 @@ void Unicorn::send( const unsigned int id, NextHop & next, const double & tickno
   if (_last_send_time > 0 && tickno - _last_send_time >= TIMEOUT_THRESHOLD) {
     printf("%lu: timeout occurred!\n", _thread_id);
     reset(tickno);
-    return;
+    // return;
   }
 
   // printf("left:%d, right:%d\n",int(_packets_sent),  (int) _largest_ack + 1 + (int) floor(_the_window));

### 2018-01-15 async_deep_reinforce/constants.py
Hopefully really make timeout work

@@ -33,7 +33,7 @@ LOG_FILE = os.path.join(ABSOLUTE_PATH,'a3c_log')
 logging.info(" ".join(map(str,("LOG_DIR:",LOG_FILE))))
 ACTOR_FACTOR = 1e0
 VALUE_FACTOR = 1e0
-GENERAL_FACTOR = 10**-2.5
+GENERAL_FACTOR = 10**-3
 # INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate
 # INITIAL_ALPHA_HIGH = 1e0*GENERAL_FACTOR   # log_uniform high limit for learning rate
 INITIAL_RATE = GENERAL_FACTOR

### 2018-01-15 src/unicorn-templates.cc

@@ -16,7 +16,7 @@ void Unicorn::send( const unsigned int id, NextHop & next, const double & tickno
   }
 
   if (_last_send_time > 0 && tickno - _last_send_time >= TIMEOUT_THRESHOLD) {
-    printf("%lu: timeout occurred!\n", _thread_id);
+    printf("%lu: timeout occurred! tickno: %f, _last_send_time: %f\n", _thread_id, tickno, _last_send_time);
     reset(tickno);
     // return;
   }

### 2018-01-15 src/unicorn.cc

@@ -235,6 +235,7 @@ double Unicorn::next_event_time( const double & tickno ) const
     /* window is currently closed */
     // return std::numeric_limits<double>::max();
     // FIXME: Is that necessary?
+
     if (_last_send_time + TIMEOUT_THRESHOLD > tickno) {
       // printf("%lu: timeout occurred!\n", _thread_id);
       // printf("timeout part: _last_send_time: %f, tickno: %f, window: %f, _packets_sent: %ld, _largest_ack: %ld\n", _last_send_time, tickno, _the_window, _packets_sent, _largest_ack);
@@ -244,6 +245,7 @@ double Unicorn::next_event_time( const double & tickno ) const
       // printf("infinite part: _last_send_time: %f, tickno: %f, window: %f, _packets_sent: %ld, _largest_ack: %ld\n", _last_send_time, tickno, _the_window, _packets_sent, _largest_ack);
       return std::numeric_limits<double>::max();
     }
+    // return std::numeric_limits<double>::max();
   }
 }
 

### 2018-01-15 async_deep_reinforce/constants.py
Set learning rate over an environment variable

@@ -33,7 +33,7 @@ LOG_FILE = os.path.join(ABSOLUTE_PATH,'a3c_log')
 logging.info(" ".join(map(str,("LOG_DIR:",LOG_FILE))))
 ACTOR_FACTOR = 1e0
 VALUE_FACTOR = 1e0
-GENERAL_FACTOR = 10**-3
+GENERAL_FACTOR = 10**float(os.environ.get('learning_rate')) if os.environ.get('learning_rate') is not None else -3
 # INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate
 # INITIAL_ALPHA_HIGH = 1e0*GENERAL_FACTOR   # log_uniform high limit for learning rate
 INITIAL_RATE = GENERAL_FACTOR

### 2018-01-15 tex/paper.tex
Tex

--- 2018-01-14 tex/paper.tex
@@
 \begin{document}
-\title{Some title}
+\title{Let's play congestion control}
 \titlenote{Produces the permission block, and
@@
 
-To this end we use an Artificial Neural Network (ANN) based approach that can be initialized with a previously trained neural network or with random weights (without any prior knowledge).
+To this end we use a neural network based approach that can be initialized with a previously trained neural network or with random weights (without any prior knowledge).
 As in congestion control the rewards used in the learning process naturally come delayed and bit by bit, we propose a novel formulation of reinforcement learning that allows us to deal with delayed and partial rewards.
@@
 
-Recent advances in Artificial Neural Network (ANN) based Reinforcement Learning (RL) for the purpose of playing video games raise the question if it is also possible to formulate sending Internet data in a fashion similar to a video game.
+Recent advances in neural network based Reinforcement Learning (RL) for the purpose of playing video games raise the question if it is also possible to formulate sending Internet data in a fashion similar to a video game.
 The rules of the ``Game of Congestion Control'' can be summarized approximately as follows: Each player wants to send \begin{enumerate*}
@@
 \includegraphics[width=\columnwidth]{figures/4_1}
-\caption{The throughput (solid line) and the loss rate (dashed line) when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1, 2 and 4 senders (from top to bottom).
+\caption{\todo{figures don't look that appealing do they?} The throughput (solid line) and the loss rate (dashed line) when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1, 2 and 4 senders (from top to bottom).
 One can see that the time to convergence increases as the number of senders increases, which is caused by the fact that the more senders use the link, the fewer packets each sender receives per unit of time for training.
@@
 maximizing throughput while minimizing packet loss).
+However, PCC doesn't maintain a congestion window but instead modulates the sending rate.
 
@@
 
-To this end we use Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn a wide range of video games and commonly outperform human players.
+To this end we use Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn to play a wide range of video games and commonly outperform human players.
 In particular it is a good choice for congestion control as it can be easily adapted to a wide range of problems and has been proven to deliver good performance.
@@
 
-There are two Artificial Neural Networks (ANNs), the Actor Network and the Value Network (the Critic part in the abbreviation stands for the Value Network).
+There are two neural networks, the Actor Network and the Value Network (the Critic part in the abbreviation stands for the Value Network).
 Given a state, the Actor Network outputs what it deems to be the optimum action to perform in that certain state.
@@
 
-With $r_t$ being the reward that was received at time $t$ and $\gamma$ being the roll-off factor, which stands for the influence that future reward has on the moving average (commonly set to $0.99$), we define the expected long-term reward as
+With $r_t$ being the reward that was received at time $t$ and $\gamma$ being a factor with $0 < \gamma \leq 1$, which stands for the influence that future reward has on the moving average, we define the expected long-term reward as
 \begin{align*}
-R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta_\text{v})\right)\left( 1-\gamma \right),
+R_t = \gamma r_{t} + (1-\gamma) R_{t+1} ,
+%R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta_\text{v})\right)\left( 1-\gamma \right),
 \end{align*}
-where $k$ is upper-bounded by $t_\text{max}$ ($t_\text{max}$ is a fixed hyperparameter that indicates how many rewards should be received before updating the neural network).
-So $R_t$ is simply a moving average of rewards at time step $t$.
-However, usual moving averages take into account values from the past while this one uses values from the future.
+
+One can see that to compute $R_t$ at time step $t$ one has to look infinitely far in the future to get all future rewards needed to accurately computer the exponentially weighted moving average at time step $t$.
+Thus, one usually only carries out $t_\text{max}$ steps (for example 20) and then uses an estimation of $R_{t_\text{max}+1}$ provided by the value network as the continuation \todo{Clearer than before?}.
+The only exception from this is when the task to carry out is over before reaching $t_\text{max}$.
+In this case, one can either use a prediction from the value network as usually or simply let the moving average end without continuation (ignoring the term weighted with $(\gamma-1)$ at the last step).
+
+%where $k$ is upper-bounded by $t_\text{max}$ ($t_\text{max}$ is a fixed hyperparameter that indicates how many rewards should be received before updating the neural network).
+So $R_t$ is the value of the exponentially weighted average of future rewards at time step $t$.
 In practice this means that we collect $t_\text{max}$ rewards, compute the expected long-term reward starting at each time step, update the neural network and start the same procedure again.
@@
 
-The loss function, which the value network tries to minimize at each time step, is the square of the difference of the actual long-term reward received and the expected long-term reward
+The loss function (not in the sense of packet loss but in the sense of the loss of a machine learning problem), which the value network tries to minimize at each time step, is the square of the difference of the actual long-term reward received and the expected long-term reward
 \begin{align*}
@@
 \end{figure}
-
-\begin{algorithm}
-\caption{Partial Action Learning -- pseudocode for each agent.
-It is possible that the agents share the global weights $\theta_{\text{a},\text{g}}$ and $\theta_{\text{v},\text{g}}$, which can be reasonable when learning off-line.
-Otherwise an individual copy of them is kept by each agent.
-All weights are initialized randomly in the beginning.
-LSTM cells are used in both the actor and the value network.}
-\label{alg:pal}
-\begin{algorithmic}[1]
-\Loop
-	\State $l_\text{actions} \gets  [] $
-	\State $l_\text{states} \gets  [] $
-	\State $l_\text{values} \gets  [] $
-	\State $l_\text{rewards} \gets  [] $
-	\State $l_\text{estimatedValues} \gets  [] $
-	\State $l_\text{snapshots} \gets  [] $
-	\State $l_\text{hiddenStates} \gets  [] $
-	\State $t \gets 0$
-	\State $s_0 \gets \text{initialState}()$
-	\State $h_\text{a} \gets \text{initialHiddenState}()$
-	\State $h_\text{v} \gets \text{initialHiddenState}()$
-	\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
-	\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
-	\Repeat
-		\State $l_\text{estimatedValues}.\text{append}(V(s_t; \theta_\text{v}))$
-		\If{$t \bmod t_\text{max} = 0$}
-			\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
-			\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
-			\State $l_\text{snapshots}.\text{append}((\theta_\text{a}, \theta_\text{v}))$
-			\State $l_\text{hiddenStates}.\text{append}((h_\text{a}, h_\text{v}))$
-		\EndIf
-		\State $l_\text{states}.\text{append}(s_t)$
-		\State Sample $a_t$ from the 
-		\StateIndent[3] Actor Network's probability distribution
-		\State $l_\text{actions}.\text{append}(a_t)$
-		\State $l_\text{values}.\text{append}(V(s_t; \theta_\text{v}))$
-		\State $l_{\text{partialActions},t} \gets \text{partialActions}(a_t)$
-		\ForAll{partial actions $a_{{\text{p},i,t}}$ in $l_{\text{partialActions},t}$}
-			\State Take partial action $a_{{\text{p},i,t}}$
-		\EndFor
-		\State Wait for the next feedback $r_{\text{p},j,t'}$
-		\StateIndent[3] where $t' \leq t$ and $0 \leq j \leq\#(l_{\text{partialActions},t'})$
-		\If{all feedback of $a_{t'}$ was received}
-			\State $r_{t'} \gets \text{reward w.r.t all feedback }r_{\text{p},k,t'}$
-					\StateIndent[4] where $0 \leq k \leq\#(l_{\text{partialActions},t'})$
-			\State $l_\text{rewards}.\text{append}(r_{t'})$
-			\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
-				\State \Call{computeGradients}{}
-			\EndIf
-		\EndIf
-		\State Generate $s_{t+1}$ using $r_{\text{p}_{t'}}$
-		\State $t \gets t+1$
-	\Until{reaching the end the episode}
-\EndLoop
-%\algstore{palalg}
-\end{algorithmic}
-\end{algorithm}
-\begin{algorithm}
-\caption{Partial Action Learning -- procedure which computes and applies the gradients.}
-\label{alg:grad}
-\begin{algorithmic}[1]%
-%\algrestore{palalg}
-\Function{computeGradients}{}
-	\State $t_\text{end} = \min\left(t_\text{max}, \#\left( l_\text{rewards} \right)\right)$
-	\State $\theta_\text{backup} \gets \left(\theta_\text{a}, \theta_\text{v}\right)$
-	\State $h_\text{backup} \gets \left(h_\text{a}, h_\text{v}\right)$
-	\State $\theta_\text{a}, \theta_\text{v} \gets l_\text{snapshots}[0]$
-	\State $h_\text{a}, h_\text{v} \gets l_\text{hiddenStates}[0]$
-	\State $R_{i+1} \gets$ last element of $l_\text{estimatedValues}$%
-	\For{$i\gets t_\text{end}-1, 0$}
-		\State $R_i \gets \left(r_i + \gamma R_{i+1}\right)\left(1-\gamma\right)$
-		\State $a \gets l_\text{actions}[i]$
-		\State $s \gets l_\text{states}[i]$
-		\State $v \gets l_\text{values}[i]$
-		\State $d\theta_\text{a} \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
-		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta_a}$
-		\State $d\theta_\text{v} \gets \frac{\partial\left(R_i - v\right)}{\partial\theta_v}$
-		\State $\theta_{\text{a},\text{g}} \gets \theta_{\text{a},\text{g}} + d\theta_\text{a}$
-		\State $\theta_{\text{v},\text{g}} \gets \theta_{\text{v},\text{g}} + d\theta_\text{v}$
-	\EndFor
-	\State Remove first $t_\text{end}$ elements from
-	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{estimatedValues}$
-	\State Remove the first element from $l_\text{snapshots}$
-	\State Remove the first element from $l_\text{hiddenStates}$
-	\State $\theta_\text{a}, \theta_\text{v} = \theta_\text{backup}$
-	\State $h_\text{a}, h_\text{v} = h_\text{backup}$
-\EndFunction
-\end{algorithmic}
-\end{algorithm}
 
@@
 
+\section*{Appendix}
+
+\begin{algorithm}[H]
+\caption{Partial Action Learning -- pseudocode for each agent.
+It is possible that the agents share the global weights $\theta_{\text{a},\text{g}}$ and $\theta_{\text{v},\text{g}}$, which can be reasonable when learning off-line.
+Otherwise an individual copy of them is kept by each agent.
+All weights are initialized randomly in the beginning.
+LSTM cells are used in both the actor and the value network.}
+\label{alg:pal}
+\begin{algorithmic}[1]
+\Loop
+	\State $l_\text{actions} \gets  [] $
+	\State $l_\text{states} \gets  [] $
+	\State $l_\text{values} \gets  [] $
+	\State $l_\text{rewards} \gets  [] $
+	\State $l_\text{estimatedValues} \gets  [] $
+	\State $l_\text{snapshots} \gets  [] $
+	\State $l_\text{hiddenStates} \gets  [] $
+	\State $t \gets 0$
+	\State $s_0 \gets \text{initialState}()$
+	\State $h_\text{a} \gets \text{initialHiddenState}()$
+	\State $h_\text{v} \gets \text{initialHiddenState}()$
+	\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
+	\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
+	\Repeat
+		\State $l_\text{estimatedValues}.\text{append}(V(s_t; \theta_\text{v}))$
+		\If{$t \bmod t_\text{max} = 0$}
+			\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
+			\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
+			\State $l_\text{snapshots}.\text{append}((\theta_\text{a}, \theta_\text{v}))$
+			\State $l_\text{hiddenStates}.\text{append}((h_\text{a}, h_\text{v}))$
+		\EndIf
+		\State $l_\text{states}.\text{append}(s_t)$
+		\State Sample $a_t$ from the 
+		\StateIndent[3] Actor Network's probability distribution
+		\State $l_\text{actions}.\text{append}(a_t)$
+		\State $l_\text{values}.\text{append}(V(s_t; \theta_\text{v}))$
+		\State $l_{\text{partialActions},t} \gets \text{partialActions}(a_t)$
+		\ForAll{partial actions $a_{{\text{p},i,t}}$ in $l_{\text{partialActions},t}$}
+			\State Take partial action $a_{{\text{p},i,t}}$
+		\EndFor
+		\State Wait for the next feedback $r_{\text{p},j,t'}$
+		\StateIndent[3] where $t' \leq t$ and $0 \leq j \leq\#(l_{\text{partialActions},t'})$
+		\If{all feedback of $a_{t'}$ was received}
+			\State $r_{t'} \gets \text{reward w.r.t all feedback }r_{\text{p},k,t'}$
+					\StateIndent[4] where $0 \leq k \leq\#(l_{\text{partialActions},t'})$
+			\State $l_\text{rewards}.\text{append}(r_{t'})$
+			\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
+				\State \Call{computeGradients}{}
+			\EndIf
+		\EndIf
+		\State Generate $s_{t+1}$ using $r_{\text{p}_{t'}}$
+		\State $t \gets t+1$
+	\Until{reaching the end the episode}
+\EndLoop
+%\algstore{palalg}
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+\caption{Partial Action Learning -- procedure which computes and applies the gradients.}
+\label{alg:grad}
+\begin{algorithmic}[1]%
+%\algrestore{palalg}
+\Function{computeGradients}{}
+	\State $t_\text{end} = \min\left(t_\text{max}, \#\left( l_\text{rewards} \right)\right)$
+	\State $\theta_\text{backup} \gets \left(\theta_\text{a}, \theta_\text{v}\right)$
+	\State $h_\text{backup} \gets \left(h_\text{a}, h_\text{v}\right)$
+	\State $\theta_\text{a}, \theta_\text{v} \gets l_\text{snapshots}[0]$
+	\State $h_\text{a}, h_\text{v} \gets l_\text{hiddenStates}[0]$
+	\State $R_{i+1} \gets$ last element of $l_\text{estimatedValues}$%
+	\For{$i\gets t_\text{end}-1, 0$}
+		\State $R_i \gets \left(r_i + \gamma R_{i+1}\right)\left(1-\gamma\right)$
+		\State $a \gets l_\text{actions}[i]$
+		\State $s \gets l_\text{states}[i]$
+		\State $v \gets l_\text{values}[i]$
+		\State $d\theta_\text{a} \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
+		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta_a}$
+		\State $d\theta_\text{v} \gets \frac{\partial\left(R_i - v\right)}{\partial\theta_v}$
+		\State $\theta_{\text{a},\text{g}} \gets \theta_{\text{a},\text{g}} + d\theta_\text{a}$
+		\State $\theta_{\text{v},\text{g}} \gets \theta_{\text{v},\text{g}} + d\theta_\text{v}$
+	\EndFor
+	\State Remove first $t_\text{end}$ elements from
+	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{estimatedValues}$
+	\State Remove the first element from $l_\text{snapshots}$
+	\State Remove the first element from $l_\text{hiddenStates}$
+	\State $\theta_\text{a}, \theta_\text{v} = \theta_\text{backup}$
+	\State $h_\text{a}, h_\text{v} = h_\text{backup}$
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
 %\subsection{Convergence}

### 2018-01-17 tex/paper.tex
Overhauled the paper

@@
 \begin{abstract}
-This paper proposes a novel method of Congestion Control that uses on-line Reinforcement Learning (RL) to maintain the optimum congestion window with respect to a given utility function.
+This paper proposes a novel method of Congestion Control that uses on-line reinforcement learning (RL) to maintain the optimum congestion window with respect to a given utility function.
 
@@
 We show that our method converges to a stable, close-to-optimum solution within the order of minutes and then commonly outperforms existing congestion control algorithms in typical networks.
-Thus, for the first time, we demonstrate that ANN based Reinforcement Learning without any preknowledge can feasibly be done on-line and can compete with hand-crafted solutions given long enough episodes.
+Thus, for the first time, we demonstrate that ANN based reinforcement learning without any preknowledge can feasibly be done on-line and can compete with hand-crafted solutions given long enough episodes.
 
@@
 
-Recent advances in neural network based Reinforcement Learning (RL) for the purpose of playing video games raise the question if it is also possible to formulate sending Internet data in a fashion similar to a video game.
+Recent advances in neural network based reinforcement learning (RL) for the purpose of playing video games raise the question if it is also possible to formulate sending Internet data in a fashion similar to a video game.
 The rules of the ``Game of Congestion Control'' can be summarized approximately as follows: Each player wants to send \begin{enumerate*}
@@
 To address these problems, we develop a new formulation of RL called \textit{Partial Action Learning} (PAL).
-PAL is a superset of Reinforcement Learning: If one uses PAL for a learning problem without delay, asynchronicity and partial actions/partial rewards, one gets classical RL.
+PAL is a superset of reinforcement learning: If one uses PAL for a learning problem without delay, asynchronicity and partial actions/partial rewards, one gets classical RL.
 
@@
 
-After having learned how to perform congestion control in a certain network scenario, our approach can compete with established algorithms and achieve superior throughput with packet loss at an average of less than 5\%, while still gradually improving and adapting its way of performing congestion control.
+After spending up to a few minutes with suboptimal performance during initial exploration (if no online pre-training was used), our approach can compete with established algorithms and achieve superior throughput with packet loss at an average of less than 5\%, while still gradually improving and adapting its way of performing congestion control.
 
@@
 These approaches as well as most others maintain a congestion window which stands for the maximum amount of data that are allowed to be traveling in the network without having been acknowledged by the receiver.
-In general, successful transmission and acknowledgement of data leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995} \todo{How much detail is necessary in the related work section?}) lead to the congestion window being lowered according to a set of fixed rules.
+In general, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995} \todo{How much detail is necessary in the related work section?}) lead to the congestion window being lowered according to a set of fixed rules.
 Algorithms such as Cubic and Compound \cite{ha_cubic:_2008, tan_compound_2006} improve congestion control specifically with respect to so-called long fat networks with a high round-trip time and bandwidth.
@@
 Otherwise -- if a packet was lost due to congestion according to the classifier -- it uses a traditional TCP congestion control algorithm.
-\citet{winstein_tcp_2013} train a machine learning solution called \textit{Remy} that finds an optimum congestion control for a given range of a range of network parameters.
-For instance, one could find an optimum congestion control for networks with a RTT of 50-100\,ms and link speeds of 100-500\,Mbit/s.
-After a lengthy training procedure Remy finds an optimum congestion control algorithm for the specified networks on a per acknowledgement basis.
+\citet{winstein_tcp_2013} train a machine learning solution called \textit{Remy} that finds an optimum congestion control for a given range of network parameters.
+For instance, one could find an optimum congestion control for networks with an RTT of 50-100\,ms and link speeds of 100-500\,Mbit/s.
+After a lengthy training procedure Remy finds an optimum congestion control algorithm for the specified networks on a per acknowledgement basis (updating the window upon receiving an acknowledgement alike traditional congestion control algorithms).
 
 Our goal is to find an optimum congestion control algorithm on a per acknowledgement basis similar to Remy that, however, can be trained on-line.
-Such a solution could be used in a purely on-line fashion, in a off-line fashion like Remy or a combination of both: One pre-trains a generic congestion control algorithm that works reasonably well for every network that then gets refined during on-line training according to the current network circumstances.
+Such a solution could be used in a purely on-line fashion, in a off-line fashion like Remy or a combination of both: One could pre-train a generic congestion control algorithm that works reasonably well for every network that then gets refined during on-line training according to the current network circumstances.
 
-To this end we use Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn to play a wide range of video games and commonly outperform human players.
-In particular it is a good choice for congestion control as it can be easily adapted to a wide range of problems and has been proven to deliver good performance.
+To this end we propose Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn to play a wide range of video games and commonly outperform human players.
+In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for playing video games and also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017} \todo{This motivation doesn't sound convincing yet, does it?}.
 Furthermore, as we will show, it is conceptually possible to use it for on-line training although -- to our knowledge -- this has not been done until now.
@@
 
-The \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) framework is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline in this section.
+The \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) framework is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline (with a focus on intuition over rigor) in this section.
 
-There are two neural networks, the Actor Network and the Value Network (the Critic part in the abbreviation stands for the Value Network).
+There are two neural networks, the Actor Network and the Critic Network.
 Given a state, the Actor Network outputs what it deems to be the optimum action to perform in that certain state.
-The Value Network estimates what long-term reward can be expected in this state.
-So an action is considered good if it achieved a long-term reward that is higher than the long-term reward expected by the Value Network and it is considered bad if the reward was lower than expected.
+The Critic Network estimates what long-term reward can be expected in this state.
+So an action is considered good if it achieved a long-term reward that is higher than the long-term reward expected by the Critic Network and it is considered bad if the reward was lower than expected.
 The long-term reward is implemented as an exponentially weighted moving average of future rewards.
@@
 
-\subsubsection{Value Network}
+\subsubsection{Critic Network}
 \label{subsubsec:genericvalue}
 
-The Value Network outputs the expected long-term reward $V(s_{t}; \theta_\text{v})$ given a state $s_t$ at time step $t$ and the parameters (neural network weights) of the value network $\theta_v$.
+The Critic Network outputs the expected long-term reward $V(s_{t}; \theta_\text{v})$ given a state $s_t$ at time step $t$ and the parameters (neural network weights) of the critic network $\theta_v$.
 
-With $r_t$ being the reward that was received at time $t$ and $\gamma$ being a factor with $0 < \gamma \leq 1$, which stands for the influence that future reward has on the moving average, we define the expected long-term reward as
+With $r_t$ being the reward that was received at time $t$ and $\gamma$ being a factor with $0 < \gamma \leq 1$, which stands for the influence that future reward has on the moving average, we define the expected long-term reward at time step $t$ as
 \begin{align*}
@@
 
-One can see that to compute $R_t$ at time step $t$ one has to look infinitely far in the future to get all future rewards needed to accurately computer the exponentially weighted moving average at time step $t$.
-Thus, one usually only carries out $t_\text{max}$ steps (for example 20) and then uses an estimation of $R_{t_\text{max}+1}$ provided by the value network as the continuation \todo{Clearer than before?}.
-The only exception from this is when the task to carry out is over before reaching $t_\text{max}$.
-In this case, one can either use a prediction from the value network as usually or simply let the moving average end without continuation (ignoring the term weighted with $(\gamma-1)$ at the last step).
+One can see that to compute $R_t$ at time step $t$ one has to look infinitely far into the future to get all future rewards needed to compute the exponentially weighted moving average at time step $t$.
+Thus, one usually only carries out $t_\text{max}$ steps (for example 20) and then uses an estimation of $R_{t_\text{max}+1}$ provided by the critic network as the continuation \todo{Clearer than before?}.
+The only exception from this is when the task to carry out is over before reaching $t_\text{max}$ (e.g.
+the TCP flow is over).
+In this case, one can either use a prediction from the critic network as a continuation as usually or simply let the moving average end without continuation (ignoring the term weighted with $(\gamma-1)$ at the last step).
 
@@
 
-The loss function (not in the sense of packet loss but in the sense of the loss of a machine learning problem), which the value network tries to minimize at each time step, is the square of the difference of the actual long-term reward received and the expected long-term reward
+The loss function (not in the sense of packet loss but in the sense of the loss of a machine learning problem), which the critic network tries to minimize at each time step, is the square of the difference of the actual long-term reward received and the expected long-term reward
 \begin{align*}
-l_{\text{v},t} = \left(R_t - V(s_t; \theta_\text{v})\right)^2.
+l_{\text{v},t} = \left(R_t - V(s_t; \theta)\right)^2.
 \end{align*}
@@
 The Actor Network outputs a probability distribution from which the action $a_t$ at time step $t$ is randomly sampled.
-We use the mean $\mu$ and the standard deviation $\sigma$, to parametrize a normal distribution.
+We use the mean $\mu$ and the standard deviation $\sigma$ to parametrize a normal distribution.
 The main idea is that the network learns to output the right mean at the right time step to maximize the future reward and that it uses the standard deviation to try out new actions, which could yield a better than expected reward.
 
-With $v_t$ being the value that the value network estimated as the future reward given the current state $s_t$ at time step $t$ ($v_t$ is just an abbreviation for $V(s_t; \theta_\text{v})$) and $\theta_a$ the parameters of the actor network and with $\beta$ being a factor that specifies the importance of the entropy $H$, $\pi$ designating the probability density function, meaning that $\pi\left( a_t \given s_t; \theta_\text{a} \right)$ is the value of the probability density function of taking action $a_t$ in state $s_t$ with the current weights of the actor network $\theta_\text{a}$, we define the loss that the actor network aims to minimize as follows:
+Each time we take an action $a_t$ we also let the critic network make an estimation $V(s_t; \theta)$, which we abbreviate as $v_t$, given the current neural network weights $\theta$ and the current state $s_t$.
+The loss function of the actor network can now be described as this: How likely was it (on a logarithmic scale) to take action $a_t$ given the probability distribution at time $t$ and how beneficial was this action when comparing to the critic's estimation?
+For example, if an action was very successful and highly unlikely, the actor network's are updated to make actions like this more likely in the future.
+
+The actor also aims at increasing the entropy of the probability density function ($\frac{1}{2} \log \left( 2 \pi e \sigma^2 \right)$) to encourage exploration: Otherwise the actor could always output similar actions but never actually take the action that yields the best performance.
+We designate the entropy as $H$ and introduce a parameter $\beta$ which specifies magnitude of the influence of the entropy.
 \begin{align*}
-l_{\text{a},t} =& -\log \left( \pi\left( a_t \given s_t; \theta_\text{a} \right)\right)\left( R_t - v_t \right)\\ 
-&- \beta H\left( \pi\left( s_t; \theta_a \right)\right).
+l_{\text{a},t} =& -\log \left( \pi\left( a_t \given s_t; \theta \right)\right)\left( R_t - v_t \right)\\ 
+&- \beta H\left( \pi\left( s_t; \theta \right)\right).
 \end{align*}
@@
 
-The key difference between PAL and previous approaches to Reinforcement Learning is that in classical Reinforcement Learning, an action is always followed by a reward and a reward is always followed by a action.
+The key difference between PAL and previous approaches to RL is that in classical RL an action is always followed by a reward and a reward is always followed by an action.
 In our proposed concept, however, it is possible to take new actions while previous actions haven't received their rewards yet.
@@
 Upon receiving feedback for a partial action, the agent determines the current state and triggers a new action.
-When all feedbacks of one action were received, the agent combines them to form the reward and updates the value and actor networks.
+When all feedbacks of one action were received, the agent combines them to form the reward and updates the critic and actor networks.
 
-In \autoref{alg:pal} we show the code that runs in each of the agents (in the congestion control scenario, one agent corresponds to a sender).
-It is possible to have several agents which share a set of weights but one can also use separate weights for each agent, which is more realistic in case of congestion control in the Internet, as different senders cannot easily share a set of neural network weights over the Internet.
+In \autoref{alg:pal} (in the appendix) we show the code that runs in each of the agents (in the congestion control scenario, one agent corresponds to a sender).
+It is possible to have several agents which share a set of neural weights (which is usually done when performing off-line training \cite{mnih_asynchronous_2016,mao_neural_2017}) but one can also use separate weights for each agent, which is more realistic in case of congestion control in the Internet, as it is not sensible that different senders share a set of neural network weights over the Internet.
 
@@
 \includesvg{Reinforcement_learning_diagram}
-\caption{The classical Reinforcement Learning approach.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
+\caption{The classical reinforcement learning approach.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
 \label{fig:reinforcement}
@@
 
-\subsection{Congestion Control specifics}
+\subsection{Congestion Control Specifics}
 
@@
 We start with one token in the state \textit{Received feedback}, which means that we can take an action right in the beginning.
-The concept is the following: If there is at least one token in the \textit{Open congestion window}, send a packet.
-The token goes to the network and after some time the acknowledgement (or timeout if the packet gets lost) for the packet is received in the \textit{Received ACK} state.
-From here one token goes to the \textit{Open congestion window} meaning that when we get an ACK (or timeout) we can send another packet (The \textit{Open congestion window} signifies the congestion window minus the packets that are currently unacknowledged.).
-Furthermore the state \textit{Received feedback} receives a token upon receiving an ACK/timeout.
+The concept is the following: If there is at least one token in the \textit{Open congestion window} send a packet (for the sake of simplicity in this example we assume that the congestion window is in packets and not bytes).
+The \textit{Open congestion window} signifies the congestion window minus the packets that are currently unacknowledged.
+The token now goes to the network and after some time the acknowledgement (ACK) (or negative-acknowledgement (NACK) or timeout if the packet gets lost) for the packet is received in the \textit{Received ACK} state.
+From here one token goes to the \textit{Open congestion window} meaning that when we get an ACK, NACK (or timeout) we can send another packet.
+Furthermore the state \textit{Received feedback} receives a token upon receiving an ACK/NACK/timeout.
 \textit{Take action} adds \textit{a} tokens to the \textit{Open congestion window}, where $a$ is a real number (possibly also negative); thus in each state there can also be a real number of tokens (e.g.~ 2.34 tokens are possible in the \textit{Open congestion window}).
-However, we define that the congestion window can never be smaller than 1 (This means that the sum of all tokens in the \textit{Open congestion window.}, \textit{Received ACK} and \textit{Packets in the network} states can never be smaller than 1.}
+However, we define that the congestion window can never be smaller than 1, which means that the sum of all tokens in the \textit{Open congestion window.}, \textit{Received ACK} and \textit{Packets in the network} states can never be smaller than 1.}
 \label{fig:petri}
@@
 
-The motivation for PAL is that classical Reinforcement Learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
-However, in the case of congestion control, it is desirable to perform a new action without having received a reward for the previous action.
-For example, imagine that we receive two acknowledgements directly after one other.
-For each of these two acknowledgements an action has to be performed but it does not seem feasible for the second action to wait until the first action has received a reward as it takes one RTT until an acknowledgement (a reward) is received.
+The motivation for PAL is that classical reinforcement learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
+\todo{Too much repetition of this?} However, in the case of congestion control, it is desirable to perform a new action without having received a reward for the previous action.
+For example, imagine that we receive an acknowledgement.
+Now we have to sample an action and wait for the reward, and it takes one RTT until the first acknowledgements (which make up the reward) are received.
+If we now receive another acknowledgement before receiving the reward of the previous action, classical RL cannot handle this situation because of the asynchronicity of actions and rewards.
 Thus the Asynchronous Actor Critic framework as described by \cite{mnih_asynchronous_2016} cannot be applied to congestion control as it assumes that actions and rewards are synchronized and so we have to use Partial Action Learning (see \autoref{subsec:pal}).
@@
 To use PAL for congestion control we first have to define the correct semantics for this specific use case and we have to explicitly state how the state, reward etc.~are defined.
-Furthermore, we have to define how the actor and value network explicitely work in case of congestion control.
+Furthermore, we have to define how the actor and critic network explicitely work in case of congestion control.
 In the following a time step $t$ corresponds to the reception of an acknowledgement.
@@
 \begin{itemize}
-\item the round-trip time of the last packet
-\item the current congestion window
 \item the time between the last two packets that were sent
 \item the time between the last two packets that were received
-\item the number of packets that were lost since the last acknowledgement was received
+\item the RTT of the last received packet
+\item whether the last packet was lost where $1$ signifies that the last packet was lost while $0$ indicates that it was received correctly
+\item the current congestion window as a real number
 \end{itemize}
+Of each of these features the feature itself is included as well as one exponentially weighted moving average (ewma) with a factor $\alpha$ of $\frac{1}{8}$ and one with an $\alpha$ of $\frac{1}{256}$ (the concept of using ewmas with these specific values of $\alpha$ was first introduced in \textit{Remy} \cite{winstein_tcp_2013}).
 Each time an acknowledgement is received, the state is updated and the actor network is asked for the next action.
-\item[$\textit{a}_\text{t}$] Based on a given state and the history of previous states (because we use LSTM cells and thus can also consider previous states), the actor network returns an action $a_t$, which is a real number that stands for the change of the congestion window.
-\item[$\textit{r}_\text{t}$] The reward is a tuple of at least one reward metric. 
-%These are discussed in more detail in \autoref{subsubsec:value}. 
-For each reward metric there is also an output of the value network that predicts the expected long-term average of this reward metric given the current state. 
-%\item[$\textit{v}_\text{t}$] The value is a tuple of the expected average reward estimated by the value network (see \autoref{subsubsec:value}) for each of the reward metrics. 
+\item[$\textit{a}_t$] Based on a given state and the history of previous states (because we use Gated Recurrent Units (GRUs) and thus can also consider previous states), the actor network returns an action $a_t$, which is a real number that stands for the change of the congestion window.
+\item[$\textit{r}_t$] The reward is a tuple of five reward metrics. 
+%These are discussed in more detail in \autoref{subsubsec:critic}. 
+For each reward metric there is also an output of the critic network that predicts the expected long-term average of this reward metric given the current state. 
+\item[$\textit{v}_t$] The value is a tuple of the expected average reward estimated by the critic network (see \autoref{subsubsec:critic}) for each of the reward metrics. 
 \end{description}
 
-We actually use the following types of reward: 
+We use the following reward metrics: 
 \begin{description}
-\item[$\textit{r}_{\text{packet},t}$] is the number of the packets that the sender sent during time step $t$ and that were not lost (so they were acknowledged at some point by the receiver).
+\item[$\textit{r}_{\text{sent},t}$] is the sum of all bytes in all packets the sender sent during time step $t$.
+\item[$\textit{r}_{\text{received},t}$] is the sum of all bytes in all packets that the sender sent during time step $t$ and that were not lost (so they were acknowledged at some point by the receiver).
 %\item[$\textit{r}_{\text{byte},t}$] is the sum of the bytes of the packets that the sender sent and that were not lost.
 Example: At time $t'$, three packets were sent with 300, 200, and 1500 bytes each so $\textit{r}_{\text{byte},t'} = 2000$
-\item[$\textit{r}_{\text{delay},t}$] is the sum of the round trip times of the packets that the sender sent and that were not lost.
-\item[$\textit{r}_{\text{duration},t}$] is the sum of the time between receiving the last packet and receiving this packet (``inter-receive time'') for the packets that the sender sent and that were not lost.
-\item[$\textit{r}_{\text{lost},t}$] is the sum of packets that were lost between receiving the last packet and receiving this packet.
+%\item[$\textit{r}_{\text{delay},t}$] is the sum of the round trip times of the packets that the sender sent and that were not lost.
+\item[$\textit{r}_{\text{duration},t}$] is the time between receiving the last packet and receiving this packet (``inter acknowledgement reception time'') summed over all packets that the sender sent and that were not lost.
+Example: The sender sent three packets, of which two were not lost.
+The time between receiving the ACK of the first packet that wasn't lost and the previous one is 5\,ms and the time between receiving the ACK of the second packet that wasn't lost and the previous ACK is 5\,ms.
+So for this time step $\textit{r}_{\text{duration},t}$ is 11\,ms.
 \end{description}
@@
 The overall structure of the neural network for both the value and the actor network is depicted in \autoref{fig:neural_network}.
+Having these reward metrics one can compose a variety of functions with their long-term averages.
+How these reward metrics can be used in reinforcement learning is described in \autoref{subsec:utility}.
 
@@
 \caption{An overview of the complete neural network being used.
-Ocher stands for the input, blue for linear layers, red for LSTM layers, green for linear layers with a softplus activation function and purple for the outputs of the neural network.
-The numbers in parentheses next to each label stand for the number of width of that layer (i.e.~the number of neurons except for the inputs and outputs).}
+Ocher stands for the input, blue for linear layers, red for GRU layers and purple for the outputs of the neural network.
+The numbers in parentheses next to each label stand for the width of that layer (i.e.~the number of neurons except for the inputs and outputs).}
 \label{fig:neural_network}
@@
 
-\subsubsection{Value Network}
-\label{subsubsec:value}
+\subsubsection{Critic Network}
+\label{subsubsec:critic}
 
-In the case of congestion control, the loss function $l_{\text{v},t}$ of the value network is actually the sum of the squares of the difference for each of the expected long-term averages and the empirically found averages for each reward metric:
+In the case of congestion control, the loss function $l_{\text{v},t}$ of the critic network is actually the sum of the squares of the difference for each of the expected long-term averages and the empirically found averages for each reward metric:
 \begin{align*}
-l_{\text{v},t} =& \left(R_{\text{packet},t} - V_\text{packet}(s_t; \theta_\text{v})\right)^2 \\
-%&+\left(R_{\text{byte},t} - V_\text{byte}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{delay},t} - V_\text{delay}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{duration},t} - V_\text{duration}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{lost},t} - V_\text{lost}(s_t; \theta_\text{v})\right)^2
+l_{\text{v},t} =& \left(R_{\text{received},t} - V_\text{received}(s_t; \theta_\text{v})\right)^2 \\
+&+\left(R_{\text{sent},t} - V_\text{sent}(s_t; \theta_\text{v})\right)^2 \\
+&+\left(R_{\text{duration},t} - V_\text{duration}(s_t; \theta_\text{v})\right)^2
 \end{align*}
@@
 
-The actor network outputs two parameters: The mean of a normal distribution $\mu$ and its standard deviation $\sigma$.
+Each time an action $a_t$ is requested, a change to the congestion window is sampled from the current normal distribution defined by the parameters $\mu$ and $\sigma$.
+However, we define that the window can never fall below $1$: $w_{t+1} = \max(w_t + a_t, 1)$.
+At the beginning of a flow, the window starts as $1$ too.
 
-Each time an action $a_t$ is requested, it is sampled from the current normal distribution defined by the parameters $\mu$ and $\sigma$.
-
-Then the window is incremented by $a_t$; however, the window can never be smaller than 1. At the beginning of a flow the window starts with 1 as well.
-
-With $H$ being the entropy, the actor network minimizes the loss
+The actor network minimizes the loss function
 %\begin{align*}
@@
 \end{align*}
-where $U$ can be any utility function defined based on some of the reward metrics.
-In the above formulation, one considers how much better the actual experienced Utility was compared to the expected one.
-The Utility is formed from the previously defined expected long-term values.
+where $U$ can be any utility function defined based on some of the (long term averages of the) reward metrics.
+Thus, in simple terms, the actor network considers if an action improved the actual experienced utility was compared to the expected one and adjusts the neural network accordingly.
 
 \subsection{Utility function}
+\label{subsec:utility}
 
-As the utility function one can choose an arbitrary function based on the reward metrics previously defined.
-A safe choice is to use PCC's \citep{dong_pcc:_2015} reward function as its convergence has been previously proven.
-It is defined as follows:
+%t%As the utility function one can choose an arbitrary function based on the reward metrics previously defined.
+It is also possible to add other reward metrics such as the delay if one wants to incorporate it in the utility function.
+
+We use a reward function that is similar to the one used by PCC \cite{dong_pcc:_2015}.
 
 \begin{align*}
-U_t = r_{\text{received},t}\text{Sigmoid}_\alpha\left(\frac{r_{\text{loss},t}}{r_{\text{sent},t}} - 0.05\right) + r_{\text{loss},t}
+U_t = \text{throughput} - \alpha \times \text{lost throughput}
 \end{align*}
-where 
+
+$\alpha$ is a parameter that determines how strongly packet loss is discouraged. We can define this utility function as follows using the previously defined reward metrics:
+
 \begin{align*}
-\text{Sigmoid}_\alpha(x) = \frac{1}{1+e^{\alpha x}},
+U_t = \frac{R_{\text{received},t}}{R_{\text{duration},t}} - \alpha \frac{R_{\text{sent},t} - R_{\text{received},t}}{R_{\text{duration},t}}
 \end{align*}
-$r_\text{received}$ is the actual throughput (in packets per second), $r_\text{loss}$ is the loss rate (in packets per second) and $r_\text{sent}$ is the sending rate (in packets per second).
-Thus $\frac{r_\text{loss}}{r_\text{sent}}$ is the ratio of data getting lost compared to those being sent, which is a number between 0 and 1.
 
-Intuitively it means that one takes the actual throughput (packets that get received by the receiver and do not get lost) times the sigmoid function of the loss ratio minus 0.05 plus the loss rate.
-The sigmoid function essentially acts as a cutoff threshold.
-As soon as the loss rate rises above 5\%, the Sigmoid function decreases very quickly and thus diminishes the Utility.
-
-It is also possible to set the cutoff threshold to 0, which means that losing packets is heavily punished; or one can increase the threshold to allow for more loss.
-
-$r_{\text{received},t}$ can be calculated as $\frac{R_\text{packet},t}{R_\text{duration},t}$, $r_{\text{loss},t}$ can be calculated as $\frac{R_\text{lost},t}{R_\text{duration},t}$ and $r_{\text{sent},t}$ can be calculated as $\frac{{R_\text{packet},t} + {R_\text{lost},t}}{R_\text{duration},t}$.
-
-Actually, we assume equally sized packets for all of the above methodology, however, it would also easily be possible to consider the number of bytes per packet as well: One would have to add an additional reward metric of the average number of bytes per packet and as well add one more output to the value network.
+The equivalent utility is also defined using the estimates of the value networks and it can be determined whether an action achieved a performance that was higher or lower than expected.
 
@@
 
+\section{Results}
+
+\section{Discussion}
+
+\subsection{Best practices for on-line reinforcement learning}
+
+\section{Conclusion}
+
 \section*{Appendix}
@@
 All weights are initialized randomly in the beginning.
-LSTM cells are used in both the actor and the value network.}
+LSTM cells are used in both the actor and the critic network.}
 \label{alg:pal}

### 2018-01-17 tex/neural_network.tex

@@
+\newcommand{\nodedistance}{1.2}
+\newcommand{\neuralnetworkwidth}{32}
+
+\pgfmathsetmacro{\shortside}{sqrt((\nodedistance * \nodedistance ) /2)}
+\pgfmathsetmacro{\solutionx}{ \shortside }
+\pgfmathsetmacro{\solutiony}{ - (3 * \shortside + 3 * \nodedistance ) }
+
+\pgfmathsetmacro{\solutiondurationx}{ \shortside }
+\pgfmathsetmacro{\solutiondurationy}{ - (2 * \shortside + 2 * \nodedistance ) }
+
 \centering
-\begin{tikzpicture}[every label/.append style={font=\tiny}, node distance=1.2cm, >=stealth',bend angle=45,auto]
+\begin{tikzpicture}[every label/.append style={font=\tiny}, node distance=\nodedistance cm, >=stealth',bend angle=45,auto]
 
@@
   \begin{scope}
-    \node [input place] (s1) [label={[align=center]left: State (12)}] {};
+    \node [input place] (s1) [label={[align=center]left: State (15)}] {};
     \node [place] (s2) [above right of=s1, label={[align=center]above: Linear}] {};
@@
 
-    \node [output place] (s17) [right of=s13, label={[align=center]above: $V_\text{packets}$ (1)}] {};
+    \node [output place] (s17) [right of=s13, label={[align=center]above: $V_\text{received}$ (1)}] {};
 
   \path[->] (s13) edge node {} (s17);
+%    
+%    \node [place] (s15) [below of=s13, label={[align=center]above: Linear}] {};
+%    
+%  \path[->] (s12) edge node {} (s15);
+%
+%    \node [output place] (s19) [right of=s15, label={[align=center]above: $V_\text{delay}$ (1)}] {};
 
-    \node [place] (s15) [below of=s13, label={[align=center]above: Linear}] {};
+%  \path[->] (s15) edge node {} (s19);
 
-  \path[->] (s12) edge node {} (s15);
-
-    \node [output place] (s19) [right of=s15, label={[align=center]above: $V_\text{delay}$ (1)}] {};
-
-  \path[->] (s15) edge node {} (s19);
-
-    \node [place] (s16) [below of=s15, label={[align=center]above: Linear}] {};
+    \node [place] (s16) [below of=s13, label={[align=center]below: Linear}] {};
 
@@
 
-    \node [output place] (s20) [right of=s16, label={[align=center]above: $V_\text{sent}$ (1)}] {};
+    \node [output place] (s20) [right of=s16, label={[align=center]below: $V_\text{sent}$ (1)}] {};
 
@@
 
-    \node [place] (s14) [below of=s16, label={[align=center]above: Linear}] {};
+    \node [place] (s40) at (\solutiondurationx cm,\solutiondurationy cm) [label={[align=center]below: Linear}] {};
 
-  \path[->] (s12) edge node {} (s14);
+  \path[->] (s1) edge node {} (s40);
 
-    \node [output place] (s21) [right of=s14, label={[align=center]above: $V_\text{duration}$ (1)}] {};
+    \node [lstm place] (s41) [right of=s40, label={[align=center]above: GRU (32)}] {};
 
-  \path[->] (s14) edge node {} (s21);
+  \path[->] (s40) edge node {} (s41);
+  \path[->] (s41) edge [loop below] node {} (s41);
+
+    \node [lstm place] (s42) [right of=s41, label={[align=center]above: GRU (32)}] {};
+
+  \path[->] (s41) edge node {} (s42);
+  \path[->] (s42) edge [loop below] node {} (s42);
+
+    \node [place] (s43) [right of=s42, label={[align=center]above: Linear}] {};
+
+  \path[->] (s42) edge node {} (s43);
+
+    \node [output place] (s44) [right of=s43, label={[align=center]above: $V_\text{duration}$ (1)}] {};
+
+  \path[->] (s43) edge node {} (s44);
 
@@
 
+%    \node [place] (s30) at (\solutionx cm,\solutiony cm) [label={[align=center]below: Linear}] {};
+%
+%  \path[->] (s1) edge node {} (s30);   
+%
+%    \node [lstm place] (s31) [right of=s30, label={[align=center]above: GRU (\neuralnetworkwidth)}] {};
+%    
+%  \path[->] (s30) edge node {} (s31);
+%  \path[->] (s31) edge [loop below] node {} (s31);
+%
+%    \node [lstm place] (s32) [right of=s31, label={[align=center]above: GRU (\neuralnetworkwidth)}] {};
+%    
+%  \path[->] (s31) edge node {} (s32);
+%  \path[->] (s32) edge [loop below] node {} (s32);
+%    
+%    \node [place] (s33) [right of=s32, label={[align=center]above: Linear}] {};
+%    
+%  \path[->] (s32) edge node {} (s33);
+%
+%    \node [output place] (s34) [right of=s33, label={[align=center]above: $V_\text{byte}$ (1)}] {};
+%    
+%  \path[->] (s33) edge node {} (s34);
+
   \end{scope}

### 2018-01-21 tex/paper.tex
Tex

@@
 \usepackage[inline]{enumitem}
+\usepackage{varwidth}
 
@@
 \makeatother
+
+\algnewcommand\True{\textbf{true}}
+\algnewcommand\False{\textbf{false}}
+\algnewcommand\Break{\textbf{break}}
 
@@
 \caption{Partial Action Learning -- pseudocode for each agent.
-It is possible that the agents share the global weights $\theta_{\text{a},\text{g}}$ and $\theta_{\text{v},\text{g}}$, which can be reasonable when learning off-line.
+It is possible that the agents share the global weights $\theta_{\text{g}}$, which is reasonable when learning off-line.
 Otherwise an individual copy of them is kept by each agent.
-All weights are initialized randomly in the beginning.
-LSTM cells are used in both the actor and the critic network.}
+By default, all weights are initialized randomly in the beginning but it is also possible to use a pre-trained neural network for initialization.
+As we use Gated Recurrent Units (GRUs) their output is stored in $h$ and fed into them at the next time step.}
 \label{alg:pal}
@@
 	\State $l_\text{rewards} \gets  [] $
-	\State $l_\text{estimatedValues} \gets  [] $
+	\State $l_\text{values'} \gets  [] $
 	\State $l_\text{snapshots} \gets  [] $
@@
 	\State $s_0 \gets \text{initialState}()$
-	\State $h_\text{a} \gets \text{initialHiddenState}()$
-	\State $h_\text{v} \gets \text{initialHiddenState}()$
-	\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
-	\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
+	\State $h \gets \text{zeros}()$
+	\State $\theta \gets \theta_{\text{g}}$
 	\Repeat
-		\State $l_\text{estimatedValues}.\text{append}(V(s_t; \theta_\text{v}))$
-		\If{$t \bmod t_\text{max} = 0$}
-			\State $\theta_\text{a} \gets \theta_{\text{a},\text{g}}$
-			\State $\theta_\text{v} \gets \theta_{\text{v},\text{g}}$
-			\State $l_\text{snapshots}.\text{append}((\theta_\text{a}, \theta_\text{v}))$
-			\State $l_\text{hiddenStates}.\text{append}((h_\text{a}, h_\text{v}))$
+		\State $l_\text{values'}.\text{append}(V(s_t; \theta))$
+		\If{\Call{updateWeights}{}}
+			\State $\theta_\text{a} \gets \theta_{\text{g}}$
+			\State $l_\text{snapshots}.\text{append}(\theta)$
+			\State $l_\text{hiddenStates}.\text{append}(h)$
 		\EndIf
@@
 		\State $l_\text{values}.\text{append}(V(s_t; \theta_\text{v}))$ 
+		\Comment{Possibly different from the}
+		\StateIndent[3] value in $l_\text{values'}$ as the weights could have 
+		\StateIndent[3] been updated in the meantime.
 		\State $l_{\text{partialActions},t} \gets \text{partialActions}(a_t)$
@@
 			\State $l_\text{rewards}.\text{append}(r_{t'})$
-			\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
+			\If{\Call{enoughRewards}{} or the episode is over}
+%			\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
 				\State \Call{computeGradients}{}
@@
 \Function{computeGradients}{}
-	\State $t_\text{end} = \min\left(t_\text{max}, \#\left( l_\text{rewards} \right)\right)$
-	\State $\theta_\text{backup} \gets \left(\theta_\text{a}, \theta_\text{v}\right)$
-	\State $h_\text{backup} \gets \left(h_\text{a}, h_\text{v}\right)$
-	\State $\theta_\text{a}, \theta_\text{v} \gets l_\text{snapshots}[0]$
-	\State $h_\text{a}, h_\text{v} \gets l_\text{hiddenStates}[0]$
-	\State $R_{i+1} \gets$ last element of $l_\text{estimatedValues}$%
-	\For{$i\gets t_\text{end}-1, 0$}
+	\State $t_\text{end} = \#\left( l_\text{rewards} \right)$
+	\State $\theta_\text{backup} \gets \theta$
+	\State $h_\text{backup} \gets h$
+	\State $\theta \gets l_\text{snapshots}[0]$
+	\State $h \gets l_\text{hiddenStates}[0]$
+	\State $R_{i+1} \gets$ last element of $l_\text{values'}$%
+	\For{$i\gets t_\text{end}-1..0$}
 		\State $R_i \gets \left(r_i + \gamma R_{i+1}\right)\left(1-\gamma\right)$
@@
 		\State $v \gets l_\text{values}[i]$
-		\State $d\theta_\text{a} \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta_a}$
-		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta_a}$
-		\State $d\theta_\text{v} \gets \frac{\partial\left(R_i - v\right)}{\partial\theta_v}$
-		\State $\theta_{\text{a},\text{g}} \gets \theta_{\text{a},\text{g}} + d\theta_\text{a}$
-		\State $\theta_{\text{v},\text{g}} \gets \theta_{\text{v},\text{g}} + d\theta_\text{v}$
+		\State $d\theta \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta}$
+		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta}$
+		\StateIndent[4] $+\frac{\partial\left(R_i - v\right)}{\partial\theta}$
+		\State $\theta_{\text{g}} \gets \theta_{\text{g}} + d\theta$
 	\EndFor
 	\State Remove first $t_\text{end}$ elements from
-	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{estimatedValues}$
+	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{values'}$
 	\State Remove the first element from $l_\text{snapshots}$
 	\State Remove the first element from $l_\text{hiddenStates}$
-	\State $\theta_\text{a}, \theta_\text{v} = \theta_\text{backup}$
-	\State $h_\text{a}, h_\text{v} = h_\text{backup}$
+	\State $\theta = \theta_\text{backup}$
+	\State $h = h_\text{backup}$
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+\caption{Procedure which determines when to synchronize the neural network weights from the global weights in case of congestion control.
+We assume that the window at time step $t$ is stored in a list $l_\text{windows}$ and that the number of bytes sent in each time step is stored in $l_\text{sent}$.
+Alternatively to this procedure one can simply synchronize the weights after each $t_\text{max}$ actions where $t_\text{max}$ is an arbitrarily chosen integer greater than 0.}
+\label{alg:update}
+\begin{algorithmic}[1]%
+%\algrestore{palalg}
+\Function{updateWeights}{}
+	\State $i \gets 0$
+	\While{$i < \#(l_\text{actions})$}
+		\State $w\gets \left\lfloor l_\text{windows}[i] + l_\text{actions}[i] \right\rfloor $
+		\For{$j\gets i..\#(l_\text{actions})-1$}
+			\State $w\gets \text{w} - l_\text{sent}[j]$
+			\If{$w\leq 0$}
+				\State $i \gets j$
+				\State \Break
+			\EndIf
+		\EndFor
+		\If{$i = \#(l_\text{windows})-1$}
+			\State \Return \True
+		\EndIf
+	\EndWhile
+	\State \Return \False
+\EndFunction
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[H]
+\caption{Procedure which determines when to start computing the gradients in case of congestion control.
+We assume that the window at time step $t$ is stored in a list $l_\text{windows}$ and that the number of bytes sent in each time step is stored in $l_\text{sent}$.
+Alternatively to this procedure one can simply compute the gradients when $t_\text{max}$ rewards have been received where $t_\text{max}$ is an arbitrarily chosen integer greater than 0.}
+\label{alg:enough}
+\begin{algorithmic}[1]%
+%\algrestore{palalg}
+\Function{enoughRewards}{}
+	\State w $\gets \left\lfloor l_\text{windows}[0] + l_\text{actions}[0] \right\rfloor$
+	\For{$i\gets 0..\#(l_\text{rewards})-1$}
+		\State w $\gets \text{w} - l_\text{sent}[i]$
+		\If{w $\leq 0$}
+			\State \Return \True
+		\EndIf
+	\EndFor
+	\State \Return \False
 \EndFunction

### 2018-01-21 src/network.cc
Add lost packets to csv logs

@@ -186,6 +186,7 @@ void Network<Gang1Type, Gang2Type>::run_simulation_with_logging_until( const dou
     }
 
   }
+
 }
 
 template <class Gang1Type, class Gang2Type>

### 2018-01-21 src/unicorn.cc

@@ -350,5 +350,6 @@ SimulationResultBuffers::SenderState Unicorn::state_DNA() const
   ret.set_window_size( _the_window );
   ret.set_intersend_time( nan("1") );
   ret.set_packets_sent(_packets_sent);
+  ret.set_packets_lost(_packets_lost);
   return ret;
 }

### 2018-01-21 src/unicornbreeder.cc

@@ -29,12 +29,12 @@ void UnicornBreeder::run(const size_t iterations)
 
     string csv_stuff = "";
 
-    csv_stuff += "time,sender_number,packets_received,packets_sent,sending_duration,total_delay\n";
+    csv_stuff += "time,sender_number,packets_received,packets_sent,packets_lost,sending_duration,total_delay\n";
     for (vector<SimulationRunData>::const_iterator it = logging.data().begin(); it != logging.data().end(); ++it) {
       for (vector<SimulationRunDataPoint>::const_iterator inner_it = (*it).data().begin(); inner_it != (*it).data().end(); ++inner_it) {
         size_t sender_index = 0;
         for (vector<SenderDataPoint>::const_iterator sender_it = (*inner_it).sender_data().begin(); sender_it != (*inner_it).sender_data().end(); ++sender_it) {
-          csv_stuff += to_string((*inner_it).seconds()) + "," + to_string(sender_index) + "," + to_string((*sender_it).utility_data.packets_received()) + "," + to_string((*sender_it).sender_state.packets_sent()) + "," + to_string((*sender_it).utility_data.sending_duration()) + "," + to_string((*sender_it).utility_data.total_delay()) + "\n";
+          csv_stuff += to_string((*inner_it).seconds()) + "," + to_string(sender_index) + "," + to_string((*sender_it).utility_data.packets_received()) + "," + to_string((*sender_it).sender_state.packets_sent()) + "," + to_string((*sender_it).sender_state.packets_lost()) +  "," + to_string((*sender_it).utility_data.sending_duration()) + "," + to_string((*sender_it).utility_data.total_delay()) + "\n";
           sender_index += 1;
         }
       }

### 2018-01-23 tex/paper.tex
Changes in the abstract

@@
 \begin{abstract}
-This paper proposes a novel method of Congestion Control that uses on-line reinforcement learning (RL) to maintain the optimum congestion window with respect to a given utility function.
+This paper proposes a new method of congestion control that uses on-line reinforcement learning to maintain the optimum congestion window with respect to a given utility function.
 
-To this end we use a neural network based approach that can be initialized with a previously trained neural network or with random weights (without any prior knowledge).
-As in congestion control the rewards used in the learning process naturally come delayed and bit by bit, we propose a novel formulation of reinforcement learning that allows us to deal with delayed and partial rewards.
+We use a neural network based approach that can be initialized with a previously trained neural network or with random weights (without any prior knowledge).
+As in congestion control the rewards used in the learning process naturally come delayed and bit by bit, we propose a specific formulation of neural network based reinforcement learning that allows us to deal with delayed and partial rewards.
 
-We show that our method converges to a stable, close-to-optimum solution within the order of minutes and then commonly outperforms existing congestion control algorithms in typical networks.
-Thus, for the first time, we demonstrate that ANN based reinforcement learning without any preknowledge can feasibly be done on-line and can compete with hand-crafted solutions given long enough episodes.
+We show that our method converges to a stable, close-to-optimum solution within the order of minutes and outperforms existing congestion control algorithms in typical networks.
+Thus, we demonstrate that neural network based reinforcement learning without any preknowledge can feasibly be done on-line and can compete with human-designed congestion-control schemes such as New Reno and Cubic.
 
@@
 Clearly, it is also desirable from an overall point of view that all participating players play the game in a way that does not give them an unfair advantage over others.
-When several players (from now on called senders) share an Internet link under the same conditions, they should receive an equal share of the reward.
+When several players (senders) share an Internet link under the same conditions, they should receive an equal share of the reward.
 
-When performing Congestion Control (CG) \todo{What should be capitalized and what shouldn't?
-Capitalize RL, CG, ANN?}, each sender uses a set of observed environment conditions to determine what action he should take to maximize his reward in the future.
+To perform Congestion Control (CG) \todo{What should be capitalized and what shouldn't?
+Capitalize RL, CG, ANN?}, each sender uses a set of observed environment conditions to determine which action he should take to maximize his reward in the future.
 The environment conditions can be any metrics that the sender can obtain from the network, for example the mean round-trip time of the last received packets, the packet loss rate etc.
@@
 \end{itemize}
-Of each of these features the feature itself is included as well as one exponentially weighted moving average (ewma) with a factor $\alpha$ of $\frac{1}{8}$ and one with an $\alpha$ of $\frac{1}{256}$ (the concept of using ewmas with these specific values of $\alpha$ was first introduced in \textit{Remy} \cite{winstein_tcp_2013}).
+Of each of these features a exponentially weighted moving average with a factor $\alpha$ of $\frac{1}{8}$ and one with an $\alpha$ of $\frac{1}{256}$ (the concept of using ewmas with these specific values of $\alpha$ was first introduced in \textit{Remy} \cite{winstein_tcp_2013}), which makes a total of 10 features.
 Each time an acknowledgement is received, the state is updated and the actor network is asked for the next action.
@@
 Ocher stands for the input, blue for linear layers, red for GRU layers and purple for the outputs of the neural network.
-The numbers in parentheses next to each label stand for the width of that layer (i.e.~the number of neurons except for the inputs and outputs).}
+The numbers in parentheses next to each label stand for the width of that layer (i.e.~the number of neurons except for the inputs and outputs).
+Parts of the neural network are shared between $V_\text{received}$ and $V_\text{sent}$ as well as $\mu$ and $\sigma$ as they output similar quantities and thus the gradients that are computed when updating the neural network are of the same order of magnitude.
+In our experiments we saw that if one uses shared weights between very different outputs, the outputs which produce significantly larger gradients would ``overwrite'' the gradients of the smaller ones.}
 \label{fig:neural_network}
@@
 \end{align*}
+
+One apparent issue of using reinforcement learning for congestion control is that in RL one usually uses a fixed parameter $\gamma$ that determines the influence of future rewards.
+However, in congestion control the larger the window is, the more packets are sent and received per unit of time.
+Thus, if one used a single fixed parameter $\gamma$, one would have the problem that one would look into the future for a shorter time the larger the window is, as more packets and thus more actions and rewards are handled per unit of time.
+Our aim is to look into the future for a more-or-less constant time span no matter the current size of the window.
+In other words, we want to define $\gamma$ so that it reflects the number of packets in the current window, so that it always looks in the future for approximately one round-trip time.
+
+A ewma with a $\gamma$ of $\frac{2}{n+1}$ has roughly the same distribution of forecast error as a regular moving average with a window size of $n$ \cite{nahmias_production_2009}.
+Furthermore, with sufficiently large $n$ this means that the first $n$ data account for $\approx 86\%$ of the total weight in the ewma \cite{boissard_applications_2012}.
+
+Thus, the idea is to make $n$ the number of packets in the current window and so we define the factor $\gamma$ as a function of the current window size and the expected amount of bytes to be sent per time step as follows:
+
+\begin{align*}
+\gamma_t = \frac{2}{1+\max(\frac{w_t+a_t}{V_\text{sent}(s_t; \theta_\text{v})},10)}
+\end{align*}
+
+We use 10 as a minimum because when $\gamma$ becomes too big, future rewards is not considered sufficiently anymore and as the variability increases instability ensues.
 
@@
 		\State $l_\text{values'}.\text{append}(V(s_t; \theta))$
-		\If{\Call{updateWeights}{}}
+		\If{\Call{updateWeights?}{}}
 			\State $\theta_\text{a} \gets \theta_{\text{g}}$
@@
 			\State $l_\text{rewards}.\text{append}(r_{t'})$
-			\If{\Call{enoughRewards}{} or the episode is over}
+			\If{\Call{gotEnoughRewards?}{} or the episode is over}
 %			\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
@@
 %\algrestore{palalg}
-\Function{updateWeights}{}
+\Function{updateWeights?}{}
 	\State $i \gets 0$
@@
 			\State $w\gets \text{w} - l_\text{sent}[j]$
-			\If{$w\leq 0$}
+			\If{$w\leq 0 \land i+1 \geq 10$} \Comment{at least 10 rewards}
 				\State $i \gets j$
@@
 %\algrestore{palalg}
-\Function{enoughRewards}{}
-	\State w $\gets \left\lfloor l_\text{windows}[0] + l_\text{actions}[0] \right\rfloor$
+\Function{gotEnoughRewards?}{}
+	\State $w \gets \left\lfloor l_\text{windows}[0] + l_\text{actions}[0] \right\rfloor$
 	\For{$i\gets 0..\#(l_\text{rewards})-1$}
-		\State w $\gets \text{w} - l_\text{sent}[i]$
-		\If{w $\leq 0$}
+		\State $w \gets \text{w} - l_\text{sent}[i]$
+		\If{$w \leq 0 \land i+1 \geq 10$} \Comment{at least 10 rewards}
 			\State \Return \True

### 2018-01-27 tex/paper.tex
Improved the whole learning thing

@@
 \begin{abstract}
-This paper proposes a new method of congestion control that uses on-line reinforcement learning to maintain the optimum congestion window with respect to a given utility function.
 
-We use a neural network based approach that can be initialized with a previously trained neural network or with random weights (without any prior knowledge).
-As in congestion control the rewards used in the learning process naturally come delayed and bit by bit, we propose a specific formulation of neural network based reinforcement learning that allows us to deal with delayed and partial rewards.
+A major challenge in congestion control is to find a set of rules that dynamically adapt the congestion window to current and anticipated future network conditions to maximize a specified utility function, e.g.
+maximum throughput or minimum packet loss.
+
+This paper proposes Reactive Adaptive eXperience based congestion control (TCP Rax), a new method of congestion control that uses online reinforcement learning to maintain an optimum congestion window with respect to a given utility function.
+We use a neural network based approach that can be initialized either with random weights (without prior knowledge) or with a previously trained neural network to improve stability and convergence time.
+The processing of rewards in congestion control depends on the arrival of acknowledgments, inducing loop delays that may lead to oscillations.
+As a remedy we propose a specific formulation of neural network based reinforcement learning that supports the proactive processing of delayed and partial rewards.
 
 We show that our method converges to a stable, close-to-optimum solution within the order of minutes and outperforms existing congestion control algorithms in typical networks.
-Thus, we demonstrate that neural network based reinforcement learning without any preknowledge can feasibly be done on-line and can compete with human-designed congestion-control schemes such as New Reno and Cubic.
+Thus, this paper demonstrates that neural network based reinforcement learning without any prior knowledge can feasibly be done on-line and can compete with human-designed congestion-control schemes such as New Reno and Cubic.
 
@@
 
-Recent advances in neural network based reinforcement learning (RL) for the purpose of playing video games raise the question if it is also possible to formulate sending Internet data in a fashion similar to a video game.
-The rules of the ``Game of Congestion Control'' can be summarized approximately as follows: Each player wants to send \begin{enumerate*}
+Recent advances in neural network based reinforcement learning (RL) for the purpose of playing video games raise the question if it is possible to use these frameworks to model end-to-end congestion control.
+In general, in congestion control each sender aims to maximize an objective function that it composed of one or more of the following to
+\begin{enumerate*}
 \item as many packets as possible

### 2018-01-27 async_deep_reinforce/a3c_training_thread.py

@@ -102,12 +102,16 @@ class A3CTrainingThread(object):
     summary_writer.flush()
 
   def start_anew(self):
+    # print("self.windows", self.windows)
     assert(len(self.windows) > 0)
     current_index = 0
     while current_index < len(self.windows):
+      # print("current_index", current_index)
       if current_index + math.floor(A3CTrainingThread.get_actual_window(self.windows[current_index]+self.actions[current_index])) == len(self.windows):
         return True
       current_index = current_index + math.floor(A3CTrainingThread.get_actual_window(self.windows[current_index]+self.actions[current_index]))
+      # print("current_index afterwards", current_index)
+
     return False
 
   def start_anew_index(self):
@@ -115,6 +119,7 @@ class A3CTrainingThread(object):
     current_index = 0
     while current_index < len(self.windows):
       current_index = current_index + math.floor(A3CTrainingThread.get_actual_window(self.windows[current_index]+self.actions[current_index]))
+    # print("current_index in start_anew", current_index)
     return current_index
 
   def action_step(self, sess, state, tickno, window):
@@ -173,19 +178,21 @@ class A3CTrainingThread(object):
     self.rewards.append((reward_throughput, reward_delay, duration, sent))
 
     # if len(self.rewards)>=LOCAL_T_MAX or (len([item for item in self.actions[:LOCAL_T_MAX] if item is not None]) == len(self.rewards) and len(self.rewards) > 0 and self.time_differences[0] is not None):
-    if ('LOCAL_T_MAX' in globals() and (len(self.rewards)>=globals()["LOCAL_T_MAX"] or (len([item for item in self.actions[:globals()["LOCAL_T_MAX"]] if item is not None]) == len(self.rewards) and len(self.rewards) > 0 and self.time_differences[0] is not None))) or (not 'LOCAL_T_MAX' in globals() and len(self.rewards)>=math.floor(A3CTrainingThread.get_actual_window(self.windows[0]+self.actions[0]))):
[…]

### 2018-01-27 async_deep_reinforce/constants.py

@@ -85,7 +85,7 @@ DURATION_FACTOR = 1e-2
 # PACKET_FACTOR = 1
 # DURATION_FACTOR = 1
 
-STATE_SIZE = int(os.environ.get('state_size')) if os.environ.get('state_size') is not None else 10
+STATE_SIZE = int(os.environ.get('state_size')) if os.environ.get('state_size') is not None else 8
 HIDDEN_SIZE = int(os.environ.get('hidden_size')) if os.environ.get('hidden_size') is not None else 32
 # ACTION_SIZE = 1 # action size
 

### 2018-01-27 src/unicorn.cc

@@ -29,7 +29,7 @@ Unicorn::Unicorn(const bool& cooperative)
     _flow_to_last_received(),
     _active_flows()
 {
-  assert( getenv("delay_delta") != NULL );
+  assert( getenv("delay_delta") != NULL && strcmp(getenv("delay_delta"), "") != );
   _delay_delta = atof(getenv("delay_delta"));
   // puts("Creating a Unicorn");
 }
@@ -164,10 +164,9 @@ void Unicorn::packets_received( const vector< remy::Packet > & packets ) {
 
 void Unicorn::reset(const double & tickno)
 {
-  printf("%lu: Fucking resetting\n", _thread_id);
+  printf("%lu: Resetting\n", _thread_id);
   _rainbow._training = _training;
   // assert(false);
-  // printf("%lu: Resetting\n", _thread_id);
   // _largest_ack -= 1;
   if (_outstanding_rewards.size() != _put_actions-_put_rewards) {
     printf("%lu: _outstanding_rewards=%lu, _put_actions=%lu, _put_rewards=%lu\n", _thread_id, _outstanding_rewards.size(), _put_actions, _put_rewards);
@@ -258,9 +257,9 @@ void Unicorn::get_action(const double& tickno) {
   const double action = _rainbow.get_action(
     _thread_id,
     {
-      // _memory._send,
-      _memory._rec_send_ewma,
-      _memory._slow_rec_send_ewma,
+      // // _memory._send,
+      // _memory._rec_send_ewma,
+      // _memory._slow_rec_send_ewma,
 
       // _memory._rec,
       _memory._rec_rec_ewma,

### 2018-01-27 src/unicorn.cc
Improved the whole learning thing

@@ -29,7 +29,7 @@ Unicorn::Unicorn(const bool& cooperative)
     _flow_to_last_received(),
     _active_flows()
 {
-  assert( getenv("delay_delta") != NULL && strcmp(getenv("delay_delta"), "") != );
+  assert( getenv("delay_delta") != NULL && strcmp(getenv("delay_delta"), "") != 0);
   _delay_delta = atof(getenv("delay_delta"));
   // puts("Creating a Unicorn");
 }

### 2018-01-28 tex/paper.tex
Add new config files change paper

@@
 \begin{document}
-\title{Let's play congestion control}
+\title{TCP Rax: Deep Reinforcement Learning for TCP Congestion Control}
 \titlenote{Produces the permission block, and
@@
 
-\author{Maximilian Bachl} 
-\affiliation{%
- \institution{Institute of Telecommunications}
- \streetaddress{Gußhausstraße 25}
- \postcode{1040}
- \city{Vienna} 
- \country{Austria}}
-\email{maximilian.bachl@tuwien.ac.at}
-
-\author{Tanja Zseby} 
-\affiliation{%
- \institution{Institute of Telecommunications}
- \streetaddress{Gußhausstraße 25}
- \postcode{1040}
- \city{Vienna} 
- \country{Austria}}
-\email{tanja.zseby@tuwien.ac.at}
-
-\author{Joachim Fabini} 
-\affiliation{%
- \institution{Institute of Telecommunications}
- \streetaddress{Gußhausstraße 25}
- \postcode{1040}
- \city{Vienna} 
- \country{Austria}}
-\email{joachim.fabini@tuwien.ac.at}
+%\author{Maximilian Bachl} 
+%\affiliation{%
+% \institution{Institute of Telecommunications}
+% \streetaddress{Gußhausstraße 25}
+% \postcode{1040}
+% \city{Vienna} 
+% \country{Austria}}
+%\email{maximilian.bachl@tuwien.ac.at}
+%
+%\author{Tanja Zseby} 
+%\affiliation{%
+% \institution{Institute of Telecommunications}
+% \streetaddress{Gußhausstraße 25}
+% \postcode{1040}
+% \city{Vienna} 
+% \country{Austria}}
+%\email{tanja.zseby@tuwien.ac.at}
+%
+%\author{Joachim Fabini} 
+%\affiliation{%
+% \institution{Institute of Telecommunications}
+% \streetaddress{Gußhausstraße 25}
+% \postcode{1040}
+% \city{Vienna} 
+% \country{Austria}}
+%\email{joachim.fabini@tuwien.ac.at}
 
 % The default list of authors is too long for headers}
-\renewcommand{\shortauthors}{Bachl, Zseby, Fabini}
+%\renewcommand{\shortauthors}{Bachl, Zseby, Fabini}
 
@@
 
-Recent advances in neural network based reinforcement learning (RL) for the purpose of playing video games raise the question if it is possible to use these frameworks to model end-to-end congestion control.
-In general, in congestion control each sender aims to maximize an objective function that it composed of one or more of the following to
+Recent advances in neural network based reinforcement learning (RL) raise the question if it is possible to use these frameworks to model end-to-end congestion control.
+In general, in congestion control each sender aims to maximize an objective function that usually comprises of
 \begin{enumerate*}
-\item as many packets as possible
+\item sending as many packets as possible
 \item with as little delay as possible
-\item while losing as few packets as possible
+\item while minimizing the amount of lost packets
 \end{enumerate*}.
 As, for instance, for some applications overall throughput is more important than little packet loss, each application can define its preference using a custom utility function.
-Clearly, it is also desirable from an overall point of view that all participating players play the game in a way that does not give them an unfair advantage over others.
-When several players (senders) share an Internet link under the same conditions, they should receive an equal share of the reward.
+Clearly, it is also desirable from an overall point of view that all senders use available resources in a way that is fair to the others: When several senders share an Internet link under the same conditions, they should receive an equal share of the reward.
+To achieve fairness, it is necessary that senders only use utility functions that
+\begin{enumerate*}
+\item converge to a stable equilibrium and
+\item provide each sender with a fair share of his subgoals
+\end{enumerate*}.
 
-To perform Congestion Control (CG) \todo{What should be capitalized and what shouldn't?
-Capitalize RL, CG, ANN?}, each sender uses a set of observed environment conditions to determine which action he should take to maximize his reward in the future.
+In an ideal machine learning based congestion control method, each sender uses a set of observed environment conditions to determine which action he should take to maximize his reward in the future.
 The environment conditions can be any metrics that the sender can obtain from the network, for example the mean round-trip time of the last received packets, the packet loss rate etc.
 An action is a change to the sending rate: For example, if a sender perceives an increase in the loss rate it might be advisable to lower the sending rate.
+To modulate the sending rate one can  either
+\begin{enumerate*}
+\item change the sending rate itself at certain points in time or
+\item maintain a congestion window that indicates the number of packets that can be in flight at a certain point in time
+\end{enumerate*}.
+It seems preferable to maintain a congestion window, as one cannot be certain that the operating system (OS) can maintain a certain constant sending rate for each application on one machine, as the OS has to schedule different applications and share the network interfaces between them.
 
-From this formulation of Congestion Control, one can observe a couple of unique features of the problem that seem to prohibit the use of RL: \begin{itemize}
-\item When playing a video game, if one pushes a button, one can see the consequences of an action immediately.
-However, in Congestion Control, the rewards are always \textbf{delayed} by one round-trip time.
-\item Due to the delay, by the time an action receives its reward, probably other actions have already been carried out in the mean time.
-Thus, contrary to classical RL, \textbf{action and rewards are out of sync}.
+While congestion control seems to be a well suited problem for reinforcement learning, there are a couple of challenges and differences from most other reinforcement learning problems:  \begin{itemize}
+\item For instance, when playing a video game, if one pushes a button, one can see the consequences of an action immediately.
+However, in congestion control, the rewards are always \textbf{delayed} by one round-trip time.
+\item Due to the delay, by the time an action receives its reward, probably other actions have already been carried out in the meantime.
+Thus, contrary to classical RL, \textbf{action and rewards are not synchronized}.
 \item \textbf{Actions and rewards are not atomic}: If an action causes three packets to be sent, we consider these three packets being sent three \textit{partial actions}.
-In case these packets are transmitted correctly, the receiver will send three acknowledgements.
+In case these packets are transmitted correctly, the receiver will send back three acknowledgements.
 Each of these acknowledgements is a \textit{partial reward} that already unveils new information on the current state of the network and enables the sender to take an action based on this new information.
-However, only if all partial rewards have been received, the sender can assemble the overall reward of the corresponding action and can perform an update of the underlying RL logic.
+However, only if all partial rewards have been received, the sender can assemble the overall reward of the corresponding action and can perform an update of the underlying reinforcement learning logic.
+\item The \textbf{number of actions and rewards} per unit of time is \textbf{variable}: If the congestion window is larger, more packets are sent and more acknowledgements are received, which means that the reinforcement learning algorithm gets more training data the larger the window is.
+This is problematic as it means that -- without proper adjustments -- the experience made at a larger congestion window will ``overwrite'' the experience made at a smaller window as with the larger window simply more training data arrives and it is thus going to be overemphasized.
 \end{itemize}
 
-To address these problems, we develop a new formulation of RL called \textit{Partial Action Learning} (PAL).
-PAL is a superset of reinforcement learning: If one uses PAL for a learning problem without delay, asynchronicity and partial actions/partial rewards, one gets classical RL.
+To address these problems, we develop a new formulation of reinforcement learning called \textit{Partial Action Learning} (PAL).
+PAL is a superset of reinforcement learning: If one uses PAL for a learning problem without delay, asynchronicity and partial actions/partial rewards, one gets classical reinforcement learning.
 
 While using PAL to train an optimum congestion control for a specific range of network scenarios in an off-line fashion is possible, it is something that has already been done previously in an approach called \textit{Remy} \citep{winstein_tcp_2013}.
-The only seizable advantage that PAL could provide here is increased training speed.
-Thus we want to show that it is  not only possible to learn congestion control by using machine learning but that it is even possible to do so on-line and without any preknowledge about the network environment.
+The only potential advantage that PAL could provide here is increased training speed.
+Thus, we want to show that it is  not only possible to learn congestion control by using machine learning but that it is even possible to do so on-line and without any preknowledge about the network environment.
 
-Using a utility function that encourages high throughput and discourages packet loss (adapted from \citep{dong_pcc:_2015}) we see that our machine learning based approach can learn congestion control that maximizes this objective given a couple of minutes of time (without any pre-training) \ref{fig:demo}, which -- to our knowledge -- is the first demonstration that neural network based on-line reinforcement learning is possible.
+Using a utility function that encourages high throughput and discourages packet loss (adapted from \citep{dong_pcc:_2015}) we see that our machine learning approach can learn congestion control that maximizes this objective given a couple of minutes of time (without any pre-training) \ref{fig:demo}.
 
@@
 \begin{minipage}{\columnwidth}
-\includegraphics[width=\columnwidth]{figures/1_4}
-\includegraphics[width=\columnwidth]{figures/2_2}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_1_2018-01-25-20:17:15_3_2018-01-25-20:17:15.0}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/4_2_2018-01-25-19:53:39_22_2018-01-25-19:53:39.0}}}
 \includegraphics[width=\columnwidth]{figures/4_1}
@@
 
-After spending up to a few minutes with suboptimal performance during initial exploration (if no online pre-training was used), our approach can compete with established algorithms and achieve superior throughput with packet loss at an average of less than 5\%, while still gradually improving and adapting its way of performing congestion control.
+After spending up to a few minutes with suboptimal performance during initial exploration (if no pre-training is used), our approach can compete with established algorithms and achieve superior throughput and packet loss of less that 1\%, while still gradually improving and adapting its way of performing congestion control.
 
@@
 These approaches as well as most others maintain a congestion window which stands for the maximum amount of data that are allowed to be traveling in the network without having been acknowledged by the receiver.
-In general, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995} \todo{How much detail is necessary in the related work section?}) lead to the congestion window being lowered according to a set of fixed rules.
+In general, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995}) lead to the congestion window being lowered according to a set of fixed rules.
 Algorithms such as Cubic and Compound \cite{ha_cubic:_2008, tan_compound_2006} improve congestion control specifically with respect to so-called long fat networks with a high round-trip time and bandwidth.
@@
 maximizing throughput while minimizing packet loss).
-However, PCC doesn't maintain a congestion window but instead modulates the sending rate.
+However, PCC doesn't maintain a congestion window but instead modulates the sending rate, using one constant sending rate for a certain time span.
 

### 2018-01-28 async_deep_reinforce/constants.py

@@ -37,6 +37,7 @@ GENERAL_FACTOR = 10**float(os.environ.get('learning_rate')) if os.environ.get('l
 # INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate
 # INITIAL_ALPHA_HIGH = 1e0*GENERAL_FACTOR   # log_uniform high limit for learning rate
 INITIAL_RATE = GENERAL_FACTOR
+print("learning_rate", INITIAL_RATE)
 
 PRECISION = tf.float32
 

### 2018-01-28 tex/paper.tex
Add config files

@@
 \includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/4_2_2018-01-25-19:53:39_22_2018-01-25-19:53:39.0}}}
-\includegraphics[width=\columnwidth]{figures/4_1}
-\caption{\todo{figures don't look that appealing do they?} The throughput (solid line) and the loss rate (dashed line) when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1, 2 and 4 senders (from top to bottom).
-One can see that the time to convergence increases as the number of senders increases, which is caused by the fact that the more senders use the link, the fewer packets each sender receives per unit of time for training.
-Also, the complexity of the learning problem increases with more senders as the environment becomes more dynamic and unpredictable.}
+\caption{The throughput and the loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1 and 2 senders (from top to bottom).
+The time needed for convergence increases with more senders, which is caused by the fact that the more senders use the link, the fewer because each sender receives fewer packets per unit of time for training and because the environment gets more complex.
+It is interesting to note that in the case of two senders, both flows complement each other and quickly adapt to the sending rate of the other flow.}
 \label{fig:demo}
@@
 
-In this section, an example of the overall procedure is outlined.
+In this section, a short example of the overall procedure is outlined.
 
@@
 \item Receive an ACK
-\item Update the internal state to reflect the information that this ACK provides (for example, the state consists of the round trip time experienced by this packet, the time since we receive the last ACK etc.)
-\item Sample an increase for the congestion window from the Actor Network (e.g. $0.24$) and add it to the current congestion window.
+\item Update the internal state to reflect the information that this ACK provides
+\item Sample an increase for the congestion window from the Actor Network and add it to the current congestion window.
 \item 
@@
 a previous action resulted in $3$ packets being sent; so if this is the ACK for the third (and last) packet of this previous action, we got all the ACKs to compute the reward), combine all partial rewards to form the actual reward for that action.
-\item If we now have $t_\text{max}$ rewards (set to 20) then we also update the neural network. 
+\item If there are now enough acknowledgements $t_\text{max}$ rewards (set to 20) then we also update the neural network. 
 \end{enumerate}

### 2018-01-29 tex/paper.tex
Adapt the whole thing once more

@@
   copyright information}
-\subtitle{Extended Abstract}
-\subtitlenote{The full version of the author's guide is available as
-  \texttt{acmart.pdf} document}
+%\subtitle{Extended Abs\texttt{tract}
+%\subtitlenote{The full version of the author's guide is available as
+%  \texttt{acmart.pdf} document}}
 
@@
 
-\subsection{Example}
+%\subsection{Example}
+%
+%In this section, a short example of the overall procedure is outlined. 
+%
+%\begin{enumerate}
+%\item Receive an ACK
+%\item Update 
+%\item Update the internal state to reflect the information that this ACK provides
+%\item Sample an increase for the congestion window from the Actor Network and add it to the current congestion window.
+%\item 
+%\begin{enumerate}
+%\item If this was the last ACK, which completes a previous action (e.g.
+a previous action resulted in $3$ packets being sent; so if this is the ACK for the third (and last) packet of this previous action, we got all the ACKs to compute the reward), combine all partial rewards to form the actual reward for that action.
+%\item If there are now enough acknowledgements $t_\text{max}$ rewards (set to 20) then we also update the neural network. 
+%\end{enumerate}
+%\end{enumerate}
 
-In this section, a short example of the overall procedure is outlined.
+\section{Evaluation}
 
-\begin{enumerate}
-\item Receive an ACK
-\item Update the internal state to reflect the information that this ACK provides
-\item Sample an increase for the congestion window from the Actor Network and add it to the current congestion window.
-\item 
-\begin{enumerate}
-\item If this was the last ACK, which completes a previous action (e.g.
-a previous action resulted in $3$ packets being sent; so if this is the ACK for the third (and last) packet of this previous action, we got all the ACKs to compute the reward), combine all partial rewards to form the actual reward for that action.
-\item If there are now enough acknowledgements $t_\text{max}$ rewards (set to 20) then we also update the neural network. 
-\end{enumerate}
+We implemented TCP Rax as an extension to both the NS-2 network simulator \cite{fall_network_2007} and as well as Remy \cite{winstein_tcp_2013}.  
+In the following evaluation we use a classical dumbbell network topology with one receiver and one or more senders and a shared bottleneck link.
 
-\end{enumerate}
+\subsection{Tradeoff between throughput and packet loss}
 
-\section{Results}
+\begin{figure}
+\begin{minipage}{\columnwidth}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_1}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_2}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4}}}
+\caption{One sender with a bandwidth/packet loss tradeoff of 1, 2 and 4 (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
+\label{fig:one_sender_loss}
+\end{minipage}
+\end{figure}
+
+The tunable tradeoff between bandwidth and packet loss rate ($\alpha$) has a clear impact on both the highest spike of packet loss that is produced in the very beginning of the online learning process as well as on the median packet loss rate later on in the learning process \autoref{fig:one_sender_loss}.
+However, the stronger packet loss is punished, the longer it takes for the sender to reach the maximum bandwidth on the link.
+
+\subsection{Number of senders}
+
+\begin{figure}
+\begin{minipage}{\columnwidth}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/2_4}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_4}}}
+\caption{One sender with a bandwidth/packet loss tradeoff of 1, 2 and 4 (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
+\label{fig:one_sender_loss}
+\end{minipage}
+\end{figure}
 

### 2018-01-29 async_deep_reinforce/constants.py

@@ -33,7 +33,7 @@ LOG_FILE = os.path.join(ABSOLUTE_PATH,'a3c_log')
 logging.info(" ".join(map(str,("LOG_DIR:",LOG_FILE))))
 ACTOR_FACTOR = 1e0
 VALUE_FACTOR = 1e0
-GENERAL_FACTOR = 10**float(os.environ.get('learning_rate')) if os.environ.get('learning_rate') is not None else -4.5
+GENERAL_FACTOR = 10**float(os.environ.get('learning_rate')) if os.environ.get('learning_rate') is not None else -3
 # INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate
 # INITIAL_ALPHA_HIGH = 1e0*GENERAL_FACTOR   # log_uniform high limit for learning rate
 INITIAL_RATE = GENERAL_FACTOR

### 2018-01-29 tex/paper.tex
Changed the thing again

@@
 \begin{minipage}{\columnwidth}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/2_4}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_4}}}
-\caption{One sender with a bandwidth/packet loss tradeoff of 1, 2 and 4 (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_2}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_2}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/new_gpu/no_loss/figures/4_2}}}
+\caption{One, two and eight senders (from top to bottom) with a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
-\label{fig:one_sender_loss}
+\label{fig:multiple_senders}
 \end{minipage}
 \end{figure}
+
+When comparing TCP Rax as the only sender on a link compared to sharing a link with more than one
 

### 2018-01-29 async_deep_reinforce/a3c_training_thread.py

@@ -14,7 +14,7 @@ from functools import reduce
 from game_ac_network import GameACLSTMNetwork
 
 # gamma_current, gamma_future = 1./(1+GAMMA), GAMMA/(1.+GAMMA)
-from constants import LOCAL_T_MAX
+# from constants import LOCAL_T_MAX
 from constants import LOG_LEVEL
 from constants import STATE_SIZE
 from constants import SIGMOID_ALPHA

### 2018-01-29 async_deep_reinforce/constants.py

@@ -31,7 +31,7 @@ logging.info(" ".join(map(str,("CHECKPOINT_DIR:",CHECKPOINT_DIR))))
 
 LOG_FILE = os.path.join(ABSOLUTE_PATH,'a3c_log')
 logging.info(" ".join(map(str,("LOG_DIR:",LOG_FILE))))
-ACTOR_FACTOR = 1e0
+ACTOR_FACTOR = 1e-1
 VALUE_FACTOR = 1e0
 GENERAL_FACTOR = 10**float(os.environ.get('learning_rate')) if os.environ.get('learning_rate') is not None else -3
 # INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate

### 2018-01-29 tex/paper.tex
Fix stuff

--- 2018-01-29 tex/paper.tex
@@
 \includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/2_4}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_4}}}
-\caption{One sender with a bandwidth/packet loss tradeoff of 1, 2 and 4 (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/new_gpu/no_loss/figures/4_4}}}
+\caption{One, two and four senders (from top to bottom) with a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
-\label{fig:one_sender_loss}
+\label{fig:multiple_senders}
 \end{minipage}
 \end{figure}
+
+While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations clearly increase when having two or four concurrent senders \autoref{fig:multiple_senders}.
+We attribute this to the fact that on a shared link $n$ can on average only get $\frac{1}{n}$ of the bandwidth and thus each sender receives fewer packets which he can use for training.
+Furthermore, as the environment becomes more dynamic with more senders, each sender has a harder time learning whether the reward changed due to his own actions or due to actions that were caused by other senders: If there is only one flow on a link, the flow ``can be sure'' that all changes in rewards were actually caused by his own actions.
+
+\subsection{Pre-training}
+
+\begin{figure}
+\begin{minipage}{\columnwidth}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4_again}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4_again}}}
+\caption{One and two senders (from top to bottom) that were trained with a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
+\label{fig:resuming_training}
+\end{minipage}
+\end{figure}
+
+To see whether pre-training is generally feasible, we used the neural networks that were produced after each 600 seconds flow shown in figure \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks.
+The flows seem to simply resume where they ended training in the flow before.
+This does not seem like a remarkable finding, however, this is unexpected: For instance, in the case of one sender on a link, the neural network starts with random weights and discovers that it achieves optimum utility at a window that results in 20\,Mbit/s of bandwidth.
+Then the sender constantly stays in this state and one would expect it to slowly ``forget'' what it learned in the very beginning when it started the flow.
+Thus it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again.
+
+\subsection{Stochastic Loss}
+
+\begin{figure}
+\begin{minipage}{\columnwidth}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-5.local/figures/1_4}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-7.local/figures/2_4}}}
+\caption{One and two senders (from top to bottom) on a link with a stochastic loss rate of 1\%, a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
+\label{fig:stochastic_loss}
+\end{minipage}
+\end{figure}
+
+To see whether pre-training is generally feasible, we used the neural networks that were produced after each 600 seconds flow shown in figure \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks.
+The flows seem to simply resume where they ended training in the flow before.
+This does not seem like a remarkable finding, however, this is unexpected: For instance, in the case of one sender on a link, the neural network starts with random weights and discovers that it achieves optimum utility at a window that results in 20\,Mbit/s of bandwidth.
+Then the sender constantly stays in this state and one would expect it to slowly ``forget'' what it learned in the very beginning when it started the flow.
+Thus it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again.
 

### 2018-01-29 async_deep_reinforce/constants.py

@@ -31,9 +31,9 @@ logging.info(" ".join(map(str,("CHECKPOINT_DIR:",CHECKPOINT_DIR))))
 
 LOG_FILE = os.path.join(ABSOLUTE_PATH,'a3c_log')
 logging.info(" ".join(map(str,("LOG_DIR:",LOG_FILE))))
-ACTOR_FACTOR = 1e-1
+ACTOR_FACTOR = 1e0
 VALUE_FACTOR = 1e0
-GENERAL_FACTOR = 10**float(os.environ.get('learning_rate')) if os.environ.get('learning_rate') is not None else -3
+GENERAL_FACTOR = 10**float(os.environ.get('learning_rate')) if os.environ.get('learning_rate') is not None else -4
 # INITIAL_ALPHA_LOW = 1e-2*GENERAL_FACTOR   # log_uniform low limit for learning rate
 # INITIAL_ALPHA_HIGH = 1e0*GENERAL_FACTOR   # log_uniform high limit for learning rate
 INITIAL_RATE = GENERAL_FACTOR

### 2018-01-29 src/unicorn.cc
Changed Unicorn

@@ -51,7 +51,7 @@ bool Unicorn::packets_lost( const std::vector< remy::Packet > & packets ) {
         _packets_lost += 1;
         _memory.lost(1);
         _memory.window(_the_window);
-        // _the_window = std::max(_the_window - 1, MIN_WINDOW_UNICORN);
+        _the_window = std::max(_the_window - 1, MIN_WINDOW_UNICORN);
         get_action(packet.tick_received);
         _largest_ack += 1;
         return true;

### 2018-01-29 async_deep_reinforce/a3c_training_thread.py
Other modification

@@ -14,7 +14,8 @@ from functools import reduce
 from game_ac_network import GameACLSTMNetwork
 
 # gamma_current, gamma_future = 1./(1+GAMMA), GAMMA/(1.+GAMMA)
-# from constants import LOCAL_T_MAX
+from constants import LOCAL_T_MAX
+from constants import GAMMA
 from constants import LOG_LEVEL
 from constants import STATE_SIZE
 from constants import SIGMOID_ALPHA
@@ -367,7 +368,7 @@ class A3CTrainingThread(object):
       # assert(False)
       # The GAMMA_FACTOR increases the influence that following observations have on this one.
 
-      GAMMA = (1 - 2/(A3CTrainingThread.get_actual_window(wi+ai) + 1))
+      # GAMMA = (1 - 2/(A3CTrainingThread.get_actual_window(wi+ai) + 1))
 
       # GAMMA = 0.99
 

### 2018-01-29 async_deep_reinforce/constants.py

@@ -76,6 +76,8 @@ SENT_OFFSET = BIAS_OFFSET
 # INTER_PACKET_ARRIVAL_TIME_OFFSET = inverse_softplus(DELAY/(1-GAMMA))
 # LOST_OFFSET = inverse_softplus(1e-10/(1-GAMMA))
 
+GAMMA = 0.99
+
 INITIAL_WINDOW_INCREASE_BIAS_OFFSET = 0
 INITIAL_WINDOW_INCREASE_WEIGHT_FACTOR = 1e-2
 PACKET_FACTOR = 1e-2

### 2018-01-29 src/unicorn.cc

@@ -51,7 +51,7 @@ bool Unicorn::packets_lost( const std::vector< remy::Packet > & packets ) {
         _packets_lost += 1;
         _memory.lost(1);
         _memory.window(_the_window);
-        _the_window = std::max(_the_window - 1, MIN_WINDOW_UNICORN);
+        // _the_window = std::max(_the_window - 1, MIN_WINDOW_UNICORN);
         get_action(packet.tick_received);
         _largest_ack += 1;
         return true;

### 2018-01-29 async_deep_reinforce/a3c_training_thread.py
Fix bug that makes things segfault

@@ -340,9 +340,11 @@ class A3CTrainingThread(object):
     R_packets, R_duration, R_sent = R_packets, 1/R_duration, R_sent
     # R_packets, R_accumulated_delay, R_duration, R_sent = (R_packets)/(1-GAMMA), (R_accumulated_delay)/(1-GAMMA), (R_duration)/(1-GAMMA), (R_sent)/(1-GAMMA)
     # logging.debug(" ".join(map(str,("exp(R_packets)", R_packets, "exp(R_accumulated_delay)", R_accumulated_delay, "exp(R_duration)", R_duration))))
-    if not (np.isfinite(R_duration) and R_duration > 0):
+    if not (R_duration > 0):
       print("R_duration", R_duration)
-    assert(np.isfinite(R_duration) and R_duration > 0)
+    if (np.isfinite(R_duration)):
+      R_duration = 0.0
+    assert(np.isfinite(R_duration)) # Pretty dumb
     assert(np.isfinite(R_packets))
     # assert(np.isfinite(R_accumulated_delay))
     assert(np.isfinite(R_sent))

### 2018-01-29 tex/paper.tex
Fix bug

@@
 While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations clearly increase when having two or four concurrent senders \autoref{fig:multiple_senders}.
-We attribute this to the fact that on a shared link $n$ can on average only get $\frac{1}{n}$ of the bandwidth and thus each sender receives fewer packets which he can use for training.
+We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which he can use for training.
 Furthermore, as the environment becomes more dynamic with more senders, each sender has a harder time learning whether the reward changed due to his own actions or due to actions that were caused by other senders: If there is only one flow on a link, the flow ``can be sure'' that all changes in rewards were actually caused by his own actions.
@@
 
-To see whether pre-training is generally feasible, we used the neural networks that were produced after each 600 seconds flow shown in figure \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks.
+To verify whether pre-training is generally feasible, we used the neural networks that were produced after each 600 seconds flow shown in figure \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks.
 The flows seem to simply resume where they ended training in the flow before.
@@
 Then the sender constantly stays in this state and one would expect it to slowly ``forget'' what it learned in the very beginning when it started the flow.
-Thus it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again.
+Thus, it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again in a completely different state than it is used to.
 

### 2018-01-29 async_deep_reinforce/a3c_training_thread.py

@@ -342,7 +342,7 @@ class A3CTrainingThread(object):
     # logging.debug(" ".join(map(str,("exp(R_packets)", R_packets, "exp(R_accumulated_delay)", R_accumulated_delay, "exp(R_duration)", R_duration))))
     if not (R_duration > 0):
       print("R_duration", R_duration)
-    if (np.isfinite(R_duration)):
+    if not np.isfinite(R_duration):
       R_duration = 0.0
     assert(np.isfinite(R_duration)) # Pretty dumb
     assert(np.isfinite(R_packets))

### 2018-01-29 tex/paper.tex
One more fix

@@
 \begin{itemize}
-\item the time between the last two packets that were sent
 \item the time between the last two packets that were received
@@
 
+It might seem simplistic to only use an additive increase to the congestion window and no more sophisticated mechanisms such as a combination of a multiplicative increase and an additive increase.
+However, as we use GRUs in the actor network, which have a non-linear output function (a hyperbolic tangent function), they can learn to become very flexible regarding their output.
+
 The actor network minimizes the loss function
@@
 
-We implemented TCP Rax as an extension to both the NS-2 network simulator \cite{fall_network_2007} and as well as Remy \cite{winstein_tcp_2013}.  
+We implemented TCP Rax as an extension to both the NS-2 network simulator \cite{fall_network_2007} and Remy \cite{winstein_tcp_2013}.  
 In the following evaluation we use a classical dumbbell network topology with one receiver and one or more senders and a shared bottleneck link.
@@
 
-The tunable tradeoff between bandwidth and packet loss rate ($\alpha$) has a clear impact on both the highest spike of packet loss that is produced in the very beginning of the online learning process as well as on the median packet loss rate later on in the learning process \autoref{fig:one_sender_loss}.
+The tunable tradeoff between bandwidth and packet loss rate ($\alpha$) has a clear impact on both the highest spike of packet loss that is produced in the very beginning of the online learning process as well as on the median packet loss rate later on in the learning process (see \autoref{fig:one_sender_loss}).
 However, the stronger packet loss is punished, the longer it takes for the sender to reach the maximum bandwidth on the link.
+
+The apparent spike of packet loss in the beginning is the following: In the very beginning the reinforcement learning learns that increasing the window is always good (as it is still below the bottleneck bandwidth).
+This leads to exponential growth of the congestion window in the beginning and as soon as the bottleneck bandwidth is reached, the reinforcement learning needs to correct its behavior to realize that increasing the congestion window is not always the appropriate action to perform.
+Potential countermeasure for this are to use pre-training (see \autoref{subsec:pretraining}) and to use a utility function that heavily penalizes excessive packet loss.
 
@@
 
-While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations clearly increase when having two or four concurrent senders \autoref{fig:multiple_senders}.
+While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations clearly increase when having two or four concurrent senders (see \autoref{fig:multiple_senders}).
 We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which he can use for training.
@@
 \subsection{Pre-training}
+\label{subsec:pretraining}
 
@@
 The flows seem to simply resume where they ended training in the flow before.
-This does not seem like a remarkable finding, however, this is unexpected: For instance, in the case of one sender on a link, the neural network starts with random weights and discovers that it achieves optimum utility at a window that results in 20\,Mbit/s of bandwidth.
+This does not seem like a remarkable finding, however, it is quite unexpected: For instance, in the case of one sender on a link, the neural network starts with random weights and discovers that it achieves optimum utility at a window that results in 20\,Mbit/s of bandwidth.
 Then the sender constantly stays in this state and one would expect it to slowly ``forget'' what it learned in the very beginning when it started the flow.
-Thus, it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again in a completely different state than it is used to.
+Thus, it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again in a completely different state than the one it is used to.
 
@@
 \caption{One and two senders (from top to bottom) on a link with a stochastic loss rate of 1\%, a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
-Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
+Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.
+It looks as if there is no stochastic loss in the very beginning of the flows, however, this is an artifact of the binning procedure: As the window is very small in the beginning, there are more flows in which no packet loss occurs whatsoever and thus the median is distorted.}
 \label{fig:stochastic_loss}
@@
 
-To see whether pre-training is generally feasible, we used the neural networks that were produced after each 600 seconds flow shown in figure \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks.
-The flows seem to simply resume where they ended training in the flow before.
-This does not seem like a remarkable finding, however, this is unexpected: For instance, in the case of one sender on a link, the neural network starts with random weights and discovers that it achieves optimum utility at a window that results in 20\,Mbit/s of bandwidth.
-Then the sender constantly stays in this state and one would expect it to slowly ``forget'' what it learned in the very beginning when it started the flow.
-Thus it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again.
-
-\section{Discussion}
+Stochastic loss does not substantially affect TCP Rax.
+This can be attributed to the fact that the reinforcement learning gets to understand that a certain minimum packet loss rate is ``normal'' and cannot be avoided and thus TCP Rax is flexible in this regard.
 
@@
 
+The goal of this paper was to determine whether it is possible to use deep reinforcement learning to learn a per-acknowledgement congestion control online.
+To this end we had to modify existing reinforcement learning algorithms to accommodate the specific needs of congestion control such as the inherent delay amongst others.
+
+We developed TCP Rax and the results of our evaluation indicate that congestion control can be learned online in principle and has the potential to outperform standard TCP variants given long enough flows.
+We came to the insight that it is possible to pre-train the neural network used in the reinforcement learning and that this improves performance especially in the beginning of the flow.
+
+Some interesting questions, however, remain.
+In what way can pre-training be utilized to compete with other TCP congestion control variants even for very short flows?
+Is it feasible and sensible that one computer uses a set of shared weights for all its flows?
+
+\todo{Ist dieser Teil mit offenen Fragen sinnlos? Hab das schon in anderen Papers gesehen...}
+
 \section*{Appendix}
 
-\begin{algorithm}[H]
+The algorithm for Partial Action Learning is described in a generic formulation that can also be used for other reinforcement learning problems with similar characteristics such as delay and asynchronous actions and rewards in \autoref{alg:pal}.
+We use custom functions to determine when to update the neural network weights and when to compute the gradients and apply them in \autoref{alg:update} and \autoref{alg:enough}.
+Intuitively, these functions ensure that the neural network gets updated approximately every round-trip time.
+For other reinforcement learning problems it is also possible to simply use a fixed step size such as proposed by \cite{mnih_asynchronous_2016}.
+
+\begin{algorithm}
 \caption{Partial Action Learning -- pseudocode for each agent.
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}
 \caption{Partial Action Learning -- procedure which computes and applies the gradients.}
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}
 \caption{Procedure which determines when to synchronize the neural network weights from the global weights in case of congestion control.
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}
 \caption{Procedure which determines when to start computing the gradients in case of congestion control.

### 2018-01-29 async_deep_reinforce/a3c_training_thread.py

@@ -14,8 +14,8 @@ from functools import reduce
 from game_ac_network import GameACLSTMNetwork
 
 # gamma_current, gamma_future = 1./(1+GAMMA), GAMMA/(1.+GAMMA)
-from constants import LOCAL_T_MAX
-from constants import GAMMA
+# from constants import LOCAL_T_MAX
+# from constants import GAMMA
 from constants import LOG_LEVEL
 from constants import STATE_SIZE
 from constants import SIGMOID_ALPHA
@@ -370,7 +370,7 @@ class A3CTrainingThread(object):
       # assert(False)
       # The GAMMA_FACTOR increases the influence that following observations have on this one.
 
-      # GAMMA = (1 - 2/(A3CTrainingThread.get_actual_window(wi+ai) + 1))
+      GAMMA = (1 - 2/(A3CTrainingThread.get_actual_window(wi+ai) + 1))
 
       # GAMMA = 0.99
 

### 2018-01-29 tex/paper.tex
Add new config file

@@
 
-\subsection{Best practices for on-line reinforcement learning}
+\subsection{Comparing to other TCP congestion control variants}
+
+Again we take the scenario of a 20\,Mbit/s shared bottleneck with a small buffer of $\frac{1}{10}$ bandwidth delay product to create a challenging scenario.
+
+\begin{figure}
+\begin{minipage}{\columnwidth}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_1_0}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_2_0}}}
+\caption{One and two senders (from top to bottom) on a link with a stochastic loss rate of 1\%, a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.
+It looks as if there is no stochastic loss in the very beginning of the flows, however, this is an artifact of the binning procedure: As the window is very small in the beginning, there are more flows in which no packet loss occurs whatsoever and thus the median is distorted.}
+\label{fig:regular_tcps}
+\end{minipage}
+\end{figure}
 

### 2018-01-29 tex/paper.tex
Add new config file

@@
 \includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_2_0}}}
-\caption{One and two senders (from top to bottom) on a link with a stochastic loss rate of 1\%, a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
-Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.
-It looks as if there is no stochastic loss in the very beginning of the flows, however, this is an artifact of the binning procedure: As the window is very small in the beginning, there are more flows in which no packet loss occurs whatsoever and thus the median is distorted.}
+\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one and two senders (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+For a flow of 600\,s the maximum possible throughput is 12\,000\,Mbit}
 \label{fig:regular_tcps}

### 2018-01-29 tex/paper.tex
Add new config file

@@
 \title{TCP Rax: Deep Reinforcement Learning for TCP Congestion Control}
-\titlenote{Produces the permission block, and
-  copyright information}
+%\titlenote{Produces the permission block, and
+%  copyright information}
 %\subtitle{Extended Abs\texttt{tract}
@@
 
-In an ideal machine learning based congestion control method, each sender uses a set of observed environment conditions to determine which action he should take to maximize his reward in the future.
+In a machine learning based congestion control method, each sender uses a set of observed environment conditions to determine which action he should take to maximize his reward in the future.
 The environment conditions can be any metrics that the sender can obtain from the network, for example the mean round-trip time of the last received packets, the packet loss rate etc.
@@
 
-Using a utility function that encourages high throughput and discourages packet loss (adapted from \citep{dong_pcc:_2015}) we see that our machine learning approach can learn congestion control that maximizes this objective given a couple of minutes of time (without any pre-training) \ref{fig:demo}.
+Using a utility function that encourages high throughput and discourages packet loss (adapted from \citep{dong_pcc:_2015}) we see that our machine learning approach can learn congestion control that maximizes this objective given a couple of minutes of time (without any pre-training) (see \autoref{fig:demo}).
 
@@
 \includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/4_2_2018-01-25-19:53:39_22_2018-01-25-19:53:39.0}}}
-\caption{The throughput and the loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1 and 2 senders (from top to bottom).
+\caption{Throughput and loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1 and 2 senders (from top to bottom).
 The time needed for convergence increases with more senders, which is caused by the fact that the more senders use the link, the fewer because each sender receives fewer packets per unit of time for training and because the environment gets more complex.
@@
 In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for playing video games and also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017} \todo{This motivation doesn't sound convincing yet, does it?}.
-Furthermore, as we will show, it is conceptually possible to use it for on-line training although -- to our knowledge -- this has not been done until now.
+Furthermore, as we will show, it is conceptually possible to use it for online training.
 
@@
 We use the mean $\mu$ and the standard deviation $\sigma$ to parametrize a normal distribution.
-The main idea is that the network learns to output the right mean at the right time step to maximize the future reward and that it uses the standard deviation to try out new actions, which could yield a better than expected reward.
+The main idea is that the neural network learns to output the right mean at the right time step to maximize the future reward and that it uses the standard deviation to try out new actions, which could yield a better than expected reward.
 
@@
 
-The motivation for PAL is that classical reinforcement learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
+%As we already mentioned, the motivation for PAL is that classical reinforcement learning assumes that a reward follows an action and vice-versa (see \autoref{fig:reinforcement}).
 \todo{Too much repetition of this?} However, in the case of congestion control, it is desirable to perform a new action without having received a reward for the previous action.
@@
 
-To use PAL for congestion control we first have to define the correct semantics for this specific use case and we have to explicitly state how the state, reward etc.~are defined.
-Furthermore, we have to define how the actor and critic network explicitely work in case of congestion control.
+To use PAL for congestion control we first have to define the correct semantics for this specific use case and define what makes up the state and the reward.
+Furthermore, we have to define how the actor and critic network explicitly work in case of congestion control.
 In the following a time step $t$ corresponds to the reception of an acknowledgement.
@@
 \item the RTT of the last received packet
-\item whether the last packet was lost where $1$ signifies that the last packet was lost while $0$ indicates that it was received correctly
+\item whether the last packet was lost, where $1$ signifies that the last packet was lost while $0$ indicates that it was received correctly
 \item the current congestion window as a real number
 \end{itemize}
-Of each of these features a exponentially weighted moving average with a factor $\alpha$ of $\frac{1}{8}$ and one with an $\alpha$ of $\frac{1}{256}$ (the concept of using ewmas with these specific values of $\alpha$ was first introduced in \textit{Remy} \cite{winstein_tcp_2013}), which makes a total of 10 features.
+Of each of these features a exponentially weighted moving average (ewma) with a factor $\alpha$ of $\frac{1}{8}$ and one with an $\alpha$ of $\frac{1}{256}$ (the concept of using ewmas with these specific values of $\alpha$ was first introduced in \textit{Remy} \cite{winstein_tcp_2013}), which makes a total of 10 features.
 Each time an acknowledgement is received, the state is updated and the actor network is asked for the next action.
 \item[$\textit{a}_t$] Based on a given state and the history of previous states (because we use Gated Recurrent Units (GRUs) and thus can also consider previous states), the actor network returns an action $a_t$, which is a real number that stands for the change of the congestion window.
-\item[$\textit{r}_t$] The reward is a tuple of five reward metrics. 
+
+It might seem simplistic to only use an additive increase to the congestion window and no more sophisticated mechanisms such as a combination of a multiplicative increase and an additive increase.
+However, as we use GRUs in the actor network, which have a nonlinear output function (a hyperbolic tangent function), they can learn to model nonlinear relationships between the input and the output.
+\item[$\textit{r}_t$] The reward is a tuple of three reward metrics. 
 %These are discussed in more detail in \autoref{subsubsec:critic}. 
@@
 \item[$\textit{r}_{\text{duration},t}$] is the time between receiving the last packet and receiving this packet (``inter acknowledgement reception time'') summed over all packets that the sender sent and that were not lost.
-Example: The sender sent three packets, of which two were not lost.
-The time between receiving the ACK of the first packet that wasn't lost and the previous one is 5\,ms and the time between receiving the ACK of the second packet that wasn't lost and the previous ACK is 5\,ms.
+Example: The sender sent three packets, two of which two were not lost.
+The time between receiving the ACK of the first packet that was not lost and the previous one is 5\,ms and the time between receiving the ACK of the second packet that was not lost and the previous ACK is 6\,ms.
 So for this time step $\textit{r}_{\text{duration},t}$ is 11\,ms.
@@
 
-In the case of congestion control, the loss function $l_{\text{v},t}$ of the critic network is actually the sum of the squares of the difference for each of the expected long-term averages and the empirically found averages for each reward metric:
+In case of congestion control, the loss function $l_{\text{v},t}$ of the critic network is actually the sum of the squares of the difference for each of the expected long-term averages and the empirically found averages for each reward metric:
 \begin{align*}
@@
 
-We use 10 as a minimum because when $\gamma$ becomes too big, future rewards is not considered sufficiently anymore and as the variability increases instability ensues.
+We use 10 as a minimum because when $\gamma$ becomes too big, future rewards is not considered sufficiently anymore and as the ensuing variability causes instability.
 
@@
 At the beginning of a flow, the window starts as $1$ too.
-
-It might seem simplistic to only use an additive increase to the congestion window and no more sophisticated mechanisms such as a combination of a multiplicative increase and an additive increase.
-However, as we use GRUs in the actor network, which have a non-linear output function (a hyperbolic tangent function), they can learn to become very flexible regarding their output.
 
@@
 where $U$ can be any utility function defined based on some of the (long term averages of the) reward metrics.
-Thus, in simple terms, the actor network considers if an action improved the actual experienced utility was compared to the expected one and adjusts the neural network accordingly.
+Thus, in simple terms, the actor network considers if an action improved the actual experienced utility compared to the expected one and adjusts the neural network accordingly.
 
@@
 
-The equivalent utility is also defined using the estimates of the value networks and it can be determined whether an action achieved a performance that was higher or lower than expected.
+The corresponding expected utility is also defined using the estimates of the value networks and it can be determined whether an action achieved a performance that was higher or lower than expected.
 
@@
 In the following evaluation we use a classical dumbbell network topology with one receiver and one or more senders and a shared bottleneck link.
+In all figures in the evaluation we use at least 50 data unless stated otherwise.
 
@@
 
-The apparent spike of packet loss in the beginning is the following: In the very beginning the reinforcement learning learns that increasing the window is always good (as it is still below the bottleneck bandwidth).
+The apparent spike of packet loss in the beginning can be explained as follows: In the very beginning the reinforcement learning learns that increasing the window is always good (as it is still below the bottleneck bandwidth).
 This leads to exponential growth of the congestion window in the beginning and as soon as the bottleneck bandwidth is reached, the reinforcement learning needs to correct its behavior to realize that increasing the congestion window is not always the appropriate action to perform.
@@
 To verify whether pre-training is generally feasible, we used the neural networks that were produced after each 600 seconds flow shown in figure \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks.
-The flows seem to simply resume where they ended training in the flow before.
+The flows seem to simply resume where they ended training in the flow before (see \autoref{fig:resuming_training}).
 This does not seem like a remarkable finding, however, it is quite unexpected: For instance, in the case of one sender on a link, the neural network starts with random weights and discovers that it achieves optimum utility at a window that results in 20\,Mbit/s of bandwidth.
@@
 
-\subsection{Comparing to other TCP congestion control variants}
+\subsection{Comparison with other TCP congestion control variants}
 
 Again we take the scenario of a 20\,Mbit/s shared bottleneck with a small buffer of $\frac{1}{10}$ bandwidth delay product to create a challenging scenario.
+Surprisingly New Reno performs by far better than more modern TCP congestion control algorithms (see \autoref{fig:regular_tcps}).
+TCP Rax achieves a significantly larger total throughput in the median though with a couple of outliers which become more significant with two concurrent senders (see \autoref{fig:rax_tcps}).
 
@@
 \caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one and two senders (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
-For a flow of 600\,s the maximum possible throughput is 12\,000\,Mbit}
+For a flow of 600\,s the maximum possible throughput is 12\,000\,Mbit.}
 \label{fig:regular_tcps}
@@
 
+\begin{figure}
+\begin{minipage}{\columnwidth}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/all_together}}}
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/all_together}}}
+\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one and two senders (from top to bottom) for TCP Rax on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+The arrow symbol indicates that the flows use a pre-trained neural network.
+For a flow of 600\,s the maximum possible throughput is 12\,000\,Mbit.}
+\label{fig:rax_tcps}
+\end{minipage}
+\end{figure}
+
+\todo{Add some kind of table with other scenarios}
+
 \section{Conclusion}
@@
 The goal of this paper was to determine whether it is possible to use deep reinforcement learning to learn a per-acknowledgement congestion control online.
-To this end we had to modify existing reinforcement learning algorithms to accommodate the specific needs of congestion control such as the inherent delay amongst others.
+To this end we had to modify existing reinforcement learning algorithms to accommodate the specific needs of congestion control such as the inherent delay.
 
@@
 
-The algorithm for Partial Action Learning is described in a generic formulation that can also be used for other reinforcement learning problems with similar characteristics such as delay and asynchronous actions and rewards in \autoref{alg:pal}.
+The algorithm for Partial Action Learning is described in \autoref{alg:pal} in a generic formulation that can also be used for other reinforcement learning problems with similar characteristics such as delay and asynchronous actions and rewards.
 We use custom functions to determine when to update the neural network weights and when to compute the gradients and apply them in \autoref{alg:update} and \autoref{alg:enough}.

### 2018-01-30 tex/paper.tex

@@
 %\setcopyright{acmlicensed}
-\setcopyright{rightsretained}
+%\setcopyright{rightsretained}
 %\setcopyright{usgov}
@@
   Paso, Texas USA} 
-\acmYear{2017}
-\copyrightyear{2017}
+\acmYear{2018}
+\copyrightyear{2018}
 
-\acmPrice{15.00}
+%\acmPrice{15.00}
 
@@
 
-The Actor Network outputs a probability distribution from which the action $a_t$ at time step $t$ is randomly sampled.
-We use the mean $\mu$ and the standard deviation $\sigma$ to parametrize a normal distribution.
+The Actor Network outputs a normal distribution parametrized by a mean $\mu$ and a standard deviation $\sigma$ from which the action $a_t$ at time step $t$ is randomly sampled.
 The main idea is that the neural network learns to output the right mean at the right time step to maximize the future reward and that it uses the standard deviation to try out new actions, which could yield a better than expected reward.
@@
 Each time we take an action $a_t$ we also let the critic network make an estimation $V(s_t; \theta)$, which we abbreviate as $v_t$, given the current neural network weights $\theta$ and the current state $s_t$.
-The loss function of the actor network can now be described as this: How likely was it (on a logarithmic scale) to take action $a_t$ given the probability distribution at time $t$ and how beneficial was this action when comparing to the critic's estimation?
-For example, if an action was very successful and highly unlikely, the actor network's are updated to make actions like this more likely in the future.
+The loss function of the actor network can now be described as this: How likely was it to take action $a_t$ given the probability distribution at time $t$ and how beneficial was this action when comparing it to the critic's estimation?
+For example, if an action was very successful and highly unlikely, the actor network's weights are updated to make actions like this more likely in the future.
 
-The actor also aims at increasing the entropy of the probability density function ($\frac{1}{2} \log \left( 2 \pi e \sigma^2 \right)$) to encourage exploration: Otherwise the actor could always output similar actions but never actually take the action that yields the best performance.
+The actor also aims at increasing the entropy of the probability density function ($\frac{1}{2} \log \left( 2 \pi e \sigma^2 \right)$) to encourage exploration: Otherwise the actor could always perform similar actions but never actually take the action that yields the best performance.
 We designate the entropy as $H$ and introduce a parameter $\beta$ which specifies magnitude of the influence of the entropy.
@@
 The key difference between PAL and previous approaches to RL is that in classical RL an action is always followed by a reward and a reward is always followed by an action.
-In our proposed concept, however, it is possible to take new actions while previous actions haven't received their rewards yet.
+In our proposed concept, however, it is possible to take new actions while previous actions haven't been rewarded yet.
 
@@
 \item[$\textit{r}_{\text{received},t}$] is the sum of all bytes in all packets that the sender sent during time step $t$ and that were not lost (so they were acknowledged at some point by the receiver).
+Example: During time step $t$, 2500 bytes are sent, of which 1250 are lost and 1250 are acknowledged at some point in time later.
+Then $\textit{r}_{\text{received},t}$ is 1250 for at time step $t$.
+However, $\textit{r}_{\text{received},t}$ can only be determined at some point in time in the future, when the sender will have received the acknowledgements for all the bytes that he sent.
 %\item[$\textit{r}_{\text{byte},t}$] is the sum of the bytes of the packets that the sender sent and that were not lost.
@@
 			\State $w\gets \text{w} - l_\text{sent}[j]$
-			\If{$w\leq 0 \land i+1 \geq 10$} \Comment{at least 10 rewards}
+			\If{$w\leq 0 \land i+1 \geq 10$} \Comment{at least 10}
 				\State $i \gets j$
@@
 \caption{Procedure which determines when to start computing the gradients in case of congestion control.
-We assume that the window at time step $t$ is stored in a list $l_\text{windows}$ and that the number of bytes sent in each time step is stored in $l_\text{sent}$.
-Alternatively to this procedure one can simply compute the gradients when $t_\text{max}$ rewards have been received where $t_\text{max}$ is an arbitrarily chosen integer greater than 0.}
+We assume that the window at time steps $t$ is stored in a list $l_\text{windows}$ and that the number of bytes sent in each time step is stored in $l_\text{sent}$ analogously to the other metrics defined in \autoref{alg:pal}.
+Alternatively to this procedure, one can simply compute the gradients when $t_\text{max}$ rewards have been received where $t_\text{max}$ is an arbitrarily chosen integer greater than 0.}
 \label{alg:enough}
@@
 		\State $w \gets \text{w} - l_\text{sent}[i]$
-		\If{$w \leq 0 \land i+1 \geq 10$} \Comment{at least 10 rewards}
+		\If{$w \leq 0 \land i+1 \geq 10$} \Comment{at least 10}
 			\State \Return \True

### 2018-01-31 tex/paper.tex
Update tex and bib tex

@@
-\documentclass[sigconf]{acmart}
+\documentclass[newfonts=false,format=sigconf,10pt,letterpaper]{acmart}
+
+\usepackage[english]{babel}
+\usepackage{subfig}
+
+% Copyright
+\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
+\setcopyright{none}
+
+\settopmatter{printacmref=false, printccs=false, printfolios=true}
 
@@
 
-A major challenge in congestion control is to find a set of rules that dynamically adapt the congestion window to current and anticipated future network conditions to maximize a specified utility function, e.g.
+A major challenge in TCP congestion control is to find a set of rules that dynamically adapt the congestion window to current and anticipated future network conditions to maximize a specified utility function, e.g.
 maximum throughput or minimum packet loss.
 
-This paper proposes Reactive Adaptive eXperience based congestion control (TCP Rax), a new method of congestion control that uses online reinforcement learning to maintain an optimum congestion window with respect to a given utility function.
+This paper proposes Reactive Adaptive eXperience based congestion control (TCP Rax), a new method of congestion control that uses online reinforcement learning to maintain an optimum congestion window with respect to a given utility function and based on current network conditions.
 We use a neural network based approach that can be initialized either with random weights (without prior knowledge) or with a previously trained neural network to improve stability and convergence time.
 The processing of rewards in congestion control depends on the arrival of acknowledgments, inducing loop delays that may lead to oscillations.
-As a remedy we propose a specific formulation of neural network based reinforcement learning that supports the proactive processing of delayed and partial rewards.
+As a remedy we propose \textit{Partial Action Learning}, a specific formulation of neural network based reinforcement learning that supports the processing of delayed and partial rewards.
 
@@
 
-Recent advances in neural network based reinforcement learning (RL) raise the question if it is possible to use these frameworks to model end-to-end congestion control.
-In general, in congestion control each sender aims to maximize an objective function that usually comprises of
+Recent advances in neural network based reinforcement learning (RL) that for the first time have the ability to learn complex tasks such as video games, raise the question if it is possible to use these frameworks to model end-to-end congestion control: As in RL, in congestion control each sender has a an observable state, performs actions and receives rewards for its actions according to the sender's objective function.
+
+In general, in congestion control each sender aims to maximize an objective function that usually comprises one or more of the following subgoals:
 \begin{enumerate*}
-\item sending as many packets as possible
+\item Send as many packets as possible
 \item with as little delay as possible
@@
 As, for instance, for some applications overall throughput is more important than little packet loss, each application can define its preference using a custom utility function.
-Clearly, it is also desirable from an overall point of view that all senders use available resources in a way that is fair to the others: When several senders share an Internet link under the same conditions, they should receive an equal share of the reward.
+Besides aiming to fulfill these subgoals, it is also desirable from an overall point of view that all senders use available resources in a way that is fair to the others: When several senders share an Internet link under the same conditions, they should receive an equal share of the available resources.
 To achieve fairness, it is necessary that senders only use utility functions that
 \begin{enumerate*}
-\item converge to a stable equilibrium and
-\item provide each sender with a fair share of his subgoals
+\item do not give an unfair advantage over others
+\item and converge to a stable equilibrium
 \end{enumerate*}.
@@
 
-While congestion control seems to be a well suited problem for reinforcement learning, there are a couple of challenges and differences from most other reinforcement learning problems:  \begin{itemize}
-\item For instance, when playing a video game, if one pushes a button, one can see the consequences of an action immediately.
-However, in congestion control, the rewards are always \textbf{delayed} by one round-trip time.
-\item Due to the delay, by the time an action receives its reward, probably other actions have already been carried out in the meantime.
-Thus, contrary to classical RL, \textbf{action and rewards are not synchronized}.
+While congestion control seems to be a well suited problem for reinforcement learning, there are a couple of unique aspects and differences that distinguish it from most other reinforcement learning problems:  \begin{itemize}
+\item \textbf{Delayed rewards}: In congestion control, rewards are always delayed by one round-trip time, while in most other areas where reinforcement learning is applied, if one performs an action, one can see the consequences of it action immediately.
+Also, due to the delay, by the time an action receives its reward, probably other actions have already been carried out in the meantime.
+Thus, contrary to most RL applications, actions and rewards are not synchronized.
 \item \textbf{Actions and rewards are not atomic}: If an action causes three packets to be sent, we consider these three packets being sent three \textit{partial actions}.
@@
 However, only if all partial rewards have been received, the sender can assemble the overall reward of the corresponding action and can perform an update of the underlying reinforcement learning logic.
-\item The \textbf{number of actions and rewards} per unit of time is \textbf{variable}: If the congestion window is larger, more packets are sent and more acknowledgements are received, which means that the reinforcement learning algorithm gets more training data the larger the window is.
+\item \textbf{Variable number of actions and rewards}: If the congestion window is larger, more packets are sent and more acknowledgements are received, which means that the reinforcement learning algorithm gets more training data the larger the window is.
 This is problematic as it means that -- without proper adjustments -- the experience made at a larger congestion window will ``overwrite'' the experience made at a smaller window as with the larger window simply more training data arrives and it is thus going to be overemphasized.
@@
 
-While using PAL to train an optimum congestion control for a specific range of network scenarios in an off-line fashion is possible, it is something that has already been done previously in an approach called \textit{Remy} \citep{winstein_tcp_2013}.
-The only potential advantage that PAL could provide here is increased training speed.
+While using PAL to train an optimum congestion control for a specific range of network scenarios in an off-line fashion is possible, offline learning of an optimum TCP congestion control is something that has already been done previously in an approach called \textit{Remy} \citep{winstein_tcp_2013}.
+The only potential advantage that PAL could provide over these previous solutions here is increased training speed.
 Thus, we want to show that it is  not only possible to learn congestion control by using machine learning but that it is even possible to do so on-line and without any preknowledge about the network environment.
@@
 \caption{Throughput and loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1 and 2 senders (from top to bottom).
-The time needed for convergence increases with more senders, which is caused by the fact that the more senders use the link, the fewer because each sender receives fewer packets per unit of time for training and because the environment gets more complex.
+The time needed for convergence increases with more senders, which is caused by the fact that the more senders use the link, the fewer packets each sender receives per unit of time for training and because of the increasing complexity of the environment.
 It is interesting to note that in the case of two senders, both flows complement each other and quickly adapt to the sending rate of the other flow.}
@@
 Besides the aforementioned human-designed algorithms there have been attempts to use machine learning to improve congestion control.
-\citet{geurts_machine_2004} train a classifier off-line to determine whether packet loss is caused by congestion or by the link layer.
+\citet{geurts_machine_2004} train a classifier off-line to determine whether packet loss is caused by congestion or by the link layer in case of wireless links.
 If it is caused by the link layer, TCP does not alter the window.
 Otherwise -- if a packet was lost due to congestion according to the classifier -- it uses a traditional TCP congestion control algorithm.
+\citet{shaio_reinforcement_2005} use online reinforcement learning to choose a sending rate for multimedia streaming.
 \citet{winstein_tcp_2013} train a machine learning solution called \textit{Remy} that finds an optimum congestion control for a given range of network parameters.
 For instance, one could find an optimum congestion control for networks with an RTT of 50-100\,ms and link speeds of 100-500\,Mbit/s.
-After a lengthy training procedure Remy finds an optimum congestion control algorithm for the specified networks on a per acknowledgement basis (updating the window upon receiving an acknowledgement alike traditional congestion control algorithms).
+After a lengthy training procedure of several CPU weeks, Remy finds an optimum congestion control algorithm for the specified networks on a per acknowledgement basis (updating the window upon receiving an acknowledgement alike traditional congestion control algorithms).
 
@@
 To this end we propose Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn to play a wide range of video games and commonly outperform human players.
-In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for playing video games and also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017} \todo{This motivation doesn't sound convincing yet, does it?}.
-Furthermore, as we will show, it is conceptually possible to use it for online training.
+In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for playing video games and also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017} as its use of neural network make it able to even learn complex tasks.
 
@@
 
-The \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) framework is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline (with a focus on intuition over rigor) in this section.
+Our proposed framework \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) framework is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline (with a focus on intuition over rigor) in this section.
 
@@
 The Critic Network estimates what long-term reward can be expected in this state.
-So an action is considered good if it achieved a long-term reward that is higher than the long-term reward expected by the Critic Network and it is considered bad if the reward was lower than expected.
-The long-term reward is implemented as an exponentially weighted moving average of future rewards.
+So an action is considered good if it achieved an empirically experienced long-term reward that is higher than the predicted long-term reward expected by the Critic Network and it is considered bad if the reward was lower than expected.
+The long-term reward is calculated as an exponentially weighted moving average of future rewards.
 So if a high reward can be achieved right now this is more favorable than if it can be achieved in the future.
@@
 
-The Critic Network outputs the expected long-term reward $V(s_{t}; \theta_\text{v})$ given a state $s_t$ at time step $t$ and the parameters (neural network weights) of the critic network $\theta_v$.
+The Critic Network outputs the expected long-term reward $V(s_{t}; \theta)$ given a state $s_t$ at time step $t$ and the parameters (neural network weights) of the critic network $\theta$.
 
-With $r_t$ being the reward that was received at time $t$ and $\gamma$ being a factor with $0 < \gamma \leq 1$, which stands for the influence that future reward has on the moving average, we define the expected long-term reward at time step $t$ as
-\begin{align*}
+With $r_t$ being the reward that was received at time $t$ and $\gamma$ being a factor with $0 < \gamma \leq 1$, which stands for the influence that future rewards have on the moving average, we define the expected long-term reward at time step $t$ as
+\begin{align}
 R_t = \gamma r_{t} + (1-\gamma) R_{t+1} ,
-%R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta_\text{v})\right)\left( 1-\gamma \right),
-\end{align*}
+%R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta)\right)\left( 1-\gamma \right),
+\end{align}
 
-One can see that to compute $R_t$ at time step $t$ one has to look infinitely far into the future to get all future rewards needed to compute the exponentially weighted moving average at time step $t$.
-Thus, one usually only carries out $t_\text{max}$ steps (for example 20) and then uses an estimation of $R_{t_\text{max}+1}$ provided by the critic network as the continuation \todo{Clearer than before?}.
+One can see that to compute $R_t$ at time step $t$ one has to look infinitely far into the future to get all future rewards as in each step one needs $R_{t+1}$.
+Thus, one usually only carries out $t_\text{max}$ steps (for example 20), waits for all the rewards and then calculates $R_t$ for each step and uses an estimation of $R_{t_\text{max}+1}$ provided by the critic network as the continuation in the future.
 The only exception from this is when the task to carry out is over before reaching $t_\text{max}$ (e.g.
 the TCP flow is over).
-In this case, one can either use a prediction from the critic network as a continuation as usually or simply let the moving average end without continuation (ignoring the term weighted with $(\gamma-1)$ at the last step).
+In this case, one can either use a prediction from the critic network as a continuation as usually or simply let the moving average end without continuation i.e.~ignoring the term weighted with $(\gamma-1)$ at the last step.
 
@@
 
-The loss function (not in the sense of packet loss but in the sense of the loss of a machine learning problem), which the critic network tries to minimize at each time step, is the square of the difference of the actual long-term reward received and the expected long-term reward
-\begin{align*}
+The loss function\footnote{not in the sense of packet loss but in the sense of the loss of a machine learning problem} (as defined by \cite{mnih_asynchronous_2016}), which the critic network tries to minimize at each time step, is the square of the difference of the actual long-term reward received and the expected long-term reward
+\begin{align}
 l_{\text{v},t} = \left(R_t - V(s_t; \theta)\right)^2.
-\end{align*}
+\end{align}
 
@@
 
-Each time we take an action $a_t$ we also let the critic network make an estimation $V(s_t; \theta)$, which we abbreviate as $v_t$, given the current neural network weights $\theta$ and the current state $s_t$.
-The loss function of the actor network can now be described as this: How likely was it to take action $a_t$ given the probability distribution at time $t$ and how beneficial was this action when comparing it to the critic's estimation?
+Each time we take an action $a_t$ we also let the critic network make an estimation $V(s_t; \theta)$, given the current neural network weights $\theta$ and the current state $s_t$.
+The loss function of the actor network can now be described as this: How likely was it to take action $a_t$ given the probability distribution ($\pi$) at time $t$ and how beneficial was this action when comparing it to the critic's estimation?
 For example, if an action was very successful and highly unlikely, the actor network's weights are updated to make actions like this more likely in the future.
 
-The actor also aims at increasing the entropy of the probability density function ($\frac{1}{2} \log \left( 2 \pi e \sigma^2 \right)$) to encourage exploration: Otherwise the actor could always perform similar actions but never actually take the action that yields the best performance.
+The actor also aims at increasing the probability density function's entropy ($\frac{1}{2} \log \left( 2 \pi e \sigma^2 \right)$) to encourage exploration: Otherwise the actor could always perform similar actions but never actually take the action that yields the best performance.
+The actor network will aim to maximize the entropy and at the same time output the best possible action: If it realizes that the entropy is too large and bad actions are taken, it learns to lower the entropy.
 We designate the entropy as $H$ and introduce a parameter $\beta$ which specifies magnitude of the influence of the entropy.
-\begin{align*}
+\begin{align}
 l_{\text{a},t} =& -\log \left( \pi\left( a_t \given s_t; \theta \right)\right)\left( R_t - v_t \right)\\ 
 &- \beta H\left( \pi\left( s_t; \theta \right)\right).
-\end{align*}
+\end{align}
 
@@
 Each partial action generates feedback upon interacting with the environment.
-Upon receiving feedback for a partial action, the agent determines the current state and triggers a new action.
+Upon receiving feedback for a partial action, the agent determines the current state and performs a new action.
 When all feedbacks of one action were received, the agent combines them to form the reward and updates the critic and actor networks.
@@
 
-\begin{figure}
-\begin{minipage}{\columnwidth}
-\includesvg{Reinforcement_learning_diagram}
-\caption{The classical reinforcement learning approach.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
-\label{fig:reinforcement}
-\end{minipage}
-\end{figure}
+%\begin{figure}
+%\begin{minipage}{\columnwidth}
+%\includesvg{Reinforcement_learning_diagram}
+%\caption{The classical reinforcement learning approach.\protect\footnote{adapted from \url{https://upload.wikimedia.org/wikipedia/commons/1/1b/Reinforcement_learning_diagram.svg}}}
+%\label{fig:reinforcement}
+%\end{minipage}
+%\end{figure}
 
@@
 \begin{minipage}{\columnwidth}
+%\input{pal.tex}
 \includesvg{Reinforcement_learning_diagram_PAL}
@@
 
-\subsection{Congestion Control Specifics}
+\subsection{Application to Congestion Control}
 
@@
 \begin{description}
-\item[$\textit{s}_\text{t}$] The state  describes the current ``congestion state''. The following features are included in it:
+\item[$\textit{s}_\text{t}$] The state  describes the current ``congestion state'' defined by:
 \begin{itemize}
@@
 \item the RTT of the last received packet
-\item whether the last packet was lost, where $1$ signifies that the last packet was lost while $0$ indicates that it was received correctly
-\item the current congestion window as a real number
+\item whether the last packet was lost, where $1$ designates that the last packet was lost while $0$ indicates that it was received correctly
+\item the current congestion window (as a real number, as actions by output by the actor network can be very small and thus the congestion window is often only changed by minuscule amounts)%
 \end{itemize}
-Of each of these features a exponentially weighted moving average (ewma) with a factor $\alpha$ of $\frac{1}{8}$ and one with an $\alpha$ of $\frac{1}{256}$ (the concept of using ewmas with these specific values of $\alpha$ was first introduced in \textit{Remy} \cite{winstein_tcp_2013}), which makes a total of 10 features.
+Of each of these features a exponentially weighted moving average (ewma) with a factor $\alpha$ of $\frac{1}{8}$ and one with an $\alpha$ of $\frac{1}{256}$ (the concept of using ewmas with these specific values of $\alpha$ was first introduced in \textit{Remy} \cite{winstein_tcp_2013}), which makes a total of 8 features.
 Each time an acknowledgement is received, the state is updated and the actor network is asked for the next action.
-\item[$\textit{a}_t$] Based on a given state and the history of previous states (because we use Gated Recurrent Units (GRUs) and thus can also consider previous states), the actor network returns an action $a_t$, which is a real number that stands for the change of the congestion window.
+\item[$\textit{a}_t$] Based on a given state and the history of previous states (because we use Gated Recurrent Units (GRUs) \cite{cho_learning_2014} and thus can also consider previous states), the actor network returns an action $a_t$, which is a real number that stands for the change of the congestion window.
 
@@
 However, as we use GRUs in the actor network, which have a nonlinear output function (a hyperbolic tangent function), they can learn to model nonlinear relationships between the input and the output.
-\item[$\textit{r}_t$] The reward is a tuple of three reward metrics. 
+\item[$\textit{r}_t$] The reward is a tuple of three reward metrics (defined in the next paragraph). 
 %These are discussed in more detail in \autoref{subsubsec:critic}. 
@@
 However, $\textit{r}_{\text{received},t}$ can only be determined at some point in time in the future, when the sender will have received the acknowledgements for all the bytes that he sent.
+Thus one has to wait one round-trip time until all rewards were received.
 %\item[$\textit{r}_{\text{byte},t}$] is the sum of the bytes of the packets that the sender sent and that were not lost.
@@
 %\item[$\textit{r}_{\text{delay},t}$] is the sum of the round trip times of the packets that the sender sent and that were not lost.
-\item[$\textit{r}_{\text{duration},t}$] is the time between receiving the last packet and receiving this packet (``inter acknowledgement reception time'') summed over all packets that the sender sent and that were not lost.
-Example: The sender sent three packets, two of which two were not lost.
-The time between receiving the ACK of the first packet that was not lost and the previous one is 5\,ms and the time between receiving the ACK of the second packet that was not lost and the previous ACK is 6\,ms.
+\item[$\textit{r}_{\text{duration},t}$] is the time between receiving the last acknowledgement and receiving this packet (``inter acknowledgement reception time'') summed over all packets that the sender sent and that were not lost.
+Example: The sender sent three packets, two of which were acknowledged at some point (one was lost).
+The time between receiving the ACK of the first packet and the previous one is 5\,ms and the time between receiving the ACK of the second packet and the previous ACK is 6\,ms.
 So for this time step $\textit{r}_{\text{duration},t}$ is 11\,ms.
@@
 In case of congestion control, the loss function $l_{\text{v},t}$ of the critic network is actually the sum of the squares of the difference for each of the expected long-term averages and the empirically found averages for each reward metric:
-\begin{align*}
-l_{\text{v},t} =& \left(R_{\text{received},t} - V_\text{received}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{sent},t} - V_\text{sent}(s_t; \theta_\text{v})\right)^2 \\
-&+\left(R_{\text{duration},t} - V_\text{duration}(s_t; \theta_\text{v})\right)^2
-\end{align*}
+\begin{align}
+l_{\text{v},t} =& \left(R_{\text{received},t} - V_\text{received}(s_t; \theta)\right)^2 \\
+&+\left(R_{\text{sent},t} - V_\text{sent}(s_t; \theta)\right)^2 \\
+&+\left(R_{\text{duration},t} - V_\text{duration}(s_t; \theta)\right)^2
+\end{align}
 
@@
 Our aim is to look into the future for a more-or-less constant time span no matter the current size of the window.
-In other words, we want to define $\gamma$ so that it reflects the number of packets in the current window, so that it always looks in the future for approximately one round-trip time.
+In other words, we want to define $\gamma$ so that it reflects the number of packets in the current window, so that it always looks in the future for approximately one round-trip time (as one full window corresponds to one round-trip time).
 
-A ewma with a $\gamma$ of $\frac{2}{n+1}$ has roughly the same distribution of forecast error as a regular moving average with a window size of $n$ \cite{nahmias_production_2009}.
+An ewma with a $\gamma$ of $\frac{2}{n+1}$ has similar characteristics as a regular moving average with a window size of $n$ \cite{nahmias_production_2009}.
 Furthermore, with sufficiently large $n$ this means that the first $n$ data account for $\approx 86\%$ of the total weight in the ewma \cite{boissard_applications_2012}.
 
-Thus, the idea is to make $n$ the number of packets in the current window and so we define the factor $\gamma$ as a function of the current window size and the expected amount of bytes to be sent per time step as follows:
+Thus, the idea is to make $n$ the number of packets in the current window and so we define the factor $\gamma$ as a function of the current window size and the expected amount of bytes to be sent per time step (the window over the expected number of bytes per packet is the expected number of packets in this window) as follows:
 
-\begin{align*}
-\gamma_t = \frac{2}{1+\max(\frac{w_t+a_t}{V_\text{sent}(s_t; \theta_\text{v})},10)}
-\end{align*}
+\begin{align}
+\gamma_t = \frac{2}{1+\max(\frac{w_t+a_t}{V_\text{sent}(s_t; \theta)},10)}
+\end{align}
 
-We use 10 as a minimum because when $\gamma$ becomes too big, future rewards is not considered sufficiently anymore and as the ensuing variability causes instability.
+We use 10 as a minimum (as proposed by \cite{dong_pcc:_2015}) because when $\gamma$ becomes too big, future rewards is not considered sufficiently anymore and as the ensuing variability causes instability.
 
@@
 The actor network minimizes the loss function
-%\begin{align*}
+%\begin{align}
 %l_{\text{a},t} =& -\log \left( \pi \left( a_t \given s_t ; \theta_\text{a} \right) \right)\\
@@
 
-\begin{align*}
+\begin{align}
 l_{\text{a},t} =& -\log \left( \pi \left( a_t \given s_t ; \theta_\text{a} \right) \right)\\
@@
 &- \beta H\left( \pi\left( s_t; \theta_a \right)\right)
-\end{align*}
+\end{align}
 where $U$ can be any utility function defined based on some of the (long term averages of the) reward metrics.
@@
 
-\begin{align*}
+\begin{align}
 U_t = \text{throughput} - \alpha \times \text{lost throughput}
-\end{align*}
+\end{align}
 
@@
 
-\begin{align*}
+\begin{align}
 U_t = \frac{R_{\text{received},t}}{R_{\text{duration},t}} - \alpha \frac{R_{\text{sent},t} - R_{\text{received},t}}{R_{\text{duration},t}}
-\end{align*}
+\end{align}
 
@@
 In the following evaluation we use a classical dumbbell network topology with one receiver and one or more senders and a shared bottleneck link.
+As our evaluation focuses on bulk transfers we use equally sized packets of 1250 bytes (TCP and IP including headers).
 In all figures in the evaluation we use at least 50 data unless stated otherwise.
@@
 \begin{minipage}{\columnwidth}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_1}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_2}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4}}}
-\caption{One sender with a bandwidth/packet loss tradeoff of 1, 2 and 4 (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\subfloat[One sender, throughput/packet loss tradeoff of 1]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_1}}}}{}
+%\caption{Blabla}
+\subfloat[One sender, throughput/packet loss tradeoff of 2]{
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_2}}}}{}
+\subfloat[One sender, throughput/packet loss tradeoff of 4]{
+\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4}}}}{}
+\caption{One sender with varying throughput/packet loss tradeoff on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
@@
 \begin{minipage}{\columnwidth}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/new_gpu/no_loss/figures/4_4}}}
-\caption{One, two and four senders (from top to bottom) with a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\subfloat[Two senders, throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4}}}}{}
+\subfloat[Four senders, throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/new_gpu/no_loss/figures/4_4}}}}{}
+\caption{Two and four senders (from top to bottom) with a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
@@
 
-While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations clearly increase when having two or four concurrent senders (see \autoref{fig:multiple_senders}).
+While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations increase when having two or four concurrent senders (see \autoref{fig:multiple_senders}).
 We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which he can use for training.
@@
 \begin{minipage}{\columnwidth}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4_again}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4_again}}}
-\caption{One and two senders (from top to bottom) that were trained with a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\subfloat[One sender, initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4_again}}}}{}
+\subfloat[Two senders, initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4_again}}}}{}
+\caption{One and two senders (from top to bottom) that were trained with a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
@@
 Then the sender constantly stays in this state and one would expect it to slowly ``forget'' what it learned in the very beginning when it started the flow.
-Thus, it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again in a completely different state than the one it is used to.
+Thus, it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again in a completely different state than the one it was used to for hundreds of seconds before.
+For two senders pre-training also greatly speeds up the initial training procedure and can avoid the spike of packet loss in the beginning that always occurs to some extent without pre-training (see \autoref{fig:one_sender_loss} and \autoref{fig:multiple_senders}).
+However, it also become apparent that variance is large when comparing to the non-pre-trained version.
+The reason for this seems to be that one sender rapidly claims all bandwidth in the very beginning leaving the other sender with nothing.
 
@@
 \begin{minipage}{\columnwidth}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-5.local/figures/1_4}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-7.local/figures/2_4}}}
-\caption{One and two senders (from top to bottom) on a link with a stochastic loss rate of 1\%, a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\subfloat[One sender sharing a link with a stochastic loss rate of 1\% with a throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-5.local/figures/1_4}}}}{}
+\subfloat[Two senders sharing a link with a stochastic loss rate of 1\% with a throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-7.local/figures/2_4}}}}{}
+\caption{One and two senders (from top to bottom) on a link with a stochastic loss rate of 1\%, a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.
@@
 
-Stochastic loss does not substantially affect TCP Rax.
-This can be attributed to the fact that the reinforcement learning gets to understand that a certain minimum packet loss rate is ``normal'' and cannot be avoided and thus TCP Rax is flexible in this regard.
+We also verify whether stochastic loss has an influence on TCP Rax.
+For this, we randomly 1\% dropped of packets on the link and it seems as if stochastic loss does not substantially affect TCP Rax.
+This can be attributed to the fact that the reinforcement learning gets to understand that a certain minimum packet loss rate is ``normal'' and cannot be avoided and thus TCP Rax can learn to distinguish packet loss that is caused by congestion from packet loss that is purely stochastic (as for example caused by errors on the physical layer).
 
@@
 Again we take the scenario of a 20\,Mbit/s shared bottleneck with a small buffer of $\frac{1}{10}$ bandwidth delay product to create a challenging scenario.
-Surprisingly New Reno performs by far better than more modern TCP congestion control algorithms (see \autoref{fig:regular_tcps}).
-TCP Rax achieves a significantly larger total throughput in the median though with a couple of outliers which become more significant with two concurrent senders (see \autoref{fig:rax_tcps}).
+Surprisingly, New Reno performs by far better than more modern TCP congestion control algorithms (see \autoref{fig:regular_tcps}).
+We attribute this to the fact that we chose a small buffer to create a challenging scenario and that Cubic and Compound are not optimized so well for this scenario but rather for long fat networks \cite{oura_fairness_2012}.
+TCP Rax achieves a significantly larger total throughput in the median though with a couple of outliers which become more significant with two concurrent senders (see \autoref{fig:rax_tcps}); when inspecting the raw data, it becomes apparent that these outliers occur usually when the neural network diverges in the first seconds of a flow.
 
@@
 \begin{minipage}{\columnwidth}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_1_0}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_2_0}}}
-\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one and two senders (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\subfloat[TCP Compound, Cubic and New Reno's total throughput and loss rate for each sender with one senders]{\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_1_0}}}}{}
+\subfloat[TCP Compound, Cubic and New Reno's total throughput and loss rate for each sender with two senders]{\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_2_0}}}}{}
+\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one and two senders of TCP Cubic/Compound/New Reno (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 For a flow of 600\,s the maximum possible throughput is 12\,000\,Mbit.}
@@
 \begin{minipage}{\columnwidth}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/all_together}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/all_together}}}
+\subfloat[TCP Rax's total throughput]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/all_together}}}}{}
+\subfloat[TCP Compound, Cubic and New Reno's average throughput with one sender]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/all_together}}}}{}
 \caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one and two senders (from top to bottom) for TCP Rax on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
@@
 
+\section{Discussion}
+
+Our results indicate that our proposed concept can do online learning of congestion control and can outperform existing congestion control algorithms in certain scenarios, although a larger number of senders make the reinforcement learning significantly more challenging.
+Regarding online learning, during the implementation of TCP Rax we came to the conclusion that a neural network consisting of LSTM cells \cite{gers_learning_1999} is not feasible as LSTM cells keep an internal state that gets unstable when it is for long training episodes (as for each received acknowledgement the LSTM state has to be updated).
+Previous work using deep reinforcement learning uses up to 50 million training steps for offline learning \cite{mnih_asynchronous_2016} while we generally use significantly less (a 20 Mbit\,s link can carry a maximum of up to one million MTU-sized packets in 10 minutes).
+Thus, when sharing the link between several senders, each sender receives only a fraction of these packets for training.
+We think that this is caused by the fact that in congestion control rewards are steadily coming in, while for other tasks, such as video games, commonly rewards are only received sparsely (e.g.
+one has to wait until the game is over to determine whether one won or lost the game).
+
 \section{Conclusion}
 
-The goal of this paper was to determine whether it is possible to use deep reinforcement learning to learn a per-acknowledgement congestion control online.
-To this end we had to modify existing reinforcement learning algorithms to accommodate the specific needs of congestion control such as the inherent delay.
+In this paper we show that it is possible to use deep reinforcement learning to learn a per-acknowledgement congestion control online.
+To this end we had to modify existing reinforcement learning algorithms to accommodate the specific needs of congestion control such as the inherent delay and proposed Partial Action Learning, a framework that allows deep reinforcement learning with delayed and partial rewards.
 
@@
 We came to the insight that it is possible to pre-train the neural network used in the reinforcement learning and that this improves performance especially in the beginning of the flow.
-
-Some interesting questions, however, remain.
-In what way can pre-training be utilized to compete with other TCP congestion control variants even for very short flows?
-Is it feasible and sensible that one computer uses a set of shared weights for all its flows?
-
-\todo{Ist dieser Teil mit offenen Fragen sinnlos? Hab das schon in anderen Papers gesehen...}
+However, TCP Rax can also compete with other congestion control variants after an initial exploration phase.
 
@@
 		\State $l_\text{actions}.\text{append}(a_t)$
-		\State $l_\text{values}.\text{append}(V(s_t; \theta_\text{v}))$ 
-		\Comment{Possibly different from the}
-		\StateIndent[3] value in $l_\text{values'}$ as the weights could have 
+		\State $l_\text{values}.\text{append}(V(s_t; \theta))$ 
+		\Comment{Possibly different from}
+		\StateIndent[3] the value in $l_\text{values'}$ as the weights could have 
 		\StateIndent[3] been updated in the meantime.
@@
 			\State $l_\text{rewards}.\text{append}(r_{t'})$
-			\If{\Call{gotEnoughRewards?}{} or the episode is over}
+			\If{\Call{gotEnoughRewards?}{}}
 %			\If{$\#\left(l_\text{rewards}\right) > t_\text{max}$ or the episode is over}
@@
 \Function{gotEnoughRewards?}{}
+	\If{the epsiode is over}
+		\State \Return \True
+	\EndIf
 	\State $w \gets \left\lfloor l_\text{windows}[0] + l_\text{actions}[0] \right\rfloor$
@@
 %
-%\begin{align*}
+%\begin{align}
 %\text{Reward} = \log(\text{throughput}) - \delta\text{delay}
-%\end{align*}
+%\end{align}
 %
@@
 %\begin{enumerate}
-%\item \begin{align*}
+%\item \begin{align}
 %\left(\frac{\partial}{\partial w}\text{throughput}\right)\left(1\right) > \left(\frac{\partial}{\partial w}\delta\text{delay}\right)\left(1\right)
-%\end{align*}
-%\item \begin{align*}
+%\end{align}
+%\item \begin{align}
 %\frac{\partial}{\partial w}\left(\delta\right)
-%\end{align*}
+%\end{align}
 %\end{enumerate}
@@
 %%We define the excessive window as 
-%%\begin{align*}
+%%\begin{align}
 %%b_\text{excess} = \max\left( \left( \sum_{j=1}^N \frac{b_j}{d_j} \right) - c, 0 \right)
-%%\end{align*}
+%%\end{align}
 %%and the queueing delay as 
-%%\begin{align*}
+%%\begin{align}
 %%d_q = \frac{b_\text{excess}}{c}.
-%%\end{align*}
+%%\end{align}
 %
@@
 %
-%\begin{align*}
+%\begin{align}
 %c = \sum_{j=1}^N \frac{b_j}{d_j + d_q(b)}
-%\end{align*}
+%\end{align}
 %
 %Then we define the utility function for each sender $i$ as
-%\begin{align*}
+%\begin{align}
 %V_i(b) \coloneqq \begin{cases}
@@
 %\end{cases}
-%\end{align*}
+%\end{align}
 %
-%\begin{align*}
+%\begin{align}
 %&\log\left(\frac{b_i}{d_i + d_q}\right) - \delta_i (d_i + d_q) \\
 %=& \log\left(b_i\right) - \log\left(d_i + d_q\right) - \delta_i (d_i + d_q)
-%\end{align*}
+%\end{align}
 %
-%\begin{align*}
+%\begin{align}
 %\frac{\partial V_i}{\partial b_i} = \begin{cases}
@@
 %\end{cases}
-%\end{align*}
+%\end{align}
 
-%\begin{align*}
+%\begin{align}
 %G(b) = \begin{pmatrix}
@@
 %\end{pmatrix} \text{if } \text{if } \sum_{j=1}^N \frac{b_i}{d_i} \leq c
-%\end{align*}
+%\end{align}
 

### 2018-01-31 tex/petri_net.tex
Update tex and bib tex

@@
   \begin{scope}
-    \node [place,tokens=1] (s1) at (0, 1) [label={[align=center]left:Receive\\ feedback}] {};
-    \node [place] (s2) at (1, 2) [label={[align=center]above:Received\\ ACK}] {};
+    \node [place,tokens=1] (s1) at (0, 1) [label={[align=center]left:Received\\ feedback}] {};
+    \node [place] (s2) at (1, 2) [label={[align=center]above:Received\\ ACKs}] {};
     \node [place] (s3) at (2, 1) [label={[align=center]right:Packets\\ in the\\ network}] {};
@@
 
-    \node [transition] (e5) at (0, 2) [label={[align=center]left:Received\\ feedback}] {}
+    \node [transition] (e5) at (0, 2) [label={[align=center]left:Receive\\ feedback}] {}
       edge [pre] node {1} (s2)

### 2018-01-31 async_deep_reinforce/a3c_training_thread.py

@@ -14,7 +14,7 @@ from functools import reduce
 from game_ac_network import GameACLSTMNetwork
 
 # gamma_current, gamma_future = 1./(1+GAMMA), GAMMA/(1.+GAMMA)
-# from constants import LOCAL_T_MAX
+from constants import LOCAL_T_MAX
 # from constants import GAMMA
 from constants import LOG_LEVEL
 from constants import STATE_SIZE

### 2018-01-31 tex/paper.tex
Update paper

@@
 
-In a machine learning based congestion control method, each sender uses a set of observed environment conditions to determine which action he should take to maximize his reward in the future.
+In a machine learning based congestion control method, each sender uses a set of observed environment conditions to determine which action it should take to maximize his reward in the future.
 The environment conditions can be any metrics that the sender can obtain from the network, for example the mean round-trip time of the last received packets, the packet loss rate etc.
@@
 
-While congestion control seems to be a well suited problem for reinforcement learning, there are a couple of unique aspects and differences that distinguish it from most other reinforcement learning problems:  \begin{itemize}
+While congestion control seems to be a well-suited problem for reinforcement learning, there are a couple of unique aspects and differences that distinguish it from most other reinforcement learning problems:  \begin{itemize}
 \item \textbf{Delayed rewards}: In congestion control, rewards are always delayed by one round-trip time, while in most other areas where reinforcement learning is applied, if one performs an action, one can see the consequences of it action immediately.
@@
 \begin{minipage}{\columnwidth}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_1_2018-01-25-20:17:15_3_2018-01-25-20:17:15.0}}}
-\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/4_2_2018-01-25-19:53:39_22_2018-01-25-19:53:39.0}}}
+\subfloat[One sender]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_1_2018-01-25-20:17:15_3_2018-01-25-20:17:15.0}}}}{}
+\subfloat[Two senders]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/4_2_2018-01-25-19:53:39_22_2018-01-25-19:53:39.0}}}}{}
 \caption{Throughput and loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1 and 2 senders (from top to bottom).
@@
 We designate the entropy as $H$ and introduce a parameter $\beta$ which specifies magnitude of the influence of the entropy.
+The loss function of the actor network is defined as follows \cite{mnih_asynchronous_2016}:
 \begin{align}
+\begin{split}
 l_{\text{a},t} =& -\log \left( \pi\left( a_t \given s_t; \theta \right)\right)\left( R_t - v_t \right)\\ 
 &- \beta H\left( \pi\left( s_t; \theta \right)\right).
+\end{split}
 \end{align}
@@
 \item the RTT of the last received packet
-\item whether the last packet was lost, where $1$ designates that the last packet was lost while $0$ indicates that it was received correctly
+\item whether the last packet was lost, where $1$ denotes that the last packet was lost while $0$ indicates that it was received correctly
 \item the current congestion window (as a real number, as actions by output by the actor network can be very small and thus the congestion window is often only changed by minuscule amounts)%
@@
 \begin{align}
+\begin{split}
 l_{\text{v},t} =& \left(R_{\text{received},t} - V_\text{received}(s_t; \theta)\right)^2 \\
@@
 &+\left(R_{\text{duration},t} - V_\text{duration}(s_t; \theta)\right)^2
+\end{split}
 \end{align}
@@
 \begin{align}
-l_{\text{a},t} =& -\log \left( \pi \left( a_t \given s_t ; \theta_\text{a} \right) \right)\\
-& \left( U_{\text{measured},t} - U_{\text{expected},t} \right)\\ 
+\begin{split}
+l_{\text{a},t} =& -\log \left( \pi \left( a_t \given s_t ; \theta_\text{a} \right) \right)\left( U_{\text{measured},t} - U_{\text{expected},t} \right)\\ 
 &- \beta H\left( \pi\left( s_t; \theta_a \right)\right)
+\end{split}
 \end{align}

### 2018-01-31 tex/paper.tex
Add all results data and finish tex

@@
 %Conference
-\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El
-  Paso, Texas USA} 
-\acmYear{2018}
-\copyrightyear{2018}
+%\acmConference[WOODSTOCK'97]{ACM Woodstock conference}{July 1997}{El
+%  Paso, Texas USA} 
+%\acmYear{2018}
+%\copyrightyear{2018}
 
@@
 \begin{document}
-\title{TCP Rax: Deep Reinforcement Learning for TCP Congestion Control}
+\title{TCP Rax: Deep Reinforcement Learning for Congestion Control}
 %\titlenote{Produces the permission block, and
@@
 
+\renewcommand{\shortauthors}{X. et al.}
+
 \begin{abstract}
@@
 The processing of rewards in congestion control depends on the arrival of acknowledgments, inducing loop delays that may lead to oscillations.
-As a remedy we propose \textit{Partial Action Learning}, a specific formulation of neural network based reinforcement learning that supports the processing of delayed and partial rewards.
+As a remedy we propose \textit{Partial Action Learning} (PAL), a specific formulation of neural network based reinforcement learning that supports the processing of delayed and partial rewards.
 
 We show that our method converges to a stable, close-to-optimum solution within the order of minutes and outperforms existing congestion control algorithms in typical networks.
-Thus, this paper demonstrates that neural network based reinforcement learning without any prior knowledge can feasibly be done on-line and can compete with human-designed congestion-control schemes such as New Reno and Cubic.
+Thus, this paper demonstrates that neural network based reinforcement learning without any prior knowledge can feasibly be done online and can compete with human-designed congestion-control schemes such as New Reno and Cubic.
 
@@
 As, for instance, for some applications overall throughput is more important than little packet loss, each application can define its preference using a custom utility function.
-Besides aiming to fulfill these subgoals, it is also desirable from an overall point of view that all senders use available resources in a way that is fair to the others: When several senders share an Internet link under the same conditions, they should receive an equal share of the available resources.
+Besides aiming to fulfill these subgoals, it is also desirable from an overall point of view that all senders use available resources in a way that is fair to others: When several senders share an Internet link under the same conditions, they should receive an equal share of the available resources.
 To achieve fairness, it is necessary that senders only use utility functions that
@@
 In a machine learning based congestion control method, each sender uses a set of observed environment conditions to determine which action it should take to maximize his reward in the future.
-The environment conditions can be any metrics that the sender can obtain from the network, for example the mean round-trip time of the last received packets, the packet loss rate etc.
-An action is a change to the sending rate: For example, if a sender perceives an increase in the loss rate it might be advisable to lower the sending rate.
+The environment conditions are metrics that the sender can obtain from the network, for example the mean round-trip time of the last received packets, the packet loss rate etc.~or combinations thereof.
+An action is a change to the sending rate: For example, if a sender perceives an increase in the loss rate it might decide to lower the sending rate.
 To modulate the sending rate one can  either
@@
 \end{enumerate*}.
-It seems preferable to maintain a congestion window, as one cannot be certain that the operating system (OS) can maintain a certain constant sending rate for each application on one machine, as the OS has to schedule different applications and share the network interfaces between them.
+As the operating system (OS) has to schedule different applications and share the network interfaces between them, maintaining a congestion window is appealing, because one cannot be sure that the OS can maintain a specific constant sending rate for each application on one machine.
 
-While congestion control seems to be a well-suited problem for reinforcement learning, there are a couple of unique aspects and differences that distinguish it from most other reinforcement learning problems:  \begin{itemize}
-\item \textbf{Delayed rewards}: In congestion control, rewards are always delayed by one round-trip time, while in most other areas where reinforcement learning is applied, if one performs an action, one can see the consequences of it action immediately.
-Also, due to the delay, by the time an action receives its reward, probably other actions have already been carried out in the meantime.
+While congestion control seems to be a well-suited task for reinforcement learning, there are a couple of unique aspects and differences that distinguish it from most other reinforcement learning problems:  \begin{itemize}
+\item \textbf{Delayed rewards}: In congestion control, rewards are always delayed by one round-trip time.
+In most other areas where reinforcement learning is applied, if an action is performed, the consequences of that action become apparent immediately.
+Also, due to the delay, by the time an action receives its reward, probably other actions have already been performed.
 Thus, contrary to most RL applications, actions and rewards are not synchronized.
@@
 In case these packets are transmitted correctly, the receiver will send back three acknowledgements.
-Each of these acknowledgements is a \textit{partial reward} that already unveils new information on the current state of the network and enables the sender to take an action based on this new information.
+Each of these acknowledgements is a \textit{partial reward} that unveils new information on the current state of the network and enables the sender to take an action based on this new information.
 However, only if all partial rewards have been received, the sender can assemble the overall reward of the corresponding action and can perform an update of the underlying reinforcement learning logic.
 \item \textbf{Variable number of actions and rewards}: If the congestion window is larger, more packets are sent and more acknowledgements are received, which means that the reinforcement learning algorithm gets more training data the larger the window is.
-This is problematic as it means that -- without proper adjustments -- the experience made at a larger congestion window will ``overwrite'' the experience made at a smaller window as with the larger window simply more training data arrives and it is thus going to be overemphasized.
+This is problematic as it means that -- without proper adjustments -- the experience made at a larger congestion window will ``overwrite'' the experience made at a smaller window.
+With the larger window more training data arrives and it is thus going to be overemphasized.
 \end{itemize}
@@
 
-While using PAL to train an optimum congestion control for a specific range of network scenarios in an off-line fashion is possible, offline learning of an optimum TCP congestion control is something that has already been done previously in an approach called \textit{Remy} \citep{winstein_tcp_2013}.
-The only potential advantage that PAL could provide over these previous solutions here is increased training speed.
-Thus, we want to show that it is  not only possible to learn congestion control by using machine learning but that it is even possible to do so on-line and without any preknowledge about the network environment.
+While using PAL to train an optimum congestion control for a specific range of network scenarios in an offline fashion is possible, offline learning of an optimum TCP congestion control is something that has already been done previously in an approach called \textit{Remy} \citep{winstein_tcp_2013}.
+The only potential advantage that PAL can provide over prior solutions is increased training speed.
+Thus, we want to show that it is  not only possible to learn congestion control by using machine learning but that it is even possible to do so online and without any preknowledge about the network environment.
 
-Using a utility function that encourages high throughput and discourages packet loss (adapted from \citep{dong_pcc:_2015}) we see that our machine learning approach can learn congestion control that maximizes this objective given a couple of minutes of time (without any pre-training) (see \autoref{fig:demo}).
+\autoref{fig:demo} shows that using a utility function that encourages high throughput and discourages packet loss (adapted from \citep{dong_pcc:_2015}) our machine learning approach can learn congestion control that maximizes this objective given a couple of minutes of time (without any pre-training).
 
@@
 \begin{minipage}{\columnwidth}
-\subfloat[One sender]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_1_2018-01-25-20:17:15_3_2018-01-25-20:17:15.0}}}}{}
-\subfloat[Two senders]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/4_2_2018-01-25-19:53:39_22_2018-01-25-19:53:39.0}}}}{}
-\caption{Throughput and loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between 1 and 2 senders (from top to bottom).
+\subfloat[One sender\label{fig:demo1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/4_1_2018-01-25-20:17:15_3_2018-01-25-20:17:15.0}}}}{}
+\subfloat[Two senders\label{fig:demo2}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/4_2_2018-01-25-19:53:39_22_2018-01-25-19:53:39.0}}}}{}
+\caption{Throughput and loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between one (\ref{fig:demo1}) and two (\ref{fig:demo2}) senders .
 The time needed for convergence increases with more senders, which is caused by the fact that the more senders use the link, the fewer packets each sender receives per unit of time for training and because of the increasing complexity of the environment.
-It is interesting to note that in the case of two senders, both flows complement each other and quickly adapt to the sending rate of the other flow.}
+In case of two senders, both flows complement each other and quickly adapt to the sending rate of the other flow.}
 \label{fig:demo}
@@
 
-After spending up to a few minutes with suboptimal performance during initial exploration (if no pre-training is used), our approach can compete with established algorithms and achieve superior throughput and packet loss of less that 1\%, while still gradually improving and adapting its way of performing congestion control.
+After spending up to a few minutes with suboptimal performance during initial exploration (if no pre-training is used), our approach can compete with established algorithms and achieve superior throughput and an average packet loss of less than 1\%, while still gradually improving and adapting its way of performing congestion control.
+
+Thus we make the three contributions:
+\begin{itemize}
+\item A window based TCP congestion control that acts on a per-acknowledgement basis that can compete with existing congestion control algorithms.
+\item The ability to learn online, offline or a combination of both offline pre-training with online learning. 
+\item A framework for deep reinforcement learning that allows dealing with delayed and partial rewards.
+\end{itemize}
 
@@
 Congestion control has been implemented in TCP since the 1980s as a series of congestion collapses dramatically decreased applications' throughput in the Internet, which lead to the development of the Tahoe and Reno algorithms \cite{jacobson_congestion_1988}.
-These approaches as well as most others maintain a congestion window which stands for the maximum amount of data that are allowed to be traveling in the network without having been acknowledged by the receiver.
-In general, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995}) lead to the congestion window being lowered according to a set of fixed rules.
+These approaches as well as most others maintain a congestion window which stands for the maximum amount of data that are allowed to be in flight in the network without having been acknowledged by the receiver.
+In most TCP congestion control algorithms, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995}) lead to the congestion window being lowered according to a set of fixed rules.
 Algorithms such as Cubic and Compound \cite{ha_cubic:_2008, tan_compound_2006} improve congestion control specifically with respect to so-called long fat networks with a high round-trip time and bandwidth.
 Proportional rate reduction \cite{dukkipati_proportional_2011} aims to make the reduction of the rate that occurs in case of congestion more smooth and steady in time to avoid the bursty behavior that occurred upon loss in previous TCP congestion control variants.
-On the other hand, BBR \cite{cardwell_bbr:_2016} does not use a set of fixed rules such as previous congestion control algorithms but instead estimates a model of the network path by using measurements of RTT and throughput and then adjusts its sending rate so that it uses the maximum bandwidth according to its network model while trying not to fill up queues.
+A different approach is taken by BBR \cite{cardwell_bbr:_2016}, which does not use a set of fixed rules such as previous congestion control algorithms.
+Instead it estimates a model of the network path by using measurements of RTT and throughput and then adjusts its sending rate so that it uses the maximum bandwidth according to its network model while trying not to fill up queues.
 A similar approach is PCC \cite{dong_pcc:_2015}, which also uses measurements to find an optimum sending rate with respect to a defined utility function (e.g.
@@
 Besides the aforementioned human-designed algorithms there have been attempts to use machine learning to improve congestion control.
-\citet{geurts_machine_2004} train a classifier off-line to determine whether packet loss is caused by congestion or by the link layer in case of wireless links.
+\citet{geurts_machine_2004} train a classifier offline to determine whether packet loss is caused by congestion or by the link layer in case of wireless links.
 If it is caused by the link layer, TCP does not alter the window.
@@
 \citet{winstein_tcp_2013} train a machine learning solution called \textit{Remy} that finds an optimum congestion control for a given range of network parameters.
-For instance, one could find an optimum congestion control for networks with an RTT of 50-100\,ms and link speeds of 100-500\,Mbit/s.
-After a lengthy training procedure of several CPU weeks, Remy finds an optimum congestion control algorithm for the specified networks on a per acknowledgement basis (updating the window upon receiving an acknowledgement alike traditional congestion control algorithms).
+For instance, one could find an optimum congestion control for networks with an RTT of 50-100\,ms and link speeds of 10-20\,Mbit/s.
+After a lengthy training procedure of several CPU weeks, Remy finds an optimum congestion control algorithm for the specified networks on a per acknowledgement basis (updating the window upon receiving an acknowledgement like traditional congestion control algorithms).
 
-Our goal is to find an optimum congestion control algorithm on a per acknowledgement basis similar to Remy that, however, can be trained on-line.
-Such a solution could be used in a purely on-line fashion, in a off-line fashion like Remy or a combination of both: One could pre-train a generic congestion control algorithm that works reasonably well for every network that then gets refined during on-line training according to the current network circumstances.
+Our goal is to find an optimum congestion control algorithm on a per acknowledgement basis similar to Remy that, however, can be trained online.
+Such a solution could be used in a purely online fashion, in a offline fashion like Remy or a combination of both: One could pre-train a generic congestion control algorithm that works reasonably well for every network that then gets refined during online training according to the current network circumstances.
 
 To this end we propose Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn to play a wide range of video games and commonly outperform human players.
-In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for playing video games and also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017} as its use of neural network make it able to even learn complex tasks.
+Its use of a neural network enable it to even learn complex tasks.
+In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for playing video games and also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017}.
 
@@
 
-Our proposed framework \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) framework is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline (with a focus on intuition over rigor) in this section.
+Our proposed framework \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline (with a focus on intuition over rigor) in this section.
 
@@
 Given a state, the Actor Network outputs what it deems to be the optimum action to perform in that certain state.
-The Critic Network estimates what long-term reward can be expected in this state.
+The Critic Network predicts what long-term reward can be expected in this state.
 So an action is considered good if it achieved an empirically experienced long-term reward that is higher than the predicted long-term reward expected by the Critic Network and it is considered bad if the reward was lower than expected.
@@
 
-The Critic Network outputs the expected long-term reward $V(s_{t}; \theta)$ given a state $s_t$ at time step $t$ and the parameters (neural network weights) of the critic network $\theta$.
+The Critic Network outputs a prediction of the long-term reward $v_t = V(s_{t}; \theta)$ given a state $s_t$ at time step $t$ and the parameters (neural network weights) of the critic network $\theta$.
 
-With $r_t$ being the reward that was received at time $t$ and $\gamma$ being a factor with $0 < \gamma \leq 1$, which stands for the influence that future rewards have on the moving average, we define the expected long-term reward at time step $t$ as
+With $r_t$ being the reward that was received at time $t$ and $\gamma$ being a factor with $0 < \gamma \leq 1$, which stands for the influence that future rewards have on the moving average, we define the actual experienced long-term reward $R_t$ at time step $t$ as
 \begin{align}
 R_t = \gamma r_{t} + (1-\gamma) R_{t+1} ,
-%R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta)\right)\left( 1-\gamma \right),
+%R_t = \left(\left(\sum_{i=0}^{k-1} \gamma^ir_{t+i}\right) + \gamma^k V(s_{t+k}; \theta)\right)\left( 1-\gamma \right).
 \end{align}
 
-One can see that to compute $R_t$ at time step $t$ one has to look infinitely far into the future to get all future rewards as in each step one needs $R_{t+1}$.
-Thus, one usually only carries out $t_\text{max}$ steps (for example 20), waits for all the rewards and then calculates $R_t$ for each step and uses an estimation of $R_{t_\text{max}+1}$ provided by the critic network as the continuation in the future.
-The only exception from this is when the task to carry out is over before reaching $t_\text{max}$ (e.g.
+To compute $R_t$ at time step $t$ it is necessary to look infinitely far into the future to get all future rewards as in each step $t$ to compute $R_t$ one needs $R_{t+1}$.
+Thus, one usually only carries out $t_\text{max}$ steps (for example 20), waits for all the rewards and then calculates $R_t$ for each step and uses an estimate of $R_{t_\text{max}+1}$ provided by the critic network as the continuation in the future.
+The only exception from this is when the task to execute ends before reaching $t_\text{max}$ (e.g.
 the TCP flow is over).
-In this case, one can either use a prediction from the critic network as a continuation as usually or simply let the moving average end without continuation i.e.~ignoring the term weighted with $(\gamma-1)$ at the last step.
+In this case, one can either use a prediction from the critic network as a continuation or alternatively let the moving average end without continuation i.e.~ignoring the term weighted with $(\gamma-1)$ at the last step.
 
@@
 
-The loss function\footnote{not in the sense of packet loss but in the sense of the loss of a machine learning problem} (as defined by \cite{mnih_asynchronous_2016}), which the critic network tries to minimize at each time step, is the square of the difference of the actual long-term reward received and the expected long-term reward
+The loss function\footnote{not in the sense of packet loss but in the sense of the loss function of a machine learning problem, which one aims to minimize} (as defined by \cite{mnih_asynchronous_2016}), which the critic network tries to minimize at each time step, is the square of the difference of the actual experienced long-term reward received and the predicted long-term reward $l_{\text{v},t}$ as
 \begin{align}
@@
 
-The Actor Network outputs a normal distribution parametrized by a mean $\mu$ and a standard deviation $\sigma$ from which the action $a_t$ at time step $t$ is randomly sampled.
-The main idea is that the neural network learns to output the right mean at the right time step to maximize the future reward and that it uses the standard deviation to try out new actions, which could yield a better than expected reward.
+The Actor Network outputs a normal distribution of the possible values for the action parametrized by a mean $\mu$ and a standard deviation $\sigma$ from which the action $a_t$ at time step $t$ is randomly sampled.
+The main idea is that the neural network learns to output the right mean of the actions at the right time step to maximize the future reward and that it uses the standard deviation to try out new actions, which could yield a better than expected reward.
 
-Each time we take an action $a_t$ we also let the critic network make an estimation $V(s_t; \theta)$, given the current neural network weights $\theta$ and the current state $s_t$.
-The loss function of the actor network can now be described as this: How likely was it to take action $a_t$ given the probability distribution ($\pi$) at time $t$ and how beneficial was this action when comparing it to the critic's estimation?
+Each time we take an action $a_t$ we also let the critic network make an estimation $v_t = V(s_t; \theta)$, given the current neural network weights $\theta$ and the current state $s_t$.
+The loss function of the actor network can be described as follows: How likely was it to take action $a_t$ given the probability distribution ($\pi$) at time $t$ and how beneficial was this action when comparing it to the critic's estimation?
 For example, if an action was very successful and highly unlikely, the actor network's weights are updated to make actions like this more likely in the future.
 
-The actor also aims at increasing the probability density function's entropy ($\frac{1}{2} \log \left( 2 \pi e \sigma^2 \right)$) to encourage exploration: Otherwise the actor could always perform similar actions but never actually take the action that yields the best performance.
+The actor also aims at increasing the probability density function's entropy ($H\left( \pi\left( s_t; \theta \right)\right) = \frac{1}{2} \log \left( 2 \pi e \sigma^2 \right)$) to encourage exploration: Otherwise the actor could always perform similar actions but never actually take the action that yields the best performance.
 The actor network will aim to maximize the entropy and at the same time output the best possible action: If it realizes that the entropy is too large and bad actions are taken, it learns to lower the entropy.
-We designate the entropy as $H$ and introduce a parameter $\beta$ which specifies magnitude of the influence of the entropy.
+We designate the entropy as $H$ and introduce a parameter $\beta$ which specifies the magnitude of the influence of the entropy.
 The loss function of the actor network is defined as follows \cite{mnih_asynchronous_2016}:
@@
 \begin{split}
-l_{\text{a},t} =& -\log \left( \pi\left( a_t \given s_t; \theta \right)\right)\left( R_t - v_t \right)\\ 
+l_{\text{a},t} =& -\log \left( \pi\left( a_t \given s_t; \theta \right)\right)\left( R_t - V(s_t; \theta) \right)\\ 
 &- \beta H\left( \pi\left( s_t; \theta \right)\right).
@@
 
-The key difference between PAL and previous approaches to RL is that in classical RL an action is always followed by a reward and a reward is always followed by an action.
+The key difference between our approach PAL and previous approaches to RL is that in classical RL an action is always followed by a reward and a reward is always followed by an action.
 In our proposed concept, however, it is possible to take new actions while previous actions haven't been rewarded yet.
@@
 In \autoref{alg:pal} (in the appendix) we show the code that runs in each of the agents (in the congestion control scenario, one agent corresponds to a sender).
-It is possible to have several agents which share a set of neural weights (which is usually done when performing off-line training \cite{mnih_asynchronous_2016,mao_neural_2017}) but one can also use separate weights for each agent, which is more realistic in case of congestion control in the Internet, as it is not sensible that different senders share a set of neural network weights over the Internet.
+It is possible to have several agents which share a set of neural weights (which is usually done when performing offline training \cite{mnih_asynchronous_2016,mao_neural_2017}) but one can also use separate weights for each agent, which is more realistic in case of congestion control in the Internet, as it is not sensible that different senders share a set of neural network weights over the Internet.
 
@@
 \caption{A petri net describing the congestion control mechanism.
-We start with one token in the state \textit{Received feedback}, which means that we can take an action right in the beginning.
-The concept is the following: If there is at least one token in the \textit{Open congestion window} send a packet (for the sake of simplicity in this example we assume that the congestion window is in packets and not bytes).
-The \textit{Open congestion window} signifies the congestion window minus the packets that are currently unacknowledged.
-The token now goes to the network and after some time the acknowledgement (ACK) (or negative-acknowledgement (NACK) or timeout if the packet gets lost) for the packet is received in the \textit{Received ACK} state.
-From here one token goes to the \textit{Open congestion window} meaning that when we get an ACK, NACK (or timeout) we can send another packet.
+It starts with one token in the state \textit{Received feedback}, which means that it can take an action right in the beginning.
+The concept is the following: If there is at least one token in the \textit{Open congestion window} it sends a packet.
+For the sake of simplicity in this example we assume that the congestion window is in packets and not bytes.
+The \textit{Open congestion window} designates the congestion window minus the packets that are currently unacknowledged.
+The token now is passed to the network and after some time the acknowledgement (ACK) (or negative-acknowledgement (NACK) or timeout if the packet gets lost) for the packet is received in the \textit{Received ACK} state.
+From here one token goes to the \textit{Open congestion window} meaning that when it receives an ACK, NACK (or timeout) it can send another packet.
 Furthermore the state \textit{Received feedback} receives a token upon receiving an ACK/NACK/timeout.
-\textit{Take action} adds \textit{a} tokens to the \textit{Open congestion window}, where $a$ is a real number (possibly also negative); thus in each state there can also be a real number of tokens (e.g.~ 2.34 tokens are possible in the \textit{Open congestion window}).
+\textit{Take action} adds \textit{a} tokens to the \textit{Open congestion window}, where $a$ is a real number (possibly also negative).
+Therefore, in each state there can be a real number of tokens (e.g.~ 2.34 tokens are possible in the \textit{Open congestion window}).
 However, we define that the congestion window can never be smaller than 1, which means that the sum of all tokens in the \textit{Open congestion window.}, \textit{Received ACK} and \textit{Packets in the network} states can never be smaller than 1.}
@@
 
-To use PAL for congestion control we first have to define the correct semantics for this specific use case and define what makes up the state and the reward.
+To use reinforcement learning for congestion control in general and our modified approach PAL specifically we first have to define the correct semantics for this specific use case and define what makes up the state and the reward.
 Furthermore, we have to define how the actor and critic network explicitly work in case of congestion control.
-In the following a time step $t$ corresponds to the reception of an acknowledgement.
-The beginning of the flow, before any packet is sent, corresponds to time step $0$.
+In the following, a time step $t$ corresponds to the reception of an acknowledgement.
+The beginning of the flow, the time when the first packet is sent, corresponds to time step $0$.
 
 \begin{description}
-\item[$\textit{s}_\text{t}$] The state  describes the current ``congestion state'' defined by:
+\item[$\textit{s}_\text{t}$] The state $s_t$ describes the current ``congestion state'' defined by:
 \begin{itemize}
-\item the time between the last two packets that were received
+\item the time between the last two acknowledgment that were received
 \item the RTT of the last received packet
-\item whether the last packet was lost, where $1$ denotes that the last packet was lost while $0$ indicates that it was received correctly
-\item the current congestion window (as a real number, as actions by output by the actor network can be very small and thus the congestion window is often only changed by minuscule amounts)%
+\item a loss indicator that shows whether the last packet was lost, where $1$ denotes that the last packet was lost while $0$ indicates that it was received correctly
+\item the current congestion window (as a real number, as actions by output by the actor network can be very small and thus the congestion window is often changed by minuscule amounts)%
 \end{itemize}
-Of each of these features a exponentially weighted moving average (ewma) with a factor $\alpha$ of $\frac{1}{8}$ and one with an $\alpha$ of $\frac{1}{256}$ (the concept of using ewmas with these specific values of $\alpha$ was first introduced in \textit{Remy} \cite{winstein_tcp_2013}), which makes a total of 8 features.
+These features are not used themselves but instead of each of these features an exponentially weighted moving average (ewma) with a factor $\alpha$ of $\frac{1}{8}$ and one with an $\alpha$ of $\frac{1}{256}$ is computed, which makes a total of 8 features.
+The concept of using ewmas with these specific values of $\alpha$ was first introduced in \textit{Remy} \cite{winstein_tcp_2013}).
 Each time an acknowledgement is received, the state is updated and the actor network is asked for the next action.
-\item[$\textit{a}_t$] Based on a given state and the history of previous states (because we use Gated Recurrent Units (GRUs) \cite{cho_learning_2014} and thus can also consider previous states), the actor network returns an action $a_t$, which is a real number that stands for the change of the congestion window.
+\item[$\textit{a}_t$] The action $a_t$ is a real number that represents the change to the congestion window.
+It is computed based on a given state and the history of previous states.
+The history of previous states as considered as our neural network uses Gated Recurrent Units (GRUs) \cite{cho_learning_2014}.
 
 It might seem simplistic to only use an additive increase to the congestion window and no more sophisticated mechanisms such as a combination of a multiplicative increase and an additive increase.
-However, as we use GRUs in the actor network, which have a nonlinear output function (a hyperbolic tangent function), they can learn to model nonlinear relationships between the input and the output.
-\item[$\textit{r}_t$] The reward is a tuple of three reward metrics (defined in the next paragraph). 
+However, GRUs have a nonlinear output function (a hyperbolic tangent function) and thus they can learn to model nonlinear relationships between the input and the output.
+\item[$\textit{r}_t$] The reward $r_t$ is a tuple of three reward metrics (defined in the next paragraph). 
 %These are discussed in more detail in \autoref{subsubsec:critic}. 
 For each reward metric there is also an output of the critic network that predicts the expected long-term average of this reward metric given the current state. 
-\item[$\textit{v}_t$] The value is a tuple of the expected average reward estimated by the critic network (see \autoref{subsubsec:critic}) for each of the reward metrics. 
+\item[$\textit{v}_t$] The value $v_t$ is a tuple of the expected average reward estimated by the critic network (see \autoref{subsubsec:critic}) for each of the reward metrics.
+Also for the value $v_t$ the current state and the history of previous states is considered.
 \end{description}
@@
 \item[$\textit{r}_{\text{sent},t}$] is the sum of all bytes in all packets the sender sent during time step $t$. 
-\item[$\textit{r}_{\text{received},t}$] is the sum of all bytes in all packets that the sender sent during time step $t$ and that were not lost (so they were acknowledged at some point by the receiver).
-Example: During time step $t$, 2500 bytes are sent, of which 1250 are lost and 1250 are acknowledged at some point in time later.
-Then $\textit{r}_{\text{received},t}$ is 1250 for at time step $t$.
+\item[$\textit{r}_{\text{received},t}$] is the sum of all bytes in all packets that the sender sent during time step $t$ and that were not lost (so they were acknowledged by the receiver).
+Example: During time step $t$, 2500 bytes are sent, of which 1250 are lost and 1250 are acknowledged.
+Then $\textit{r}_{\text{received},t}$ is 1250 at time step $t$.
 However, $\textit{r}_{\text{received},t}$ can only be determined at some point in time in the future, when the sender will have received the acknowledgements for all the bytes that he sent.
@@
 %\item[$\textit{r}_{\text{delay},t}$] is the sum of the round trip times of the packets that the sender sent and that were not lost.
-\item[$\textit{r}_{\text{duration},t}$] is the time between receiving the last acknowledgement and receiving this packet (``inter acknowledgement reception time'') summed over all packets that the sender sent and that were not lost.
+\item[$\textit{r}_{\text{duration},t}$] is the time between receiving the last acknowledgement and receiving this packet (``inter acknowledgement reception time'') summed over all packets that the sender sent and that were acknowledged.
 Example: The sender sent three packets, two of which were acknowledged at some point (one was lost).
-The time between receiving the ACK of the first packet and the previous one is 5\,ms and the time between receiving the ACK of the second packet and the previous ACK is 6\,ms.
+The time between receiving the ACK of the first packet and the previous one (received at an earlier time step) is 5\,ms and the time between receiving the ACK of the second packet and the previous ACK is 6\,ms.
 So for this time step $\textit{r}_{\text{duration},t}$ is 11\,ms.
@@
 \begin{figure}
+\begin{minipage}{\columnwidth}
 
@@
 
-\caption{An overview of the complete neural network being used.
+\caption{An overview of the complete neural network being used in our experiments.
 Ocher stands for the input, blue for linear layers, red for GRU layers and purple for the outputs of the neural network.
 The numbers in parentheses next to each label stand for the width of that layer (i.e.~the number of neurons except for the inputs and outputs).
-Parts of the neural network are shared between $V_\text{received}$ and $V_\text{sent}$ as well as $\mu$ and $\sigma$ as they output similar quantities and thus the gradients that are computed when updating the neural network are of the same order of magnitude.
+Parts of the neural network are shared between $V_\text{received}$ and $V_\text{sent}$ as well as $\mu$ and $\sigma$ as they output quantities of similar magnitude \protect\footnote{The expected number of bytes sent and the expected number of bytes received are usually of similar size.} and thus the gradients that are computed when updating the neural network are of the same order of magnitude.
 In our experiments we saw that if one uses shared weights between very different outputs, the outputs which produce significantly larger gradients would ``overwrite'' the gradients of the smaller ones.}
 \label{fig:neural_network}
+\end{minipage}
 \end{figure}
@@
 
-In case of congestion control, the loss function $l_{\text{v},t}$ of the critic network is actually the sum of the squares of the difference for each of the expected long-term averages and the empirically found averages for each reward metric:
+In case of congestion control, the loss function $l_{\text{v},t}$ of the critic network is actually the sum of the squares of the difference for each of the predicted long-term averages and the empirically found averages for each reward metric:
 \begin{align}
@@
 However, in congestion control the larger the window is, the more packets are sent and received per unit of time.
-Thus, if one used a single fixed parameter $\gamma$, one would have the problem that one would look into the future for a shorter time the larger the window is, as more packets and thus more actions and rewards are handled per unit of time.
-Our aim is to look into the future for a more-or-less constant time span no matter the current size of the window.
+Thus, the use of a single fixed parameter $\gamma$ incurs the problem that larger congestion windows result in shorter future intervals to be considered for congestion control.
+The reason is that the larger the window, the more packets and thus more actions and rewards are handled per unit of time.
+Our aim is to look into the future for a more-or-less constant time span independent of the current size of the window.
 In other words, we want to define $\gamma$ so that it reflects the number of packets in the current window, so that it always looks in the future for approximately one round-trip time (as one full window corresponds to one round-trip time).
 
-An ewma with a $\gamma$ of $\frac{2}{n+1}$ has similar characteristics as a regular moving average with a window size of $n$ \cite{nahmias_production_2009}.
-Furthermore, with sufficiently large $n$ this means that the first $n$ data account for $\approx 86\%$ of the total weight in the ewma \cite{boissard_applications_2012}.
+An ewma with a $\gamma$ of $\frac{2}{n+1}$ has similar characteristics as a regular moving average with a window size of $n$.
+Furthermore, with sufficiently large $n$ this means that the first $n$ data account for $\approx 86\%$ of the total weight in the ewma \cite{nahmias_production_2009}.
 
-Thus, the idea is to make $n$ the number of packets in the current window and so we define the factor $\gamma$ as a function of the current window size and the expected amount of bytes to be sent per time step (the window over the expected number of bytes per packet is the expected number of packets in this window) as follows:
+Thus, the idea is to make $n$ the number of packets in the current window and so we define the factor $\gamma_t$ as a function of the current window size $w_t$ and the expected amount of bytes to be sent per time step (the window divided by the expected number of bytes per packet is the expected number of packets in this window) as follows:
 
 \begin{align}
-\gamma_t = \frac{2}{1+\max(\frac{w_t+a_t}{V_\text{sent}(s_t; \theta)},10)}
+\gamma_t = \frac{2}{1+\max(\frac{w_t}{V_\text{sent}(s_t; \theta)},10)}
 \end{align}
 
-We use 10 as a minimum (as proposed by \cite{dong_pcc:_2015}) because when $\gamma$ becomes too big, future rewards is not considered sufficiently anymore and as the ensuing variability causes instability.
+We use 10 as a minimum (as proposed by \cite{dong_pcc:_2015}) because when $\gamma$ becomes too large, future rewards are not considered sufficiently anymore and as the ensuing variability causes instability.
 
@@
 
-The actor network minimizes the loss function
+The actor network aims to minimize the loss function
 %\begin{align}
@@
 where $U$ can be any utility function defined based on some of the (long term averages of the) reward metrics.
-Thus, in simple terms, the actor network considers if an action improved the actual experienced utility compared to the expected one and adjusts the neural network accordingly.
+In other words, the actor network considers if an action improved the actual experienced utility compared to the expected one and adjusts the neural network accordingly.
 
@@
 
-$\alpha$ is a parameter that determines how strongly packet loss is discouraged. We can define this utility function as follows using the previously defined reward metrics:
+$\alpha$ is a parameter that determines how strongly packet loss is discouraged (throughput/packet loss tradeoff factor).
+We can define this utility function as follows using the previously defined reward metrics:
 
 \begin{align}
-U_t = \frac{R_{\text{received},t}}{R_{\text{duration},t}} - \alpha \frac{R_{\text{sent},t} - R_{\text{received},t}}{R_{\text{duration},t}}
+U_{\text{measured},t} = \frac{R_{\text{received},t}}{R_{\text{duration},t}} - \alpha \frac{R_{\text{sent},t} - R_{\text{received},t}}{R_{\text{duration},t}}
 \end{align}
 
-The corresponding expected utility is also defined using the estimates of the value networks and it can be determined whether an action achieved a performance that was higher or lower than expected.
+The corresponding expected utility $U_{\text{expected},t}$ is also defined using the estimates of the value networks and it can be determined whether an action achieved a performance that was higher or lower than expected.
 
@@
 
-We implemented TCP Rax as an extension to both the NS-2 network simulator \cite{fall_network_2007} and Remy \cite{winstein_tcp_2013}.  
+We implemented TCP Rax as an extension to both the ns-2 network simulator \cite{fall_network_2007} and Remy \cite{winstein_tcp_2013}.  
 In the following evaluation we use a classical dumbbell network topology with one receiver and one or more senders and a shared bottleneck link.
 As our evaluation focuses on bulk transfers we use equally sized packets of 1250 bytes (TCP and IP including headers).
-In all figures in the evaluation we use at least 50 data unless stated otherwise.
+In all figures in the evaluation we use data from at least 50 simulations unless stated otherwise.
+In case of multiple senders per simulation we randomly sample one sender from each simulation and leave out the others.
+The reason is that in each simulation the behavior of the senders is correlated.
 
@@
 \begin{minipage}{\columnwidth}
-\subfloat[One sender, throughput/packet loss tradeoff of 1]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_1}}}}{}
+\subfloat[One sender, throughput/packet loss tradeoff of 1\label{fig:one_sender_loss1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_1}}}}{}
 %\caption{Blabla}
-\subfloat[One sender, throughput/packet loss tradeoff of 2]{
+\subfloat[One sender, throughput/packet loss tradeoff of 2\label{fig:one_sender_loss2}]{
 \includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_2}}}}{}
-\subfloat[One sender, throughput/packet loss tradeoff of 4]{
+\subfloat[One sender, throughput/packet loss tradeoff of 4\label{fig:one_sender_loss3}]{
 \includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4}}}}{}
@@
 This leads to exponential growth of the congestion window in the beginning and as soon as the bottleneck bandwidth is reached, the reinforcement learning needs to correct its behavior to realize that increasing the congestion window is not always the appropriate action to perform.
-Potential countermeasure for this are to use pre-training (see \autoref{subsec:pretraining}) and to use a utility function that heavily penalizes excessive packet loss.
+Potential countermeasures for this are to either use pre-training (see \autoref{subsec:pretraining}) or to use a utility function that heavily penalizes excessive packet loss.
 
@@
 \begin{minipage}{\columnwidth}
-\subfloat[Two senders, throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4}}}}{}
-\subfloat[Four senders, throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/new_gpu/no_loss/figures/4_4}}}}{}
-\caption{Two and four senders (from top to bottom) with a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\subfloat[Two senders, throughput/packet loss tradeoff of 2\label{fig:multiple_senders1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_2}}}}{}
+\subfloat[Four senders, throughput/packet loss tradeoff of 2\label{fig:multiple_senders2}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/new_gpu/no_loss/figures/4_2}}}}{}
+\caption{Two (\ref{fig:multiple_senders1}) and four (\ref{fig:multiple_senders2}) senders with a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
@@
 
-While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations increase when having two or four concurrent senders (see \autoref{fig:multiple_senders}).
-We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which he can use for training.
+While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations and training time needed to achieve good performance increase when having two or four concurrent senders (see \autoref{fig:multiple_senders}).
+We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which it can use for training.
 Furthermore, as the environment becomes more dynamic with more senders, each sender has a harder time learning whether the reward changed due to his own actions or due to actions that were caused by other senders: If there is only one flow on a link, the flow ``can be sure'' that all changes in rewards were actually caused by his own actions.
@@
 \begin{minipage}{\columnwidth}
-\subfloat[One sender, initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4_again}}}}{}
-\subfloat[Two senders, initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4_again}}}}{}
-\caption{One and two senders (from top to bottom) that were trained with a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\subfloat[One sender, initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4\label{fig:resuming_training1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4_again}}}}{}
+\subfloat[Two senders, initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4\label{fig:resuming_training2}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4_again}}}}{}
+\caption{One (\ref{fig:resuming_training1}) and two (\ref{fig:resuming_training2}) senders that were trained with a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
@@
 For two senders pre-training also greatly speeds up the initial training procedure and can avoid the spike of packet loss in the beginning that always occurs to some extent without pre-training (see \autoref{fig:one_sender_loss} and \autoref{fig:multiple_senders}).
-However, it also become apparent that variance is large when comparing to the non-pre-trained version.
-The reason for this seems to be that one sender rapidly claims all bandwidth in the very beginning leaving the other sender with nothing.
+However, it also becomes apparent that variance is large when comparing to the non-pre-trained version.
+We argue that the reason for this is that one sender rapidly claims all bandwidth in the very beginning leaving the other senders with nothing.
 
@@
 \begin{minipage}{\columnwidth}
-\subfloat[One sender sharing a link with a stochastic loss rate of 1\% with a throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-5.local/figures/1_4}}}}{}
-\subfloat[Two senders sharing a link with a stochastic loss rate of 1\% with a throughput/packet loss tradeoff of 4]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-7.local/figures/2_4}}}}{}
-\caption{One and two senders (from top to bottom) on a link with a stochastic loss rate of 1\%, a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\subfloat[One sender sharing a link with a stochastic loss rate of 1\% with a throughput/packet loss tradeoff of 4\label{fig:stochastic_loss1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-5.local/figures/1_4}}}}{}
+\subfloat[Two senders sharing a link with a stochastic loss rate of 1\% with a throughput/packet loss tradeoff of 4\label{fig:stochastic_loss2}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-7.local/figures/2_4}}}}{}
+\caption{One (\ref{fig:stochastic_loss1}) and two (\ref{fig:stochastic_loss2}) senders on a link with a stochastic loss rate of 1\%, a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.
-It looks as if there is no stochastic loss in the very beginning of the flows, however, this is an artifact of the binning procedure: As the window is very small in the beginning, there are more flows in which no packet loss occurs whatsoever and thus the median is distorted.}
+Apparently there is no stochastic loss in the very beginning of the flows, however, this is an artifact of the binning procedure (we use bins of one second): As the window is very small in the beginning, there are many flows in which no packet loss occurs whatsoever in the first bins and thus the median is distorted.}
 \label{fig:stochastic_loss}
@@
 We also verify whether stochastic loss has an influence on TCP Rax.
-For this, we randomly 1\% dropped of packets on the link and it seems as if stochastic loss does not substantially affect TCP Rax.
-This can be attributed to the fact that the reinforcement learning gets to understand that a certain minimum packet loss rate is ``normal'' and cannot be avoided and thus TCP Rax can learn to distinguish packet loss that is caused by congestion from packet loss that is purely stochastic (as for example caused by errors on the physical layer).
+For this, we randomly dropped 1\% of packets on the link and it seems as if stochastic loss does not substantially affect TCP Rax.
+This can be attributed to the fact that the reinforcement learning realizes that a certain minimum packet loss rate is ``normal'' and cannot be avoided and thus TCP Rax can learn to distinguish packet loss that is caused by congestion from packet loss that is purely stochastic (as for example caused by errors on the physical layer).
 
@@
 
-Again we take the scenario of a 20\,Mbit/s shared bottleneck with a small buffer of $\frac{1}{10}$ bandwidth delay product to create a challenging scenario.
+For reasons of comparability we take the same scenario of a 20\,Mbit/s shared bottleneck with a small buffer of $\frac{1}{10}$ bandwidth delay product to create a challenging scenario.
 Surprisingly, New Reno performs by far better than more modern TCP congestion control algorithms (see \autoref{fig:regular_tcps}).
 We attribute this to the fact that we chose a small buffer to create a challenging scenario and that Cubic and Compound are not optimized so well for this scenario but rather for long fat networks \cite{oura_fairness_2012}.
-TCP Rax achieves a significantly larger total throughput in the median though with a couple of outliers which become more significant with two concurrent senders (see \autoref{fig:rax_tcps}); when inspecting the raw data, it becomes apparent that these outliers occur usually when the neural network diverges in the first seconds of a flow.
+TCP Rax achieves a significantly larger total throughput on average though with a couple of outliers which become more significant with two concurrent senders (see \autoref{fig:rax_tcps}); when inspecting the raw data, it becomes apparent that these outliers occur usually when the neural network diverges in the first seconds of a flow.
 
@@
 \begin{minipage}{\columnwidth}
-\subfloat[TCP Compound, Cubic and New Reno's total throughput and loss rate for each sender with one senders]{\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_1_0}}}}{}
-\subfloat[TCP Compound, Cubic and New Reno's total throughput and loss rate for each sender with two senders]{\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_2_0}}}}{}
-\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one and two senders of TCP Cubic/Compound/New Reno (from top to bottom) on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
-For a flow of 600\,s the maximum possible throughput is 12\,000\,Mbit.}
+\subfloat[TCP Compound, Cubic and New Reno's and Rax's total throughput and loss rate for each sender with one senders in a simulation of 600\,s\label{fig:regular_tcps1}
+]{\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_1_0}}}}{}
+\subfloat[TCP Compound, Cubic and New Reno and Rax's total throughput and loss rate for each sender with two senders in a simulation of 600\,s\label{fig:regular_tcps2}
+]{\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_2_0}}}}{}
+\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one (\ref{fig:regular_tcps1}) and two (\ref{fig:regular_tcps2})senders of TCP Cubic/Compound/New Reno and Rax with a throughput/packet loss tradeoff of 4 without pre-training on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+For a flow of 600\,s the maximum possible throughput is 12\,000\,Mbit.
+For TCP Rax we use the mean over 50 simulation runs, while for the other TCP congestion control variants we use only one run because we use a deterministic simulator.}
 \label{fig:regular_tcps}
@@
 \begin{minipage}{\columnwidth}
-\subfloat[TCP Rax's total throughput]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/all_together}}}}{}
-\subfloat[TCP Compound, Cubic and New Reno's average throughput with one sender]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/all_together}}}}{}
-\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one and two senders (from top to bottom) for TCP Rax on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\subfloat[TCP Rax's total throughput for one sender showing different throughput/packet loss tradeoffs\label{fig:rax_tcps1}
+]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/all_together}}}}{}
+\subfloat[TCP Rax's total throughput for two senders showing different throughput/packet loss tradeoffs\label{fig:rax_tcps2}
+]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/all_together}}}}{}
+\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one (\ref{fig:rax_tcps1}) and two senders (\ref{fig:rax_tcps2}) for TCP Rax on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 The arrow symbol indicates that the flows use a pre-trained neural network.
@@
 
-\todo{Add some kind of table with other scenarios}
-
 \section{Discussion}
 
-Our results indicate that our proposed concept can do online learning of congestion control and can outperform existing congestion control algorithms in certain scenarios, although a larger number of senders make the reinforcement learning significantly more challenging.
-Regarding online learning, during the implementation of TCP Rax we came to the conclusion that a neural network consisting of LSTM cells \cite{gers_learning_1999} is not feasible as LSTM cells keep an internal state that gets unstable when it is for long training episodes (as for each received acknowledgement the LSTM state has to be updated).
-Previous work using deep reinforcement learning uses up to 50 million training steps for offline learning \cite{mnih_asynchronous_2016} while we generally use significantly less (a 20 Mbit\,s link can carry a maximum of up to one million MTU-sized packets in 10 minutes).
-Thus, when sharing the link between several senders, each sender receives only a fraction of these packets for training.
-We think that this is caused by the fact that in congestion control rewards are steadily coming in, while for other tasks, such as video games, commonly rewards are only received sparsely (e.g.
+Our results demonstrate that our proposed concept can perform online learning of congestion control and can outperform existing congestion control algorithms in certain scenarios, although a larger number of senders make the reinforcement learning significantly more challenging and time consuming.
+Regarding online learning, during the implementation of TCP Rax we came to the conclusion that a neural network consisting of LSTM cells \cite{gers_learning_1999} is not feasible.
+LSTM cells keep an internal state that gets unstable when it is used for long training episodes, as for each received acknowledgement the LSTM state has to be updated.
+Previous work using deep reinforcement learning uses up to 50 million training steps for offline learning \cite{mnih_asynchronous_2016} while we generally use significantly less (a 20 Mbit\,s link can carry a maximum of up to one million packets of 1250 bytes size in 10 minutes).
+When sharing the link between several senders, each sender receives only a fraction of these packets for training.
+For example, with two senders on one link (see Figure \ref{fig:multiple_senders2}) one sender can receive a maximum of 600\,000 packets during a ten minute flow for training.
+We argue that the reason for TCP Rax needing fewer training samples than other reinforcement learning tasks is that in congestion control rewards are steadily coming in, while for other tasks, such as video games, commonly rewards are only received sparsely (e.g.
 one has to wait until the game is over to determine whether one won or lost the game).
@@
 
-We developed TCP Rax and the results of our evaluation indicate that congestion control can be learned online in principle and has the potential to outperform standard TCP variants given long enough flows.
+We developed TCP Rax and the results of our evaluation indicate that congestion control can be learned online and has the potential to outperform standard TCP variants for significantly large flows.
 We came to the insight that it is possible to pre-train the neural network used in the reinforcement learning and that this improves performance especially in the beginning of the flow.
-However, TCP Rax can also compete with other congestion control variants after an initial exploration phase.
+However, TCP Rax can also compete with other congestion control variants after an initial exploration phase without pre-training.
 
@@
 \caption{Partial Action Learning -- pseudocode for each agent.
-It is possible that the agents share the global weights $\theta_{\text{g}}$, which is reasonable when learning off-line.
+The global weights $\theta_{\text{g}}$ can be shared by multiple agents, which is reasonable when learning offline.
 Otherwise an individual copy of them is kept by each agent.
-By default, all weights are initialized randomly in the beginning but it is also possible to use a pre-trained neural network for initialization.
+By default, all weights are initialized randomly but it is also possible to use a pre-trained neural network for initialization.
 As we use Gated Recurrent Units (GRUs) their output is stored in $h$ and fed into them at the next time step.}
@@
 \Loop
-	\State $l_\text{actions} \gets  [] $
-	\State $l_\text{states} \gets  [] $
-	\State $l_\text{values} \gets  [] $
-	\State $l_\text{rewards} \gets  [] $
-	\State $l_\text{values'} \gets  [] $
-	\State $l_\text{snapshots} \gets  [] $
-	\State $l_\text{hiddenStates} \gets  [] $
+	\State $l_\text{actions}, l_\text{states}, l_\text{values}, l_\text{rewards}, l_\text{values'} \gets  [] $
+	\State $l_\text{snapshots}, l_\text{hiddenStates} \gets  [] $
 	\State $t \gets 0$
@@
 		\State $t \gets t+1$
-	\Until{reaching the end the episode}
+	\Until{reaching the end of the episode}
 \EndLoop
@@
 	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{values'}$
-	\State Remove the first element from $l_\text{snapshots}$
-	\State Remove the first element from $l_\text{hiddenStates}$
+	\State Remove the first element from $l_\text{snapshots}$, $l_\text{hiddenStates}$
 	\State $\theta = \theta_\text{backup}$

### 2018-01-31 async_deep_reinforce/a3c_training_thread.py

@@ -14,7 +14,7 @@ from functools import reduce
 from game_ac_network import GameACLSTMNetwork
 
 # gamma_current, gamma_future = 1./(1+GAMMA), GAMMA/(1.+GAMMA)
-from constants import LOCAL_T_MAX
+# from constants import LOCAL_T_MAX
 # from constants import GAMMA
 from constants import LOG_LEVEL
 from constants import STATE_SIZE

### 2018-01-31 src/unicorn-templates.cc

@@ -39,11 +39,12 @@ void Unicorn::send( const unsigned int id, NextHop & next, const double & tickno
       _flow_to_last_received[_flow_id] = tickno;
     }
 
-    _last_send_time = tickno;
     _id_to_sent_during_action[_packets_sent] = _put_actions;
     _outstanding_rewards[_put_actions]["sent"] += 1;
     _outstanding_rewards[_put_actions]["intersend_duration_acc"] += tickno - _last_send_time;
     _packets_sent++;
+
+    _last_send_time = tickno;
 
     _memory.packet_sent( p );
     next.accept( p, tickno );

### 2018-01-31 src/unicorn.cc

@@ -52,7 +52,7 @@ bool Unicorn::packets_lost( const std::vector< remy::Packet > & packets ) {
         _memory.lost(1);
         _memory.window(_the_window);
         // _the_window = std::max(_the_window - 1, MIN_WINDOW_UNICORN);
-        // get_action(packet.tick_received);
+        get_action(packet.tick_received);
         _largest_ack += 1;
         return true;
       }

### 2018-02-03 tex/paper.tex
Fixed assertion bug

@@
 
-In a machine learning based congestion control method, each sender uses a set of observed environment conditions to determine which action it should take to maximize his reward in the future.
+In a machine learning based congestion control method, each sender uses a set of observed environment conditions to determine which action it should take to maximize its reward in the future.
 The environment conditions are metrics that the sender can obtain from the network, for example the mean round-trip time of the last received packets, the packet loss rate etc.~or combinations thereof.
@@
 
-Thus we make the three contributions:
+Thus we make three contributions:
 \begin{itemize}
-\item A window based TCP congestion control that acts on a per-acknowledgement basis that can compete with existing congestion control algorithms.
-\item The ability to learn online, offline or a combination of both offline pre-training with online learning. 
-\item A framework for deep reinforcement learning that allows dealing with delayed and partial rewards.
+\item TCP Rax: A reinforcement learning based TCP congestion control algorithm that helps senders to learn a good congestion control strategy autonomously.
+Our method acts on a per acknowledgement basis and can compete with existing congestion control algorithms.
+\item The ability to learn online, offline or a combination of both: offline pre-training with online learning.
+\item Partial Action Learning (PAL): A new method for deep reinforcement learning that allows dealing with delayed and partial rewards.
 \end{itemize}
@@
 These approaches as well as most others maintain a congestion window which stands for the maximum amount of data that are allowed to be in flight in the network without having been acknowledged by the receiver.
-In most TCP congestion control algorithms, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995}) lead to the congestion window being lowered according to a set of fixed rules.
+In most TCP congestion control algorithms, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995}) leads to the congestion window being lowered according to a set of fixed rules.
 Algorithms such as Cubic and Compound \cite{ha_cubic:_2008, tan_compound_2006} improve congestion control specifically with respect to so-called long fat networks with a high round-trip time and bandwidth.
 Proportional rate reduction \cite{dukkipati_proportional_2011} aims to make the reduction of the rate that occurs in case of congestion more smooth and steady in time to avoid the bursty behavior that occurred upon loss in previous TCP congestion control variants.
-A different approach is taken by BBR \cite{cardwell_bbr:_2016}, which does not use a set of fixed rules such as previous congestion control algorithms.
+A different approach is taken by BBR \cite{cardwell_bbr:_2016}, which does not use a set of fixed rules like previous congestion control algorithms.
 Instead it estimates a model of the network path by using measurements of RTT and throughput and then adjusts its sending rate so that it uses the maximum bandwidth according to its network model while trying not to fill up queues.
@@
 To this end we propose Partial Action Learning which is based on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}, which has been demonstrated to be able to learn to play a wide range of video games and commonly outperform human players.
-Its use of a neural network enable it to even learn complex tasks.
+Its use of a neural network enables it to even learn complex tasks.
 In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for playing video games and also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017}.
@@
 
-To compute $R_t$ at time step $t$ it is necessary to look infinitely far into the future to get all future rewards as in each step $t$ to compute $R_t$ one needs $R_{t+1}$.
-Thus, one usually only carries out $t_\text{max}$ steps (for example 20), waits for all the rewards and then calculates $R_t$ for each step and uses an estimate of $R_{t_\text{max}+1}$ provided by the critic network as the continuation in the future.
+To compute $R_t$ at time step $t$ it is necessary to look infinitely far into the future to get all future rewards as in each step $t$ computing $R_{t+1}$ relies on knowing $R_{t+1}$.
+Thus, the algorithm takes $t_\text{max}$ steps (for example 20), waits for all the rewards and then calculates $R_t$ for each step and uses an estimate of $R_{t_\text{max}+1}$ provided by the critic network as the continuation in the future.
 The only exception from this is when the task to execute ends before reaching $t_\text{max}$ (e.g.
 the TCP flow is over).
-In this case, one can either use a prediction from the critic network as a continuation or alternatively let the moving average end without continuation i.e.~ignoring the term weighted with $(\gamma-1)$ at the last step.
+In this case, one can either use a prediction from the critic network as a continuation or alternatively let the moving average end without continuation i.e.~ignore the term weighted with $(\gamma-1)$ at the last step.
 
@@
 
-The Actor Network outputs a normal distribution of the possible values for the action parametrized by a mean $\mu$ and a standard deviation $\sigma$ from which the action $a_t$ at time step $t$ is randomly sampled.
-The main idea is that the neural network learns to output the right mean of the actions at the right time step to maximize the future reward and that it uses the standard deviation to try out new actions, which could yield a better than expected reward.
+The Actor Network outputs a normal distribution of possible values for the action parametrized by a mean $\mu$ and a standard deviation $\sigma$.
+From this distribution each action $a_t$ at time step $t$ is randomly sampled.
+The main idea is that the neural network learns to output the right mean of the actions at the right time step to maximize future rewards and that it uses the standard deviation to try out new actions, which could yield a better than expected reward.
 
@@
 
-The actor also aims at increasing the probability density function's entropy ($H\left( \pi\left( s_t; \theta \right)\right) = \frac{1}{2} \log \left( 2 \pi e \sigma^2 \right)$) to encourage exploration: Otherwise the actor could always perform similar actions but never actually take the action that yields the best performance.
+The actor also aims at increasing the probability density function's entropy ($H( \pi( s_t; \theta )) = \frac{1}{2} \log ( 2 \pi e \sigma^2 )$) to encourage exploration: Otherwise the actor could always perform similar actions but never actually take the action that yields the best performance.
 The actor network will aim to maximize the entropy and at the same time output the best possible action: If it realizes that the entropy is too large and bad actions are taken, it learns to lower the entropy.
@@
 In the following evaluation we use a classical dumbbell network topology with one receiver and one or more senders and a shared bottleneck link.
-As our evaluation focuses on bulk transfers we use equally sized packets of 1250 bytes (TCP and IP including headers).
+As our evaluation focuses on bulk transfers we use equally sized packets of 1250 bytes (including TCP and IP headers).
 In all figures in the evaluation we use data from at least 50 simulations unless stated otherwise.
@@
 \caption{One sender with varying throughput/packet loss tradeoff on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
-Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
+Median in dark color and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in a light color.}
 \label{fig:one_sender_loss}
@@
 \caption{Two (\ref{fig:multiple_senders1}) and four (\ref{fig:multiple_senders2}) senders with a bandwidth/packet loss tradeoff of 2 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
-Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
+Median in dark color and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in a light color.}
 \label{fig:multiple_senders}
@@
 While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations and training time needed to achieve good performance increase when having two or four concurrent senders (see \autoref{fig:multiple_senders}).
-We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which it can use for training.
+We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which it can use for training and therefore training takes longer.
 Furthermore, as the environment becomes more dynamic with more senders, each sender has a harder time learning whether the reward changed due to his own actions or due to actions that were caused by other senders: If there is only one flow on a link, the flow ``can be sure'' that all changes in rewards were actually caused by his own actions.
@@
 \begin{minipage}{\columnwidth}
-\subfloat[One sender, initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4\label{fig:resuming_training1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4_again}}}}{}
-\subfloat[Two senders, initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4\label{fig:resuming_training2}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4_again}}}}{}
+\subfloat[One sender initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4\label{fig:resuming_training1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-4.local/figures/1_4_again}}}}{}
+\subfloat[Two senders initialized with a neural network from a previous flow with a throughput/packet loss tradeoff of 4\label{fig:resuming_training2}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-6.local/figures/2_4_again}}}}{}
 \caption{One (\ref{fig:resuming_training1}) and two (\ref{fig:resuming_training2}) senders that were trained with a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
-Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.}
+Median in dark color and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in a light color.}
 \label{fig:resuming_training}
@@
 
-To verify whether pre-training is generally feasible, we used the neural networks that were produced after each 600 seconds flow shown in figure \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks.
+To verify whether pre-training is generally feasible, we used the neural networks that were produced after each 600 second flow shown in \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks.
 The flows seem to simply resume where they ended training in the flow before (see \autoref{fig:resuming_training}).
@@
 For two senders pre-training also greatly speeds up the initial training procedure and can avoid the spike of packet loss in the beginning that always occurs to some extent without pre-training (see \autoref{fig:one_sender_loss} and \autoref{fig:multiple_senders}).
-However, it also becomes apparent that variance is large when comparing to the non-pre-trained version.
-We argue that the reason for this is that one sender rapidly claims all bandwidth in the very beginning leaving the other senders with nothing.
+Pre-training also works for more than one flow and although TCP Rax learns to minimize packet loss even further (see Figure \ref{fig:multiple_senders2}) some variability in throughput remains.
+We argue that this is caused by the stochastic behavior of the reinforcement learning: While each sender aims to maximize its utility, at the same time it also tries out new actions to check whether they could improve its utility even further and this behavior results in fluctuations.
 
@@
 \begin{minipage}{\columnwidth}
-\subfloat[One sender sharing a link with a stochastic loss rate of 1\% with a throughput/packet loss tradeoff of 4\label{fig:stochastic_loss1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-5.local/figures/1_4}}}}{}
+\subfloat[One sender on a link with a stochastic loss rate of 1\% with a throughput/packet loss tradeoff of 4\label{fig:stochastic_loss1}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-5.local/figures/1_4}}}}{}
 \subfloat[Two senders sharing a link with a stochastic loss rate of 1\% with a throughput/packet loss tradeoff of 4\label{fig:stochastic_loss2}]{\includegraphics[width=\columnwidth]{{{/home/max/repos/remy/compute-0-7.local/figures/2_4}}}}{}
 \caption{One (\ref{fig:stochastic_loss1}) and two (\ref{fig:stochastic_loss2}) senders on a link with a stochastic loss rate of 1\%, a bandwidth/packet loss tradeoff of 4 on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
-Median in black and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in gray.
-Apparently there is no stochastic loss in the very beginning of the flows, however, this is an artifact of the binning procedure (we use bins of one second): As the window is very small in the beginning, there are many flows in which no packet loss occurs whatsoever in the first bins and thus the median is distorted.}
+Median in dark color and $\frac{1}{4}$ and $\frac{3}{4}$ quantile band in a light color.
+Apparently there is no stochastic loss in the very beginning of the flows, however, this is an artifact of the binning procedure (we use bins of one second): As the window is very small in the beginning, there are many flows in which no packet loss occurs whatsoever in the first bins and thus the median is distorted towards 0.}
 \label{fig:stochastic_loss}
@@
 
-We also verify whether stochastic loss has an influence on TCP Rax.
+We also verify whether stochastic packet loss on the link has an influence on TCP Rax.
 For this, we randomly dropped 1\% of packets on the link and it seems as if stochastic loss does not substantially affect TCP Rax.
@@
 ]{\includegraphics[width=\columnwidth]{{{/home/max/repos/ns-allinone-2.35/ns-2.35/tcl/ex/congctrl/regular-tcp/nam_traces/figures/all_together_2_0}}}}{}
-\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one (\ref{fig:regular_tcps1}) and two (\ref{fig:regular_tcps2})senders of TCP Cubic/Compound/New Reno and Rax with a throughput/packet loss tradeoff of 4 without pre-training on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
+\caption{Total throughput and loss rate ($\frac{\text{all packets lost}}{\text{all packets sent}}$) for one (\ref{fig:regular_tcps1}) and two (\ref{fig:regular_tcps2}) senders of TCP Cubic/Compound/New Reno and Rax with a throughput/packet loss tradeoff of 4 without pre-training on a link with 20\,Mbit/s bottleneck speed and 50\,ms RTT and a $\frac{1}{10}$ bandwidth delay buffer.
 For a flow of 600\,s the maximum possible throughput is 12\,000\,Mbit.
@@
 
-Our results demonstrate that our proposed concept can perform online learning of congestion control and can outperform existing congestion control algorithms in certain scenarios, although a larger number of senders make the reinforcement learning significantly more challenging and time consuming.
-Regarding online learning, during the implementation of TCP Rax we came to the conclusion that a neural network consisting of LSTM cells \cite{gers_learning_1999} is not feasible.
+Our results demonstrate that our proposed concept can perform online learning of congestion control and can outperform existing congestion control algorithms in certain scenarios.
+Our results also show that with an increasing number of senders the reinforcement learning becomes significantly more challenging and time consuming.
+Regarding online learning, during the implementation of TCP Rax it became apparent that a neural network consisting of LSTM cells \cite{gers_learning_1999} is not feasible.
 LSTM cells keep an internal state that gets unstable when it is used for long training episodes, as for each received acknowledgement the LSTM state has to be updated.
-Previous work using deep reinforcement learning uses up to 50 million training steps for offline learning \cite{mnih_asynchronous_2016} while we generally use significantly less (a 20 Mbit\,s link can carry a maximum of up to one million packets of 1250 bytes size in 10 minutes).
+Previous work using deep reinforcement learning uses up to 50 million training steps for offline learning \cite{mnih_asynchronous_2016} while we generally use significantly less (a 20 Mbit/s link can carry a maximum of up to one million packets of 1250 bytes size in 10 minutes).
 When sharing the link between several senders, each sender receives only a fraction of these packets for training.
@@
 
-In this paper we show that it is possible to use deep reinforcement learning to learn a per-acknowledgement congestion control online.
+%In this paper we show that it is possible to use deep reinforcement learning to learn a per acknowledgement congestion control online.
 To this end we had to modify existing reinforcement learning algorithms to accommodate the specific needs of congestion control such as the inherent delay and proposed Partial Action Learning, a framework that allows deep reinforcement learning with delayed and partial rewards.
 
-We developed TCP Rax and the results of our evaluation indicate that congestion control can be learned online and has the potential to outperform standard TCP variants for significantly large flows.
-We came to the insight that it is possible to pre-train the neural network used in the reinforcement learning and that this improves performance especially in the beginning of the flow.
+In this paper we show that it is possible to use deep reinforcement learning to learn a per acknowledgement congestion control strategy online.
+In order to achieve this we had to modify existing reinforcement learning algorithms to accommodate the specific needs of congestion control such as the inherent delay and proposed Partial Action Learning (PAL), a new method that allows deep reinforcement learning with delayed and partial rewards.
+
+We developed TCP Rax and the results of our evaluation show that congestion control can be learned online and that it can outperform standard TCP variants for large flows.
+With TCP Rax it is possible to pre-train the neural network used in the reinforcement learning and this improves performance especially in the beginning of the flow.
 However, TCP Rax can also compete with other congestion control variants after an initial exploration phase without pre-training.

### 2018-02-03 async_deep_reinforce/a3c_training_thread.py

@@ -14,7 +14,7 @@ from functools import reduce
 from game_ac_network import GameACLSTMNetwork
 
 # gamma_current, gamma_future = 1./(1+GAMMA), GAMMA/(1.+GAMMA)
-# from constants import LOCAL_T_MAX
+from constants import LOCAL_T_MAX
 # from constants import GAMMA
 from constants import LOG_LEVEL
 from constants import STATE_SIZE
@@ -313,12 +313,12 @@ class A3CTrainingThread(object):
     assert(len(states) > 0)
     assert(len(rewards) > 0)
     assert(len(values) > 0)
-    if not (len(actions) == len(ticknos) == len(windows) == len(states) == len(rewards) == len(values)):
-      print(len(self.actions), len(self.ticknos), len(self.windows), len(self.states), len(self.rewards), len(self.values))
-      print(len(actions), len(ticknos), len(windows), len(states), len(rewards), len(values))
-      print(self.actions, self.ticknos, self.windows, self.states, self.rewards, self.values)
-      print(actions, ticknos, windows, states, rewards, values)
-    assert(len(actions) == len(ticknos) == len(windows) == len(states) == len(rewards) == len(values))
+    if not (len(actions) == len(ticknos) == len(windows) == len(states) == len(values)):
+      print(len(self.actions), len(self.ticknos), len(self.windows), len(self.states), len(self.values))
+      print(len(actions), len(ticknos), len(windows), len(states), len(values))
+      print(self.actions, self.ticknos, self.windows, self.states, self.values)
+      print(actions, ticknos, windows, states, values)
+     assert(len(actions) == len(ticknos) == len(windows) == len(states) == len(values))
 
     # if not len(self.actions) == len(self.ticknos) == len(self.windows) == len(self.states) == len(self.values) == len(self.estimated_values):
     #   print("In thread:", self.thread_index, "rewards:", len(self.rewards), ";", len(self.actions), len(self.ticknos), len(self.windows), len(self.states), len(self.values), len(self.estimated_values))
@@ -370,9 +370,9 @@ class A3CTrainingThread(object):
       # assert(False)
       # The GAMMA_FACTOR increases the influence that following observations have on this one.
 
-      GAMMA = (1 - 2/(A3CTrainingThread.get_actual_window(wi+ai) + 1))
[…]

### 2018-06-21 tex/paper.tex
Final version

@@
-\documentclass[newfonts=false,format=sigconf,10pt,letterpaper]{acmart}
+\documentclass[10pt,sigconf,anonymous]{acmart}
 
@@
 
-% Copyright
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
-\setcopyright{none}
+%% Copyright
+%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
+%\setcopyright{none}
+
+%Conference Info
+\acmYear{2018}
+\copyrightyear{2018}
+\setcopyright{acmcopyright}
+\acmConference{CoNEXT '18}{December 4--7, 2018}{Heraklion/Crete, Greece}
+\acmPrice{TBA}
+\acmDOI{TBA}
+\acmISBN{TBA}
 
@@
 \begin{document}
-\title{TCP Rax: Deep Reinforcement Learning for Congestion Control}
+\title{Rax: Deep Reinforcement Learning for Congestion Control}
 %\titlenote{Produces the permission block, and
@@
 
-A major challenge in TCP congestion control is to find a set of rules that dynamically adapt the congestion window to current and anticipated future network conditions to maximize a specified utility function, e.g.
+A major challenge in congestion control is to find a set of rules that dynamically adapt the congestion window to current and anticipated future network conditions to maximize a specified utility function, e.g.
 maximum throughput or minimum packet loss.
 
-This paper proposes Reactive Adaptive eXperience based congestion control (TCP Rax), a new method of congestion control that uses online reinforcement learning to maintain an optimum congestion window with respect to a given utility function and based on current network conditions.
+This paper proposes Reactive Adaptive eXperience based congestion control (Rax), a new method of congestion control that uses online reinforcement learning to maintain an optimum congestion window with respect to a given utility function and based on current network conditions.
 We use a neural network based approach that can be initialized either with random weights (without prior knowledge) or with a previously trained neural network to improve stability and convergence time.
-The processing of rewards in congestion control depends on the arrival of acknowledgments, inducing loop delays that may lead to oscillations.
-As a remedy we propose \textit{Partial Action Learning} (PAL), a specific formulation of neural network based reinforcement learning that supports the processing of delayed and partial rewards.
+As the processing of rewards in congestion control depends on the arrival of acknowledgments, which are delayed and come piece by piece, the problem is not suitable for current implementations of deep reinforcement learning.
+As a remedy we propose \textit{Partial Action Learning} (PAL), a specific formulation of deep reinforcement learning that supports the processing of delayed and partial rewards.
 
 We show that our method converges to a stable, close-to-optimum solution within the order of minutes and outperforms existing congestion control algorithms in typical networks.
-Thus, this paper demonstrates that neural network based reinforcement learning without any prior knowledge can feasibly be done online and can compete with human-designed congestion-control schemes such as New Reno and Cubic.
+Thus, this paper demonstrates that neural network based reinforcement learning can be done online and can compete with human-designed congestion-control schemes such as New Reno and Cubic.
 
@@
 
-Recent advances in neural network based reinforcement learning (RL) that for the first time have the ability to learn complex tasks such as video games, raise the question if it is possible to use these frameworks to model end-to-end congestion control: As in RL, in congestion control each sender has a an observable state, performs actions and receives rewards for its actions according to the sender's objective function.
+\begin{figure}
+
+\input{concept.tex}
+
+\caption{Hybrid learning approach to congestion control: After initial training in a simulator (Offline Learning), the congestion controller is deployed in a real network where it performs congestion control while at the same time still learning ways to improve (Online Learning).
+After sufficient training it is also possible to disable the online learning and use the previously learned knowledge (Static Operation).
+This can help reduce computational resources.}
+\label{fig:concept}
+\end{figure}
+
+Congestion control has been a long-standing problem in networking.
+In the past decades various attempts have been made to devise a set of rules that performs optimum congestion control.
+More recently, researchers have realized that such a set of rules might be impossible to find due to the complexity of the current Internet (caused by various types middleboxes, different congestion control algorithms and Active Queue Management) and the fast-paced evolution of Internet technologies.
+Two principal solutions have been proposed recently: \begin{enumerate*}
+\item Remy, a machine learning approach that is trained offline in a simulator and then statically deployed \cite{winstein_tcp_2013}, and \item PCC, which learns online but forgets all it learned after each flow \cite{dong_pcc:_2015} \end{enumerate*}.
+We argue that the solution lies in a hybrid approach: A machine learning based congestion control that receives some initial pre-training and then uses online learning to constantly refine and optimize its behavior (\autoref{fig:concept}).
+
+Recent advances in neural network based reinforcement learning (RL), which for the first time have the ability to learn complex tasks such as video games, seem to be an ideal tool to model end-to-end congestion control: As in RL, in congestion control each sender has a an observable state, performs actions and receives rewards for its actions according to the sender's objective function.
 
@@
 
-To address these problems, we develop a new formulation of reinforcement learning called \textit{Partial Action Learning} (PAL).
-PAL is a superset of reinforcement learning: If one uses PAL for a learning problem without delay, asynchronicity and partial actions/partial rewards, one gets classical reinforcement learning.
+To address these problems, we develop a new implementation of deep reinforcement learning called \textit{Partial Action Learning} (PAL) which we base on the Asynchronous Advantage Actor Critic framework \cite{mnih_asynchronous_2016}.
 
@@
 \caption{Throughput and loss rate when sharing a bottleneck link of 20\,Mbit/s with a two-way end-to-end delay of 50\,ms and a very small buffer of $\frac{1}{10}$ bandwidth delay product between one (\ref{fig:demo1}) and two (\ref{fig:demo2}) senders .
-The time needed for convergence increases with more senders, which is caused by the fact that the more senders use the link, the fewer packets each sender receives per unit of time for training and because of the increasing complexity of the environment.
+The time needed for convergence increases with more senders, because the more senders use the link, the fewer packets each sender receives per unit of time for training and because of the increasing complexity of the environment.
 In case of two senders, both flows complement each other and quickly adapt to the sending rate of the other flow.}
@@
 
-Congestion control has been implemented in TCP since the 1980s as a series of congestion collapses dramatically decreased applications' throughput in the Internet, which lead to the development of the Tahoe and Reno algorithms \cite{jacobson_congestion_1988}.
-These approaches as well as most others maintain a congestion window which stands for the maximum amount of data that are allowed to be in flight in the network without having been acknowledged by the receiver.
+Early congestion control algorithms such as Tahoe and Reno \cite{jacobson_congestion_1988} pioneered the idea to maintain a congestion window which stands for the maximum amount of data that are allowed to be in flight in the network without having been acknowledged by the receiver.
 In most TCP congestion control algorithms, receiving acknowledgements of data that were successfully transmitted leads to the congestion window being increased while packet loss (or an increase in delay as in the Vegas algorithm \cite{brakmo_tcp_1995}) leads to the congestion window being lowered according to a set of fixed rules.
 Algorithms such as Cubic and Compound \cite{ha_cubic:_2008, tan_compound_2006} improve congestion control specifically with respect to so-called long fat networks with a high round-trip time and bandwidth.
-Proportional rate reduction \cite{dukkipati_proportional_2011} aims to make the reduction of the rate that occurs in case of congestion more smooth and steady in time to avoid the bursty behavior that occurred upon loss in previous TCP congestion control variants.
 A different approach is taken by BBR \cite{cardwell_bbr:_2016}, which does not use a set of fixed rules like previous congestion control algorithms.
@@
 Its use of a neural network enables it to even learn complex tasks.
-In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for playing video games and also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017}.
+In particular, it is a good choice for congestion control as it has been proven to deliver good performance and high training speed for not only playing video games but also for networking related tasks such as Adaptive Bitrate Algorithms \cite{mao_neural_2017} and Traffic Engineering \cite{xu_experience-driven_2018}.
 
@@
 
-Our proposed framework \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline (with a focus on intuition over rigor) in this section.
+Actor Critic methods of reinforcement learning were first studied by \cite{sutton_policy_1999} and recently became popular for deep learning.
+Our proposed framework \textit{Partial Action Learning} (PAL) (see \ref{subsec:pal}) is based on the Actor Critic framework for neural networks proposed by \citet{mnih_asynchronous_2016}, which we outline in this section.
 
@@
 the TCP flow is over).
-In this case, one can either use a prediction from the critic network as a continuation or alternatively let the moving average end without continuation i.e.~ignore the term weighted with $(\gamma-1)$ at the last step.
+In this case, one can let the moving average end without continuation i.e.~ignore the term weighted with $(\gamma-1)$ at the last step.
 
@@
 
-The key difference between our approach PAL and previous approaches to RL is that in classical RL an action is always followed by a reward and a reward is always followed by an action.
-In our proposed concept, however, it is possible to take new actions while previous actions haven't been rewarded yet.
-
+The key difference between our approach PAL and previous implementations of deep RL is that in our approach we can handle the delay of rewards.
 Another major difference in PAL is that one action generates a number of partial actions ($\geq 0$) (see \autoref{fig:pal}).
@@
 Upon receiving feedback for a partial action, the agent determines the current state and performs a new action.
-When all feedbacks of one action were received, the agent combines them to form the reward and updates the critic and actor networks.
+When all feedback of one action was received, the agent combines it to form the reward and updates the critic and actor networks.
 
 In \autoref{alg:pal} (in the appendix) we show the code that runs in each of the agents (in the congestion control scenario, one agent corresponds to a sender).
-It is possible to have several agents which share a set of neural weights (which is usually done when performing offline training \cite{mnih_asynchronous_2016,mao_neural_2017}) but one can also use separate weights for each agent, which is more realistic in case of congestion control in the Internet, as it is not sensible that different senders share a set of neural network weights over the Internet.
+It is possible to have several agents which share a set of neural network weights (which is usually done when performing offline training \cite{mnih_asynchronous_2016,mao_neural_2017}) but one can also use separate weights for each agent, which is more realistic in case of congestion control in the Internet, as it is not sensible that different senders share a set of neural network weights over the Internet.
 
@@
 
-We use a reward function that is similar to the one used by PCC \cite{dong_pcc:_2015}.
+We use a reward function that is similar to the one used by PCC, whose convergence for multiple senders has been proven previously \cite{dong_pcc:_2015}.
 
@@
 
-The tunable tradeoff between bandwidth and packet loss rate ($\alpha$) has a clear impact on both the highest spike of packet loss that is produced in the very beginning of the online learning process as well as on the median packet loss rate later on in the learning process (see \autoref{fig:one_sender_loss}).
+The tunable tradeoff between bandwidth and packet loss rate ($\alpha$) has a clear impact on both the highest spike of packet loss that is produced in the very beginning of the online learning process as well as on the median packet loss rate later on in the learning process (\autoref{fig:one_sender_loss}).
 However, the stronger packet loss is punished, the longer it takes for the sender to reach the maximum bandwidth on the link.
@@
 
-While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations and training time needed to achieve good performance increase when having two or four concurrent senders (see \autoref{fig:multiple_senders}).
+While TCP Rax generally does not exhibit much variance after a few minutes of training when it is the only flow on a link, fluctuations and training time needed to achieve good performance increase when having two or four concurrent senders (\autoref{fig:multiple_senders}).
 We attribute this to the fact that on a shared link with $n$ senders each can only get $\frac{1}{n}$ of the bandwidth on average and thus each sender receives fewer packets which it can use for training and therefore training takes longer.
@@
 To verify whether pre-training is generally feasible, we used the neural networks that were produced after each 600 second flow shown in \autoref{fig:multiple_senders} and started flows of equal length with the pre-trained neural networks.
-The flows seem to simply resume where they ended training in the flow before (see \autoref{fig:resuming_training}).
+The flows seem to simply resume where they ended training in the flow before (\autoref{fig:resuming_training}).
 This does not seem like a remarkable finding, however, it is quite unexpected: For instance, in the case of one sender on a link, the neural network starts with random weights and discovers that it achieves optimum utility at a window that results in 20\,Mbit/s of bandwidth.
@@
 Thus, it is interesting that after hundreds of seconds of only seeing one same state, the neural network still acts correctly when starting with a small congestion window again in a completely different state than the one it was used to for hundreds of seconds before.
-For two senders pre-training also greatly speeds up the initial training procedure and can avoid the spike of packet loss in the beginning that always occurs to some extent without pre-training (see \autoref{fig:one_sender_loss} and \autoref{fig:multiple_senders}).
-Pre-training also works for more than one flow and although TCP Rax learns to minimize packet loss even further (see Figure \ref{fig:multiple_senders2}) some variability in throughput remains.
+For two senders pre-training also greatly speeds up the initial training procedure and can avoid the spike of packet loss in the beginning that always occurs to some extent without pre-training (\autoref{fig:one_sender_loss} and \autoref{fig:multiple_senders}).
+Pre-training also works for more than one flow and although TCP Rax learns to minimize packet loss even further (Figure \ref{fig:multiple_senders2}) some variability in throughput remains.
 We argue that this is caused by the stochastic behavior of the reinforcement learning: While each sender aims to maximize its utility, at the same time it also tries out new actions to check whether they could improve its utility even further and this behavior results in fluctuations.
@@
 For reasons of comparability we take the same scenario of a 20\,Mbit/s shared bottleneck with a small buffer of $\frac{1}{10}$ bandwidth delay product to create a challenging scenario.
-Surprisingly, New Reno performs by far better than more modern TCP congestion control algorithms (see \autoref{fig:regular_tcps}).
+Surprisingly, New Reno performs by far better than more modern TCP congestion control algorithms (\autoref{fig:regular_tcps}).
 We attribute this to the fact that we chose a small buffer to create a challenging scenario and that Cubic and Compound are not optimized so well for this scenario but rather for long fat networks \cite{oura_fairness_2012}.
-TCP Rax achieves a significantly larger total throughput on average though with a couple of outliers which become more significant with two concurrent senders (see \autoref{fig:rax_tcps}); when inspecting the raw data, it becomes apparent that these outliers occur usually when the neural network diverges in the first seconds of a flow.
+TCP Rax achieves a significantly larger total throughput on average though with a couple of outliers which become more significant with two concurrent senders (\autoref{fig:rax_tcps}); when inspecting the raw data, it becomes apparent that these outliers occur usually when the neural network diverges in the first seconds of a flow.
 
@@
 When sharing the link between several senders, each sender receives only a fraction of these packets for training.
-For example, with two senders on one link (see Figure \ref{fig:multiple_senders2}) one sender can receive a maximum of 600\,000 packets during a ten minute flow for training.
+For example, with two senders on one link (Figure \ref{fig:multiple_senders2}) one sender can receive a maximum of 600\,000 packets during a ten minute flow for training.
 We argue that the reason for TCP Rax needing fewer training samples than other reinforcement learning tasks is that in congestion control rewards are steadily coming in, while for other tasks, such as video games, commonly rewards are only received sparsely (e.g.
@@
 
-\begin{algorithm}
-\caption{Partial Action Learning -- procedure which computes and applies the gradients.}
-\label{alg:grad}
-\begin{algorithmic}[1]%
-%\algrestore{palalg}
-\Function{computeGradients}{}
-	\State $t_\text{end} = \#\left( l_\text{rewards} \right)$
-	\State $\theta_\text{backup} \gets \theta$
-	\State $h_\text{backup} \gets h$
-	\State $\theta \gets l_\text{snapshots}[0]$
-	\State $h \gets l_\text{hiddenStates}[0]$
-	\State $R_{i+1} \gets$ last element of $l_\text{values'}$%
-	\For{$i\gets t_\text{end}-1..0$}
-		\State $R_i \gets \left(r_i + \gamma R_{i+1}\right)\left(1-\gamma\right)$
-		\State $a \gets l_\text{actions}[i]$
-		\State $s \gets l_\text{states}[i]$
-		\State $v \gets l_\text{values}[i]$
-		\State $d\theta \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta}$
-		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta}$
-		\StateIndent[4] $+\frac{\partial\left(R_i - v\right)}{\partial\theta}$
-		\State $\theta_{\text{g}} \gets \theta_{\text{g}} + d\theta$
-	\EndFor
-	\State Remove first $t_\text{end}$ elements from
-	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{values'}$
-	\State Remove the first element from $l_\text{snapshots}$, $l_\text{hiddenStates}$
-	\State $\theta = \theta_\text{backup}$
-	\State $h = h_\text{backup}$
-\EndFunction
-\end{algorithmic}
-\end{algorithm}
+%\begin{algorithm}
+%\caption{Partial Action Learning -- procedure which computes and applies the gradients.}
+%\label{alg:grad}
+%\begin{algorithmic}[1]%
+%%\algrestore{palalg}
+%\Function{computeGradients}{}
+%	\State $t_\text{end} = \#\left( l_\text{rewards} \right)$
+%	\State $\theta_\text{backup} \gets \theta$
+%	\State $h_\text{backup} \gets h$
+%	\State $\theta \gets l_\text{snapshots}[0]$
+%	\State $h \gets l_\text{hiddenStates}[0]$
+%	\State $R_{i+1} \gets$ last element of $l_\text{values'}$%
+%	\For{$i\gets t_\text{end}-1..0$}
+%		\State $R_i \gets \left(r_i + \gamma R_{i+1}\right)\left(1-\gamma\right)$
+%		\State $a \gets l_\text{actions}[i]$
+%		\State $s \gets l_\text{states}[i]$
+%		\State $v \gets l_\text{values}[i]$
+%		\State $d\theta \gets -\frac{\partial\log \left( \pi\left( a \given s; \theta_\text{a} \right)\right)\left(R_i - v\right)}{\partial\theta}$
+%		\StateIndent[4] $-\frac{\partial\beta H\left( \pi\left( s; \theta_a \right)\right)}{\partial\theta}$
+%		\StateIndent[4] $+\frac{\partial\left(R_i - v\right)}{\partial\theta}$
+%		\State $\theta_{\text{g}} \gets \theta_{\text{g}} + d\theta$
+%	\EndFor
+%	\State Remove first $t_\text{end}$ elements from
+%	\StateIndent[2] $l_\text{actions}$, $l_\text{states}$, $l_\text{values}$, $l_\text{rewards}$, $l_\text{values'}$
+%	\State Remove the first element from $l_\text{snapshots}$, $l_\text{hiddenStates}$
+%	\State $\theta = \theta_\text{backup}$
+%	\State $h = h_\text{backup}$
+%\EndFunction
+%\end{algorithmic}
+%\end{algorithm}
 
@@
 
-\begin{algorithm}
-\caption{Procedure which determines when to start computing the gradients in case of congestion control.
+%\begin{algorithm}
+%\caption{Procedure which determines when to start computing the gradients in case of congestion control.
 We assume that the window at time steps $t$ is stored in a list $l_\text{windows}$ and that the number of bytes sent in each time step is stored in $l_\text{sent}$ analogously to the other metrics defined in \autoref{alg:pal}.
 Alternatively to this procedure, one can simply compute the gradients when $t_\text{max}$ rewards have been received where $t_\text{max}$ is an arbitrarily chosen integer greater than 0.}
-\label{alg:enough}
-\begin{algorithmic}[1]%
-%\algrestore{palalg}
-\Function{gotEnoughRewards?}{}
-	\If{the epsiode is over}
-		\State \Return \True
-	\EndIf
-	\State $w \gets \left\lfloor l_\text{windows}[0] + l_\text{actions}[0] \right\rfloor$
-	\For{$i\gets 0..\#(l_\text{rewards})-1$}
-		\State $w \gets \text{w} - l_\text{sent}[i]$
-		\If{$w \leq 0 \land i+1 \geq 10$} \Comment{at least 10}
-			\State \Return \True
-		\EndIf
-	\EndFor
-	\State \Return \False
-\EndFunction
-\end{algorithmic}
-\end{algorithm}
+%\label{alg:enough}
+%\begin{algorithmic}[1]%
+%%\algrestore{palalg}
+%\Function{gotEnoughRewards?}{}
+%	\If{the epsiode is over}
+%		\State \Return \True
+%	\EndIf
+%	\State $w \gets \left\lfloor l_\text{windows}[0] + l_\text{actions}[0] \right\rfloor$
+%	\For{$i\gets 0..\#(l_\text{rewards})-1$}
+%		\State $w \gets \text{w} - l_\text{sent}[i]$
+%		\If{$w \leq 0 \land i+1 \geq 10$} \Comment{at least 10}
+%			\State \Return \True
+%		\EndIf
+%	\EndFor
+%	\State \Return \False
+%\EndFunction
+%\end{algorithmic}
+%\end{algorithm}
 

### 2018-06-21 async_deep_reinforce/game_ac_network.py

@@ -27,6 +27,7 @@ import numpy.random
 import itertools
 
 tiny = 1e-10
+really_tiny = 1e-20
 
 from tensorflow.python.ops import variable_scope as vs
 
